### 2020-11-12 main.tex
Initial Overleaf Import

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,amsfonts,clevref,
 aps,
%pra,
%prb,
%rmp,
prstab,
%prstper,
%floatfix,
]{revtex4-2}
\usepackage{wasysym}
\usepackage{graphicx, subcaption}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage[allcolors=blue]{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

% \usepackage[active, tightpage]{preview}
% \setlength\PreviewBorder{0pt}
% \PreviewSnarfEnvironment[{[]}]{figure}

%\usepackage[active,tightpage,floats]{preview}

%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}

%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
% \usepackage{epic,eepic,pspicture,mathptmx,times,graphpap,siunitx, pgf, import}
% For algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[ruled,vlined]{algorithm2e}

\usepackage{siunitx, pgf, import}
\usepackage[super]{nth}
\newcommand{\shirinkimage}[2]{\resizebox{#1\textwidth}{!}{\input{#2}}}
\usepackage{cleveref}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning} 
\tikzset{
    %Define standard arrow tip
    >=stealth',
    %Define style for boxes
    punkt/.style={
           rectangle,
           rounded corners,
           draw=black, very thick,
           text width=3.5em,
           minimum height=2em,
           text centered},
    % Define arrow style
    pil/.style={
           ->,
           thick,
           shorten <=2pt,
           shorten >=2pt,}
}
%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}
\usepackage{wasysym}
%\usepackage[backend=biber,style=numeric,sorting=none]{biblatex}

\newcommand{\NB}[1]{\textcolor{red}{#1}}

\begin{document}

\preprint{APS/123-QED}

\title{Model based reinforcement learning at FERMI FEL}

\author{Simon Hirlaender}
%\thanks{These two authors contributed equally}
% \affiliation{University of Malta}
\affiliation{Univertsity of Salzburg, Salzburg, Austria}
\author{Niky Bruchon}%, Gianfranco Fenu, Felice Andrea Pellegrino, Erica Salvato}
\affiliation{University of Trieste, Trieste, IT}
% \author{Gianluca Valentino}
% \affiliation{University of Malta, Msida, MT}
% \author{Marco Lonza, Giulio Gaio}
% \affiliation{Elettra Sincrotrone Trieste, Trieste, IT}
% \author{Verena Kain}
% %\thanks{These two authors contributed equally}
% \email{verena.kain@cern.ch}

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
In this paper we discuss a model based reinforcement learning approach in comparison to a model free reinforcement learning approach applied at the FERMI FEL system. Both algorithms and approaches are new in this context and the main purpose of this paper is to show that reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems. In terms of sample-complexity the model-based approach is faster, while the finale performance of the model free method is superior - the so called asymptotic performance. The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency
\end{abstract}
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

%\tableofcontents
\section{Introduction and Motivation}
In particle accelerators one main goal is to provide stable and reproducible performance. In order to achieve this, a number of control problems have to be considered simultaneously. Especially if there is no way to model the physics, one might use optimisation techniques as e.g. derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes. Another promising way is to apply reinforcement learning which has several advantages over optimisation methods:
\begin{itemize}
    \item Covers a larger class of problems.
    \item Solves the problem not always completely from the beginning as an DFO.
    \item Ability to use existing data.
    \item The underlying structure of the problem might be deduced.
\end{itemize}
RL optimises a sequence of decisions. 
In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste. Three algorithms were applied successfully showing different advantages, which will e discussed in the following.

% The methods proposed are the iterative Linear Quadratic Regulator (iLQR) - an extension to non-linear systems of the fundamental control theory Linear Quadratic Regulator (LQR) problem -  and the Normalized Advantage Function (NAF) - a continuous variant of the Q-learning algorithm from reinforcement learning field. An exhaustive presentation of both techniques is provided in the following sections.

In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator. 
\section{The physical set-up}
At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation. Nevertheless, in the last years various approaches have been studied to investigate their applicability in tuning operations. 
%%% Last introduction paragraph
The paper is organised as follows. 
\begin{itemize}
    \item Description of the problem set-up at FERMI
    \item Overview of RL
    \item Details of the implementations used in these studies and theoretical concerns
    \item Results 
    \item Summary and outlook
\end{itemize}

\section{Deep Reinforcement learning}
Assume states $s$ the set of all states $S$ and actions $a$ all actions $a$ and rewards $r$ and the set of all rewards $R$ an initial state distribution $d_0$.
Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
\begin{align}
\max J(\pi)  =  \mathbb E_{\tau \thicksim p_{\pi}}\left[\sum_{t=0}^{H}\gamma^t r(s_t,a_t)\right],
\end{align}
where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \mathcal{B}^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \mathcal{B}^*\vec Q_\theta(s_t,a_t) = r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{NAF}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s),
\end{align}
the maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
     \min_\theta\left( (r(s_t,a_t)+\gamma V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
\end{align}
$\theta\prime$ denotes the weights of a previous iteration.
\subsection{New NAF network architecture}

\begin{algorithm}[ht]
\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
\begin{algorithmic}[1]
\State initialise $\theta_0$
\For{iteration $k \in [0, \dots, K]$}
\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
\EndFor
\end{algorithmic}
\end{algorithm}

\section{Dyna-style reinforcement learning}

\subsection{Operational deployment}
\NB{Niky} % come abbiamo interagito con FERMI
\subsection{\NB{short FERMI introduction}}
The Free Electron laser Radiation for Multidisciplinary Investigation  better known as the FERMI, is the seeded free-electron laser facility next to the third-generation synchrotron facility ELETTRA at the Elettra Sinctrotrone Trieste laboratories.
A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure. The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.

The external optical laser signal provide is contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy. The modulation in energy is the converted into a charge modulation in the dispersive section, and finally the density modulated beams radiation is amplified in the radiators section. The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.

For this reason the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.

\subsection{\NB{our system: the modulator and the seed laser}}
A more detailed description of the alignment process in the modulator is here provided.
The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
The final problem faced consists in optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.

\section{Proof-of-principle application of RL Agent at FERMI trajectory correction}
\NB{Niky}

\subsection{Experiment results from FERMI RL tests}
Niky and Simon

\section{Discussion and outlook}

\section{Conclusions}

\appendix
\section{NAF2 details}\label{appendix:pernaf}

\begin{figure}[!h]
  \centering
  \input{Figures/SL_Alignment_Scheme}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_verification.pdf}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}
\begin{figure}[!h]
  \centering
  \includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_episodes.pdf}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}
\begin{figure}[!h]
  \centering
  \includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
  \caption{The training .}
  \label{fig:comparsion_per}
\end{figure}

% The \nocite command causes all entries in a bibliography to be printed out
% whether or not they are actually referenced in the text. This is appropriate
% for the sample file to show the different styles of references, but authors
% most likely will not want to use it.
% \nocite{*}
% \bibliography{apssamp}% Produces the bibliography via BibTeX.

% \bibliography{bibliography}
\addcontentsline{toc}{chapter}{Bibliography}

\begin{thebibliography}{9}

\bibitem{lsa} D. Jacquet, R. Gorbonosov, G. Kruk, "LSA - The High Level Application Software of the LHC and its Performance during the first 3 years of Operation", ICALEPS, San Francisco, CA, USA, 6 - 11 Oct 2013, pp.thppc058.

\bibitem{multiobjective-optimization} A. Edelen, N. Neveu, M. Frey, Y. Huber, C. Mayes, and A. Adelmann, "Machine learning for orders of magnitude speedup in multiobjective optimization of particle accelerator systems", Phys. Rev. Accel. Beams 23, 044601, 2020.

\bibitem{collimator-alignment} G. Azzopardi, A. Muscat, G. Valentino, S. Redaelli, B. Salvachua, "Operational results of LHC collimator alignment using machine learning", Proc. IPAC'19, Melbourne, Australia, pp. 1208-1211, 2019.

\bibitem{SPSAlignment} %InProceedings (Hirlaender2019)
 S. Hirlaender, M. Fraser, B. Goddard , V. Kain, J. Prieto, L. Stoel, M. Szakaly, F. Velotti, "Automatisation of the SPS ElectroStatic Septa Alignment",  
in 10th Int. Partile Accelerator Conf.(IPAC'19), 2019, p. 4001-4004.

\bibitem{liu} E. Shaposhnikova et al., "LHC Injectors Upgrade (LIU) Project at CERN", 7$^{th}$ International Particle Accelerator Conference, Busan, Korea, 8 - 13 May 2016, pp.MOPOY059, \url{https://doi.org/10.18429/JACoW-IPAC2016-MOPOY059}.

\bibitem{awake} E. Adli, A. Ahuja, O. Apsimon, R. Apsimon, A.-M. Bachmann, D. Barrientos, F. Batsch, J. Bauche, V. B. Olsen, M. Bernardini \textit{et al.}, “Acceleration of electrons in the plasma wakefield of a proton bunch,” Nature 561, 363–367 (2018), \url{https://doi.org/10.1038/s41586-018-0485-4)}.

\bibitem{scheinker} A. Scheinker \textit{et al.}, "Online multi-objective particle accelerator optimization of the AWAKE electron beam line for simultaneous emittance and orbit control", AIP Advances 10, 055320 (2020), \url{https://doi.org/10.1063/5.0003423}.

\bibitem{barto-sutton} R. Sutton, A. Barto, "Introduction to Reinforcement Learning",  MIT Press, Cambridge, MA, USA, 2018.

\bibitem{LQR} H. Kwakernaak, R. Sivan, "Linear Optimal Control Systems", Wiley-Interscience, 1972.

\bibitem{SBP} A. Nagabandi, G. Kahn, R. S. Fearing, S. Levine, "Neural Network Dynamics for Model-Based Deep Reinforcement Learning with Model-Free Fine-Tuning", arXiv:1708.02596, 2017.

\bibitem{DYNA} R. Sutton, "Dyna, an integrated architecture for learning, planning, and reacting", AAAI Spring Symposium, pp. 151-155, 1991, \url{https://doi.org/10.1145/122344.122377}.

\bibitem{benchmarkmbrl} T. Wang et al., “Benchmarking Model-Based Reinforcement Learning”,  \url{https://arxiv.org/abs/1907.02057v1}, 2019-07-03.

\bibitem{dqn} V. Mnih \textit{et al.}, "Human-level control through deep reinforcement learning", Nature 518, 529–533 (2015).

\bibitem{ddpg} T. Lillicrap~\textit{et al.}, "Continuous control with deep reinforcement learning", Proc. ICLR 2016 \url{https://arxiv.org/abs/1509.02971}.

\bibitem{naf} S. Gu, T. Lillicrap, I. Sutskever, S. Levine, "Continuous Deep Q-Learning with Model-based Acceleration", in Proc. 33rd International Conference on Machine Learning, New York, NY, USA, 2016.

\bibitem{td3} S. Fujimoto, H. van Hoof, D. Meger, "Addressing Function Approximation Error in Actor-Critic Methods", arXiv:1802.09477, 2018.

\bibitem{deeppilco} Y. Gal, R. McAllister, and C. E. Rasmussen, “Improving PILCO with Bayesian neural network dynamics models”, in Data-Efficient Machine Learning workshop, ICML, 2016, vol. 4, p. 34.

\bibitem{per} T. Schaul, J. Quan, I. Antonoglou, D. Silver, "Prioritized Experience Replay", arXiv:1511.05952, 2015-11-18.

\bibitem{pernafgit} S. Hirlaender, \textit{PER-NAF} available at \url{https://github.com/MathPhysSim/PER-NAF/} and
\url{https://pypi.org/project/pernaf/}.
\bibitem{pyjapc} "pyjapc" available at \url{https://pypi.org/project/pyjapc/}.

\bibitem{openai_gym} \url{http://gym.openai.com}

\bibitem{madx} MAD-X documentation and source code available at \url{https://mad.web.cern.ch/mad/}.

\bibitem{linac4} G. Bellodi, "Linac4 Commissioning Status and Challenges to Nominal Operation", 61$^st$ ICFA Advanced Beam Dynamics Workshop on High-Intensity and High-Brightness Hadron Beams, Daejeon, Korea, 17 - 22 Jun 2018, pp.MOA1PL03, \url{https://doi.org/10.18429/JACoW-HB2018-MOA1PL03}.

\bibitem{auto_matching} F. Velotti, B. Goddard et al., "Automatic AWAKE electron beamline setup using unsupervised machine learning", to be submitted to Phys. Rev. Accel. Beams.

\bibitem{Kumar_1} A. Kumar, A. Gupta, S. Levine, “DisCor: Corrective Feedback in Reinforcement Learning via Distribution Correction,” arXiv:2003.07305, 2020-03-16.

\bibitem{Kumar_2} A. Kumar, J. Fu, G. Tucker, S. Levine, “Stabilizing Off-Policy Q-Learning via Bootstrapping Error Reduction,” arXiv:1906.00949, 2019-06-03.

% NEW REFERENCES
\bibitem[bruchon2020basic]{Bruchon, Niky, et al. "Basic Reinforcement Learning Techniques to Control the Intensity of a Seeded Free-Electron Laser." Electronics 9.5 (2020): 781.}
% https://www.mdpi.com/2079-9292/9/5/781/htm

\end{thebibliography}

%%%%% CLEAR DOUBLE PAGE!
\newpage{\pagestyle{empty}\cleardoublepage}

\end{document}
%
% ****** End of file apssamp.tex ******

### 2020-11-12 AE_dyna_results.py
- initial commit

@@ -1,0 +1,192 @@
+import os
+import pickle
+import matplotlib.pyplot as plt
+import numpy as np
+
+label = "ME-TRPO"
+label = "AE-DYNA"
+if label == "ME-TRPO":
+    # ME-TRPO results
+    project_directory = 'Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/run2/'
+else:
+    # AE-Dyna results
+    project_directory = 'Data_Experiments/2020_11_05_AE_Dyna@FERMI/-nr_steps_25-cr_lr-n_ep_13-m_bs_100-sim_steps_3000-m_iter_35-ensnr_3-init_200/'
+
+def read_rewards(rewards):
+    iterations_all = []
+    final_rews_all = []
+    mean_rews_all = []
+    stds = []
+
+
+
+    iterations = []
+    final_rews = []
+    mean_rews = []
+    for i in range(len(rewards)):
+        if len(rewards[i]) > 0:
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+            iterations.append(len(rewards[i]))
+            try:
+                mean_rews.append(np.sum(rewards[i][1:]))
+            except:
+                mean_rews.append([])
+            stds.append(np.std(rewards[i][1:]))
+
+    # iterations = np.mean(np.array(iterations_all), axis=0)
+    # final_rews = np.mean(np.array(final_rews_all), axis=0)
+    # mean_rews = np.mean(np.array(mean_rews_all), axis=0)
+
+    return np.array(iterations), np.array(final_rews), np.array(mean_rews), np.array(stds)
+
+
+
+def plot_results(data, label='Verification', **kwargs):
+        '''plotting'''
+        rewards = data['rews']
+        # iterations = []
+        # finals = []
+        # means = []
+        # stds = []
+        #
+        # for i in range(len(rewards)):
+        #     if (len(rewards[i]) > 1):
+        #         finals.append(rewards[i][-1])
+        #         means.append(np.mean(rewards[i][1:]))
+        #         stds.append(np.std(rewards[i][1:]))
+        #         iterations.append(len(rewards[i]))
+        #
+        # x = range(len(iterations))
+        # iterations = np.array(iterations)
+        # finals = np.array(finals)
+        # means = np.array(means)
+        # stds = np.array(stds)
+
+        iterations, finals, means, stds = read_rewards(rewards)
[…]

### 2020-11-12 AEdyna_clean.py

@@ -1,0 +1,1249 @@
+import os
+import pickle
+from datetime import datetime
+
+import gym
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+# from stable_baselines.common.noise import NormalActionNoise
+
+from stable_baselines.sac.policies import MlpPolicy
+# from stable_baselines.td3.policies import MlpPolicy
+# from stable_baselines.common.policies import MlpPolicy
+from stable_baselines import SAC as Agent
+# from stable_baselines import TD3 as Agent
+# from stable_baselines import PPO2 as Agent
+# from stable_baselines import TRPO as Agent
+import tensorflow as tf
+
+from local_fel_simulated_env import FelLocalEnv
+from simulated_tango import SimTangoConnection
+
+# from naf2 import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+config = tf.ConfigProto(
+    device_count={'GPU': 0}
+)
+# config = None
+
+tango = SimTangoConnection()
+real_env = FelLocalEnv(tango=tango)
+
+# Hyper papameters
+steps_per_env = 25
+init_random_steps = 100
+total_steps = 200
+num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
+
+print('Number of epochs: ', num_epochs)
+
+hidden_sizes = [100, 100]
+
+max_training_iterations = 20
+delay_before_convergence_check = 1
+
+algorithm = 'SAC'
+
+# minibatch_size = 100
+simulated_steps = 2500
+
+model_batch_size = 25
+num_ensemble_models = 5
+model_iter = 30
+
+model_training_iterations = 10
+network_size = 15
+# Set the priors for the anchor method:
+#  TODO: How to set these correctly?
+init_params = dict(init_stddev_1_w=np.sqrt(1),
+                   init_stddev_1_b=np.sqrt(1),
+                   init_stddev_2_w=1.0 / np.sqrt(network_size))
+
+data_noise = 0.000001  # estimated noise variance
+lambda_anchor = data_noise / (np.array([init_params['init_stddev_1_w'],
+                                        init_params['init_stddev_1_b'],
+                                        init_params['init_stddev_2_w']]) ** 2)
+
+# How often to check the progress of the network training
+# e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
[…]

### 2020-11-12 AEdyna_clean_tests.py

@@ -1,0 +1,1334 @@
+import os
+import pickle
+from datetime import datetime
+import itertools as it
+import gym
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+# from stable_baselines.common.noise import NormalActionNoise
+
+# from stable_baselines.sac.policies import MlpPolicy
+# from stable_baselines import SAC as Agent
+# from stable_baselines.td3.policies import MlpPolicy
+# from stable_baselines import TD3 as Agent
+from stable_baselines.common.policies import MlpPolicy
+from stable_baselines import PPO2 as Agent
+
+import tensorflow as tf
+
+from local_fel_simulated_env import FelLocalEnv
+from simulated_tango import SimTangoConnection
+
+# from naf2 import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+config = tf.ConfigProto(
+    device_count={'GPU': 0}
+)
+# config = None
+
+tango = SimTangoConnection()
+real_env = FelLocalEnv(tango=tango)
+
+# Hyper papameters
+steps_per_env = 25
+init_random_steps = 100
+total_steps = 250
+num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
+
+print('Number of epochs: ', num_epochs)
+
+hidden_sizes = [100, 100]
+
+max_training_iterations = 50
+delay_before_convergence_check = 1
+
+algorithm = 'SAC'
+
+# minibatch_size = 100
+simulated_steps = 3000
+
+model_batch_size = 5
+num_ensemble_models = 5
+
+early_stopping = True
+model_iter = 30
+
+# model_training_iterations = 10
+network_size = 15
+# Set the priors for the anchor method:
+#  TODO: How to set these correctly?
+init_params = dict(init_stddev_1_w=np.sqrt(1),
+                   init_stddev_1_b=np.sqrt(1),
+                   init_stddev_2_w=1 / np.sqrt(network_size))
+
+data_noise = 0.00000  # estimated noise variance
+lambda_anchor = data_noise / (np.array([init_params['init_stddev_1_w'],
+                                        init_params['init_stddev_1_b'],
+                                        init_params['init_stddev_2_w']]) ** 2)
+
+# How often to check the progress of the network training
+# e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
[…]

### 2020-11-12 AEdyna_versuch.py

@@ -1,0 +1,1763 @@
+import os
+import pickle
+from datetime import datetime
+
+import gym
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+# from stable_baselines.common.noise import NormalActionNoise
+
+from stable_baselines.sac.policies import MlpPolicy
+# from stable_baselines.td3.policies import MlpPolicy
+# from stable_baselines.common.policies import MlpPolicy
+from stable_baselines import SAC as Agent
+# from stable_baselines import TD3 as Agent
+# from stable_baselines import PPO2 as Agent
+# from stable_baselines import TRPO as Agent
+import tensorflow as tf
+
+from local_fel_simulated_env import FelLocalEnv
+from simulated_tango import SimTangoConnection
+
+# from naf2 import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+# tf.random.set_seed(random_seed)
+# random.seed(random_seed)
+config = tf.ConfigProto(
+    device_count={'GPU': 0}
+)
+# config = None
+
+tango = SimTangoConnection()
+real_env = FelLocalEnv(tango=tango)
+
+# Hyper papameters
+steps_per_env = 25
+init_random_steps = 100  # For niky: You can change the number of steps -------------------------------------------------------------------
+total_steps = 200
+num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
+
+print('Number of epochs: ', num_epochs)
+
+hidden_sizes = [100, 100]
+cr_lr = 1e-5
+gamma = 0.9999
+lam = 0.95
+
+max_training_iterations = 20
+# max_training_iterations = 10
+delay_before_convergence_check = 1
+
+number_envs = 1
+critic_iter = 80
+delta = 0.05
+algorithm = 'TRPO'
+conj_iters = 40
+minibatch_size = 100
+simulated_steps = 5000
+
+mb_lr = 1e-4
+model_batch_size = 5
+# For niky: You can change the number of models here -------------------------------------------------------------------
+num_ensemble_models = 5
+# ----------------------------------------------------------------------------------------------------------------------
+model_iter = 2
+
+model_training_iterations = 10
+lr_training = 1e-4
+network_size = 15
+# Set the priors for the anchor method:
+#  TODO: How to set these correctly?
+init_params = dict(init_stddev_1_w=np.sqrt(1),
+                   init_stddev_1_b=np.sqrt(1),
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/basic_test_script.py

@@ -1,0 +1,197 @@
+import pickle
+import random
+import random
+import sys
+
+import matplotlib.pyplot as plt
+import numpy as np
+import tensorflow as tf
+
+from local_fel_simulated_env import FelLocalEnv
+from pernaf.pernaf.naf import NAF
+from pernaf.pernaf.utils.statistic import Statistic
+
+# from pendulum import PendulumEnv as simpleEnv
+# set random seed
+from simulated_tango import SimTangoConnection
+
+random_seed = 123
+# set random seed
+tf.set_random_seed(random_seed)
+np.random.seed(random_seed)
+random.seed(random_seed)
+
+tango = SimTangoConnection()
+env = FelLocalEnv(tango=tango)
+
+directory = "checkpoints/test_implementation/"
+
+
+def plot_results(env, label):
+    # plotting
+    print('now plotting')
+    rewards = env.rewards
+    initial_states = env.initial_conditions
+
+    iterations = []
+    final_rews = []
+    # starts = []
+    sum_rews = []
+    mean_rews = []
+    # init_states = pd.read_pickle('/Users/shirlaen/PycharmProjects/DeepLearning/spinningup/Environments/initData')
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 0):
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+            # starts.append(-np.sqrt(np.mean(np.square(initial_states[i]))))
+            iterations.append(len(rewards[i]))
+            sum_rews.append(np.sum(rewards[i]))
+            mean_rews.append(np.mean(rewards[i]))
+    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
+
+    fig, axs = plt.subplots(2, 1, constrained_layout=True)
+
+    ax = axs[0]
+    ax.plot(iterations)
+    ax.set_title('Iterations' + plot_suffix)
+    fig.suptitle(label, fontsize=12)
+
+    ax = axs[1]
+    ax.plot(final_rews, 'r--')
+
+    ax.set_title('Final reward per episode')  # + plot_suffix)
+    ax.set_xlabel('Episodes (1)')
+
+    # ax1 = plt.twinx(ax)
+    # color = 'lime'
+    # ax1.set_ylabel('starts', color=color)  # we already handled the x-label with ax1
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/run_naf2.py

@@ -1,0 +1,172 @@
+import os
+import pickle
+import random
+import matplotlib.pyplot as plt
+import numpy as np
+import tensorflow as tf
+
+# from local_fel_simulated_env import FelLocalEnv
+from laser_trajectory_control_env import LaserTrajectoryControlEnv
+# from simulated_tango import SimTangoConnection
+from tango_connection import TangoConnection
+from naf2 import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+random.seed(random_seed)
+
+# tango = SimTangoConnection()
+# env = FelLocalEnv(tango=tango)
+conf_file = '/home/niky/FERMI/2020_07_20/configuration/conf_fel.json'
+tango = TangoConnection(conf_file=conf_file)
+env = LaserTrajectoryControlEnv(tango=tango)
+
+def plot_results(env, file_name):
+    # plotting
+    print('Now plotting')
+    rewards = env.rewards
+    initial_rewards = env.init_rewards
+    # print('initial_rewards :', initial_rewards)
+
+    iterations = []
+    final_rews = []
+    starts = []
+    sum_rews = []
+    mean_rews = []
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 0):
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+            starts.append(initial_rewards[i])
+            iterations.append(len(rewards[i]))
+            sum_rews.append(np.sum(rewards[i]))
+            mean_rews.append(np.mean(rewards[i]))
+    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
+
+    fig, axs = plt.subplots(2, 1)
+
+    ax = axs[0]
+    color = 'blue'
+    ax.plot(iterations, c=color)
+    ax.set_ylabel('steps', color=color)
+    ax.tick_params(axis='y', labelcolor=color)
+    ax1 = plt.twinx(ax)
+    color = 'k'
+    ax1.plot(np.cumsum(iterations), c=color)
+    ax1.set_ylabel('cumulative steps', color=color)
+    ax.set_title('Iterations' + plot_suffix)
+    # fig.suptitle(label, fontsize=12)
+
+    ax = axs[1]
+    color = 'red'
+    ax.plot(starts, c=color)
+    ax.set_ylabel('starts', color=color)
+    ax.axhline(-0.05, ls=':', color='r')
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/run_test_naf2.py

@@ -1,0 +1,172 @@
+import os
+import pickle
+import random
+import matplotlib.pyplot as plt
+import numpy as np
+import tensorflow as tf
+
+# from local_fel_simulated_env import FelLocalEnv
+from laser_trajectory_control_env import LaserTrajectoryControlEnv
+# from simulated_tango import SimTangoConnection
+from tango_connection import TangoConnection
+from naf2 import NAF
+
+# set random seed
+random_seed = 123
+np.random.seed(random_seed)
+random.seed(random_seed)
+
+# tango = SimTangoConnection()
+# env = FelLocalEnv(tango=tango)
+conf_file = '/home/niky/FERMI/2020_07_20/configuration/conf_fel.json'
+tango = TangoConnection(conf_file=conf_file)
+env = LaserTrajectoryControlEnv(tango=tango)
+
+def plot_results(env, file_name):
+    # plotting
+    print('Now plotting')
+    rewards = env.rewards
+    initial_rewards = env.init_rewards
+    # print('initial_rewards :', initial_rewards)
+
+    iterations = []
+    final_rews = []
+    starts = []
+    sum_rews = []
+    mean_rews = []
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 0):
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+            starts.append(initial_rewards[i])
+            iterations.append(len(rewards[i]))
+            sum_rews.append(np.sum(rewards[i]))
+            mean_rews.append(np.mean(rewards[i]))
+    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
+
+    fig, axs = plt.subplots(2, 1)
+
+    ax = axs[0]
+    color = 'blue'
+    ax.plot(iterations, c=color)
+    ax.set_ylabel('steps', color=color)
+    ax.tick_params(axis='y', labelcolor=color)
+    ax1 = plt.twinx(ax)
+    color = 'k'
+    ax1.plot(np.cumsum(iterations), c=color)
+    ax1.set_ylabel('cumulative steps', color=color)
+    ax.set_title('Iterations' + plot_suffix)
+    # fig.suptitle(label, fontsize=12)
+
+    ax = axs[1]
+    color = 'red'
+    ax.plot(starts, c=color)
+    ax.set_ylabel('starts', color=color)
+    ax.axhline(-0.05, ls=':', color='r')
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/spinning_up.py

@@ -1,0 +1,108 @@
+import random
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import tensorflow as tf
+# import PyQt5
+from spinup import td3_tf1 as td3
+from spinup import sac_tf1 as sac
+from spinup.algos.tf1.ppo.ppo import ppo
+from spinup.algos.tf1.trpo.trpo import trpo
+
+from local_fel_simulated_env import FelLocalEnv
+# from pendulum import PendulumEnv as simpleEnv
+# set random seed
+from simulated_tango import SimTangoConnection
+
+random_seed = 111
+# set random seed
+tf.set_random_seed(random_seed)
+np.random.seed(random_seed)
+random.seed(random_seed)
+
+tango = SimTangoConnection()
+env = FelLocalEnv(tango=tango)
+env_fn = lambda: env
+
+label = 'Sim_Tango'
+directory = "checkpoints/test_implementation/"
+
+
+def plot_results(env, label='def'):
+    # plotting
+    print('now plotting')
+    rewards = env.rewards
+    initial_states = env.initial_conditions
+
+    iterations = []
+    final_rews = []
+    # starts = []
+    sum_rews=[]
+    mean_rews = []
+    # init_states = pd.read_pickle('/Users/shirlaen/PycharmProjects/DeepLearning/spinningup/Environments/initData')
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 0):
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+            # starts.append(-np.sqrt(np.mean(np.square(initial_states[i]))))
+            iterations.append(len(rewards[i]))
+            sum_rews.append(np.sum(rewards[i]))
+            mean_rews.append(np.mean(rewards[i]))
+    plot_suffix = ""#f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
+
+    fig, axs = plt.subplots(2, 1, constrained_layout=True)
+
+    ax=axs[0]
+    ax.plot(iterations)
+    ax.set_title('Iterations' + plot_suffix)
+    fig.suptitle(label, fontsize=12)
+
+    ax = axs[1]
+    ax.plot(final_rews, 'r--')
+
[…]

### 2020-11-12 Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/AE_TRPO_stable_tests.py

@@ -1,0 +1,1720 @@
+import os
+import pickle
+from datetime import datetime
+
+import gym
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import tensorflow as tf
+
+from laser_trajectory_control_env import LaserTrajectoryControlEnv
+from tango_connection import TangoConnection
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+
+
+conf_file = '/home/niky/FERMI/2020_10_06/configuration/conf_fel2.json'
+tango = TangoConnection(conf_file=conf_file)
+real_env = LaserTrajectoryControlEnv(tango=tango)
+
+# Todo: Add info to buffer and test ensemble advantage
+# We will take the avarage cumulative reward and the entropy of the trajectories as an observable,
+# as well as the reward on the measured data during the data collection
+
+
+# For Niky Hyper parameters:
+#------------------------------------------------------
+steps_per_env = 15
+init_random_steps = 50
+# num_epochs = int((300 - init_random_steps) / (steps_per_env)) + 1
+num_epochs = 20
+# num_epochs = int((125-init_random_steps)/(steps_per_env))
+print('Number of epoches: ', num_epochs)
+
+hidden_sizes = [100, 100]
+cr_lr = 1e-3
+gamma = 0.9999
+lam = 0.95
+
+max_training_iterations = 50
+
+delay_before_convergence_check = 1
+
+# number_envs = 1
+critic_iter = 40
+delta = 0.01
+algorithm = 'TRPO'
+conj_iters = 15
+minibatch_size = 100
+simulated_steps = 1000
+
+mb_lr = 1e-3
+model_batch_size = 5  # train the dynamics model
+
+num_ensemble_models = 3
+model_iter = 25
+
+# How often to check the progress of the network training
+# e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
+# dynamic_wait_time = lambda it, ep: True  # (it + 1) % 1 == 0  #
+dynamic_wait_time = lambda it, ep: (it + 1) % 2 == 0  #
+
+# Set the priors for the anchor method:
+#  TODO: How to set these correctly?
+init_params = dict(init_stddev_1_w=np.sqrt(1),
+                   init_stddev_1_b=np.sqrt(1),
+                   init_stddev_2_w=1 / np.sqrt(100))
+
+data_noise = 1e-4  # estimated noise variance
+
[…]

### 2020-11-12 Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/ME_TRPO.py

@@ -1,0 +1,1432 @@
+import os
+import pickle
+
+import numpy as np
+import tensorflow as tf
+import gym
+from datetime import datetime
+import pandas as pd
+import matplotlib.pyplot as plt
+
+from laser_trajectory_control_env import LaserTrajectoryControlEnv
+from tango_connection import TangoConnection
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+
+conf_file = '/home/niky/FERMI/2020_10_06/configuration/conf_fel2.json'
+tango = TangoConnection(conf_file=conf_file)
+real_env = LaserTrajectoryControlEnv(tango=tango)
+
+# Hyper papameters
+steps_per_env = 20
+init_random_steps = 100
+num_epochs = int((500-init_random_steps)/(steps_per_env)) # increase 500 to 700
+# num_epochs = int((125-init_random_steps)/(steps_per_env))
+print('num of epoches', num_epochs)
+
+
+hidden_sizes = [100, 100]
+cr_lr = 1e-3
+gamma = 0.9999
+lam = 0.95
+
+max_training_iterations = 100
+# max_training_iterations = 10
+delay_before_convergence_check = 2
+
+number_envs = 1
+critic_iter = 80
+delta = 0.05
+algorithm = 'TRPO'
+conj_iters = 10
+minibatch_size = 500
+simulated_steps = 2000
+
+#mb_lr = 5e-4
+mb_lr = 1e-3
+model_batch_size = 100
+num_ensemble_models = 5
+model_iter = 15
+
+# How often to check the progress of the network training
+# e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
+# dynamic_wait_time = lambda it, ep: (it + 1) % 5 == 0  #
+dynamic_wait_time = lambda it, ep: (it + 1) % 3 == 0  #
+# dynamic_wait_time = lambda it, ep: (it + 1) % 1 == 0  #
+
+# Create the logging directory:
+project_directory = 'Data_logging/ME_TRPO/mt_1/'
+
+
+hyp_str_all = '-nr_steps_' + str(steps_per_env) + '-cr_lr' + str(cr_lr) + '-crit_it_' + str(
+    critic_iter) + '-d_' + str(delta) + '-conj_iters_' + str(conj_iters) + '-n_ep_' + str(num_epochs) + \
+              '-mini_bs_' + str(minibatch_size) + '-m_bs_' + str(model_batch_size) + \
[…]

### 2020-11-12 Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/ME_TRPO_read_data.py

@@ -1,0 +1,129 @@
+import os
+import pickle
+import matplotlib.pyplot as plt
+import numpy as np
+project_directory = 'Data_logging/run_5/-nr_steps_20-cr_lr0.001-crit_it_80-d_0.05-conj_iters_10-n_ep_20-mini_bs_500-m_bs_100-mb_lr_0.001-sim_steps_2000-m_iter_15-ensnr_5-init_100/'
+#'Data_logging/run_1/-nr_steps_25-cr_lr0.001-crit_it_80-d_0.05-conj_iters_10-n_ep_16-mini_bs_500-m_bs_100-mb_lr_0.0005-sim_steps_2000-m_iter_15-ensnr_5-init_300/'
+
+# for file in os.listdir(project_directory):
+#     filename = file
+# print(filename)
+# # filename = '18_17_49_35_default'
+# filehandler = open(project_directory + filename, 'rb')
+# object = pickle.load(filehandler)
+
+'''
+filenames = []
+for file in os.listdir(project_directory):
+    if 'test' in file:
+        filenames.append(file)
+
+filenames.sort()
+
+# filename = '09_25_19_18_04_training_observables'
+filename = filenames[-1]
+print(filename)
+
+filehandler = open(project_directory + filename, 'rb')
+object = pickle.load(filehandler)
+'''
+
+def plot_results(data, label='test', **kwargs):
+    # plotting
+    print('now plotting')
+    rewards = data
+    print(rewards)
+
+    # initial_states = env.initial_conditions
+
+    iterations = []
+    finals = []
+    means = []
+
+    # init_states = pd.read_pickle('/Users/shirlaen/PycharmProjects/DeepLearning/spinningup/Environments/initData')
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 1):
+            # finals.append(rewards[i][len(rewards[i]) - 1])
+            finals.append(rewards[i][-1])
+            means.append(np.mean(rewards[i][1:]))
+            iterations.append(len(rewards[i]))
+    print(finals)
+    plot_suffix = label  # , Fermi time: {env.TOTAL_COUNTER / 600:.1f} h'
+
+    fig, axs = plt.subplots(2, 1, sharex=True)
+
+    ax = axs[0]
+    ax.plot(iterations)
+    ax.set_title(plot_suffix)
+    # fig.suptitle(label, fontsize=12)
+
+    ax = axs[1]
+    color = 'blue'
[…]

### 2020-11-12 Data_Experiments/2020_11_05_AE_Dyna@FERMI/AEdyna_clean_test.py

@@ -1,0 +1,1353 @@
+import os
+import pickle
+from datetime import datetime
+import itertools as it
+import gym
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+# from stable_baselines.common.noise import NormalActionNoise
+
+from stable_baselines.sac.policies import MlpPolicy
+# from stable_baselines.td3.policies import MlpPolicy
+# from stable_baselines.common.policies import MlpPolicy
+from stable_baselines import SAC as Agent
+# from stable_baselines import TD3 as Agent
+# from stable_baselines import PPO2 as Agent
+
+import tensorflow as tf
+
+# from local_fel_simulated_env import FelLocalEnv
+# from simulated_tango import SimTangoConnection
+from laser_trajectory_control_env import LaserTrajectoryControlEnv
+from tango_connection import TangoConnection
+
+# from naf2 import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+config = tf.ConfigProto(
+    device_count={'GPU': 0}
+)
+# config = None
+
+# tango = SimTangoConnection()
+# real_env = FelLocalEnv(tango=tango)
+conf_file = '/home/niky/FERMI/2020_11_05/configuration/conf_fel2.json'
+tango = TangoConnection(conf_file=conf_file)
+real_env = LaserTrajectoryControlEnv(tango=tango)
+
+# Hyper papameters
+steps_per_env = 25
+init_random_steps = 200
+total_steps = 500  # 350  # 300  # 250
+num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
+
+print('Number of epochs: ', num_epochs)
+
+hidden_sizes = [100, 100]
+
+max_training_iterations = 50
+delay_before_convergence_check = 1
+
+algorithm = 'SAC'
+
+# minibatch_size = 100
+simulated_steps = 3000  # 2500
+
+model_batch_size = 100
+num_ensemble_models = 3
+
+early_stopping = True
+model_iter = 35  # 30
+
+model_training_iterations = 10
+network_size = 15
+# Set the priors for the anchor method:
+#  TODO: How to set these correctly?
+init_params = dict(init_stddev_1_w=np.sqrt(1),
+                   init_stddev_1_b=np.sqrt(1),
+                   init_stddev_2_w=1.0 / np.sqrt(network_size))
+
+data_noise = 0.00000  # estimated noise variance
+lambda_anchor = data_noise / (np.array([init_params['init_stddev_1_w'],
[…]

### 2020-11-12 local_fel_simulated_env.py

@@ -1,0 +1,377 @@
+import pickle
+
+import numpy as np
+import gym
+# import pybobyqa
+import tensorflow as tf
+import matplotlib.pyplot as plt
+import pandas as pd
+
+from simulated_tango import SimTangoConnection
+
+
+class FelLocalEnv(gym.Env):
+
+    def __init__(self, tango, **kwargs):
+        self.max_steps = 10
+        print('init env ' * 20)
+        self.init_rewards = []
+        self.done = False
+        self.current_length = 0
+        self.__name__ = 'FelLocalEnv'
+
+        self.curr_episode = -1
+        self.TOTAL_COUNTER = -1
+
+        self.rewards = []
+        self.states = []
+        self.actions = []
+        self.dones = []
+
+        self.initial_conditions = []
+
+        # tango = SimTangoConnection() simulates the behaviour of the system we want to control
+        self.tango = tango
+
+        # some information from tango
+        self.state_size = self.tango.state_size
+        self.action_size = self.tango.action_size
+
+        self.target_state = self.tango.target_state
+        self.target_intensity = self.tango.target_intensity
+
+        # current state
+        self.init_state = self.tango.state
+
+        # scaling factor definition
+        if 'half_range' in kwargs:
+            self.half_range = kwargs.get('half_range')
+        else:
+            self.half_range = 3000
+
+        self.state_range = self.get_range()
+        self.state_scale = 2 * self.half_range
+
+        # state, intensity and reward first definition
+        self.state = self.scale(self.init_state)
+        self.intensity = self.get_intensity()
+        self.reward = self.get_reward()
+
+        # max action allowed
+        if 'max_action' in kwargs:
+            max_action = kwargs.get('max_action')
+        else:
+            max_action = 500
+            # max_action = 6000
+        self.max_action = max_action / self.state_scale
+
+        print('max_action', max_action)
+
+        # state space definition
+        self.observation_space = gym.spaces.Box(low=0.0,
+                                                high=1.0,
+                                                shape=(self.state_size,),
+                                                dtype=np.float64)
+
+        # action space definition
+        self.action_space = gym.spaces.Box(low=-self.max_action,
+                                           high=self.max_action,
+                                           shape=(self.state_size,),
+                                           dtype=np.float64)
+        self.test = False
+
[…]

### 2020-11-12 simulated_tango.py

@@ -1,0 +1,115 @@
+import numpy as np
+
+
+class SimTangoConnection:
+    # Seed laser control simulator
+
+    def __init__(self, **kwargs):
+
+        self.system = 'LocalSimulator'
+
+        # actuators: 2 piezo motors with 2 degree of freedom each
+        self.actuators_num = 2
+        self.actuators_attr_num = 2
+
+        # sensors: 2 Coupled Charge Device
+        self.sensors_num = 2
+
+        # state: piezo-motor inputs
+        self.state_size = self.actuators_num * self.actuators_attr_num
+
+        # action: variation of piezo-motor inputs
+        self.action_size = self.actuators_num * self.actuators_attr_num
+
+        # state with highest intensity
+        if 'target_state' in kwargs:
+            self.target_state = kwargs.get('target_state')
+        else:
+            self.target_state = 131072 * np.ones(self.state_size)
+            # piezo-motor range: integers in [0, 262144]
+
+        # response matrix
+        self.rm = self.get_respmatrix()
+
+        # state definition
+        self.state = self.target_state.copy()
+
+        # position on CCD when the highest intensity is reached
+        self.target_position = self.get_position()
+
+        # position definition
+        self.position = self.target_position.copy()
+
+        # highest intensity
+        self.target_intensity = self.get_intensity()
+
+        # intensity definition
+        self.intensity = self.target_intensity.copy()
+
+
+
+
+    def get_respmatrix(self):
+        # response matrix
+        # position on CCDs =  response matrix * piezo-motor inputs
+        rm = np.array([[-1.5570540161682593E-5, +3.2428289038152253E-7, +0.0000000000000000E-0, +0.0000000000000000E-0],
+                       [+1.7061003855705661E-6, +1.3362442319301898E-5, +0.0000000000000000E-0, +0.0000000000000000E-0],
+                       [+3.6504472405940234E-5, -2.7883739555787350E-8, +2.1631117516360490E-5, -1.0906340491205139E-6],
+                       [-2.9017613830940000E-6, -2.6667704592363296E-5, -5.2804805443334150E-7,
+                        +8.0338913621924470E-6]])
+        return rm
+
+    def get_position(self):
+        # return current position on CCD
+        position = self.rm.dot(self.state)
+        return position
+
+    def set_state(self, state):
+        # set state in simulator
+        self.state = state
+
+    def get_state(self):
+        # return current state
+        state = self.state
+        return state
+
+    def get_intensity(self):
+        # return current intensity
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/basic_test_script.py
Updates from Overleaf

@@ -1,197 +1,0 @@
-import pickle
-import random
-import random
-import sys
-
-import matplotlib.pyplot as plt
-import numpy as np
-import tensorflow as tf
-
-from local_fel_simulated_env import FelLocalEnv
-from pernaf.pernaf.naf import NAF
-from pernaf.pernaf.utils.statistic import Statistic
-
-# from pendulum import PendulumEnv as simpleEnv
-# set random seed
-from simulated_tango import SimTangoConnection
-
-random_seed = 123
-# set random seed
-tf.set_random_seed(random_seed)
-np.random.seed(random_seed)
-random.seed(random_seed)
-
-tango = SimTangoConnection()
-env = FelLocalEnv(tango=tango)
-
-directory = "checkpoints/test_implementation/"
-
-
-def plot_results(env, label):
-    # plotting
-    print('now plotting')
-    rewards = env.rewards
-    initial_states = env.initial_conditions
-
-    iterations = []
-    final_rews = []
-    # starts = []
-    sum_rews = []
-    mean_rews = []
-    # init_states = pd.read_pickle('/Users/shirlaen/PycharmProjects/DeepLearning/spinningup/Environments/initData')
-
-    for i in range(len(rewards)):
-        if (len(rewards[i]) > 0):
-            final_rews.append(rewards[i][len(rewards[i]) - 1])
-            # starts.append(-np.sqrt(np.mean(np.square(initial_states[i]))))
-            iterations.append(len(rewards[i]))
-            sum_rews.append(np.sum(rewards[i]))
-            mean_rews.append(np.mean(rewards[i]))
-    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
-
-    fig, axs = plt.subplots(2, 1, constrained_layout=True)
-
-    ax = axs[0]
-    ax.plot(iterations)
-    ax.set_title('Iterations' + plot_suffix)
-    fig.suptitle(label, fontsize=12)
-
-    ax = axs[1]
-    ax.plot(final_rews, 'r--')
-
-    ax.set_title('Final reward per episode')  # + plot_suffix)
-    ax.set_xlabel('Episodes (1)')
-
-    # ax1 = plt.twinx(ax)
-    # color = 'lime'
-    # ax1.set_ylabel('starts', color=color)  # we already handled the x-label with ax1
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/laser_trajectory_control_env.py

@@ -1,208 +1,0 @@
-import numpy as np
-import gym
-
-# from tango_connection import TangoConnection
-
-class LaserTrajectoryControlEnv(gym.Env):
-
-    def __init__(self, tango, **kwargs):
-        self.init_rewards = []
-        self.done = False
-        self.current_length = 0
-        self.__name__ = 'LaserTrajectoryControlEnv'
-        
-        self.curr_episode = -1
-        self.TOTAL_COUNTER = -1
-        self.rewards = []
-        self.initial_conditions = []
-
-        #
-        self.tango = tango
-
-        # some information from tango
-        self.system = self.tango.system
-
-        self.state_size = self.tango.state_size
-        self.action_size = self.tango.action_size
-
-        self.init_state = self.tango.init_state
-        self.init_intensity = self.tango.init_intensity
-
-        # scaling factor definition
-        if 'half_range' in kwargs:
-            self.half_range = kwargs.get('half_range')
-        else:
-            self.half_range = 3000
-            if self.system == 'eos':
-                self.half_range = 30000  # 30000
-
-        self.state_range = self.get_range()
-        self.state_scale = 2 * self.half_range
-
-        # target intensity
-        if 'target_intensity' in kwargs:
-            self.target_intensity = kwargs.get('target_intensity')
-        else:
-            self.target_intensity = self.init_intensity
-
-        # state, intensity and reward definition
-        self.init_state_norm = self.scale(self.init_state)
-        self.init_intensity_norm = self.get_intensity()
-        self.state = self.init_state_norm.copy()
-        self.intensity = self.init_intensity_norm.copy()
-        self.reward = self.get_reward()
-
-        ## max action allowed
-        if 'max_action' in kwargs:
-            max_action = kwargs.get('max_action')
-        else:
-            max_action = 500
-            if self.system == 'eos':
-                max_action = 5000  # 2500  # 5000
-        
-        self.max_action = max_action/self.state_scale
-
-        # observation space definition
-        self.observation_space = gym.spaces.Box(low=0.0 + self.max_action,
-                                         high=1.0 - self.max_action,
-                                         shape=(self.state_size,),
-                                         dtype=np.float64)
-
-        # action spacec definition
-        self.action_space = gym.spaces.Box(low=-self.max_action,
-                                          high=self.max_action,
-                                          shape=(self.action_size,),
-                                          dtype=np.float64)
-
-    def get_range(self):
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/naf2_notsoold.py

@@ -1,965 +1,0 @@
-import os
-import pickle
-import time
-
-import tensorflow as tf
-from tensorflow import keras
-
-tf.keras.backend.set_floatx('float64')
-import numpy as np
-from tqdm import tqdm
-
-
-class ReplayBuffer:
-    """
-    A simple FIFO experience replay buffer for NAF_debug agents.
-    """
-
-    def __init__(self, obs_dim, act_dim, size):
-        self.obs1_buf = np.zeros([size, obs_dim], dtype=np.float64)
-        self.obs2_buf = np.zeros([size, obs_dim], dtype=np.float64)
-        self.acts_buf = np.zeros([size, act_dim], dtype=np.float64)
-        self.rews_buf = np.zeros(size, dtype=np.float64)
-        self.done_buf = np.zeros(size, dtype=np.float64)
-        self.ptr, self.size, self.max_size = 0, 0, size
-
-    def store(self, obs, act, rew, next_obs, done):
-        self.obs1_buf[self.ptr] = obs
-        self.obs2_buf[self.ptr] = next_obs
-        self.acts_buf[self.ptr] = act
-        self.rews_buf[self.ptr] = rew
-        self.done_buf[self.ptr] = done
-        self.ptr = (self.ptr + 1) % self.max_size
-        self.size = min(self.size + 1, self.max_size)
-
-    def sample_batch(self, batch_size=32):
-        if self.size < batch_size:
-            idxs = np.arange(self.size)
-        else:
-            idxs = np.random.randint(0, self.size, size=batch_size)
-
-        return dict(obs1=self.obs1_buf[idxs],
-                    obs2=self.obs2_buf[idxs],
-                    acts=self.acts_buf[idxs],
-                    rews=self.rews_buf[idxs],
-                    done=self.done_buf[idxs])
-
-    def save_to_pkl(self, name, directory):
-        idxs = np.arange(self.size)
-        buffer_data = dict(obs1=self.obs1_buf[idxs],
-                    obs2=self.obs2_buf[idxs],
-                    acts=self.acts_buf[idxs],
-                    rews=self.rews_buf[idxs],
-                    done=self.done_buf[idxs])
-        f = open(directory + name, "wb")
-        pickle.dump(buffer_data, f)
-        f.close()
-
-    def read_from_pkl(self, name, directory):
-        with open(directory + name, 'rb') as f:
-            buffer_data = pickle.load(f)
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/new_test_file.py

@@ -1,177 +1,0 @@
-import os
-import pickle
-import pandas as pd
-import random
-
-import time
-
-import gym
-import matplotlib.pyplot as plt
-import numpy as np
-import tensorflow as tf
-
-#from local_fel_simulated_env import FelLocalEnv
-from laser_trajectory_control_env import LaserTrajectoryControlEnv
-from naf2 import NAF
-from pernaf.pernaf.utils.statistic import Statistic
-
-# from simple_environment import simpleEnv
-
-# from pendulum import PendulumEnv as simpleEnv
-# set random seed
-random_seed = 111
-# set random seed
-
-np.random.seed(random_seed)
-random.seed(random_seed)
-
-# from simulated_tango import SimTangoConnection
-from tango_connection import TangoConnection
-
-random_seed = 123
-# set random seed
-# tf.set_random_seed(random_seed)
-np.random.seed(random_seed)
-random.seed(random_seed)
-
-# tango = SimTangoConnection()
-# env = FelLocalEnv(tango=tango)
-conf_file = '/home/niky/FERMI/2020_07_20/configuration/conf_eos.json'
-tango = TangoConnection(conf_file=conf_file)
-env = LaserTrajectoryControlEnv(tango=tango)
-
-directory = "checkpoints/test_implementation/"
-
-label = 'New NAF'
-
-directory = "checkpoints/new_test/"
-
-
-# TODO: Test the loading
-
-def plot_results(env, label):
-    # plotting
-    print('now plotting')
-    rewards = env.rewards
-    initial_states = env.initial_conditions
-
-    iterations = []
-    final_rews = []
-    # starts = []
-    sum_rews = []
-    mean_rews = []
-    # init_states = pd.read_pickle('/Users/shirlaen/PycharmProjects/DeepLearning/spinningup/Environments/initData')
-
-    for i in range(len(rewards)):
-        if (len(rewards[i]) > 0):
-            final_rews.append(rewards[i][len(rewards[i]) - 1])
-            # starts.append(-np.sqrt(np.mean(np.square(initial_states[i]))))
-            iterations.append(len(rewards[i]))
-            sum_rews.append(np.sum(rewards[i]))
-            mean_rews.append(np.mean(rewards[i]))
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/run_naf2.py

@@ -1,172 +1,0 @@
-import os
-import pickle
-import random
-import matplotlib.pyplot as plt
-import numpy as np
-import tensorflow as tf
-
-# from local_fel_simulated_env import FelLocalEnv
-from laser_trajectory_control_env import LaserTrajectoryControlEnv
-# from simulated_tango import SimTangoConnection
-from tango_connection import TangoConnection
-from naf2 import NAF
-
-# set random seed
-random_seed = 111
-np.random.seed(random_seed)
-random.seed(random_seed)
-
-# tango = SimTangoConnection()
-# env = FelLocalEnv(tango=tango)
-conf_file = '/home/niky/FERMI/2020_07_20/configuration/conf_fel.json'
-tango = TangoConnection(conf_file=conf_file)
-env = LaserTrajectoryControlEnv(tango=tango)
-
-def plot_results(env, file_name):
-    # plotting
-    print('Now plotting')
-    rewards = env.rewards
-    initial_rewards = env.init_rewards
-    # print('initial_rewards :', initial_rewards)
-
-    iterations = []
-    final_rews = []
-    starts = []
-    sum_rews = []
-    mean_rews = []
-
-    for i in range(len(rewards)):
-        if (len(rewards[i]) > 0):
-            final_rews.append(rewards[i][len(rewards[i]) - 1])
-            starts.append(initial_rewards[i])
-            iterations.append(len(rewards[i]))
-            sum_rews.append(np.sum(rewards[i]))
-            mean_rews.append(np.mean(rewards[i]))
-    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
-
-    fig, axs = plt.subplots(2, 1)
-
-    ax = axs[0]
-    color = 'blue'
-    ax.plot(iterations, c=color)
-    ax.set_ylabel('steps', color=color)
-    ax.tick_params(axis='y', labelcolor=color)
-    ax1 = plt.twinx(ax)
-    color = 'k'
-    ax1.plot(np.cumsum(iterations), c=color)
-    ax1.set_ylabel('cumulative steps', color=color)
-    ax.set_title('Iterations' + plot_suffix)
-    # fig.suptitle(label, fontsize=12)
-
-    ax = axs[1]
-    color = 'red'
-    ax.plot(starts, c=color)
-    ax.set_ylabel('starts', color=color)
-    ax.axhline(-0.05, ls=':', color='r')
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/run_test_naf2.py

@@ -1,172 +1,0 @@
-import os
-import pickle
-import random
-import matplotlib.pyplot as plt
-import numpy as np
-import tensorflow as tf
-
-# from local_fel_simulated_env import FelLocalEnv
-from laser_trajectory_control_env import LaserTrajectoryControlEnv
-# from simulated_tango import SimTangoConnection
-from tango_connection import TangoConnection
-from naf2 import NAF
-
-# set random seed
-random_seed = 123
-np.random.seed(random_seed)
-random.seed(random_seed)
-
-# tango = SimTangoConnection()
-# env = FelLocalEnv(tango=tango)
-conf_file = '/home/niky/FERMI/2020_07_20/configuration/conf_fel.json'
-tango = TangoConnection(conf_file=conf_file)
-env = LaserTrajectoryControlEnv(tango=tango)
-
-def plot_results(env, file_name):
-    # plotting
-    print('Now plotting')
-    rewards = env.rewards
-    initial_rewards = env.init_rewards
-    # print('initial_rewards :', initial_rewards)
-
-    iterations = []
-    final_rews = []
-    starts = []
-    sum_rews = []
-    mean_rews = []
-
-    for i in range(len(rewards)):
-        if (len(rewards[i]) > 0):
-            final_rews.append(rewards[i][len(rewards[i]) - 1])
-            starts.append(initial_rewards[i])
-            iterations.append(len(rewards[i]))
-            sum_rews.append(np.sum(rewards[i]))
-            mean_rews.append(np.mean(rewards[i]))
-    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
-
-    fig, axs = plt.subplots(2, 1)
-
-    ax = axs[0]
-    color = 'blue'
-    ax.plot(iterations, c=color)
-    ax.set_ylabel('steps', color=color)
-    ax.tick_params(axis='y', labelcolor=color)
-    ax1 = plt.twinx(ax)
-    color = 'k'
-    ax1.plot(np.cumsum(iterations), c=color)
-    ax1.set_ylabel('cumulative steps', color=color)
-    ax.set_title('Iterations' + plot_suffix)
-    # fig.suptitle(label, fontsize=12)
-
-    ax = axs[1]
-    color = 'red'
-    ax.plot(starts, c=color)
-    ax.set_ylabel('starts', color=color)
-    ax.axhline(-0.05, ls=':', color='r')
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/show_progress_presi_gsi.py

@@ -1,214 +1,0 @@
-import os
-import pickle
-import numpy as np
-import matplotlib.pyplot as plt
-
-
-def load_pickle_logging(file_name):
-    directory = 'checkpoints/' + file_name + '/'
-    files = []
-    directory = directory + 'data/'
-    for f in os.listdir(directory):
-        if 'trajectory_data' in f and 'pkl' in f:
-            files.append(f)
-    files.sort()
-    print(files[-1])
-
-    with open(directory + files[-1], 'rb') as f:
-        states = pickle.load(f)
-        actions = pickle.load(f)
-        rewards = pickle.load(f)
-        dones = pickle.load(f)
-    return states, actions, rewards, dones
-
-
-def load_pickle_final(file_name):
-    directory = 'checkpoints/' + file_name + '/'
-    file = 'plot_data_0.pkl'
-
-    with open(directory + file, 'rb') as f:
-        rews = pickle.load(f)
-        inits = pickle.load(f)
-        losses = pickle.load(f)
-        v_s = pickle.load(f)
-    return rews, inits, losses, v_s
-
-
-file_name = 'FEL_training_100_double_q_Tango_11'
-states_0, actions_0, rewards_0, dones_0 = load_pickle_logging(file_name)
-file_name = 'FEL_training_100_double_q_Tango_11_bis'
-states_1, actions_1, rewards_1, dones_1 = load_pickle_logging(file_name)
-rewards = [rewards_0, rewards_1]
-
-file_name_s = 'FEL_training_100_single_q_Tango_11'
-states_s0, actions_s0, rewards_s0, dones_s0 = load_pickle_logging(file_name_s)
-file_name_s = 'FEL_training_100_single_q_Tango_11_bis'
-states_s, actions_s, rewards_s, dones_s = load_pickle_logging(file_name_s)
-rewards_s = [rewards_s, rewards_s0]
-
-
-def read_rewards(rewards0):
-    iterations_all = []
-    final_rews_all = []
-    mean_rews_all = []
-    for k in range(len(rewards0)):
-        rewards = rewards0[k]
-
-        iterations = []
-        final_rews = []
-        mean_rews = []
-        for i in range(len(rewards)):
-            if len(rewards[i]) > 0:
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/spinning_up.py

@@ -1,108 +1,0 @@
-import random
-
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import tensorflow as tf
-# import PyQt5
-from spinup import td3_tf1 as td3
-from spinup import sac_tf1 as sac
-from spinup.algos.tf1.ppo.ppo import ppo
-from spinup.algos.tf1.trpo.trpo import trpo
-
-from local_fel_simulated_env import FelLocalEnv
-# from pendulum import PendulumEnv as simpleEnv
-# set random seed
-from simulated_tango import SimTangoConnection
-
-random_seed = 111
-# set random seed
-tf.set_random_seed(random_seed)
-np.random.seed(random_seed)
-random.seed(random_seed)
-
-tango = SimTangoConnection()
-env = FelLocalEnv(tango=tango)
-env_fn = lambda: env
-
-label = 'Sim_Tango'
-directory = "checkpoints/test_implementation/"
-
-
-def plot_results(env, label='def'):
-    # plotting
-    print('now plotting')
-    rewards = env.rewards
-    initial_states = env.initial_conditions
-
-    iterations = []
-    final_rews = []
-    # starts = []
-    sum_rews=[]
-    mean_rews = []
-    # init_states = pd.read_pickle('/Users/shirlaen/PycharmProjects/DeepLearning/spinningup/Environments/initData')
-
-    for i in range(len(rewards)):
-        if (len(rewards[i]) > 0):
-            final_rews.append(rewards[i][len(rewards[i]) - 1])
-            # starts.append(-np.sqrt(np.mean(np.square(initial_states[i]))))
-            iterations.append(len(rewards[i]))
-            sum_rews.append(np.sum(rewards[i]))
-            mean_rews.append(np.mean(rewards[i]))
-    plot_suffix = ""#f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
-
-    fig, axs = plt.subplots(2, 1, constrained_layout=True)
-
-    ax=axs[0]
-    ax.plot(iterations)
-    ax.set_title('Iterations' + plot_suffix)
-    fig.suptitle(label, fontsize=12)
-
-    ax = axs[1]
-    ax.plot(final_rews, 'r--')
-
[…]

### 2020-11-12 Data_Experiments/2020_07_20_NAF@FERMI/tango_connection.py

@@ -1,247 +1,0 @@
-import json
-import time
-import numpy as np
-import PyTango as tango
-
-class TangoConnection:
-
-    def __init__(self, conf_file, **kwargs):
-
-        # load json configuration file
-        with open(conf_file) as f:
-            self.conf_data = json.load(f)
-
-        self.system = self.conf_data['system']
-
-        # get actuators data
-        conf_actuators = self.conf_data['actuators']
-        self.actuators_data= self.get_confdata(conf_actuators)
-        self.actuators_device_num = self.actuators_data[0]
-        self.actuators_device_list = self.actuators_data[1]
-        self.actuators_device_attr_num = self.actuators_data[2]
-        self.actuators_device_attr_list = self.actuators_data[3]
-
-        self.actuators_size = np.sum(self.actuators_device_attr_num)
-        self.state_size = self.actuators_size.copy()
-        self.action_size = self.actuators_size.copy()
-        self.state = np.zeros(self.state_size)
-
-        # get sensors data
-        conf_sensors = self.conf_data['sensors']
-        self.sensors_data = self.get_confdata(conf_sensors)
-        self.sensors_device_num = self.sensors_data[0]
-        self.sensors_device_list = self.sensors_data[1]
-        self.sensors_device_attr_num = self.sensors_data[2]
-        self.sensors_device_attr_list = self.sensors_data[3]
-
-        self.sensors_size = np.sum(self.sensors_device_attr_num)
-        self.intensity = np.zeros(1)
-
-        # get spectrometer data
-        conf_spectrometer = self.conf_data['spectrometer']
-        self.spectrometer_data = self.get_confdata(conf_spectrometer)
-        self.spectrometer_device_num = self.spectrometer_data[0]
-        self.spectrometer_device_list = self.spectrometer_data[1]
-        self.spectrometer_device_attr_num = self.spectrometer_data[2]
-        self.spectrometer_device_attr_list = self.spectrometer_data[3]
-
-        # get security data
-        conf_security = self.conf_data['security']
-        self.security_data = self.get_confdata(conf_security)
-        self.security_device_num = self.security_data[0]
-        self.security_device_list = self.security_data[1]
-        self.security_device_attr_num = self.security_data[2]
-        self.security_device_attr_list = self.security_data[3]
-        self.security_threshold = 100.
-
-        if 'num_samples' in kwargs:
-            self.num_samples = kwargs.get('num_samples')
-        else:
-            self.num_samples = 11  # 25  # 51  # 25
-
[…]

### 2020-11-12 Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/AE_TRPO_stable_tests.py

@@ -1,1720 +1,0 @@
-import os
-import pickle
-from datetime import datetime
-
-import gym
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import tensorflow as tf
-
-from laser_trajectory_control_env import LaserTrajectoryControlEnv
-from tango_connection import TangoConnection
-
-# set random seed
-random_seed = 111
-np.random.seed(random_seed)
-
-
-conf_file = '/home/niky/FERMI/2020_10_06/configuration/conf_fel2.json'
-tango = TangoConnection(conf_file=conf_file)
-real_env = LaserTrajectoryControlEnv(tango=tango)
-
-# Todo: Add info to buffer and test ensemble advantage
-# We will take the avarage cumulative reward and the entropy of the trajectories as an observable,
-# as well as the reward on the measured data during the data collection
-
-
-# For Niky Hyper parameters:
-#------------------------------------------------------
-steps_per_env = 15
-init_random_steps = 50
-# num_epochs = int((300 - init_random_steps) / (steps_per_env)) + 1
-num_epochs = 20
-# num_epochs = int((125-init_random_steps)/(steps_per_env))
-print('Number of epoches: ', num_epochs)
-
-hidden_sizes = [100, 100]
-cr_lr = 1e-3
-gamma = 0.9999
-lam = 0.95
-
-max_training_iterations = 50
-
-delay_before_convergence_check = 1
-
-# number_envs = 1
-critic_iter = 40
-delta = 0.01
-algorithm = 'TRPO'
-conj_iters = 15
-minibatch_size = 100
-simulated_steps = 1000
-
-mb_lr = 1e-3
-model_batch_size = 5  # train the dynamics model
-
-num_ensemble_models = 3
-model_iter = 25
-
-# How often to check the progress of the network training
-# e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
-# dynamic_wait_time = lambda it, ep: True  # (it + 1) % 1 == 0  #
-dynamic_wait_time = lambda it, ep: (it + 1) % 2 == 0  #
-
-# Set the priors for the anchor method:
-#  TODO: How to set these correctly?
-init_params = dict(init_stddev_1_w=np.sqrt(1),
-                   init_stddev_1_b=np.sqrt(1),
-                   init_stddev_2_w=1 / np.sqrt(100))
-
-data_noise = 1e-4  # estimated noise variance
-
[…]

### 2020-11-12 Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/ME_TRPO.py

@@ -1,1432 +1,0 @@
-import os
-import pickle
-
-import numpy as np
-import tensorflow as tf
-import gym
-from datetime import datetime
-import pandas as pd
-import matplotlib.pyplot as plt
-
-from laser_trajectory_control_env import LaserTrajectoryControlEnv
-from tango_connection import TangoConnection
-
-# set random seed
-random_seed = 111
-np.random.seed(random_seed)
-
-conf_file = '/home/niky/FERMI/2020_10_06/configuration/conf_fel2.json'
-tango = TangoConnection(conf_file=conf_file)
-real_env = LaserTrajectoryControlEnv(tango=tango)
-
-# Hyper papameters
-steps_per_env = 20
-init_random_steps = 100
-num_epochs = int((500-init_random_steps)/(steps_per_env)) # increase 500 to 700
-# num_epochs = int((125-init_random_steps)/(steps_per_env))
-print('num of epoches', num_epochs)
-
-
-hidden_sizes = [100, 100]
-cr_lr = 1e-3
-gamma = 0.9999
-lam = 0.95
-
-max_training_iterations = 100
-# max_training_iterations = 10
-delay_before_convergence_check = 2
-
-number_envs = 1
-critic_iter = 80
-delta = 0.05
-algorithm = 'TRPO'
-conj_iters = 10
-minibatch_size = 500
-simulated_steps = 2000
-
-#mb_lr = 5e-4
-mb_lr = 1e-3
-model_batch_size = 100
-num_ensemble_models = 5
-model_iter = 15
-
-# How often to check the progress of the network training
-# e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
-# dynamic_wait_time = lambda it, ep: (it + 1) % 5 == 0  #
-dynamic_wait_time = lambda it, ep: (it + 1) % 3 == 0  #
-# dynamic_wait_time = lambda it, ep: (it + 1) % 1 == 0  #
-
-# Create the logging directory:
-project_directory = 'Data_logging/ME_TRPO/mt_1/'
-
-
-hyp_str_all = '-nr_steps_' + str(steps_per_env) + '-cr_lr' + str(cr_lr) + '-crit_it_' + str(
-    critic_iter) + '-d_' + str(delta) + '-conj_iters_' + str(conj_iters) + '-n_ep_' + str(num_epochs) + \
-              '-mini_bs_' + str(minibatch_size) + '-m_bs_' + str(model_batch_size) + \
[…]

### 2020-11-12 Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/ME_TRPO_stable.py

@@ -1,1573 +1,0 @@
-import os
-import pickle
-from datetime import datetime
-
-import gym
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import tensorflow as tf
-
-# from local_fel_simulated_env import FelLocalEnv
-# from simulated_tango import SimTangoConnection
-from laser_trajectory_control_env import LaserTrajectoryControlEnv
-from tango_connection import TangoConnection
-
-# set random seed
-random_seed = 111
-np.random.seed(random_seed)
-
-conf_file = '/home/niky/FERMI/2020_10_06/configuration/conf_fel2.json'
-tango = TangoConnection(conf_file=conf_file)
-real_env = LaserTrajectoryControlEnv(tango=tango)
-
-# Hyper papameters
-steps_per_env = 15
-init_random_steps = 45  # 50  # 15
-# For niky: You can change the number of steps -------------------------------------------------------------------
-total_steps = 450  # 400  # 300 # 
-num_epochs = int((total_steps - init_random_steps) / (steps_per_env)) + 1
-
-print('Number of epoches: ', num_epochs)
-
-hidden_sizes = [100, 100]
-cr_lr = 1e-4 # try 1e-3
-gamma = 0.9999
-lam = 0.95
-
-max_training_iterations = 50  # 25  # 100
-# max_training_iterations = 10
-delay_before_convergence_check = 2
-
-number_envs = 1
-critic_iter = 15
-delta = 0.05
-algorithm = 'TRPO'
-conj_iters = 15
-minibatch_size = 500
-simulated_steps = 2000  # 2500  # 2000  # 5000  # in case to increase (Niky)
-
-mb_lr = 1e-3  # 1e-2  # 1e-4  # 1e-3
-model_batch_size = 5 # in case change to 15 or 25 (Niky)
-# For niky: You can change the number of models here -------------------------------------------------------------------
-num_ensemble_models = 9  # 1  # 5
-# ----------------------------------------------------------------------------------------------------------------------
-model_iter = 15
-
-# How often to check the progress of the network training
-# e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
-dynamic_wait_time = lambda it, ep: (it + 1) % 3 == 0  #
-# dynamic_wait_time = lambda it, ep: (it + 1) % 1 == 0  #
-
-# Create the logging directory:
-project_directory = 'Data_logging/ME_TRPO_stable/2020_10_06_ME_TRPO_stable@FERMI/'
-
[…]

### 2020-11-12 Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/new_ME_TRPO_read_data_presi_gsi.py

@@ -1,132 +1,0 @@
-import os
-import pickle
-import matplotlib.pyplot as plt
-import numpy as np
-project_directory = 'Data_logging/run_test/-nr_steps_25-cr_lr0.001-crit_it_80-d_0.05-conj_iters_10-n_ep_1-mini_bs_500' \
-                    '-m_bs_100-mb_lr_0.0005-sim_steps_2000-m_iter_15-ensnr_5-init_100/'
-project_directory = 'Data_logging/ME_TRPO_stable/2020_10_06_ME_TRPO_stable@FERMI/run2/'
-# for file in os.listdir(project_directory):
-#     filename = file
-# print(filename)
-# # filename = '18_17_49_35_default'
-# filehandler = open(project_directory + filename, 'rb')
-# object = pickle.load(filehandler)
-
-filenames = []
-for file in os.listdir(project_directory):
-    if 'final' in file:
-        filenames.append(file)
-
-filenames.sort()
-
-# filename = '09_25_19_18_04_training_observables'
-filename = filenames[-1]
-print(filename)
-
-filehandler = open(project_directory + filename, 'rb')
-object = pickle.load(filehandler)
-
-def plot_results(data, label='Verification', **kwargs):
-    # plotting
-    print('now plotting')
-    rewards = data
-    print(rewards)
-
-    # initial_states = env.initial_conditions
-
-    iterations = []
-    finals = []
-    means = []
-
-    # init_states = pd.read_pickle('/Users/shirlaen/PycharmProjects/DeepLearning/spinningup/Environments/initData')
-
-    for i in range(len(rewards)):
-        if (len(rewards[i]) > 1):
-            # finals.append(rewards[i][len(rewards[i]) - 1])
-            finals.append(rewards[i][-1])
-            means.append(np.sum(rewards[i][1:]))
-            iterations.append(len(rewards[i]))
-    print(finals)
-    plot_suffix = label  # , Fermi time: {env.TOTAL_COUNTER / 600:.1f} h'
-
-    fig, axs = plt.subplots(2, 1, sharex=True)
-
-    ax = axs[0]
-    ax.plot(iterations)
-    ax.set_ylabel('steps')
-    ax.set_title(plot_suffix)
-    # fig.suptitle(label, fontsize=12)
-
-    ax = axs[1]
-    color = 'blue'
[…]

### 2020-11-12 Data_Experiments/2020_11_05_AE_Dyna@FERMI/AEdyna_clean_test.py

@@ -1,1353 +1,0 @@
-import os
-import pickle
-from datetime import datetime
-import itertools as it
-import gym
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-# from stable_baselines.common.noise import NormalActionNoise
-
-from stable_baselines.sac.policies import MlpPolicy
-# from stable_baselines.td3.policies import MlpPolicy
-# from stable_baselines.common.policies import MlpPolicy
-from stable_baselines import SAC as Agent
-# from stable_baselines import TD3 as Agent
-# from stable_baselines import PPO2 as Agent
-
-import tensorflow as tf
-
-# from local_fel_simulated_env import FelLocalEnv
-# from simulated_tango import SimTangoConnection
-from laser_trajectory_control_env import LaserTrajectoryControlEnv
-from tango_connection import TangoConnection
-
-# from naf2 import NAF
-
-# set random seed
-random_seed = 111
-np.random.seed(random_seed)
-config = tf.ConfigProto(
-    device_count={'GPU': 0}
-)
-# config = None
-
-# tango = SimTangoConnection()
-# real_env = FelLocalEnv(tango=tango)
-conf_file = '/home/niky/FERMI/2020_11_05/configuration/conf_fel2.json'
-tango = TangoConnection(conf_file=conf_file)
-real_env = LaserTrajectoryControlEnv(tango=tango)
-
-# Hyper papameters
-steps_per_env = 25
-init_random_steps = 200
-total_steps = 500  # 350  # 300  # 250
-num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
-
-print('Number of epochs: ', num_epochs)
-
-hidden_sizes = [100, 100]
-
-max_training_iterations = 50
-delay_before_convergence_check = 1
-
-algorithm = 'SAC'
-
-# minibatch_size = 100
-simulated_steps = 3000  # 2500
-
-model_batch_size = 100
-num_ensemble_models = 3
-
-early_stopping = True
-model_iter = 35  # 30
-
-model_training_iterations = 10
-network_size = 15
-# Set the priors for the anchor method:
-#  TODO: How to set these correctly?
-init_params = dict(init_stddev_1_w=np.sqrt(1),
-                   init_stddev_1_b=np.sqrt(1),
-                   init_stddev_2_w=1.0 / np.sqrt(network_size))
-
-data_noise = 0.00000  # estimated noise variance
-lambda_anchor = data_noise / (np.array([init_params['init_stddev_1_w'],
[…]

### 2020-11-12 local_fel_simulated_env.py

@@ -1,377 +1,0 @@
-import pickle
-
-import numpy as np
-import gym
-# import pybobyqa
-import tensorflow as tf
-import matplotlib.pyplot as plt
-import pandas as pd
-
-from simulated_tango import SimTangoConnection
-
-
-class FelLocalEnv(gym.Env):
-
-    def __init__(self, tango, **kwargs):
-        self.max_steps = 10
-        print('init env ' * 20)
-        self.init_rewards = []
-        self.done = False
-        self.current_length = 0
-        self.__name__ = 'FelLocalEnv'
-
-        self.curr_episode = -1
-        self.TOTAL_COUNTER = -1
-
-        self.rewards = []
-        self.states = []
-        self.actions = []
-        self.dones = []
-
-        self.initial_conditions = []
-
-        # tango = SimTangoConnection() simulates the behaviour of the system we want to control
-        self.tango = tango
-
-        # some information from tango
-        self.state_size = self.tango.state_size
-        self.action_size = self.tango.action_size
-
-        self.target_state = self.tango.target_state
-        self.target_intensity = self.tango.target_intensity
-
-        # current state
-        self.init_state = self.tango.state
-
-        # scaling factor definition
-        if 'half_range' in kwargs:
-            self.half_range = kwargs.get('half_range')
-        else:
-            self.half_range = 3000
-
-        self.state_range = self.get_range()
-        self.state_scale = 2 * self.half_range
-
-        # state, intensity and reward first definition
-        self.state = self.scale(self.init_state)
-        self.intensity = self.get_intensity()
-        self.reward = self.get_reward()
-
-        # max action allowed
-        if 'max_action' in kwargs:
-            max_action = kwargs.get('max_action')
-        else:
-            max_action = 500
-            # max_action = 6000
-        self.max_action = max_action / self.state_scale
-
-        print('max_action', max_action)
-
-        # state space definition
-        self.observation_space = gym.spaces.Box(low=0.0,
-                                                high=1.0,
-                                                shape=(self.state_size,),
-                                                dtype=np.float64)
-
-        # action space definition
-        self.action_space = gym.spaces.Box(low=-self.max_action,
-                                           high=self.max_action,
-                                           shape=(self.state_size,),
-                                           dtype=np.float64)
-        self.test = False
-
[…]

### 2020-11-16 main.tex
Updates from Overleaf

@@
 \end{align}
-$\theta\prime$ denotes the weights of a previous iteration.
+$\theta\prime$ denotes the weights of a previous iteration. The optimal policy is given by:
+\begin{align}
+    \pi^*(s)\arg\!\max
+\end{align}
 \subsection{New NAF network architecture}
 
-\begin{algorithm}[ht]
-\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
-\begin{algorithmic}[1]
-\State initialise $\theta_0$
-\For{iteration $k \in [0, \dots, K]$}
-\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
-\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
-\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
-\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
-\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
-\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
-\EndFor
-\end{algorithmic}
-\end{algorithm}
+% \begin{algorithm}[ht]
+% \caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
+% \begin{algorithmic}[1]
+% \State initialise $\theta_0$
+% \For{iteration $k \in [0, \dots, K]$}
+% \State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
+% \State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
+% \State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
+% \State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
+% \State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
+% \State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
+% \EndFor
+% \end{algorithmic}
+% \end{algorithm}
 

### 2020-11-16 Data_Experiments/2020_11_05_AE_Dyna@FERMI/laser_trajectory_control_env.py

@@ -1,240 +1,0 @@
-import numpy as np
-import gym
-
-# from tango_connection import TangoConnection
-
-class LaserTrajectoryControlEnv(gym.Env):
-
-    def __init__(self, tango, **kwargs):
-        self.init_rewards = []
-        self.done = False
-        self.current_length = 0
-        self.__name__ = 'LaserTrajectoryControlEnv'
-        
-        self.curr_episode = -1
-        self.TOTAL_COUNTER = -1
-        self.rewards = []
-        self.actions = []
-        self.states = []
-        self.dones = []
-        self.initial_conditions = []
-        
-        self.max_length = 25
-        self.max_steps = 10
-
-        #
-        self.tango = tango
-
-        # some information from tango
-        self.system = self.tango.system
-
-        self.state_size = self.tango.state_size
-        self.action_size = self.tango.action_size
-
-        self.init_state = self.tango.init_state
-        self.init_intensity = self.tango.init_intensity
-
-        # scaling factor definition
-        if 'half_range' in kwargs:
-            self.half_range = kwargs.get('half_range')
-        else:
-            self.half_range = 3000
-            if self.system == 'eos':
-                self.half_range = 30000  # 30000
-
-        self.state_range = self.get_range()
-        self.state_scale = 2 * self.half_range
-
-        # target intensity
-        if 'target_intensity' in kwargs:
-            self.target_intensity = kwargs.get('target_intensity')
-        else:
-            self.target_intensity = self.init_intensity
-
-        # state, intensity and reward definition
-        self.init_state_norm = self.scale(self.init_state)
-        self.init_intensity_norm = self.get_intensity()
-        self.state = self.init_state_norm.copy()
-        self.intensity = self.init_intensity_norm.copy()
-        self.reward = self.get_reward()
-
-        ## max action allowed
-        if 'max_action' in kwargs:
-            max_action = kwargs.get('max_action')
-        else:
-            max_action = 500
-            # bigger max_action... evalueate if the size is correct!
-            # max_action = 6000 # 3000
-            if self.system == 'eos':
-                max_action = 5000  # 2500  # 5000
-        
-        self.max_action = max_action/self.state_scale
-        
-        # observation space definition
-        self.observation_space = gym.spaces.Box(low=0.0, #+ self.max_action,
-                                         high=1.0, #- self.max_action,
-                                         shape=(self.state_size,),
-                                         dtype=np.float64)
-
-        # action spacec definition
[…]

### 2020-11-16 AEDYNA.py
- initial commit

@@ -1,24 +1,14 @@
 import os
 import pickle
 from datetime import datetime
-import itertools as it
+
 import gym
 import matplotlib.pyplot as plt
 import numpy as np
-import pandas as pd
-# from stable_baselines.common.noise import NormalActionNoise
-
-# from stable_baselines.sac.policies import MlpPolicy
-# from stable_baselines import SAC as Agent
-# from stable_baselines.td3.policies import MlpPolicy
-# from stable_baselines import TD3 as Agent
-from stable_baselines.common.policies import MlpPolicy
-from stable_baselines import PPO2 as Agent
-
 import tensorflow as tf
+from inverted_pendulum import PendulumEnv
 
-from local_fel_simulated_env import FelLocalEnv
-from simulated_tango import SimTangoConnection
+# from stable_baselines.common.noise import NormalActionNoise
 
 # from naf2 import NAF
 
@@ -28,37 +18,52 @@ np.random.seed(random_seed)
 config = tf.ConfigProto(
     device_count={'GPU': 0}
 )
-# config = None
+config = None
 
-tango = SimTangoConnection()
-real_env = FelLocalEnv(tango=tango)
+############################################################
+# Loading the environment
+############################################################
 
-# Hyper papameters
-steps_per_env = 25
-init_random_steps = 100
-total_steps = 250
+real_env = env = PendulumEnv()
+
+############################################################
+# Hyperparameters
+############################################################
+
+
+steps_per_env = 200
+init_random_steps = 200
+total_steps = 2500
 num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
 
 print('Number of epochs: ', num_epochs)
 
-hidden_sizes = [100, 100]
 
 max_training_iterations = 50
 delay_before_convergence_check = 1
 
-algorithm = 'SAC'
+# Select the rl-algorithm
 
-# minibatch_size = 100
-simulated_steps = 3000
+algorithm = 'PPO'
+# algorithm = 'PPO'
 
-model_batch_size = 5
-num_ensemble_models = 5
+if algorithm == 'SAC':
+    from stable_baselines.sac.policies import MlpPolicy
+    from stable_baselines import SAC as Agent
+else:
+    from stable_baselines.common.policies import MlpPolicy
+    from stable_baselines import PPO2 as Agent
+
+simulated_steps = 2500
+
+model_batch_size = 100
+num_ensemble_models = 3
 
 early_stopping = True
-model_iter = 30
+model_iter = 2
 
 # model_training_iterations = 10
[…]

### 2020-11-16 AEdyna_release.py

@@ -1,0 +1,1231 @@
+import os
+import pickle
+from datetime import datetime
+import itertools as it
+import gym
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+# from stable_baselines.common.noise import NormalActionNoise
+
+import tensorflow as tf
+
+from local_fel_simulated_env import FelLocalEnv
+from simulated_tango import SimTangoConnection
+
+# from naf2 import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+config = tf.ConfigProto(
+    device_count={'GPU': 0}
+)
+# config = None
+
+############################################################
+# Loading the environment
+############################################################
+tango = SimTangoConnection()
+real_env = FelLocalEnv(tango=tango)
+
+############################################################
+# Hyperparameters
+############################################################
+
+
+steps_per_env = 25
+init_random_steps = 100
+total_steps = 250
+num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
+
+print('Number of epochs: ', num_epochs)
+
+hidden_sizes = [100, 100]
+
+max_training_iterations = 50
+delay_before_convergence_check = 1
+
+# Select the rl-algorithm
+
+algorithm = 'SAC'
+# algorithm = 'PPO'
+
+
+# from stable_baselines.td3.policies import MlpPolicy
+# from stable_baselines import TD3 as Agent
+if algorithm == 'SAC':
+    from stable_baselines.sac.policies import MlpPolicy
+    from stable_baselines import SAC as Agent
+else:
+    from stable_baselines.common.policies import MlpPolicy
+    from stable_baselines import PPO2 as Agent
+
+simulated_steps = 3000
+
+model_batch_size = 5
+num_ensemble_models = 5
+
+early_stopping = True
+model_iter = 30
+
+# model_training_iterations = 10
+network_size = 15
+# Set the priors for the anchor method:
+#  TODO: How to set these correctly?
+init_params = dict(init_stddev_1_w=np.sqrt(1),
+                   init_stddev_1_b=np.sqrt(1),
+                   init_stddev_2_w=1 / np.sqrt(network_size))
+
+data_noise = 0.00000  # estimated noise variance
+lambda_anchor = data_noise / (np.array([init_params['init_stddev_1_w'],
+                                        init_params['init_stddev_1_b'],
+                                        init_params['init_stddev_2_w']]) ** 2)
+
+# How often to check the progress of the network training
[…]

### 2020-11-16 inverted_pendulum.py

@@ -1,0 +1,103 @@
+import gym
+from gym import spaces
+from gym.utils import seeding
+import numpy as np
+from os import path
+
+
+class PendulumEnv(gym.Env):
+    metadata = {
+        'render.modes': ['human', 'rgb_array'],
+        'video.frames_per_second': 30
+    }
+
+    def __init__(self, g=10.0):
+        self.max_speed = 8
+        self.max_torque = 2.
+        self.dt = .05
+        self.g = g
+        self.m = 1.
+        self.l = 1.
+        self.viewer = None
+
+        high = np.array([1., 1., self.max_speed], dtype=np.float32)
+        self.action_space = spaces.Box(
+            low=-self.max_torque,
+            high=self.max_torque, shape=(1,),
+            dtype=np.float32
+        )
+        self.observation_space = spaces.Box(
+            low=-high,
+            high=high,
+            dtype=np.float32
+        )
+
+        self.seed()
+
+    def seed(self, seed=None):
+        self.np_random, seed = seeding.np_random(seed)
+        return [seed]
+
+    def step(self, u):
+        th, thdot = self.state  # th := theta
+
+        g = self.g
+        m = self.m
+        l = self.l
+        dt = self.dt
+
+        u = np.clip(u, -self.max_torque, self.max_torque)[0]
+        self.last_u = u  # for rendering
+        costs = angle_normalize(th) ** 2 + .1 * thdot ** 2 + .001 * (u ** 2)
+
+        newthdot = thdot + (-3 * g / (2 * l) * np.sin(th + np.pi) + 3. / (m * l ** 2) * u) * dt
+        newth = th + newthdot * dt
+        newthdot = np.clip(newthdot, -self.max_speed, self.max_speed)
+
+        self.state = np.array([newth, newthdot])
+
+        return self._get_obs(), -costs, False, {}
+
+    def reset(self):
+        high = np.array([np.pi, 1])
+        self.state = self.np_random.uniform(low=-high, high=high)
+        self.last_u = None
+        return self._get_obs()
+
+    def _get_obs(self):
+        theta, thetadot = self.state
+        return np.array([np.cos(theta), np.sin(theta), thetadot])
+
+    def render(self, mode='human'):
+        if self.viewer is None:
+            from gym.envs.classic_control import rendering
+            self.viewer = rendering.Viewer(500, 500)
+            self.viewer.set_bounds(-2.2, 2.2, -2.2, 2.2)
[…]

### 2020-11-16 run_naf2.py

@@ -1,0 +1,176 @@
+import os
+import pickle
+import random
+import matplotlib.pyplot as plt
+import numpy as np
+import tensorflow as tf
+
+
+
+
+# from laser_trajectory_control_env import LaserTrajectoryControlEnv
+# from tango_connection import TangoConnection
+from local_fel_simulated_env import FelLocalEnv as Ta
+from simulated_tango import SimTangoConnection
+
+from naf2 import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+random.seed(random_seed)
+
+# tango = SimTangoConnection()
+# env = FelLocalEnv(tango=tango)
+conf_file = '/home/niky/FERMI/2020_07_20/configuration/conf_fel.json'
+tango = TangoConnection(conf_file=conf_file)
+env = LaserTrajectoryControlEnv(tango=tango)
+
+def plot_results(env, file_name):
+    # plotting
+    print('Now plotting')
+    rewards = env.rewards
+    initial_rewards = env.init_rewards
+    # print('initial_rewards :', initial_rewards)
+
+    iterations = []
+    final_rews = []
+    starts = []
+    sum_rews = []
+    mean_rews = []
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 0):
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+            starts.append(initial_rewards[i])
+            iterations.append(len(rewards[i]))
+            sum_rews.append(np.sum(rewards[i]))
+            mean_rews.append(np.mean(rewards[i]))
+    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
+
+    fig, axs = plt.subplots(2, 1)
+
+    ax = axs[0]
+    color = 'blue'
+    ax.plot(iterations, c=color)
+    ax.set_ylabel('steps', color=color)
+    ax.tick_params(axis='y', labelcolor=color)
+    ax1 = plt.twinx(ax)
+    color = 'k'
+    ax1.plot(np.cumsum(iterations), c=color)
+    ax1.set_ylabel('cumulative steps', color=color)
+    ax.set_title('Iterations' + plot_suffix)
+    # fig.suptitle(label, fontsize=12)
+
+    ax = axs[1]
+    color = 'red'
+    ax.plot(starts, c=color)
+    ax.set_ylabel('starts', color=color)
[…]

### 2020-11-16 main.tex
txs auto checkin

@@
 \begin{align}
-     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s),
+     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
 \end{align}
+ The optimal policy is given by:
+\begin{align}
+    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
+\end{align}
+One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-fitting:
 the maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
 \begin{align}
-     \min_\theta\left( (r(s_t,a_t)+\gamma V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
+	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
 \end{align}
-$\theta\prime$ denotes the weights of a previous iteration. The optimal policy is given by:
-\begin{align}
-    \pi^*(s)\arg\!\max
-\end{align}
+$\theta\prime$ denotes the weights of a previous iteration.
 \subsection{New NAF network architecture}

### 2020-11-16 main.tex
txs auto checkin

@@
 \end{align}
-One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-fitting:
-the maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
+One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-fitting.
+The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
 \begin{align}

### 2020-11-16 main.tex
txs auto checkin

@@
 \end{align}
-One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-fitting.
+One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
 The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
 \begin{align}
-	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
+	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
 \end{align}

### 2020-11-16 main.tex
txs auto checkin

@@
 One way to reduce the effect is to take a simple analytical form of the $Q$-function.
-\subsection{NAF}
+\subsection{Value based}
 If a specific quadratic form of the $Q$ function is assumed:
@@
 $\theta\prime$ denotes the weights of a previous iteration.
-\subsection{New NAF network architecture}
+%\subsection{New NAF network architecture}
 
@@
 
-\section{Dyna-style reinforcement learning}
+\section{Uncertainty aware Dyna-style reinforcement learning}
+The original Dyna algorithm \cite{DYNA} is modified here in several aspects.
+Generally, Dyna style algorithms denote algorithms, where a MFRL algorithm is trained on purely synthetic data from an approximate dynamics model or on a mixture of synthetic and real data.
+We use only synthetic data to reduce the interaction with the real environment to a minimum.
 

### 2020-11-16 main.tex
txs auto checkin

@@
 \end{itemize}
-RL optimises a sequence of decisions. 
+RL optimises a sequence of decisions. On critical aspect using RL is the number of iteration to train a controller.
 In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.
@@
 We use only synthetic data to reduce the interaction with the real environment to a minimum.
+An overview of the used method is shown in \cref{fig:MBRL_overview}.
+At the beginning the data is collected or read in from a pre-collection.
+An uncertainty aware model is trained, using anchored ensembles on the data, which allows to take the allegorical (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
+Subsequently a MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data, by taking the uncertainty into account.
+After a defined number of training steps the controller is testes on each model individually.
+If there is no improvement in a specific ratio<1 of models, the controller is tested on the real environment.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/MBRL_overview}
+	\caption{A schematic overview of the AE-Dyna approach used in this paper.}
+	\label{fig:MBRL_overview}
+\end{figure}
 

### 2020-11-16 main.tex
txs auto checkin

@@
 \subsection{\NB{short FERMI introduction}}
-The Free Electron laser Radiation for Multidisciplinary Investigation  better known as the FERMI, is the seeded free-electron laser facility next to the third-generation synchrotron facility ELETTRA at the Elettra Sinctrotrone Trieste laboratories.
+The Free Electron laser Radiation for Multidisciplinary Investigation  better known as the FERMI, is the seeded free-electron laser facility next to the third-generation synchrotron facility ELETTRA at the Elettra Sinchrotrone Trieste laboratories.
 A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure.

### 2020-11-16 Data_Experiments/2020_11_05_AE_Dyna@FERMI/AEdyna_clean_test.py
- data update

@@ -1,14 +1,26 @@
 import os
 import pickle
 from datetime import datetime
-
+import itertools as it
 import gym
 import matplotlib.pyplot as plt
 import numpy as np
-import tensorflow as tf
-from inverted_pendulum import PendulumEnv
-
+import pandas as pd
 # from stable_baselines.common.noise import NormalActionNoise
+
+from stable_baselines.sac.policies import MlpPolicy
+# from stable_baselines.td3.policies import MlpPolicy
+# from stable_baselines.common.policies import MlpPolicy
+from stable_baselines import SAC as Agent
+# from stable_baselines import TD3 as Agent
+# from stable_baselines import PPO2 as Agent
+
+import tensorflow as tf
+
+# from local_fel_simulated_env import FelLocalEnv
+# from simulated_tango import SimTangoConnection
+from laser_trajectory_control_env import LaserTrajectoryControlEnv
+from tango_connection import TangoConnection
 
 # from naf2 import NAF
 
@@ -18,57 +30,45 @@ np.random.seed(random_seed)
 config = tf.ConfigProto(
     device_count={'GPU': 0}
 )
-config = None
+# config = None
 
-############################################################
-# Loading the environment
-############################################################
+# tango = SimTangoConnection()
+# real_env = FelLocalEnv(tango=tango)
+conf_file = '/home/niky/FERMI/2020_11_05/configuration/conf_fel2.json'
+tango = TangoConnection(conf_file=conf_file)
+real_env = LaserTrajectoryControlEnv(tango=tango)
 
-real_env = env = PendulumEnv()
-
-############################################################
-# Hyperparameters
-############################################################
-
-
-steps_per_env = 201
-init_random_steps = 201
-total_steps = 10000
+# Hyper papameters
+steps_per_env = 25
+init_random_steps = 200
+total_steps = 500  # 350  # 300  # 250
 num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
 
 print('Number of epochs: ', num_epochs)
 
+hidden_sizes = [100, 100]
 
 max_training_iterations = 50
 delay_before_convergence_check = 1
 
-# Select the rl-algorithm
+algorithm = 'SAC'
 
-algorithm = 'PPO'
-# algorithm = 'PPO'
-
-if algorithm == 'SAC':
-    from stable_baselines.sac.policies import MlpPolicy
-    from stable_baselines import SAC as Agent
-else:
-    from stable_baselines.common.policies import MlpPolicy
[…]

### 2020-11-16 main.tex
txs auto checkin

@@
 After a defined number of training steps the controller is testes on each model individually.
-If there is no improvement in a specific ratio<1 of models, the controller is tested on the real environment.
+If there is no improvement in a specific ratio$<1$ of models, the controller is tested on the real environment.
 
@@
 \end{figure}
-
+\subsection{What was tested?}
+\begin{itemize}
+	\item The ANN - including the prior - number of models
+	\item The number of data points and the policy
+	\item The uncertainty - how is this included
+	\item The episodic design
+	\item ...
+\end{itemize}
 \subsection{Operational deployment}

### 2020-11-16 main.tex
txs auto checkin

@@
 After a defined number of training steps the controller is testes on each model individually.
-If there is no improvement in a specific ratio$<1$ of models, the controller is tested on the real environment.
+If there is no improvement in a specific ratio$<1$ of models, the controller is tested on the real environment for a number of episodes.
+The training is stopped if the controller solves the problem on the real environement.
 
@@
 \begin{itemize}
-	\item The ANN - including the prior - number of models
+	\item The ANN - including the prior - number of models - noise level
 	\item The number of data points and the policy

### 2020-11-16 main.tex
txs auto checkin

@@
 If there is no improvement in a specific ratio$<1$ of models, the controller is tested on the real environment for a number of episodes.
-The training is stopped if the controller solves the problem on the real environement.
+The training is stopped if the controller solves the problem on the real environment.
 
@@
 \appendix
+\section{A non-linear standard control problem}
+To provide some transparency of these studies we provide results on a famous classical standard control problem, the inverted pendulum. It is a non-linear low dimensional continuous control problem. 
 \section{NAF2 details}\label{appendix:pernaf}

### 2020-11-16 main.tex
txs auto checkin

@@
 
-% \begin{algorithm}[ht]
-% \caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
-% \begin{algorithmic}[1]
-% \State initialise $\theta_0$
-% \For{iteration $k \in [0, \dots, K]$}
-% \State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
-% \State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
-% \State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
-% \State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
-% \State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
-% \State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
-% \EndFor
-% \end{algorithmic}
-% \end{algorithm}
+\begin{algorithm}[ht]
+\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
+\begin{algorithmic}[1]
+\State initialise $\theta_0$
+\For{iteration $k \in [0, \dots, K]$}
+\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
+\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
+\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
+\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
+\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
+\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
+\EndFor
+\end{algorithmic}
+\end{algorithm}
 
@@
 \section{A non-linear standard control problem}
-To provide some transparency of these studies we provide results on a famous classical standard control problem, the inverted pendulum. It is a non-linear low dimensional continuous control problem. 
+To provide some transparency of these studies we provide results on a famous classical standard control problem, the inverted pendulum.
+It is a non-linear low dimensional unsolved continuous control problem, which means there is no threshold for the reward to terminate an episode.
+The episode length is set to 200 steps.
 \section{NAF2 details}\label{appendix:pernaf}

### 2020-11-16 main.tex
txs auto checkin

@@
 \usepackage{algpseudocode}
-\usepackage[ruled,vlined]{algorithm2e}
+%\usepackage[ruled,vlined]{algorithm2e}
 
@@
 \newcommand{\NB}[1]{\textcolor{red}{#1}}
+\include{tex/defs}
 
@@
 One way to reduce the effect is to take a simple analytical form of the $Q$-function.
-\subsection{Value based}
+\subsection{NAF}
 If a specific quadratic form of the $Q$ function is assumed:
@@
 \State initialise $\theta_0$
-\For{iteration $k \in [0, \dots, K]$}
-\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
-\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
-\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
-\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
-\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
-\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
-\EndFor
-\end{algorithmic}
+%\For{iteration $k \in [0, \dots, K]$}
+%\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
+%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
+%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
+%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
+%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
+%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
+%\EndFor
+%\end{algorithmic}
 \end{algorithm}

### 2020-11-16 main.tex
txs auto checkin

@@
 % For algorithms
-\usepackage{algorithm}
+%\usepackage{algorithm}
 \usepackage{algpseudocode}
-%\usepackage[ruled,vlined]{algorithm2e}
+\usepackage[ruled,vlined]{algorithm2e}
 
@@
 \section{Deep Reinforcement learning}
-Assume states $s$ the set of all states $S$ and actions $a$ all actions $a$ and rewards $r$ and the set of all rewards $R$ an initial state distribution $d_0$.
+Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\rewards$ an initial state distribution $d_0$.
 Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
@@
 \State initialise $\theta_0$
-%\For{iteration $k \in [0, \dots, K]$}
+\For{iteration $k \in [0, \dots, K]$}
 %\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
@@
 %\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
-%\EndFor
-%\end{algorithmic}
+\EndFor
+\end{algorithmic}
 \end{algorithm}

### 2020-11-16 main.tex
txs auto checkin

@@
 \section{Deep Reinforcement learning}
-Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\rewards$ an initial state distribution $d_0$.
+Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $d_0$.
 Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:

### 2020-11-16 main.tex
txs auto checkin

@@
 \end{align}
-$\theta\prime$ denotes the weights of a previous iteration.
+$\theta\prime$ denotes the weights of a previous iteration. To stabilize the network training a small artificial noise is added to the actions in the update.
 %\subsection{New NAF network architecture}
@@
 \For{iteration $k \in [0, \dots, K]$}
-%\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
+\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
 %\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
@@
 %\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
-\EndFor
+%\EndFor
 \end{algorithmic}
@@
 \begin{itemize}
-	\item The ANN - including the prior - number of models - noise level
-	\item The number of data points and the policy
-	\item The uncertainty - how is this included
-	\item The episodic design
-	\item ...
+	\item The ANN - including the prior - number of models - noise level.
+	\item The number of data points and the policy.
+	\item The uncertainty - how is this included.
+	\item The episodic design.
+	\item The MFRL agent as e.g. the TRPO and PPO or the SAC.
 \end{itemize}

### 2020-11-16 main.tex
txs auto checkin

@@
 \end{figure}
-\subsection{What was tested?}
+\subsection{Critical design decisions}
+In the following the most important aspects of a successful application of a Dyna-AE algorithm are listed:
 \begin{itemize}
@@
 	\item The number of data points and the policy.
-	\item The uncertainty - how is this included.
-	\item The episodic design.
+	\item The uncertainty - how is this included?
+	\item The episodic design - avoid long trajectories.
 	\item The MFRL agent as e.g. the TRPO and PPO or the SAC.
 \end{itemize}
+We try to discuss all items in some detail in the following.
+The anchoring ensemble methods usually yields good results already with a small number of different networks.
+The idea behind it, is to mimic the posterior probability of the dynamics model to capture first of all its own uncertainty due to the lack of data.
+Empirical results show that three to five models were sufficient to see clear advantages over a single network approach.
 \subsection{Operational deployment}

### 2020-11-16 main.tex
txs auto checkin

@@
 The idea behind it, is to mimic the posterior probability of the dynamics model to capture first of all its own uncertainty due to the lack of data.
-Empirical results show that three to five models were sufficient to see clear advantages over a single network approach.
+Empirical results show that three to five models were sufficient to see clear advantages over a single network approach and the main goal is not to calculate the exact posterior of the model.
+A small two-layer network with around 15-25 nodes and the $tanh$ activation were used.
+The last layer was linear and the inputs were normalized to the interval [-1,1].
 \subsection{Operational deployment}

### 2020-11-16 main.tex
txs auto checkin

@@
 The last layer was linear and the inputs were normalized to the interval [-1,1].
+The prior is controlled by the initialization of the weights, which were normalized by the number of nodes in each layer.
+The noise level is also respected during the initialization.
+For the moment only homeoscetastic errors are used.\\
+One of the most crucial points is how this uncertainties are taken into account by the RL algorithms.
+Several different approaches were tried.
+It is possible to take the average of the models and add Gaussian noise leveled by the standard deviation of the models.
+Another straight forward way is to randomly select a model to provide the next step.
+
 \subsection{Operational deployment}

### 2020-11-16 main.tex
txs auto checkin

@@
 It is possible to take the average of the models and add Gaussian noise leveled by the standard deviation of the models.
-Another straight forward way is to randomly select a model to provide the next step.
+Another straight forward way is to randomly select a model to provide at each training step, which is the original implementation of the ME-TRPO algorithm and was used in the experiments labeled \it{ME-TRPO}.
+A pessimistic setting only would select the model resulting in the lowest predicted reward.
+Good results were obtained by following a randomly selected single model each full episode and were used in the experiments labeled \it{AE-Dyna}.
 

### 2020-11-16 main.tex
txs auto checkin

@@
 \begin{itemize}
-	\item The ANN - including the prior - number of models - noise level.
+	\item The ANN - including the prior - number of models - noise level - early stopping
 	\item The number of data points and the policy.
@@
 The noise level is also respected during the initialization.
-For the moment only homeoscetastic errors are used.\\
+For the moment only homeoscetastic errors are used.
+Several methods to set the training method were implemented.
+Early stopping, a standard techniques showed good results, mainly to the fact, that the number of training steps is increasing with more data.
+The advantage is that the uncertainty at regions without data are shrink in this way.
+A fixed loss value threshold was also tested, as also used in the original anchore-ensemble technique.
+In the experiments a combination of both was taken.\\
 One of the most crucial points is how this uncertainties are taken into account by the RL algorithms.

### 2020-11-16 main.tex
txs auto checkin

@@
 It is possible to take the average of the models and add Gaussian noise leveled by the standard deviation of the models.
-Another straight forward way is to randomly select a model to provide at each training step, which is the original implementation of the ME-TRPO algorithm and was used in the experiments labeled \it{ME-TRPO}.
+Another straight forward way is to randomly select a model to provide at each training step, which is the original implementation of the ME-TRPO algorithm and was used in the experiments labeled \emph{ME-TRPO}.
 A pessimistic setting only would select the model resulting in the lowest predicted reward.
-Good results were obtained by following a randomly selected single model each full episode and were used in the experiments labeled \it{AE-Dyna}.
+Good results were obtained by following a randomly selected single model each full episode and were used in the experiments labeled \emph{AE-Dyna}.
 

### 2020-11-16 main.tex
txs auto checkin

@@
 A pessimistic setting only would select the model resulting in the lowest predicted reward.
-Good results were obtained by following a randomly selected single model each full episode and were used in the experiments labeled \emph{AE-Dyna}.
+Good results were obtained by following a randomly selected single model each full episode and were used in the experiments labeled \emph{AE-Dyna}.\\
+The number of data-points taken every time the model as improved was firstly determined by the number of initial data points, collect using a random policy.
+The initialization phase has to be chosen not to small to minimize the chance of getting trapped in a local minimum for too long.
+Afterwards at least on full episode should be taken.
+We decided to use a short total episode length to reduce the impact of the compound error, the accumulation error following a wrong model.
+The maximal number of steps was ten.
 

### 2020-11-16 main.tex
txs auto checkin

@@
 	\item The MFRL agent as e.g. the TRPO and PPO or the SAC.
+	\item Tuning the algorithm on a model.
 \end{itemize}
@@
 The maximal number of steps was ten.
+During the data collected on the real system, the latest policy was taken with some small noise added to improve the exploration.\\
+To decide which MFRL algorithm to use, two main algorithm classes have to be considered: the on and the off-policy algorithms.
+Online algorithm show a more stable convergence in general, while off-policy algorithms have the advantage that the data buffer can be filled with real data as well.
+A very attractive off-policy algorithm is the Soft-actor-critic.
+This algorithm tries to maximize the entropy of the actions to find a good trade-off between exploration and exploitation.
+On the other hand the on-policy algorithm TRPO provides some theoretical improvement guarantees.
+In the latter case the policy was improved over the whole training, while for the SAC the policy was reset, to take advantage of the exploration features.
 

### 2020-11-16 main.tex
txs auto checkin

@@
 \section{Deep Reinforcement learning}
-Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $d_0$.
-Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
+Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
+Goal of reinforcement learning is to find a $\policy\mapsto a$, which is the solution of:
 \begin{align}
-\max J(\pi)  =  \mathbb E_{\tau \thicksim p_{\pi}}\left[\sum_{t=0}^{H}\gamma^t r(s_t,a_t)\right],
+\max J (\policy)  =  \mathbb E_{\traj \thicksim p_{\olicyi}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],
 \end{align}
@@
 \begin{align}
-    \min_\theta \left(\vec Q_\theta - \mathcal{B}^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
+    \min_\theta \left(\vec Q_\theta - \bellman^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
 \end{align}
@@
 \begin{align}
- \mathcal{B}^*\vec Q_\theta(s_t,a_t) = r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
+ \bellman^*\vec Q_\theta(s_t,a_t) := r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
 \end{align}
@@
 \end{align}
-$\theta\prime$ denotes the weights of a previous iteration. To stabilize the network training a small artificial noise is added to the actions in the update.
+$\theta\prime$ denotes the weights of a previous iteration.
+To stabilize the network training a small artificial noise is added to the actions in the update.
+This algorithm has an extremely good sample-efficiency for suitable problems as can be found in accelerators.
 %\subsection{New NAF network architecture}
@@
 	\item The episodic design - avoid long trajectories.
-	\item The MFRL agent as e.g. the TRPO and PPO or the SAC.
+	\item The MFRL agent as e.g. the TRPO and PPO or the SAC and its training
 	\item Tuning the algorithm on a model.
@@
 Afterwards at least on full episode should be taken.
-We decided to use a short total episode length to reduce the impact of the compound error, the accumulation error following a wrong model.
+We decided to use a short total episode length to reduce the impact of the compound error, the accumulation error following a wrong model, as well known for Dyna-style approaches.
 The maximal number of steps was ten.

### 2020-11-16 main.tex
txs auto checkin

@@
 \section{Deep Reinforcement learning}
-Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
+Assume states $s$ the set of all states $\bs$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
 Goal of reinforcement learning is to find a $\policy\mapsto a$, which is the solution of:
 \begin{align}
-\max J (\policy)  =  \mathbb E_{\traj \thicksim p_{\olicyi}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],
+\max J (\policy)  =  \mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],
 \end{align}
-where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
+where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
 In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g.

### 2020-11-16 main.tex
txs auto checkin

@@
 \end{align}
-where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
+where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, a_1,\dots s_\horizon,a_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
 In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g.

### 2020-11-16 main.tex
txs auto checkin

@@
 where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, a_1,\dots s_\horizon,a_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
-In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g.
+In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
 a deep neural network.
@@
 \subsection{Approximate dynamic programming}
-The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator.
+The state-value function $Q(\bs,\ba)$ is expressed as $Q_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
 By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:

### 2020-11-16 main.tex
txs auto checkin

@@
 The state-value function $Q(\bs,\ba)$ is expressed as $Q_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
-By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
+By satisfying the Bellmann-optimality operator the optimal $Q^*(\bs,\ba)$ can be found:
 \begin{align}
@@
 \begin{align}
- \bellman^*\vec Q_\theta(s_t,a_t) := r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
+ \bellman^*\vec Q_\theta(\bs_t,\ba_t) := r(\bs_t,\ba_t)+\gamma\max_\ba\left( Q_\theta(\bs_{t+1},\ba_t) - Q_\theta(\bs_t,\ba_t)\right).
 \end{align}
@@
 \begin{align}
-     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
+     Q_\theta(\bs,\ba) = -\frac{1}{2}(\ba-\mu_\theta(\bs))P_\theta(\ba-\mu_\theta(\bs))^T+V_\theta(s).
 \end{align}
- The optimal policy is given by:
+ The policy is given by:
 \begin{align}
-    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
+    \policy_\theta(\bs)=(\delta(\ba)-\arg\!\max_\ba Q_\theta(\bs,\ba)).
 \end{align}
@@
 \begin{align}
-	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
+	\min_\theta\left( (\reward(\bs_t,\ba_t)+\gamma \min_{\hat\theta,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
 \end{align}
-$\theta\prime$ denotes the weights of a previous iteration.
+$\hat\theta$ are the weights of a previous iteration or the inits.
 To stabilize the network training a small artificial noise is added to the actions in the update.

### 2020-11-16 main.tex
txs auto checkin

@@
 \end{align}
-$\hat\theta$ are the weights of a previous iteration or the inits.
+$\hat\theta$ are the weights of a previous iteration or the initialized weights.
 To stabilize the network training a small artificial noise is added to the actions in the update.
-This algorithm has an extremely good sample-efficiency for suitable problems as can be found in accelerators.
+This algorithm has an extremely good sample-efficiency for suitable problems as can be found in accelerators and is used as the baseline for the considered control problem.
 %\subsection{New NAF network architecture}

### 2020-11-17 main.tex
txs auto checkin

@@
 \maketitle
+\begin{itemize}
+	\item Noise benchmarks with pendulum
+	\item Benchmark NAF2 with pendulum
 
+\end{itemize}
 %\tableofcontents
@@
 \begin{align}
- \bellman^*\vec Q_\theta(\bs_t,\ba_t) := r(\bs_t,\ba_t)+\gamma\max_\ba\left( Q_\theta(\bs_{t+1},\ba_t) - Q_\theta(\bs_t,\ba_t)\right).
+ \bellman^*\vec Q_\theta(\bs_t,\ba_t) := r(\bs_t,\ba_t)+\gamma\max_\ba\left( Q_\theta(\bs_{t+1},\ba) - Q_\theta(\bs_t,\ba_t)\right).
 \end{align}
@@
 \begin{align}
-	\min_\theta\left( (\reward(\bs_t,\ba_t)+\gamma \min_{\hat\theta,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
+	\min_\theta\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
 \end{align}
-$\hat\theta$ are the weights of a previous iteration or the initialized weights.
+$\theta_\text{targ}$ are the weights of a target network, which is softly updated.
 To stabilize the network training a small artificial noise is added to the actions in the update.
@@
 Early stopping, a standard techniques showed good results, mainly to the fact, that the number of training steps is increasing with more data.
-The advantage is that the uncertainty at regions without data are shrink in this way.
+The advantage is that the uncertainty at regions without data shrink in this way leading to a better training performance.
 A fixed loss value threshold was also tested, as also used in the original anchore-ensemble technique.

### 2020-11-17 main.tex
txs auto checkin

@@
 	\item Benchmark NAF2 with pendulum
-
+	\item What is our improvement concerning the algorithm
+	\item The first time model based on accelerator problem, 
 \end{itemize}
@@
 One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
-The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
+The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality}:
 \begin{align}
@@
 The episode length is set to 200 steps.
-\section{NAF2 details}\label{appendix:pernaf}
+\subsection{The impact of noise}
+\section{NAF2 details}\label{appendix:naf2}
+\section{AE training details}
+\section{Theoretical aspects on the AE-DYNA on FERMI FEL}
 

### 2020-11-17 AEDYNA.py
- for tests

@@ -53,13 +53,13 @@ else:
     from stable_baselines.common.policies import MlpPolicy
     from stable_baselines import PPO2 as Agent
 
-simulated_steps = 20000
+simulated_steps = 30000
 
 model_batch_size = 100
 num_ensemble_models = 3
 
 early_stopping = True
-model_iter = 2
+model_iter = 10
 
 # model_training_iterations = 10
 network_size = 25
@@ -491,6 +491,7 @@ class NN:
 class NetworkEnv(gym.Wrapper):
     '''
     Wrapper to handle the network interaction
+    Here you can change the treatment of the uncertainty
     '''
 
     def __init__(self, env, model_func=None, done_func=None, number_models=1, **kwargs):
@@ -523,7 +524,8 @@ class NetworkEnv(gym.Wrapper):
             obs, rew = self.model_func(self.obs, [np.squeeze(action)])
         else:
             # Can be activated to randomize each step
-            current_model = np.random.randint(0, max(self.number_models, 1)) # self.current_model
+            # current_model = np.random.randint(0, max(self.number_models, 1)) # self.current_model
+            current_model = self.current_model
             obs, rew = self.model_func(self.obs, [np.squeeze(action)], current_model)
         # obs, rew, _, _ = self.env.step(action)
         self.obs = np.clip(obs.copy(), -1, 1)
@@ -537,7 +539,7 @@ class NetworkEnv(gym.Wrapper):
 
         return self.obs, rew, self.done, dict()
 
-    def visualize(self, data=None, label=None):
+    def visualize(self, data=None, label=None, **kwargs):
         action = [np.zeros(self.env.action_space.shape)]
         state = np.zeros(self.env.observation_space.shape)
         maximum = 0
@@ -550,6 +552,8 @@ class NetworkEnv(gym.Wrapper):
         y = np.arange(-1, 1, delta)
         X, Y = np.meshgrid(x, y)
 
+        if 'data_points' in kwargs:
+            data_points = kwargs.get('data_points')
         if self.number_models == num_ensemble_models:
             Nr = 1
             Nc = 1
@@ -588,6 +592,7 @@ class NetworkEnv(gym.Wrapper):
             #             # plt.plot(np.array(states, dtype=object)[:, 1],)
             #         # images.append(axs[i, j].contour(X, Y, (rewards - 1) / 2, 25, alpha=1))
             #         # axs[i, j].label_outer()
+
             plt.title(maximum)
             # plt.title(label)
[…]

### 2020-11-17 show_models.py
-test

@@ -1,0 +1,63 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+# data = dict(data=data,
+#             model=model_nr,
+#             rews=rews,
+#             X=X,
+#             Y=Y)
+import pickle
+
+project_directory = 'Data_logging/Test_plots/'+'-nr_steps_25-cr_lr-n_ep_7-m_bs_100-sim_steps_2500-m_iter_30-ensnr_3-init_100/'
+files = []
+inner_list = []
+for f in os.listdir(project_directory):
+    if 'plot_model' in f:
+        if f[-1] == '0':
+            if len(inner_list)>0:
+                files.append(inner_list)
+            inner_list = []
+            inner_list.append(f)
+        else:
+            inner_list.append(f)
+
+# print(files)
+number = 3
+with open(project_directory+files[number][0], 'rb') as f:
+    file_data = pickle.load(f)
+# print(file_data)
+
+maximum = 0
+data = file_data['data']
+if data is not None:
+    action = [data[1]]
+    state = data[0]
+    maximum = (data[2] - 1) / 2
+else:
+    action = np.zeros(4)
+    state = np.zeros(4)
+
+delta = 0.05
+x = np.arange(-1, 1, delta)
+y = np.arange(-1, 1, delta)
+X, Y = np.meshgrid(x, y)
+Nr = 1
+Nc = 1
+fig, axs = plt.subplots(Nr, Nc)
+fig.subplots_adjust(hspace=0.3)
+images = []
+for pos in range(len(files[0])):
+    with open(project_directory+files[number][pos], 'rb') as f:
+        file_data = pickle.load(f)
+    rewards= file_data['rews']
+
+    # print(self.number_models)
+    for i1 in range(len(x)):
+        for j1 in range(len(y)):
+            state[0] = x[i1]
+            state[1] = y[j1]
+
+    axs.contour(X, Y, (rewards - 1) / 2, alpha=1)
+
+# plt.colorbar()
+fig.show()

### 2020-11-17 run_naf2.py
- NAF studies...

@@ -1,6 +1,8 @@
 import os
 import pickle
 import random
+
+import gym
 import matplotlib.pyplot as plt
 import numpy as np
 import tensorflow as tf
@@ -15,7 +17,65 @@ random_seed = 111
 np.random.seed(random_seed)
 random.seed(random_seed)
 
-env = PendulumEnv()
+
+class MonitoringEnv(gym.Wrapper):
+    '''
+    Gym Wrapper to store information for scaling to correct scpace and for post analysis.
+    '''
+
+    def __init__(self, env, **kwargs):
+        gym.Wrapper.__init__(self, env)
+        self.rewards = []
+        self.init_rewards = []
+        self.current_episode = -1
+        self.current_step = -1
+
+        self.obs_dim = self.env.observation_space.shape
+        self.obs_high = self.env.observation_space.high
+        self.obs_low = self.env.observation_space.high
+        self.act_dim = self.env.action_space.shape
+        self.act_high = self.env.action_space.high
+        self.act_low = self.env.action_space.low
+
+        # state space definition
+        self.observation_space = gym.spaces.Box(low=-1.0,
+                                                high=1.0,
+                                                shape=self.obs_dim,
+                                                dtype=np.float64)
+
+        # action space definition
+        self.action_space = gym.spaces.Box(low=-1.0,
+                                           high=1.0,
+                                           shape=self.act_dim,
+                                           dtype=np.float64)
+
+    def scale_state_env(self, ob):
+        scale = (self.env.observation_space.high - self.env.observation_space.low)
+        return (2 * ob - (self.env.observation_space.high + self.env.observation_space.low)) / scale
+
+    def scale_rew(self, rew):
+        return rew/100
+
+    def reset(self,**kwargs):
+        self.current_step = 0
+        self.current_episode += 1
+        self.rewards.append([])
+        return self.scale_state_env(self.env.reset(**kwargs))
+
+    def step(self, action):
+        self.current_step += 1
+        ob, reward, done, info = self.env.step(action)
+        self.rewards[self.current_episode].append(reward)
+        if self.current_step>=200:
+            done = True
+        ob = self.scale_state_env(ob)
+        reward = self.scale_rew(reward)
+
+        return ob, reward, done, info
+
+env = MonitoringEnv(env=PendulumEnv())
+
+
 
[…]

### 2020-11-17 main.tex
txs auto checkin

@@
 	\item What is our improvement concerning the algorithm
-	\item The first time model based on accelerator problem, 
+	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
+	\item Tailored to accelerators - short horizons are standard?
+	\item Improvements of the algorithms:
+	\begin{itemize}
+		\item Adding the noise feature
+		\item Jumping feature using SAC
+		\item Noise in the policy
+		\item discuss different approaches to e.g. dynamic waiting time
+	\end{itemize}
 \end{itemize}
@@
 % \bibliography{bibliography}
-\addcontentsline{toc}{chapter}{Bibliography}
-
-\begin{thebibliography}{9}
-
-\bibitem{lsa} D.
+%\addcontentsline{toc}{chapter}{Bibliography}
+%
+%\begin{thebibliography}{9}
+%
+%\bibitem{lsa} D.
 Jacquet, R.
@@
 Kruk, "LSA - The High Level Application Software of the LHC and its Performance during the first 3 years of Operation", ICALEPS, San Francisco, CA, USA, 6 - 11 Oct 2013, pp.thppc058.
-
-\bibitem{multiobjective-optimization} A.
+%
+%\bibitem{multiobjective-optimization} A.
 Edelen, N.
@@
 Beams 23, 044601, 2020.
-
-\bibitem{collimator-alignment} G.
+%
+%\bibitem{collimator-alignment} G.
 Azzopardi, A.
@@
 1208-1211, 2019.
-
-\bibitem{SPSAlignment} %InProceedings (Hirlaender2019)
- S. Hirlaender, M. Fraser, B. Goddard , V. Kain, J. Prieto, L. Stoel, M. Szakaly, F. Velotti, "Automatisation of the SPS ElectroStatic Septa Alignment",  
-in 10th Int. Partile Accelerator Conf.(IPAC'19), 2019, p. 4001-4004.
-
-\bibitem{liu} E.
+%
+%\bibitem{SPSAlignment} %InProceedings (Hirlaender2019)
+% S. Hirlaender, M. Fraser, B. Goddard , V. Kain, J. Prieto, L. Stoel, M. Szakaly, F. Velotti, "Automatisation of the SPS ElectroStatic Septa Alignment",  
+%in 10th Int. Partile Accelerator Conf.(IPAC'19), 2019, p. 4001-4004.
+%
+%\bibitem{liu} E.
 Shaposhnikova et al., "LHC Injectors Upgrade (LIU) Project at CERN", 7$^{th}$ International Particle Accelerator Conference, Busan, Korea, 8 - 13 May 2016, pp.MOPOY059, \url{https://doi.org/10.18429/JACoW-IPAC2016-MOPOY059}.
-
-\bibitem{awake} E.
+%
+%\bibitem{awake} E.
 Adli, A.
@@
 Bernardini \textit{et al.}, “Acceleration of electrons in the plasma wakefield of a proton bunch,” Nature 561, 363–367 (2018), \url{https://doi.org/10.1038/s41586-018-0485-4)}.
-
-\bibitem{scheinker} A.
+%
+%\bibitem{scheinker} A.
 Scheinker \textit{et al.}, "Online multi-objective particle accelerator optimization of the AWAKE electron beam line for simultaneous emittance and orbit control", AIP Advances 10, 055320 (2020), \url{https://doi.org/10.1063/5.0003423}.
-
-\bibitem{barto-sutton} R. Sutton, A. Barto, "Introduction to Reinforcement Learning",  MIT Press, Cambridge, MA, USA, 2018.
-
-\bibitem{LQR} H. Kwakernaak, R. Sivan, "Linear Optimal Control Systems", Wiley-Interscience, 1972.
-
-\bibitem{SBP} A. Nagabandi, G. Kahn, R. S. Fearing, S. Levine, "Neural Network Dynamics for Model-Based Deep Reinforcement Learning with Model-Free Fine-Tuning", arXiv:1708.02596, 2017.
-
-\bibitem{DYNA} R. Sutton, "Dyna, an integrated architecture for learning, planning, and reacting", AAAI Spring Symposium, pp. 151-155, 1991, \url{https://doi.org/10.1145/122344.122377}.
-
-\bibitem{benchmarkmbrl} T. Wang et al., “Benchmarking Model-Based Reinforcement Learning”,  \url{https://arxiv.org/abs/1907.02057v1}, 2019-07-03.
-
-\bibitem{dqn} V. Mnih \textit{et al.}, "Human-level control through deep reinforcement learning", Nature 518, 529–533 (2015).
-
-\bibitem{ddpg} T. Lillicrap~\textit{et al.}, "Continuous control with deep reinforcement learning", Proc. ICLR 2016 \url{https://arxiv.org/abs/1509.02971}.
-
-\bibitem{naf} S.
+%
+%\bibitem{barto-sutton} R. Sutton, A. Barto, "Introduction to Reinforcement Learning",  MIT Press, Cambridge, MA, USA, 2018.
+%
+%\bibitem{LQR} H. Kwakernaak, R. Sivan, "Linear Optimal Control Systems", Wiley-Interscience, 1972.
+%
+%\bibitem{SBP} A. Nagabandi, G. Kahn, R. S. Fearing, S. Levine, "Neural Network Dynamics for Model-Based Deep Reinforcement Learning with Model-Free Fine-Tuning", arXiv:1708.02596, 2017. 
+%
+%\bibitem{DYNA} R. Sutton, "Dyna, an integrated architecture for learning, planning, and reacting", AAAI Spring Symposium, pp. 151-155, 1991, \url{https://doi.org/10.1145/122344.122377}.
+%
+%\bibitem{benchmarkmbrl} T. Wang et al., “Benchmarking Model-Based Reinforcement Learning”,  \url{https://arxiv.org/abs/1907.02057v1}, 2019-07-03.
+%
+%\bibitem{dqn} V. Mnih \textit{et al.}, "Human-level control through deep reinforcement learning", Nature 518, 529–533 (2015).
+%
+%\bibitem{ddpg} T. Lillicrap~\textit{et al.}, "Continuous control with deep reinforcement learning", Proc. ICLR 2016 \url{https://arxiv.org/abs/1509.02971}.
+%
+%\bibitem{naf} S.
 Gu, T.
@@
 33rd International Conference on Machine Learning, New York, NY, USA, 2016.
-
-\bibitem{td3} S. Fujimoto, H. van Hoof, D. Meger, "Addressing Function Approximation Error in Actor-Critic Methods", arXiv:1802.09477, 2018.
-
-\bibitem{deeppilco} Y. Gal, R. McAllister, and C. E. Rasmussen, “Improving PILCO with Bayesian neural network dynamics models”, in Data-Efficient Machine Learning workshop, ICML, 2016, vol. 4, p. 34.
-
-\bibitem{per} T. Schaul, J. Quan, I. Antonoglou, D. Silver, "Prioritized Experience Replay", arXiv:1511.05952, 2015-11-18.
-
-\bibitem{pernafgit} S. Hirlaender, \textit{PER-NAF} available at \url{https://github.com/MathPhysSim/PER-NAF/} and
-\url{https://pypi.org/project/pernaf/}.
-\bibitem{pyjapc} "pyjapc" available at \url{https://pypi.org/project/pyjapc/}.
-
-\bibitem{openai_gym} \url{http://gym.openai.com}
-
-\bibitem{madx} MAD-X documentation and source code available at \url{https://mad.web.cern.ch/mad/}.
-
-\bibitem{linac4} G.
+%
+%\bibitem{td3} S. Fujimoto, H. van Hoof, D. Meger, "Addressing Function Approximation Error in Actor-Critic Methods", arXiv:1802.09477, 2018.
+%
+%\bibitem{deeppilco} Y. Gal, R. McAllister, and C. E. Rasmussen, “Improving PILCO with Bayesian neural network dynamics models”, in Data-Efficient Machine Learning workshop, ICML, 2016, vol. 4, p. 34.
+%
+%\bibitem{per} T. Schaul, J. Quan, I. Antonoglou, D. Silver, "Prioritized Experience Replay", arXiv:1511.05952, 2015-11-18.
+%
+%\bibitem{pernafgit} S. Hirlaender, \textit{PER-NAF} available at \url{https://github.com/MathPhysSim/PER-NAF/} and
+%\url{https://pypi.org/project/pernaf/}.
+%\bibitem{pyjapc} "pyjapc" available at \url{https://pypi.org/project/pyjapc/}.
+%
+%\bibitem{openai_gym} \url{http://gym.openai.com}
+%
+%\bibitem{madx} MAD-X documentation and source code available at \url{https://mad.web.cern.ch/mad/}.
+%
+%\bibitem{linac4} G.
 Bellodi, "Linac4 Commissioning Status and Challenges to Nominal Operation", 61$^st$ ICFA Advanced Beam Dynamics Workshop on High-Intensity and High-Brightness Hadron Beams, Daejeon, Korea, 17 - 22 Jun 2018, pp.MOA1PL03, \url{https://doi.org/10.18429/JACoW-HB2018-MOA1PL03}.
-
-\bibitem{auto_matching} F. Velotti, B. Goddard et al., "Automatic AWAKE electron beamline setup using unsupervised machine learning", to be submitted to Phys. Rev. Accel. Beams.
-
-\bibitem{Kumar_1} A. Kumar, A. Gupta, S. Levine, “DisCor: Corrective Feedback in Reinforcement Learning via Distribution Correction,” arXiv:2003.07305, 2020-03-16.
-
-\bibitem{Kumar_2} A. Kumar, J. Fu, G. Tucker, S. Levine, “Stabilizing Off-Policy Q-Learning via Bootstrapping Error Reduction,” arXiv:1906.00949, 2019-06-03.
-
-% NEW REFERENCES
-\bibitem[bruchon2020basic]{Bruchon, Niky, et al. "Basic Reinforcement Learning Techniques to Control the Intensity of a Seeded Free-Electron Laser." Electronics 9.5 (2020): 781.}
-% https://www.mdpi.com/2079-9292/9/5/781/htm
-
-\end{thebibliography}
+%
+%\bibitem{auto_matching} F. Velotti, B. Goddard et al., "Automatic AWAKE electron beamline setup using unsupervised machine learning", to be submitted to Phys. Rev. Accel. Beams.
+%
+%\bibitem{Kumar_1} A. Kumar, A. Gupta, S. Levine, “DisCor: Corrective Feedback in Reinforcement Learning via Distribution Correction,” arXiv:2003.07305, 2020-03-16.
+%
+%\bibitem{Kumar_2} A. Kumar, J. Fu, G. Tucker, S. Levine, “Stabilizing Off-Policy Q-Learning via Bootstrapping Error Reduction,” arXiv:1906.00949, 2019-06-03.
+%
+%% NEW REFERENCES
+%\bibitem[bruchon2020basic]{Bruchon, Niky, et al. "Basic Reinforcement Learning Techniques to Control the Intensity of a Seeded Free-Electron Laser." Electronics 9.5 (2020): 781.}
+%% https://www.mdpi.com/2079-9292/9/5/781/htm
+%
+%
+%
+%\end{thebibliography}
 

### 2020-11-17 main.tex
txs auto checkin

@@
 \section{Deep Reinforcement learning}
-Assume states $s$ the set of all states $\bs$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
-Goal of reinforcement learning is to find a $\policy\mapsto a$, which is the solution of:
+Assume states $\s$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
+Goal of reinforcement learning is to find a $\policy$ $\states\mapsto \actions$, which is the solution of:
 \begin{align}
@@
 \end{align}
-where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, a_1,\dots s_\horizon,a_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
+where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
 In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.

### 2020-11-17 main.tex
txs auto checkin

@@
 \section{Deep Reinforcement learning}
-Assume states $\s$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
+Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
 Goal of reinforcement learning is to find a $\policy$ $\states\mapsto \actions$, which is the solution of:

### 2020-11-17 main.tex
txs auto checkin

@@
 Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
-Goal of reinforcement learning is to find a $\policy$ $\states\mapsto \actions$, which is the solution of:
+Goal of reinforcement learning is to find a $\policy: \bs\mapsto \ba$, which is the solution of:
 \begin{align}
@@
 The state-value function $Q(\bs,\ba)$ is expressed as $Q_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
-By satisfying the Bellmann-optimality operator the optimal $Q^*(\bs,\ba)$ can be found:
+By satisfying the Bellmann-optimality equation the optimal $Q^*(\bs,\ba)$ can be found:
 \begin{align}

### 2020-11-18 main.tex
txs auto checkin
-changes in NAF2

@@
 The state-value function $Q(\bs,\ba)$ is expressed as $Q_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
-By satisfying the Bellmann-optimality equation the optimal $Q^*(\bs,\ba)$ can be found:
+By satisfying the Bellmann-optimality equation $Q_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$:
 \begin{align}
@@
 \end{align}
-This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
+And $\policy$ can be calculated via:
+\begin{align}
+	\policy_\theta(\bs)=(\delta(\ba)-\arg\!\max_\ba Q_\theta(\bs,\ba)).
+\end{align}
+The Bellman operator $\bellman^*$ has a unique fixed point but is non-linear, due to the $\max$ - operator:
 \begin{align}
@@
 \end{align}
- The policy is given by:
-\begin{align}
-    \policy_\theta(\bs)=(\delta(\ba)-\arg\!\max_\ba Q_\theta(\bs,\ba)).
-\end{align}
+
 One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
+It is motivated by double Q-learning \cite{NIPS2010_091d584f}.
 The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality}:
@@
 \section{Uncertainty aware Dyna-style reinforcement learning}
-The original Dyna algorithm \cite{DYNA} is modified here in several aspects.
-Generally, Dyna style algorithms denote algorithms, where a MFRL algorithm is trained on purely synthetic data from an approximate dynamics model or on a mixture of synthetic and real data.
+The original Dyna algorithm \cite{Kurutach2018} is modified here in several aspects.
+Generally, Dyna style algorithms \cite{Sutton1991} denote algorithms, where a MFRL algorithm is trained on purely synthetic data from an approximate dynamics model or on a mixture of synthetic and real data.
 We use only synthetic data to reduce the interaction with the real environment to a minimum.
@@
 At the beginning the data is collected or read in from a pre-collection.
-An uncertainty aware model is trained, using anchored ensembles on the data, which allows to take the allegorical (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
+An uncertainty aware model is trained, using anchored ensembles \cite{Pearce2018} on the data, which allows to take the allegorical (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
 Subsequently a MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data, by taking the uncertainty into account.
@@
 % \bibliography{apssamp}% Produces the bibliography via BibTeX.
+ \bibliography{tex/Bibliography}
 

### 2020-11-18 AEDYNA.py

@@ -33,7 +33,7 @@ real_env = env = PendulumEnv()
 
 steps_per_env = 201
 init_random_steps = 201
-total_steps = 10000
+total_steps = 1004
 num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
 
 print('Number of epochs: ', num_epochs)
@@ -53,10 +53,10 @@ else:
     from stable_baselines.common.policies import MlpPolicy
     from stable_baselines import PPO2 as Agent
 
-simulated_steps = 30000
+simulated_steps = 25000
 
 model_batch_size = 100
-num_ensemble_models = 3
+num_ensemble_models = 5
 
 early_stopping = True
 model_iter = 10
@@ -524,8 +524,8 @@ class NetworkEnv(gym.Wrapper):
             obs, rew = self.model_func(self.obs, [np.squeeze(action)])
         else:
             # Can be activated to randomize each step
-            # current_model = np.random.randint(0, max(self.number_models, 1)) # self.current_model
-            current_model = self.current_model
+            current_model = np.random.randint(0, max(self.number_models, 1)) # self.current_model
+            # current_model = self.current_model
             obs, rew = self.model_func(self.obs, [np.squeeze(action)], current_model)
         # obs, rew, _, _ = self.env.step(action)
         self.obs = np.clip(obs.copy(), -1, 1)
@@ -1206,9 +1206,9 @@ def aedyna(env_name, cr_lr=5e-3, num_epochs=50,
                 plot_observables(data=data, label=label)
 
                 # stop training if the policy hasn't improved
-                if (np.sum(best_sim_test >= sim_rewards) > int(num_ensemble_models * 0.7)):
+                if (np.sum(best_sim_test >= sim_rewards) >= int(num_ensemble_models * 0.7)):
                     if it > delay_before_convergence_check and ep < num_epochs - 1:
-                        print('break-no improvement')
+                        print('break-no improvement in', int(num_ensemble_models * 0.7), ' models')
                         break
                 else:
                     best_sim_test = sim_rewards

### 2020-11-18 naf2.py

@@ -414,7 +414,7 @@ class QModel:
             del kwargs['discount']
         else:
             self.polyak = 0.999
-        batch = self.replay_buffer.sample_batch(batch_size=1000000)
+        batch = self.replay_buffer.sample_batch(batch_size=20000)
         # batch, prios = self.replay_buffer.sample_batch(batch_size=batch_size)
         # nr = self.replay_buffer.size
         #
@@ -648,10 +648,10 @@ class NAF(object):
         self.losses = []
         self.pretune = pretune
         # self.prio_info = prio_info
-        self.prio_info = dict()
-        self.per_flag = bool(self.prio_info)
-        # self.per_flag = False
-        print('PER is:', self.per_flag)
+        # self.prio_info = dict()
+        # self.per_flag = bool(self.prio_info)
+        # # self.per_flag = False
+        # print('PER is:', self.per_flag)
 
         self.env = env
 
@@ -766,7 +766,6 @@ class NAF(object):
     def predict(self, model, state, is_train):
 
         if is_train and model.replay_buffer.size < self.warm_up_steps:
-            print(10 * 'inits ')
             action = model.de_normalize(np.random.uniform(-1, 1, self.action_size), model.act_box)
             # print(action)
             return np.array(action)
@@ -865,7 +864,8 @@ class NAF(object):
                 o = o2
                 d = False if t == self.max_steps - 1 else d
 
-                if t % self.initial_episode_length == 0 and self.q_main_model_1.replay_buffer.size <= self.warm_up_steps:
+                if t % self.initial_episode_length == 0 and \
+                        self.q_main_model_1.replay_buffer.size <= self.warm_up_steps:
                     o = self.env.reset()
                     self.init_trajectory_data(state=o)
                     print('Initial reset at ', t)
@@ -889,6 +889,7 @@ class NAF(object):
         #     else:
         #         batch, priority_info = model.replay_buffer.sample_normal(batch_size=self.batch_size)
         # else:
+        # Generate batch for monitoring the performance
         batch = model.replay_buffer.sample_batch(200)
 
         # o = batch['obs1']

### 2020-11-18 run_naf2.py

@@ -7,7 +7,6 @@ import matplotlib.pyplot as plt
 import numpy as np
 import tensorflow as tf
 
-
 from inverted_pendulum import PendulumEnv
 
 from naf2 import NAF
@@ -54,9 +53,9 @@ class MonitoringEnv(gym.Wrapper):
         return (2 * ob - (self.env.observation_space.high + self.env.observation_space.low)) / scale
 
     def scale_rew(self, rew):
-        return rew/100
+        return rew / 100
 
-    def reset(self,**kwargs):
+    def reset(self, **kwargs):
         self.current_step = 0
         self.current_episode += 1
         self.rewards.append([])
@@ -66,15 +65,15 @@ class MonitoringEnv(gym.Wrapper):
         self.current_step += 1
         ob, reward, done, info = self.env.step(action)
         self.rewards[self.current_episode].append(reward)
-        if self.current_step>=200:
+        if self.current_step >= 200:
             done = True
         ob = self.scale_state_env(ob)
         reward = self.scale_rew(reward)
 
         return ob, reward, done, info
 
-env = MonitoringEnv(env=PendulumEnv())
 
+env = MonitoringEnv(env=PendulumEnv())
 
 
 def plot_results(env, file_name):
@@ -152,6 +151,7 @@ def plot_results(env, file_name):
     plt.savefig(file_name + '_rewards.pdf')
     plt.show()
 
+
 def plot_convergence(agent, file_name):
     losses, vs = agent.losses, agent.vs
     # losses2, vs2 = agent.losses2, agent.vs2
@@ -178,6 +178,7 @@ def plot_convergence(agent, file_name):
     plt.savefig(file_name + '_convergence' + '.pdf')
     plt.show()
 
+
 if __name__ == '__main__':
 
     directory = "checkpoints/pendulum/"
@@ -189,11 +190,11 @@ if __name__ == '__main__':
     nafnet_kwargs = dict(hidden_sizes=[100, 100], activation=tf.nn.tanh
                          , weight_init=tf.random_uniform_initializer(-0.05, 0.05, seed=random_seed))
 
-    noise_info = dict(noise_function=lambda nr: max(0., 1-(nr/25)))
+    noise_info = dict(noise_function=lambda nr: max(0., 1 - (nr / 25)))
 
     # the target network is updated at the end of each episode
     # the number of episodes is executed each step in the environment
[…]

### 2020-11-18 main.tex
- converging NAF2

@@
 In this paper we discuss a model based reinforcement learning approach in comparison to a model free reinforcement learning approach applied at the FERMI FEL system.
-Both algorithms and approaches are new in this context and the main purpose of this paper is to show that reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems.
-In terms of sample-complexity the model-based approach is faster, while the finale performance of the model free method is superior - the so called asymptotic performance.
+Both algorithms and approaches are new in this context and the main purpose of this paper is to show how reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems.
+In terms of sample-complexity the model-based approach is faster, while the final performance of the model free method is superior - the so called asymptotic performance.
 The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency
@@
 	\item Benchmark NAF2 with pendulum
-	\item What is our improvement concerning the algorithm
 	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
@@
 Especially if there is no way to model the physics, one might use optimisation techniques as e.g.
-derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes.
-Another promising way is to apply reinforcement learning which has several advantages over optimisation methods:
+derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes to restore or maintain the performance.
+Another promising way is to apply reinforcement learning which shows several advantages over optimisation methods:
 \begin{itemize}
-    \item Covers a larger class of problems.
-    \item Solves the problem not always completely from the beginning as an DFO.
-    \item Ability to use existing data.
+    \item It covers a larger class of problems, RL optimises a sequence of decisions.
+    \item It memorizes the problem and does not start always from scratch as an DFO.
+    \item Existing data can be used.
     \item The underlying structure of the problem might be deduced.
 \end{itemize}
-RL optimises a sequence of decisions. On critical aspect using RL is the number of iteration to train a controller.
+On critical aspect using RL is the number of iteration to train a controller.
 In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.

### 2020-11-18 naf2.py

@@ -166,9 +166,9 @@ class QModel:
         h = self.normalize(inputs_state, box=self.obs_box)
         for hidden_dim in hidden_sizes:
             h = self.fc(h, hidden_dim)
-        V = self.fc(h, 1, name='V')
+        V = self.fc(h, 1, activation=None, name='V')
 
-        l = self.fc(h, (self.act_dim * (self.act_dim + 1) / 2))
+        l = self.fc(h, (self.act_dim * (self.act_dim + 1) / 2), activation=None, name='l')
         mu = self.fc(h, self.act_dim, name='mu')
 
         # action = inputs[:, obs_dim:]
@@ -241,21 +241,23 @@ class QModel:
     def normalize(self, input, box):
         low = tf.convert_to_tensor(box.low, dtype=tf.float64)
         high = tf.convert_to_tensor(box.high, dtype=tf.float64)
-        return tf.math.scalar_mul(tf.convert_to_tensor(2, dtype=tf.float64),
-                                  tf.math.add(tf.convert_to_tensor(-0.5, dtype=tf.float64),
-                                              tf.multiply(tf.math.add(input, -low), 1 / (high - low))))
+        # return tf.math.scalar_mul(tf.convert_to_tensor(2, dtype=tf.float64),
+        #                           tf.math.add(tf.convert_to_tensor(-0.5, dtype=tf.float64),
+        #                                       tf.multiply(tf.math.add(input, -low), 1 / (high - low))))
+        return input
 
     def de_normalize(self, input, box):
         low = tf.convert_to_tensor(box.low, dtype=tf.float64)
         high = tf.convert_to_tensor(box.high, dtype=tf.float64)
         input = tf.convert_to_tensor(input, dtype=tf.float64)
-        return tf.math.add(
-            tf.multiply(tf.math.add(tf.math.scalar_mul(tf.convert_to_tensor(0.5, dtype=tf.float64), input),
-                                    tf.convert_to_tensor(0.5, dtype=tf.float64)),
-                        (high - low)), low)
+        # return tf.math.add(
+        #     tf.multiply(tf.math.add(tf.math.scalar_mul(tf.convert_to_tensor(0.5, dtype=tf.float64), input),
+        #                             tf.convert_to_tensor(0.5, dtype=tf.float64)),
+        #                 (high - low)), low)
+        return input
 
-    def fc(self, x, hidden_size, name=None):
[…]

### 2020-11-18 run_naf2.py

@@ -17,65 +17,6 @@ np.random.seed(random_seed)
 random.seed(random_seed)
 
 
-class MonitoringEnv(gym.Wrapper):
-    '''
-    Gym Wrapper to store information for scaling to correct scpace and for post analysis.
-    '''
-
-    def __init__(self, env, **kwargs):
-        gym.Wrapper.__init__(self, env)
-        self.rewards = []
-        self.init_rewards = []
-        self.current_episode = -1
-        self.current_step = -1
-
-        self.obs_dim = self.env.observation_space.shape
-        self.obs_high = self.env.observation_space.high
-        self.obs_low = self.env.observation_space.high
-        self.act_dim = self.env.action_space.shape
-        self.act_high = self.env.action_space.high
-        self.act_low = self.env.action_space.low
-
-        # state space definition
-        self.observation_space = gym.spaces.Box(low=-1.0,
-                                                high=1.0,
-                                                shape=self.obs_dim,
-                                                dtype=np.float64)
-
-        # action space definition
-        self.action_space = gym.spaces.Box(low=-1.0,
-                                           high=1.0,
-                                           shape=self.act_dim,
-                                           dtype=np.float64)
-
-    def scale_state_env(self, ob):
-        scale = (self.env.observation_space.high - self.env.observation_space.low)
-        return (2 * ob - (self.env.observation_space.high + self.env.observation_space.low)) / scale
-
-    def scale_rew(self, rew):
-        return rew / 100
-
-    def reset(self, **kwargs):
-        self.current_step = 0
-        self.current_episode += 1
-        self.rewards.append([])
-        return self.scale_state_env(self.env.reset(**kwargs))
-
-    def step(self, action):
-        self.current_step += 1
-        ob, reward, done, info = self.env.step(action)
-        self.rewards[self.current_episode].append(reward)
-        if self.current_step >= 200:
-            done = True
-        ob = self.scale_state_env(ob)
-        reward = self.scale_rew(reward)
-
-        return ob, reward, done, info
-
-
-env = MonitoringEnv(env=PendulumEnv())
-
-
 def plot_results(env, file_name):
     # plotting
     print('Now plotting')
@@ -179,9 +120,113 @@ def plot_convergence(agent, file_name):
     plt.show()
 
 
+class MonitoringEnv(gym.Wrapper):
+    '''
[…]

### 2020-11-23 main.tex
- converging NAF2

@@
 \section{A non-linear standard control problem}
-To provide some transparency of these studies we provide results on a famous classical standard control problem, the inverted pendulum.
-It is a non-linear low dimensional unsolved continuous control problem, which means there is no threshold for the reward to terminate an episode.
+To provide some transparency of these studies for other labs we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.
+It is a non-linear low dimensional unsolved continuous control problem.
+Unsolved means there is no threshold for the reward to terminate an episode.
 The episode length is set to 200 steps.
+In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly in terms of noise handling.
+This is of importance when dealing with measurements on real systems.
 \subsection{The impact of noise}
+We add artificial Gaussian noise $\gauss(0,0.05)$ on the state with and compare the different \emph{NAF} variants: \emph{smoothing-double}, \emph{smoothing single}, \emph{standart}.
 \section{NAF2 details}\label{appendix:naf2}

### 2020-11-23 Benchmark_NAF.py

@@ -1,0 +1,83 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pickle
+
+main_dir = 'checkpoints'
+
+
+def read_rews(rewards):
+    iterations = []
+    final_rews = []
+    sum_rews = []
+    mean_rews = []
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 0):
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+            iterations.append(len(rewards[i]))
+            sum_rews.append(np.sum(rewards[i]))
+            mean_rews.append(np.mean(rewards[i]))
+    return iterations, final_rews, sum_rews, mean_rews
+
+
+def plot_results(rewards_list, file_name=None):
+    fig, ax = plt.subplots(1, 1)
+    # ax1 = plt.twinx(ax)
+
+    for rewards in rewards_list:
+        iterations, final_rews, sum_rews, mean_rews = read_rews(rewards)
+        ax.plot(sum_rews)
+        # ax1.plot(final_rews)
+    # color = 'lime'
+    # ax1.set_ylabel('finals', color=color)
+    # ax1.tick_params(axis='y', labelcolor=color)
+
+    ax.set_ylabel('cum. reward')
+    ax.set_xlabel('# episode')
+    plt.show()
+
+
+def plot_convergence(agent, file_name):
+    losses, vs = agent.losses, agent.vs
+    # losses2, vs2 = agent.losses2, agent.vs2
+
+    fig, ax = plt.subplots()
+    # ax.set_title(label)
+    ax.set_xlabel('# steps')
+
+    color = 'tab:blue'
+
+    # ax.semilogy(losses2, color=color)
+    ax.tick_params(axis='y', labelcolor=color)
+    ax.set_ylabel('td_loss', color=color)
+    ax.semilogy(losses, color=color)
+    # ax.set_ylim(0, 1)
+
+    ax1 = plt.twinx(ax)
+    # ax1.set_ylim(-2, 1)
+    color = 'lime'
+    ax1.set_ylabel('V', color=color)  # we already handled the x-label with ax1
+    ax1.tick_params(axis='y', labelcolor=color)
+    ax1.plot(vs, color=color)
+    # ax1.plot(vs2, color=color)
+    plt.savefig(file_name + '_convergence' + '.pdf')
+    plt.show()
+
+def load_pickle_final(test_name):
+    directory = 'checkpoints/' + test_name + '/'
[…]

### 2020-11-23 modelling_keras.py

@@ -1,0 +1,74 @@
+import numpy as np
+import tensorflow as tf
+from tensorflow import keras
+from tensorflow.keras import layers
+
+
+def fc(x, hidden_size, activation=tf.nn.tanh,
+       kernel_initializer=tf.compat.v1.random_uniform_initializer(-0.01, 0.01),
+       name=None):
+    layer = keras.layers.Dense(hidden_size, activation=activation,
+                               kernel_initializer=kernel_initializer,
+                               kernel_regularizer=None,
+                               bias_initializer=tf.compat.v1.constant_initializer(0.0), name=name)
+    return layer(x)
+
+
+kernel_initializer = tf.compat.v1.random_uniform_initializer(-0.01, 0.01)
+
+obs_dim = 2
+act_dim = 2
+hidden_sizes = (100, 100)
+inputs_state = keras.Input(shape=(obs_dim,), name="state_input")
+
+# h = inputs[:, 0:obs_dim]
+h = inputs_state
+for hidden_dim in hidden_sizes:
+    h = fc(h, hidden_dim, kernel_initializer=kernel_initializer)
+V = fc(h, 1, activation=None, kernel_initializer=kernel_initializer, name='V')
+
+l = fc(h, (act_dim * (act_dim + 1) / 2),
+       kernel_initializer=kernel_initializer, name='l')
+mu = fc(h, act_dim, kernel_initializer=kernel_initializer, name='mu')
+state_out_put = [V, l, mu]
+
+value_model = keras.Model(inputs_state, V, name='value_model')
+action_model = keras.Model(inputs_state, mu, name='action_model')
+
+inputs_action = keras.Input(shape=(act_dim,), name="action_input")
+
+pivot = 0
+rows = []
+for idx in range(act_dim):
+    count = act_dim - idx
+    diag_elem = tf.exp(tf.slice(l, (0, pivot), (-1, 1)))
+    non_diag_elems = tf.slice(l, (0, pivot + 1), (-1, count - 1))
+    row = tf.pad(tensor=tf.concat((diag_elem, non_diag_elems), 1), paddings=((0, 0), (idx, 0)))
+    rows.append(row)
+    pivot += count
+L = tf.transpose(a=tf.stack(rows, axis=1), perm=(0, 2, 1))
+P = tf.matmul(L, tf.transpose(a=L, perm=(0, 2, 1)))
[…]

### 2020-11-23 naf2_new.py

@@ -1,8 +1,8 @@
 import os
 import pickle
 import shutil
-import time
-
+import matplotlib.pyplot as plt
+import gym
 import tensorflow as tf
 from tensorflow import keras
 
@@ -69,7 +69,53 @@ def basic_loss_function(y_true, y_pred):
     return tf.math.reduce_mean(y_true - y_pred)
 
 
+class NormalizeEnv(gym.Wrapper):
+    '''
+    Gym Wrapper to normalize the environment
+    '''
+
+    def __init__(self, env, **kwargs):
+        gym.Wrapper.__init__(self, env)
+
+        self.obs_dim = self.env.observation_space.shape
+        self.obs_high = self.env.observation_space.high
+        self.obs_low = self.env.observation_space.high
+        self.act_dim = self.env.action_space.shape
+        self.act_high = self.env.action_space.high
+        self.act_low = self.env.action_space.low
+
+        # state space definition
+        self.observation_space = gym.spaces.Box(low=-1.0,
+                                                high=1.0,
+                                                shape=self.obs_dim,
+                                                dtype=np.float64)
+
+        # action space definition
+        self.action_space = gym.spaces.Box(low=-1.0,
+                                           high=1.0,
+                                           shape=self.act_dim,
+                                           dtype=np.float64)
+
+    def reset(self, **kwargs):
+        return self.scale_state_env(self.env.reset(**kwargs))
+
+    def step(self, action):
+        # TODO: check the dimensions
+        ob, reward, done, info = self.env.step(self.descale_action_env(action)[0])
+        return self.scale_state_env(ob), reward, done, info
+
+    def descale_action_env(self, act):
+        scale = (self.env.action_space.high - self.env.action_space.low)
+        return_value = (scale * act + self.env.action_space.high + self.env.action_space.low) / 2
+        return return_value
+
+    def scale_state_env(self, ob):
+        scale = (self.env.observation_space.high - self.env.observation_space.low)
+        return (2 * ob - (self.env.observation_space.high + self.env.observation_space.low)) / scale
+
+
 class QModel:
+    """Artificial neural net holding the state-action value function in a simple analytical form"""
 
     def __init__(self, obs_dim=2, act_dim=2, **kwargs):
         if 'directory' in kwargs:
@@ -95,11 +141,22 @@ class QModel:
             self.__name__ = kwargs.get('name')
             print(self.__name__)
 
+        if 'learning_rate' in kwargs:
[…]

### 2020-11-23 read_paper_tests.py

@@ -1,0 +1,54 @@
+import os
+import pickle
+import sys
+
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+
+def read_experiment(root_dir):
+    data_all = pd.DataFrame()
+    n_average = 15
+    for root, dirs, files in os.walk(root_dir):
+        for file in files:
+            if file == 'plot_data_0.pkl':
+                file_name = root + os.sep + file
+                print(file_name)
+                openfile = open(file_name, 'rb')
+                data = np.sum(pickle.load(openfile), axis=1)
+                # data_frame = pd.DataFrame(data, columns=[root.split('/')[0]])
+                data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
+                data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
+                data_all = pd.concat([data_all, data_frame.T], sort=False)
+                openfile.close()
+
+    return data_all
+
+
+try:
+    root_dir = sys.argv[1]
+except:
+    # root_dir = "SINGLE/"
+    # single = read_experiment(root_dir=root_dir)
+    # root_dir = "DOUBLE/"
+    # double = read_experiment(root_dir=root_dir)
+
+    # data_all = pd.concat([single, double], sort=False)
+    root_dir = "Experiments-150-200-50/"
+    data_all = read_experiment(root_dir=root_dir)
+
+figure_name = 'Smoothing_comparison.pdf'
+current_test_name = ''
+
+fig, axs = plt.subplots(1, 1)
+
+# data_all = data_all.loc[:, :length]
+print(data_all)
+
+sns.lineplot(ax=axs, data=data_all.T)
+plt.ylabel(' cum reward (arb. units.)')
+plt.xlabel('no. episode')
+plt.savefig(root_dir + figure_name)
+plt.show()

### 2020-11-23 run_naf2.py

@@ -1,6 +1,7 @@
 import os
 import pickle
 import random
+import sys
 
 import gym
 import matplotlib.pyplot as plt
@@ -92,7 +93,6 @@ def plot_results(env, file_name):
     plt.savefig(file_name + '_rewards.pdf')
     plt.show()
 
-
 def plot_convergence(agent, file_name):
     losses, vs = agent.losses, agent.vs
     # losses2, vs2 = agent.losses2, agent.vs2
@@ -126,6 +126,10 @@ class MonitoringEnv(gym.Wrapper):
     '''
 
     def __init__(self, env, **kwargs):
+        self.plot_label = False
+        if 'plot_progress' in kwargs:
+            self.plot_label = kwargs.get('plot_progress')
+
         gym.Wrapper.__init__(self, env)
         self.rewards = []
         self.init_rewards = []
@@ -172,7 +176,9 @@ class MonitoringEnv(gym.Wrapper):
         self.rewards[self.current_episode].append(reward)
         if self.current_step >= 200:
             done = True
-            self.plot_results(self.current_episode)
+
+            if self.plot_label:
+                self.plot_results(self.current_episode)
         ob = self.scale_state_env(ob)
         reward = self.scale_rew(reward)
 
@@ -203,8 +209,6 @@ class MonitoringEnv(gym.Wrapper):
                 sum_rews.append(np.sum(rewards[i]))
                 mean_rews.append(np.mean(rewards[i]))
 
-        plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
-
         fig, ax = plt.subplots(1, 1)
         ax.set_title(label=label)
         color = 'blue'
@@ -212,34 +216,59 @@ class MonitoringEnv(gym.Wrapper):
         ax.set_ylabel('cum. reward', color=color)
         ax.set_xlabel('# episode')
         ax.tick_params(axis='y', labelcolor=color)
-        ax1 = plt.twinx(ax)
-        color = 'lime'
-        ax1.plot(mean_rews, c=color)
-        ax1.set_ylabel('mean reward', color=color)  # we already handled the x-label with ax1
-        ax1.tick_params(axis='y', labelcolor=color)
-        # plt.savefig(file_name + '_rewards.pdf')
         plt.show()
 
 
-env = MonitoringEnv(env=PendulumEnv())
[…]

### 2020-11-23 run_paper_tests.py

@@ -1,0 +1,36 @@
+import os
+import shutil
+
+name = 'random_seeds'
+# root_dir = "DOUBLE/"
+root_dir = "Experiments-150-200-50/"
+
+if os.path.exists(root_dir):
+    shutil.rmtree(root_dir)
+
+# Parameters for the test
+# parameter_list = [
+#     'True-not-noisy',
+#     'True-low-noisy',
+#     'True-high-noisy',
+#     'False-not-noisy',
+#     'False-low-noisy',
+#     'False-high-noisy'
+# ]
+parameter_list = [
+    'Clipping',
+    'No-clipping-clipping-no-smoothing',
+    'No-clipping-clipping-smoothing',
+]
+
+for index in range(len(parameter_list)):
+    name = 'test-'
+    name += '_' + str(parameter_list[index])
+    print(name)
+    for seed in [10]:
+        command = "python run_naf2_single.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
+        os.system(command)
+
+command = "python read_paper_tests.py " + root_dir
+
+os.system(command)

### 2020-11-24 main.tex
- paper changers

@@
 
-\title{Model based reinforcement learning at FERMI FEL}
+\title{Model based and model free reinforcement learning at FERMI FEL}
 
@@
 \begin{align}
-\max J (\policy)  =  \mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],
+\max J (\policy)  =  \mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
 \end{align}
@@
 \begin{align}
-     Q_\theta(\bs,\ba) = -\frac{1}{2}(\ba-\mu_\theta(\bs))P_\theta(\ba-\mu_\theta(\bs))^T+V_\theta(s).
+     Q_\theta(\bs,\ba) = -\frac{1}{2}(\ba-\mu_\theta(\bs))P_\theta(\ba-\mu_\theta(\bs))^T+V_\theta(\bs).\label{eq:state-action-value-approxiation}
 \end{align}
 
-One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
+One modification, which is used in these tests is a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
 It is motivated by double Q-learning \cite{NIPS2010_091d584f}.
@@
 To stabilize the network training a small artificial noise is added to the actions in the update.
-This algorithm has an extremely good sample-efficiency for suitable problems as can be found in accelerators and is used as the baseline for the considered control problem.
+This algorithm has an extremely good sample-efficiency for suitable problems as can be found often in accelerators and is used as the baseline for the considered control problem, as it shows very good results.
 %\subsection{New NAF network architecture}
@@
 \subsection{Experiment results from FERMI RL tests}
-Niky and Simon
-
-\section{Discussion and outlook}
-
-\section{Conclusions}
-
-\appendix
-\section{A non-linear standard control problem}
-To provide some transparency of these studies for other labs we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.
-It is a non-linear low dimensional unsolved continuous control problem.
-Unsolved means there is no threshold for the reward to terminate an episode.
-The episode length is set to 200 steps.
-In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly in terms of noise handling.
-This is of importance when dealing with measurements on real systems.
-\subsection{The impact of noise}
-We add artificial Gaussian noise $\gauss(0,0.05)$ on the state with and compare the different \emph{NAF} variants: \emph{smoothing-double}, \emph{smoothing single}, \emph{standart}.
-\section{NAF2 details}\label{appendix:naf2}
-\section{AE training details}
-\section{Theoretical aspects on the AE-DYNA on FERMI FEL}
+As discussed in the previous sections, several tests were performed on the FERMI XFEL.
+The main purpose was to test the newly implemented algorithms on a real system to evaluate their operational feasibility.
+The \emph{NAF2} algorithm, as a representative for highly sample efficient MFRL algorithms, was tested first.
+\subsection{MFRL tests}
+In total four tests were carried out, two using a single network and two using the double network architecture. In both cases \emph{smoothing} was applied, as described in \cref{appendix:naf2}.\\
+\Cref{fig:NAF_training} displays the results, averaged over the two test.
+A training of 100 episodes (shades area) is followed by a verification of 50 steps.
+in the upper figure the number of iterations per episode is plotted including the cumulative number of steps.
+In the verification phase both algorithms show a similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\
+Additionally the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps.
+The blue curves shows the Bellmann error, which is comparable.
+The state-value function, which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network tends to overestimate the value.
+In this case convergence is reached ~1400 steps.\\
+\subsection{MBRL tests}
+The second test campaign was employing the \emph{AE-DYNA} algorithm, as representative for pure a MBRL algorithms.
+Two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA SAC} variant.
+The first uses the trust region policy optimization -\emph{TRPO}- \cite{Schulman2015} to train the controller.
+The TRPO monotonically converges to better policy and this property is exploited in the training.
+The convergence property can be seen in \cref{fig:ME-TRPO_observables}.
+In the upper figure the total reward per batch is shown, as well as the number of data points used in the training of the dynamics model.
+In the lower plot the average cumulative reward as achieved by the TRPO on the individual models in the ensemble on a number of 10 episodes is drawn.
+The shade area shows the corresponding standard deviation to indicate the uncertainty of the dynamics model.
+\\
+Secondly, the \emph{AE-DYNA SAC} was tested, which uses the soft actor critic -SAC- algorithm \cite{Haarnoja2018a}.
+The SAC not only tries to maximise $J$ \cref{eq:cumulative_reward}, but also simultaneously but weights on the maximisation of the entropy in the action space.
+In this way exploration is encouraged to avoid getting stuck in a local optimum.
+In this test the controller is reset each time, when new data for the dynamics model training is acquired.
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_episodes.pdf}
+	\caption{The training on the \emph{NAF2} in the FERMI FEL.}
+	\label{fig:NAF_training}
+\end{figure}
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
+	\caption{The training .}
+	\label{fig:NAF_convergence}
+\end{figure}
 
@@
 	\input{Figures/SL_Alignment_Scheme}
-  \caption{The training .}
-  \label{fig:comparsion_per}
-\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
 	\caption{The training .}
@@
 	\caption{The training .}
-  \label{fig:comparsion_per}
+	\label{fig:ME-TRPO_observables}
 \end{figure}
@@
 	\caption{The training .}
-  \label{fig:comparsion_per}
+	\label{fig:ME-TRPO_verification}
 \end{figure}
+
+\section{Discussion and outlook}
+
+\section{Conclusions}
+
+\appendix
+\section{A non-linear standard control problem}
+To provide some transparency of these studies for other labs we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.
+It is a non-linear low dimensional unsolved continuous control problem.
+Unsolved means there is no threshold for the reward to terminate an episode.
+The episode length is set to 200 steps.
+In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly in terms of noise handling.
+This is of importance when dealing with measurements on real systems.
+
+\subsection{NAF2 details}\label{appendix:naf2}
+%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
+ We compare the different \emph{NAF} variants: \emph{smoothing-double}, \emph{smoothing single}, \emph{standart}.
+The smoothing adds a small clipped noise on the actions to stabilize the network training as:
+ \begin{equation}
+ 	a'(s') = \text{clip}\left(\mu_{\theta_{\text{targ}}}(s') + \text{clip}(\epsilon,-c,c), a_{Low}, a_{High}\right),
+ \end{equation}
+ 	where $\epsilon \sim \mathcal{N}(0, \sigma)$. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
+ The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
+ \begin{align}
+y(r,s',d) = r + \gamma (1 - d) \min_{i=1,2} Q_{\phi_{i, \text{targ}}}(s', a'(s'))
+ \end{align}
+and then both are learned by regressing to this target:
+\begin{align}
+	L(\phi_1, {\mathcal D}) = \mathbb E_{(s,a,r,s',d) \sim {\mathcal D}}{
+		\Bigg( Q_{\phi_1}(s,a) - y(r,s',d) \Bigg)^2
+	}
+\end{align}
+and the policy is obtained via:
+\begin{align}
+\max_{\theta} \underset{s \sim {\mathcal D}}{{\mathrm E}}\left[ Q_{\phi_1}(s, \mu_{\theta}(s)) \right].
+\end{align}
+The results are shown in \cref{fig:comparsion_smoothing}.
+One sees the cumulative reward per episode for a training of a total of 50 episodes.
+The curve labelled \emph{clipping} corresponds to the double network including the smoothing method and shows the best stability during the training yielding quickly a high reward.
+Also the single network shows good and comparable performance, except for the stability.
+The worst performance is achieved without smoothing and a single network, nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
+	\caption{Cumulative reward of different NAF implementations as discussed in the text.}
+	\label{fig:comparsion_smoothing}
+\end{figure}
+\subsection{The impact of noise}
+A test adding artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma)$ with $\sigma=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}.
+There the difference of the three methods becomes even more evident.
 
@@
 	\centering
-  \includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_episodes.pdf}
-  \caption{The training .}
-  \label{fig:comparsion_per}
+	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise}
+	\caption{Cumulative reward of different NAF implementations with artificial noise as discussed in the text.}
+	\label{fig:comparsion_noise}
 \end{figure}
+
+\section{AE training details}
+\section{Theoretical aspects on the AE-DYNA on FERMI FEL}
 \begin{figure}[!h]
 	\centering
-  \includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
+	\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
 	\caption{The training .}
-  \label{fig:comparsion_per}
+	\label{fig:Learning_evolution}
 \end{figure}

### 2020-11-24 Benchmark_NAF.py

@@ -4,7 +4,6 @@ import pickle
 
 main_dir = 'checkpoints'
 
-
 def read_rews(rewards):
     iterations = []
     final_rews = []
@@ -18,7 +17,6 @@ def read_rews(rewards):
             sum_rews.append(np.sum(rewards[i]))
             mean_rews.append(np.mean(rewards[i]))
     return iterations, final_rews, sum_rews, mean_rews
-
 
 def plot_results(rewards_list, file_name=None):
     fig, ax = plt.subplots(1, 1)
@@ -35,7 +33,6 @@ def plot_results(rewards_list, file_name=None):
     ax.set_ylabel('cum. reward')
     ax.set_xlabel('# episode')
     plt.show()
-
 
 def plot_convergence(agent, file_name):
     losses, vs = agent.losses, agent.vs

### 2020-11-24 naf2_new.py

@@ -3,13 +3,15 @@ import pickle
 import shutil
 import matplotlib.pyplot as plt
 import gym
+
 import tensorflow as tf
 from tensorflow import keras
+# Turn off warnings form tensorflow
+tf.get_logger().setLevel('ERROR')
 
 tf.keras.backend.set_floatx('float64')
 import numpy as np
 from tqdm import tqdm
-
 
 class ReplayBuffer:
     """
@@ -63,7 +65,6 @@ class ReplayBuffer:
         for i in range(len(obs1s)):
             self.store(obs1s[i], acts[i], rews[i], obs2s[i], dones[i])
         # print(self.size)
-
 
 def basic_loss_function(y_true, y_pred):
     return tf.math.reduce_mean(y_true - y_pred)

### 2020-11-24 read_paper_tests.py

@@ -26,7 +26,6 @@ def read_experiment(root_dir):
 
     return data_all
 
-
 try:
     root_dir = sys.argv[1]
 except:
@@ -36,19 +35,19 @@ except:
     # double = read_experiment(root_dir=root_dir)
 
     # data_all = pd.concat([single, double], sort=False)
-    root_dir = "Experiments-150-200-50/"
+    root_dir = "Experiments-150-200-50-noisy/"
     data_all = read_experiment(root_dir=root_dir)
 
-figure_name = 'Smoothing_comparison.pdf'
+figure_name = '../Figures/Comparison_noise.pdf'
 current_test_name = ''
 
 fig, axs = plt.subplots(1, 1)
 
 # data_all = data_all.loc[:, :length]
 print(data_all)
-
+sns.set_palette(sns.color_palette("bright", 8))
 sns.lineplot(ax=axs, data=data_all.T)
-plt.ylabel(' cum reward (arb. units.)')
+plt.ylabel('cum. reward (arb. units.)')
 plt.xlabel('no. episode')
 plt.savefig(root_dir + figure_name)
 plt.show()

### 2020-11-24 run_paper_tests.py

@@ -3,10 +3,10 @@ import shutil
 
 name = 'random_seeds'
 # root_dir = "DOUBLE/"
-root_dir = "Experiments-150-200-50/"
+root_dir = "Experiments-150-200-50-noisy/"
 
-if os.path.exists(root_dir):
-    shutil.rmtree(root_dir)
+# if os.path.exists(root_dir):
+#     shutil.rmtree(root_dir)
 
 # Parameters for the test
 # parameter_list = [
@@ -18,16 +18,16 @@ if os.path.exists(root_dir):
 #     'False-high-noisy'
 # ]
 parameter_list = [
-    'Clipping',
-    'No-clipping-clipping-no-smoothing',
-    'No-clipping-clipping-smoothing',
+    'Clipping-noisy',
+    'NocNoS-noisy',
+    'NoCS-noisy',
 ]
 
 for index in range(len(parameter_list)):
-    name = 'test-'
+    name = 'Setting'
     name += '_' + str(parameter_list[index])
     print(name)
-    for seed in [10]:
+    for seed in [10, 20, 30, 40, 50]:
         command = "python run_naf2_single.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
         os.system(command)
 

### 2020-11-24 main.tex
- paper changers

@@
 In terms of sample-complexity the model-based approach is faster, while the final performance of the model free method is superior - the so called asymptotic performance.
-The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency
+The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency.
+The implementation is available at \cite{Hirlaender2020}.
+
 \end{abstract}
@@
 One way to reduce the effect is to take a simple analytical form of the $Q$-function.
-\subsection{NAF}
+\subsection{Normalized advantage function}
 If a specific quadratic form of the $Q$ function is assumed:
@@
 \end{figure}
-\subsection{Critical design decisions}
+\subsection{Critical design decisions}\label{ss:critical_design}
 In the following the most important aspects of a successful application of a Dyna-AE algorithm are listed:
@@
 Two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA SAC} variant.
-The first uses the trust region policy optimization -\emph{TRPO}- \cite{Schulman2015} to train the controller.
+In both cases an ensemble of three network was used and the epistemic and aleatoric uncertainty is measured via the anchor-ensemble technique as discussed in \cref{ss:critical_design}.
+On top early stopping was used.
+\\ The first uses the trust region policy optimization -\emph{TRPO}- \cite{Schulman2015} to train the controller.
 The TRPO monotonically converges to better policy and this property is exploited in the training.
@@
 The shade area shows the corresponding standard deviation to indicate the uncertainty of the dynamics model.
+To measure the convergence of the TRPO the logarithm of the standard deviation of the distribution drawn trajectories $p_\traj$ is also visualized.
+Here the training was stopped after 450 steps acquiring 25 steps each dynamics training.
+In the verification, as can be seen in \cref{fig:ME-TRPO_verification}, all of the 50 episodes where successfully finished after a few steps.
 \\
@@
 In this test the controller is reset each time, when new data for the dynamics model training is acquired.
+Consequently, the performance drops each time, the dynamics model was retrained, as shown in \cref{fig:AE-DYNA_observables}.
+Chances to get trapped in a local optimum are smaller by using this strategy of training.
+In difference to the first MBRL test, the batches of when acquiring new data are 50 with an initial random walk of 100 steps.
+The number of initial steps was chosen carefully large enough, because the convergence is slowed down and there is the risk that the training becomes unfeasibly.
+The training was stopped after the acquisition of 500 data points and a verification was done as in the first test.
+Again the success is 100\%, but the number of iterations per step is smaller.
+It might be a result of the bigger number of data points, but in general, this method showed a better asymptotic performance than the $ME-TRPO$ variant.
 \begin{figure}[!h]
@@
 	\caption{The training .}
-	\label{fig:comparsion_per}
+	\label{fig:AE-DYNA_observables}
 \end{figure}
@@
 	\caption{The training .}
-	\label{fig:comparsion_per}
+	\label{fig:AE-DYNA_verification}
 \end{figure}
@@
 
-\section{AE training details}
 \section{Theoretical aspects on the AE-DYNA on FERMI FEL}
+In this section we want to discuss some theoretical aspects concerning the AE-DYNA approach. A simulation was used to obtain the presented analysis. \\
+One issue of MBRL methods is the asymptotic performance.
+Altough good results were obtained using the AE-DYNA, there is a difference between the MFRL and MBRL.
+We call it the \emph{reality gap}.
+There are ideas to attack the problem e.g.
+by learning a meta-policy \cite{Clavera2018} followed by a fine tuning on the residual physics, which minimizes the \emph{reality gap} ot training on a simulator and closing than with a small amount of training iterations on the real system the gap \cite{Zeng2019}.
 \begin{figure}[!h]

### 2020-11-24 AE_dyna_results.py

@@ -6,21 +6,26 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 label = "ME-TRPO"
-# label = "AE-DYNA"
+label = "AE-DYNA"
+
+label = 'simulation'
+
 if label == "ME-TRPO":
     # ME-TRPO results
     project_directory = 'Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/run2/'
-else:
+elif label=="AE-DYNA":
     # AE-Dyna results
-    project_directory = 'Data_Experiments/2020_11_05_AE_Dyna@FERMI/-nr_steps_25-cr_lr-n_ep_13-m_bs_100-sim_steps_3000-m_iter_35-ensnr_3-init_200/'
+    project_directory = 'Data_Experiments/2020_11_05_AE_Dyna@FERMI/-nr_steps_25-cr_lr-n_ep_13-m_bs_100-sim_steps_3000' \
+                        '-m_iter_35-ensnr_3-init_200/ '
+else:
+    project_directory = 'Data_logging/Simulation/-nr_steps_201-n_ep_12-m_bs_100-sim_steps_2500-m_iter_2-ensnr_3-init_201/'
+
 
 def read_rewards(rewards):
     iterations_all = []
     final_rews_all = []
     mean_rews_all = []
     stds = []
-
-
 
     iterations = []
     final_rews = []
@@ -72,7 +77,7 @@ def plot_results(data, label='Verification', **kwargs):
         ax = axs[0]
         x = range(len(iterations))
         ax.plot(x, iterations)
-        ax.set_ylabel('Iterations (1)')
+        ax.set_ylabel('no. iterations')
         ax.set_title(plot_suffix)
         # fig.suptitle(label, fontsize=12)
         if 'data_number' in kwargs:
@@ -89,7 +94,7 @@ def plot_results(data, label='Verification', **kwargs):
         ax.plot(x, finals, color=color)
 
         ax.set_title('Final reward per episode')  # + plot_suffix)
-        ax.set_xlabel('Episodes (1)')
+        ax.set_xlabel('no. episodes')
 
         ax1 = plt.twinx(ax)
         color = 'lime'
@@ -120,14 +125,14 @@ def plot_observables(data, label='Experiment', **kwargs):
     ax.step(x, batch_rews_all[0])
     ax.fill_between(x, batch_rews_all[0] - batch_rews_all[1], batch_rews_all[0] + batch_rews_all[1],
[…]

### 2020-11-24 NAF_results.py

@@ -93,14 +93,14 @@ def plot_results(rewards, rewards_single, **kwargs):
     color = 'blue'
     ax.plot(iterations, c=color)
     ax.plot(iterations_s, c=color, ls=':')
-    ax.set_ylabel('steps', color=color)
+    ax.set_ylabel('no. iterations', color=color)
     ax.tick_params(axis='y', labelcolor=color)
     ax1 = plt.twinx(ax)
     color = 'k'
     ax1.plot(np.cumsum(iterations), c=color)
     ax1.plot(np.cumsum(iterations_s), c=color, ls=':')
-    ax1.set_ylabel('cumulative steps', color=color)
-    ax.set_title('Iterations' + plot_suffix)
+    ax1.set_ylabel('no. cumulative steps', color=color)
+    ax.set_title('no. iterations' + plot_suffix)
     # fig.suptitle(label, fontsize=12)
 
     ax = axs[1]
@@ -109,18 +109,18 @@ def plot_results(rewards, rewards_single, **kwargs):
     # ax.plot(starts, c=color)
     ax.plot(mean_rews, c=color)
     ax.plot(mean_rews_s, c=color, ls=':')
-    ax.set_ylabel('cum. return', color=color)
+    ax.set_ylabel('cum. return (arb. units)', color=color)
     # ax.axhline(-0.05, ls=':', color='r')
     ax.tick_params(axis='y', labelcolor=color)
-    ax.set_title('Reward per episode')  # + plot_suffix)
-    ax.set_xlabel('episodes')
+    ax.set_title('reward per episode (arb. units)')  # + plot_suffix)
+    ax.set_xlabel('no. episodes')
 
     ax1 = plt.twinx(ax)
     color = 'lime'
     ax1.plot(final_rews[:-1], color=color)
     ax1.plot(final_rews_s[:-1], color=color, ls=':')
 
-    ax1.set_ylabel('final return', color=color)
+    ax1.set_ylabel('final return (arb. units)', color=color)
     ax1.axhline(-0.05, ls=':', color=color)
     ax1.tick_params(axis='y', labelcolor=color)
 
@@ -175,20 +175,20 @@ rewards_s = [rews0, rews1]
 def plot_convergence(losses, v_s, losses_s, v_s_s, label, **kwargs):
     fig, ax = plt.subplots()
     ax.set_title(label)
-    ax.set_xlabel('steps')
+    ax.set_xlabel('no. steps')
 
     color = 'tab:blue'
[…]

### 2020-11-24 AE_dyna_results.py
- paper changers

@@ -5,20 +5,18 @@ import pickle
 import matplotlib.pyplot as plt
 import numpy as np
 
-label = "ME-TRPO"
-label = "AE-DYNA"
-
-label = 'simulation'
+label = ["ME-TRPO", "AE-DYNA",'simulation'][1]
 
 if label == "ME-TRPO":
     # ME-TRPO results
     project_directory = 'Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/run2/'
 elif label=="AE-DYNA":
     # AE-Dyna results
-    project_directory = 'Data_Experiments/2020_11_05_AE_Dyna@FERMI/-nr_steps_25-cr_lr-n_ep_13-m_bs_100-sim_steps_3000' \
-                        '-m_iter_35-ensnr_3-init_200/ '
+    project_directory = 'Data_Experiments/2020_11_05_AE_Dyna@FERMI/-nr_steps_25-cr_lr-n_ep_13-m_bs_100-sim_steps_3000-m_iter_35-ensnr_3-init_200/'
 else:
-    project_directory = 'Data_logging/Simulation/-nr_steps_201-n_ep_12-m_bs_100-sim_steps_2500-m_iter_2-ensnr_3-init_201/'
+    project_directory = 'Data_logging/Simulation/'+\
+                        '-nr_steps_25-n_ep_97-m_bs_5-sim_steps_3000-m_iter_30-ensnr_5-init_100'+\
+                        '/'
 
 
 def read_rewards(rewards):
@@ -93,12 +91,12 @@ def plot_results(data, label='Verification', **kwargs):
         ax.tick_params(axis='y', labelcolor=color)
         ax.plot(x, finals, color=color)
 
-        ax.set_title('Final reward per episode')  # + plot_suffix)
+        # ax.set_title('final reward per episode (arb. units)')  # + plot_suffix)
         ax.set_xlabel('no. episodes')
 
         ax1 = plt.twinx(ax)
         color = 'lime'
-        ax1.set_ylabel('Mean reward', color=color)  # we already handled the x-label with ax1
+        ax1.set_ylabel('cum. reward (arb. units)', color=color)  # we already handled the x-label with ax1
         ax1.tick_params(axis='y', labelcolor=color)
         ax1.fill_between(x, means - stds, means + stds,
                          alpha=0.5, edgecolor=color, facecolor='#FF9848')

### 2020-11-25 main.tex
- paper changers

@@
 The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency.
-The implementation is available at \cite{Hirlaender2020}.
+%The implementation is available at \cite{Hirlaender2020}.
 
@@
 \begin{itemize}
-	\item Noise benchmarks with pendulum
-	\item Benchmark NAF2 with pendulum
+	\item Noise benchmarks with pendulum - done
+	\item Benchmark NAF2 with pendulum - done
+	\item Compare NAF2 and AE-DYNA
 	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
 	\item Tailored to accelerators - short horizons are standard?
+	\item Long waiting time in MBRL
+	\item Noise sensivity as mentioned in Benchmark paper
 	\item Improvements of the algorithms:
@@
 		\item discuss different approaches to e.g. dynamic waiting time
+		\item usually prior not discussed in papers
 	\end{itemize}
@@
 \end{itemize}
-On critical aspect using RL is the number of iteration to train a controller.
+One critical aspect using RL is the number of iterations needed to train a controller and the a second critical aspect is the robustness of the training itself.\\
 In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.
-Three algorithms were applied successfully showing different advantages, which will e discussed in the following.
-
-% The methods proposed are the iterative Linear Quadratic Regulator (iLQR) - an extension to non-linear systems of the fundamental control theory Linear Quadratic Regulator (LQR) problem -  and the Normalized Advantage Function (NAF) - a continuous variant of the Q-learning algorithm from reinforcement learning field.
-An exhaustive presentation of both techniques is provided in the following sections.
-
-In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator. 
-\section{The physical set-up}
-At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation.
-Nevertheless, in the last years various approaches have been studied to investigate their applicability in tuning operations.
-%%% Last introduction paragraph
-The paper is organised as follows. 
+Three different algorithms were applied successfully to the FERMI FEL problem and reveal different advantages.
+The critical aspects are addressed and tailored solutions, generally applicable to accelerator control problems are presented.
+The paper is organised as follows:
 \begin{itemize}
@@
 \end{itemize}
+% The methods proposed are the iterative Linear Quadratic Regulator (iLQR) - an extension to non-linear systems of the fundamental control theory Linear Quadratic Regulator (LQR) problem -  and the Normalized Advantage Function (NAF) - a continuous variant of the Q-learning algorithm from reinforcement learning field.
+An exhaustive presentation of both techniques is provided in the following sections.
+
+\section{The physical set-up of the studied problem}
+\begin{figure}[!h]
+	\centering
+	\input{Figures/SL_Alignment_Scheme}
+	\caption{A schematic view on the set-up of the FERMI FEL.}
+	\label{fig:schematic_FEL}
+\end{figure}
+In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator.\\
+At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation.
+Nevertheless, in the last years various approaches have been studied to investigate their applicability in tuning operations.
+\\
+A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure.
+The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.\\
+The external optical laser signal provide is contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy.
+The modulation in energy is the converted into a charge modulation in the dispersive section, and finally the density modulated beams radiation is amplified in the radiators section.
+The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.\\
+For this reason the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
+\subsection{\NB{our system: the modulator and the seed laser}}
+
+The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
+The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
+A schematic overview of the set-up is provided in \cref{fig:schematic_FEL}.
+Two mirrors, TT1 and TT2, are used to control the trajectory of the laser by tilting and  inclining, which gives a total of four degrees of freedom (DOF).
+In turn the laser overlaps with the electron beam between the two screens, CCD1 and CCD2.
+Lastly the monitor measures the intensity $I_0$.
+The final problem faced consists in optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.
+
+%\section{Proof-of-principle application of RL Agent at FERMI trajectory correction}
+%\NB{Niky}
 
@@
 Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
-Goal of reinforcement learning is to find a $\policy: \bs\mapsto \ba$, which is the solution of:
+Goal of reinforcement learning is to find a $\policy^*: \bs\mapsto \ba$, which is the solution of:
 \begin{align}
-\max J (\policy)  =  \mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
+\policy^*\arg\!\max J (\policy)  = 
+\mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
 \end{align}
 where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
+$\transitions(\bs_{t+1}|\bs_t,\ba_t)$ characterises the probability to end in a state $\bs_{t+1}$ if an action $\ba_t$ is taken at state $\bs_t$.
 In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
 a deep neural network.
-In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
+In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{RichardS.UniversityofAlbertaSutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}.
+In the latter about approximate dynamic programming, which we now discuss.
+\section{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
+We do not provide an exhaustive treatment of state of the art MFRL algorithms.
+Only the main principles are covered for the later statements and an overview is given e.g.
+in \cite{RichardS.UniversityofAlbertaSutton2018,Levine2020}.
+We discuss in detail the normalised advantage function algorithm, which has good characteristics for accelerator control problems as shown in \cref{KAIN}.
 \subsection{Approximate dynamic programming}
-The state-value function $Q(\bs,\ba)$ is expressed as $Q_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
-By satisfying the Bellmann-optimality equation $Q_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$:
+The state-value function $Q(\bs,\ba)$:
+\begin{align}
+	Q^\policy(\bs,\ba) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
+\end{align}
+ is expressed as $Q^\policy_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
+By satisfying the Bellmann-optimality equation $Q^\policy_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$:
 \begin{align}
@@
 \end{align}
-The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
+The form of this equation can cause overestimation and other complications, when using a function approximator.
+Several methods exist which try to mitigate the problems \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
 One way to reduce the effect is to take a simple analytical form of the $Q$-function.
 \subsection{Normalized advantage function}
-If a specific quadratic form of the $Q$ function is assumed:
+If a specific quadratic form of the $Q$ function is assumed \cite{Gu2016}:
 \begin{align}
@@
 One modification, which is used in these tests is a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
-It is motivated by double Q-learning \cite{NIPS2010_091d584f}.
+It is motivated by double Q-learning \cite{NIPS2010_091d584f,Hasselt2015}.
 The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality}:
@@
 This algorithm has an extremely good sample-efficiency for suitable problems as can be found often in accelerators and is used as the baseline for the considered control problem, as it shows very good results.
+An illustration and additional changes to the original proposal \cite{Gu2016} and a previous implementation used for experiments using a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
 %\subsection{New NAF network architecture}
 
-\begin{algorithm}[ht]
-\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
-\begin{algorithmic}[1]
-\State initialise $\theta_0$
-\For{iteration $k \in [0, \dots, K]$}
-\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
-%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
-%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
-%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
-%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
-%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
-%\EndFor
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}[ht]
+%\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
+%\begin{algorithmic}[1]
+%\State initialise $\theta_0$
+%\For{iteration $k \in [0, \dots, K]$}
+%\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
+%%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
+%%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
+%%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
+%%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
+%%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
+%%\EndFor
+%\end{algorithmic}
+%\end{algorithm}
 
@@
 
-\subsection{Operational deployment}
-\NB{Niky} % come abbiamo interagito con FERMI
-\subsection{\NB{short FERMI introduction}}
-The Free Electron laser Radiation for Multidisciplinary Investigation  better known as the FERMI, is the seeded free-electron laser facility next to the third-generation synchrotron facility ELETTRA at the Elettra Sinchrotrone Trieste laboratories.
-A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure.
-The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.
-
-The external optical laser signal provide is contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy.
-The modulation in energy is the converted into a charge modulation in the dispersive section, and finally the density modulated beams radiation is amplified in the radiators section.
-The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.
-
-For this reason the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
-
-\subsection{\NB{our system: the modulator and the seed laser}}
-A more detailed description of the alignment process in the modulator is here provided.
-The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
-The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
-The final problem faced consists in optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.
-
-\section{Proof-of-principle application of RL Agent at FERMI trajectory correction}
-\NB{Niky}
-
 \subsection{Experiment results from FERMI RL tests}
@@
 \Cref{fig:NAF_training} displays the results, averaged over the two test.
-A training of 100 episodes (shades area) is followed by a verification of 50 steps.
-in the upper figure the number of iterations per episode is plotted including the cumulative number of steps.
-In the verification phase both algorithms show a similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\
+A training of 100 episodes was done.
+In the upper figure the number of iterations per episode is plotted including the cumulative number of steps.
+In the verification \cref{fig:NAF_verification} both algorithms show a similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\
 Additionally the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps.
@@
 	\centering
-	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_episodes.pdf}
-	\caption{The training on the \emph{NAF2} in the FERMI FEL.}
+	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
+	\caption{The training of the \emph{NAF2} on the FERMI FEL.}
 	\label{fig:NAF_training}
 \end{figure}
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
+	\caption{The verification of the \emph{NAF2} on the FERMI FEL.}
+	\label{fig:NAF_verification}
+\end{figure}
+
 \begin{figure}[!h]
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
-	\caption{The training .}
+	\caption{The verification of the AE-DYNA-SAC on the FERMI FEL.}
 	\label{fig:NAF_convergence}
@@
 	\centering
-	\input{Figures/SL_Alignment_Scheme}
-	\caption{The training .}
-	\label{fig:comparsion_per}
-\end{figure}
-
-\begin{figure}[!h]
-	\centering
 	\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
-	\caption{The training .}
+	\caption{The training of the AE-DYNA-SAC on the FERMI FEL.}
 	\label{fig:AE-DYNA_observables}
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_verification.pdf}
-	\caption{The training .}
+	\caption{The verification of the AE-DYNA-SAC on the FERMI FEL.}
 	\label{fig:AE-DYNA_verification}
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
-	\caption{The training .}
+	\caption{The training of the ME-TRPO on the FERMI FEL.}
 	\label{fig:ME-TRPO_observables}
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
-	\caption{The training .}
+	\caption{The verification of the ME-TRPO on the FERMI FEL.}
 	\label{fig:ME-TRPO_verification}
@@
 A test adding artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma)$ with $\sigma=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}.
-There the difference of the three methods becomes even more evident.
-
+There the difference of the three methods becomes even more evident.The results are shown in \cref{fig:comparsion_noise}.
+After around 65 episodes the single network without smoothing decreases before reaching the final performance at around 95 episodes, while the smoothing prevents this performance drop.
 \begin{figure}[!h]
@@
 	\label{fig:comparsion_noise}
+\end{figure}
+Using the anchors in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}.
+It the upper plot the mean cumulative reward on 10 test episodes on the real environment is shown during the training.
+It should indicate the result if the training is stopped at this training iteration.
+One cannot observe this during a real training, unless one does costly performance measurements during the training.
+The blue curve reaches the target much quicker exhibiting less variation.\\
+The lower plot of \cref{fig:comparsion_noise_ae_dyna} shows the batch rewards a measured during the data collection, which is observable during the training.
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise_ae_dyna}
+	\caption{Cumulative reward of AE-DYNA with artificial noise using the anchor.}
+	\label{fig:comparsion_noise_ae_dyna}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_ae_dyna}
+	\caption{The comparison of the NAF2 and the AE-DYNA on the noisy pendulum.}
+	\label{fig:comparsion_noise_ae_dyna}
 \end{figure}

### 2020-11-25 AEDYNA.py

@@ -1,5 +1,6 @@
 import os
 import pickle
+import sys
 from datetime import datetime
 
 import gym
@@ -21,19 +22,13 @@ config = tf.ConfigProto(
 config = None
 
 ############################################################
-# Loading the environment
-############################################################
-
-real_env = env = PendulumEnv()
-
-############################################################
 # Hyperparameters
 ############################################################
 
 
 steps_per_env = 201
 init_random_steps = 201
-total_steps = 1004
+total_steps = 2009
 num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
 
 print('Number of epochs: ', num_epochs)
@@ -53,7 +48,7 @@ else:
     from stable_baselines.common.policies import MlpPolicy
     from stable_baselines import PPO2 as Agent
 
-simulated_steps = 25000
+simulated_steps = 2500
 
 model_batch_size = 100
 num_ensemble_models = 5
@@ -62,17 +57,6 @@ early_stopping = True
 model_iter = 10
 
 # model_training_iterations = 10
-network_size = 25
-# Set the priors for the anchor method:
-#  TODO: How to set these correctly?
-init_params = dict(init_stddev_1_w=np.sqrt(1),
-                   init_stddev_1_b=np.sqrt(1),
-                   init_stddev_2_w=1 / np.sqrt(network_size))
-
-data_noise = 0.00000  # estimated noise variance
-lambda_anchor = data_noise / (np.array([init_params['init_stddev_1_w'],
-                                        init_params['init_stddev_1_b'],
-                                        init_params['init_stddev_2_w']]) ** 2)
 
 # How often to check the progress of the network training
 # e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
@@ -83,29 +67,8 @@ lr_start = 1e-3
 lr_end = 1e-3
 lr = lambda ep: max(lr_start + ep / 30 * (lr_end - lr_start), lr_end)
 
-# Create the logging directory:
-project_directory = 'Data_logging/Simulation/'
-
-hyp_str_all = '-nr_steps_' + str(steps_per_env) + '-n_ep_' + str(num_epochs) + \
-              '-m_bs_' + str(model_batch_size) + \
-              '-sim_steps_' + str(simulated_steps) + \
-              '-m_iter_' + str(model_iter) + '-ensnr_' + str(num_ensemble_models) + '-init_' + str(
[…]

### 2020-11-25 AE_dyna_results.py

@@ -5,18 +5,16 @@ import pickle
 import matplotlib.pyplot as plt
 import numpy as np
 
-label = ["ME-TRPO", "AE-DYNA",'simulation'][2]
+label = ["ME-TRPO", "AE-DYNA", 'Simulation'][0]
 
 if label == "ME-TRPO":
     # ME-TRPO results
     project_directory = 'Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/run2/'
-elif label=="AE-DYNA":
+elif label == "AE-DYNA":
     # AE-Dyna results
     project_directory = 'Data_Experiments/2020_11_05_AE_Dyna@FERMI/-nr_steps_25-cr_lr-n_ep_13-m_bs_100-sim_steps_3000-m_iter_35-ensnr_3-init_200/'
 else:
-    project_directory = 'Data_logging/Simulation/'+\
-                        '-nr_steps_201-n_ep_49-m_bs_100-sim_steps_20000-m_iter_10-ensnr_3-init_201'+\
-                        '/'
+    project_directory = 'Data_logging/temp/-nr_steps_25-cr_lr-n_ep_7-m_bs_100-sim_steps_2500-m_iter_30-ensnr_5-init_100/'
 
 
 def read_rewards(rewards):
@@ -45,68 +43,120 @@ def read_rewards(rewards):
     return np.array(iterations), np.array(final_rews), np.array(mean_rews), np.array(stds)
 
 
+# def plot_results(data, label='Verification', **kwargs):
+#         '''plotting'''
+#         rewards = data['rews']
+#         # iterations = []
+#         # finals = []
+#         # means = []
+#         # stds = []
+#         #
+#         # for i in range(len(rewards)):
+#         #     if (len(rewards[i]) > 1):
+#         #         finals.append(rewards[i][-1])
+#         #         means.append(np.mean(rewards[i][1:]))
+#         #         stds.append(np.std(rewards[i][1:]))
+#         #         iterations.append(len(rewards[i]))
+#         #
+#         # x = range(len(iterations))
+#         # iterations = np.array(iterations)
+#         # finals = np.array(finals)
+#         # means = np.array(means)
+#         # stds = np.array(stds)
+#
+#         iterations, finals, means, stds = read_rewards(rewards)
+#         plot_suffix = label  # , Fermi time: {env.TOTAL_COUNTER / 600:.1f} h'
+#
+#         fig, axs = plt.subplots(2, 1, sharex=True)
[…]

### 2020-11-25 AEdyna_release.py

@@ -1107,7 +1107,7 @@ def aedyna(env_name, cr_lr=5e-3, num_epochs=50,
         mb_lr = lr(ep)
         print('mb_lr: ', mb_lr)
         if early_stopping:
-            model_buffer.generate_random_dataset(ratio=0.1)
+            model_buffer.generate_random_dataset(ratio=0.2)
         else:
             model_buffer.generate_random_dataset()
 

### 2020-11-25 Data_logging/temp/AEdyna_clean_tests.py

@@ -24,11 +24,12 @@ from simulated_tango import SimTangoConnection
 
 # set random seed
 random_seed = 111
+tf.set_random_seed(random_seed)
 np.random.seed(random_seed)
 config = tf.ConfigProto(
     device_count={'GPU': 0}
 )
-# config = None
+config = None
 
 tango = SimTangoConnection()
 real_env = FelLocalEnv(tango=tango)
@@ -52,7 +53,7 @@ algorithm = 'SAC'
 simulated_steps = 2500
 
 model_batch_size = 100
-num_ensemble_models = 5
+num_ensemble_models = 3
 
 early_stopping = True
 model_iter = 30

### 2020-11-25 NAF_results.py

@@ -1,6 +1,5 @@
 '''This script generates the plots for the NAF2 tests at FERMI FEL'''
 
-
 import os
 import pickle
 import numpy as np
@@ -9,7 +8,7 @@ import matplotlib.pyplot as plt
 
 def load_pickle_logging(file_name):
     # directory = 'checkpoints/' + file_name + '/'
-    directory =  file_name + '/'
+    directory = file_name + '/'
     files = []
     directory = directory + 'data/'
     for f in os.listdir(directory):
@@ -28,7 +27,7 @@ def load_pickle_logging(file_name):
 
 def load_pickle_final(file_name):
     # directory = 'checkpoints/' + file_name + '/'
-    directory =  file_name + '/'
+    directory = file_name + '/'
     file = 'plot_data_0.pkl'
 
     with open(directory + file, 'rb') as f:
@@ -52,7 +51,7 @@ states_s, actions_s, rewards_s, dones_s = load_pickle_logging(file_name_s)
 rewards_s = [rewards_s, rewards_s0]
 
 
-def read_rewards(rewards_in):
+def read_rewards(rewards_in, data_range_in=None):
     iterations_all = []
     final_rews_all = []
     mean_rews_all = []
@@ -62,7 +61,12 @@ def read_rewards(rewards_in):
         iterations = []
         final_rews = []
         mean_rews = []
-        for i in range(len(rewards)):
+        if data_range_in is None:
+            data_range = range(len(rewards))
+        else:
+            data_range = range(data_range_in[0], min(len(rewards), data_range_in[1]))
+        for i in data_range:
+            print(i)
             if len(rewards[i]) > 0:
                 final_rews.append(rewards[i][len(rewards[i]) - 1])
                 iterations.append(len(rewards[i]))
@@ -80,16 +84,20 @@ def read_rewards(rewards_in):
     return iterations, final_rews, mean_rews
 
 
-def plot_results(rewards, rewards_single, **kwargs):
+def plot_results(rewards, rewards_single,label=None, **kwargs):
+    if 'data_range_in' in kwargs:
+        data_range_in = kwargs.get('data_range_in')
+    else:
+        data_range_in = None
 
-    iterations, final_rews, mean_rews = read_rewards(rewards)
[…]

### 2020-11-25 read_naf_tests.py

@@ -1,0 +1,54 @@
+import os
+import pickle
+import sys
+
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+
+def read_experiment(root_dir):
+    data_all = pd.DataFrame()
+    n_average = 15
+    for root, dirs, files in os.walk(root_dir):
+        for file in files:
+            if file == 'plot_data_0.pkl':
+                file_name = root + os.sep + file
+                print(file_name)
+                openfile = open(file_name, 'rb')
+                data = np.sum(pickle.load(openfile), axis=1)
+                # data_frame = pd.DataFrame(data, columns=[root.split('/')[0]])
+                data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
+                data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
+                data_all = pd.concat([data_all, data_frame.T], sort=False)
+                openfile.close()
+
+    return data_all
+
+try:
+    root_dir = sys.argv[1]
+except:
+    # root_dir = "SINGLE/"
+    # single = read_experiment(root_dir=root_dir)
+    # root_dir = "DOUBLE/"
+    # double = read_experiment(root_dir=root_dir)
+
+    # data_all = pd.concat([single, double], sort=False)
+    root_dir = "Experiments-150-200-50-noisy/"
+    data_all = read_experiment(root_dir=root_dir)
+
+
+figure_name = '../Figures/Comparison_noise.pdf'
+current_test_name = ''
+
+fig, axs = plt.subplots(1, 1)
+
+# data_all = data_all.loc[:, :length]
+print(data_all)
+sns.set_palette(sns.color_palette("bright", 8))
+sns.lineplot(ax=axs, data=data_all.T)
+plt.ylabel('cum. reward (arb. units.)')
+plt.xlabel('no. episode')
+plt.savefig(root_dir + figure_name)
+plt.show()

### 2020-11-25 read_paper_tests.py

@@ -8,46 +8,117 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 
 
-def read_experiment(root_dir):
+def read_experiment_observables(root_dir, key = 'tests_all'):
     data_all = pd.DataFrame()
-    n_average = 15
+    n_average = 5
     for root, dirs, files in os.walk(root_dir):
+        file_names = []
         for file in files:
-            if file == 'plot_data_0.pkl':
-                file_name = root + os.sep + file
-                print(file_name)
-                openfile = open(file_name, 'rb')
-                data = np.sum(pickle.load(openfile), axis=1)
-                # data_frame = pd.DataFrame(data, columns=[root.split('/')[0]])
-                data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
-                data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
-                data_all = pd.concat([data_all, data_frame.T], sort=False)
-                openfile.close()
+            if 'observables' in file and not 'Noise-off' in root:
+                file_names.append(file)
+        if file_names:
+            file_names.sort()
+            file_name = file_names[-1]
+            file_name = root + os.sep + file_name
+            openfile = open(file_name, 'rb')
+            if not key in ['step_counts_all']:
+                data = pd.DataFrame(pickle.load(openfile)[key]).sum()
+                # print(data)
+                data_frame = pd.DataFrame(data, columns=[root.split('_')[4]])
+                # print(data_frame)
+            else:
+                data = pickle.load(openfile)[key]
+                data_frame = pd.DataFrame(data, columns=[root.split('_')[4]])
+
+            data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
+            data_all = pd.concat([data_all, data_frame.T], sort=False)
+            openfile.close()
 
     return data_all
 
-try:
-    root_dir = sys.argv[1]
-except:
-    # root_dir = "SINGLE/"
-    # single = read_experiment(root_dir=root_dir)
-    # root_dir = "DOUBLE/"
-    # double = read_experiment(root_dir=root_dir)
+def read_rewards(rewards):
+    stds = []
 
-    # data_all = pd.concat([single, double], sort=False)
[…]

### 2020-11-25 run_aedyna_noise_test_pendulum.py

@@ -1,0 +1,27 @@
+import os
+import shutil
+
+name = 'random_seeds'
+# root_dir = "DOUBLE/"
+root_dir = 'Data/Simulation/noisy/'
+
+# if os.path.exists(root_dir):
+#     shutil.rmtree(root_dir)
+
+parameter_list = [
+    'Noise-off-non-aleatoric',
+    'Noise-on-non-aleatoric',
+    'Noise-on-aleatoric',
+]
+
+for index in range(len(parameter_list)):
+    name = 'Setting'
+    name += '_' + str(parameter_list[index])
+    print(name)
+    for seed in [10, 20, 30, 40, 50]:
+        command = "python AEDYNA.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
+        os.system(command)
+
+# command = "python read_paper_tests.py " + root_dir
+
+os.system(command)

### 2020-11-25 run_paper_naf_tests.py

@@ -28,7 +28,7 @@ for index in range(len(parameter_list)):
     name += '_' + str(parameter_list[index])
     print(name)
     for seed in [10, 20, 30, 40, 50]:
-        command = "python run_naf2_single.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
+        command = "python run_naf2_for_tests.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
         os.system(command)
 
 command = "python read_paper_tests.py " + root_dir

### 2020-11-30 main.tex
- paper changers

@@
 
-\title{Model based and model free reinforcement learning at FERMI FEL}
+	\title{Model based and model-free reinforcement learning at FERMI FEL}
 
@@
 	\begin{abstract}
-In this paper we discuss a model based reinforcement learning approach in comparison to a model free reinforcement learning approach applied at the FERMI FEL system.
-Both algorithms and approaches are new in this context and the main purpose of this paper is to show how reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems.
-In terms of sample-complexity the model-based approach is faster, while the final performance of the model free method is superior - the so called asymptotic performance.
-The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency.
+		In this paper, we compare a model-based reinforcement learning approach in comparison to a model-free reinforcement learning approach applied at the FERMI FEL system.
+Both approaches hold tremendous promise and the main purpose of this paper is to show how reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems.
+In terms of sample-complexity, the model-based approach is faster, while asymptotic performance of the model-free method is sightly superior.
+The model-based algorithm is done in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
+In both cases existing methods were adapted to exhibit increased noise resistivity as incumbent in accelerator control problems.
+Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.
 		%The implementation is available at \cite{Hirlaender2020}.
@@
 	\maketitle
-\begin{itemize}
-	\item Noise benchmarks with pendulum - done
-	\item Benchmark NAF2 with pendulum - done
-	\item Compare NAF2 and AE-DYNA
-	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
-	\item Tailored to accelerators - short horizons are standard?
-	\item Long waiting time in MBRL
-	\item Noise sensivity as mentioned in Benchmark paper
-	\item Improvements of the algorithms:
-	\begin{itemize}
-		\item Adding the noise feature
-		\item Jumping feature using SAC
-		\item Noise in the policy
-		\item discuss different approaches to e.g. dynamic waiting time
-		\item usually prior not discussed in papers
-	\end{itemize}
-\end{itemize}
+
 	%\tableofcontents
 	\section{Introduction and Motivation}
-In particle accelerators one main goal is to provide stable and reproducible performance.
-In order to achieve this, a number of control problems have to be considered simultaneously.
-Especially if there is no way to model the physics, one might use optimisation techniques as e.g.
-derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes to restore or maintain the performance.
-Another promising way is to apply reinforcement learning which shows several advantages over optimisation methods:
+	In particle accelerator operation, one main goal is to provide stable and reproducible performance.
+In order to achieve this, several control problems have to be considered simultaneously \cite{Kain2020}.
+Especially if there is no way to model the physics a priori, one might use optimisation techniques as, e.g.
+derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015} or model-based optimisations as Gaussian processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.
+Another promising way is to apply reinforcement learning \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020} which unveils several advantages over optimisation methods:
 	\begin{itemize}
-    \item It covers a larger class of problems, RL optimises a sequence of decisions.
-    \item It memorizes the problem and does not start always from scratch as an DFO.
-    \item Existing data can be used.
+		\item It covers a larger class of problems, as RL optimises a sequence of decisions.
+		\item It memorises the problem and does not always begin from zero as a DFO.
+		\item Already existing data can be used.
 		\item The underlying structure of the problem might be deduced.
 	\end{itemize}
-One critical aspect using RL is the number of iterations needed to train a controller and the a second critical aspect is the robustness of the training itself.\\
-In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.
-Three different algorithms were applied successfully to the FERMI FEL problem and reveal different advantages.
-The critical aspects are addressed and tailored solutions, generally applicable to accelerator control problems are presented.
+	One demanding aspect using RL is the number of iterations needed to train a controller, since RL methods are known to be data-hungry \cite{Sutton2018}, and a second critical aspect is the robustness of the training in real world problems itself.\\
+	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.\\
+	 Two different algorithm classes were applied successfully to the FERMI FEL problem and reveal different advantages.
+The critical aspects are addressed, and tailored solutions, generally applicable to problems showing a sample complexity as in accelerator control problems are presented.
 	The paper is organised as follows:
@@
 		\item Overview of RL
-	\item Details of the implementations used in these studies and theoretical concerns
+		\item Design decisions of the implementations used in these studies and theoretical concerns
 		\item Results 
@@
 	\end{itemize}
-% The methods proposed are the iterative Linear Quadratic Regulator (iLQR) - an extension to non-linear systems of the fundamental control theory Linear Quadratic Regulator (LQR) problem -  and the Normalized Advantage Function (NAF) - a continuous variant of the Q-learning algorithm from reinforcement learning field.
+	% The methods proposed are the iterative Linear Quadratic Regulator (iLQR) - an extension to non-linear systems of the fundamental control theory Linear Quadratic Regulator (LQR) problem -  and the Normalised Advantage Function (NAF) - a continuous variant of the Q-learning algorithm from reinforcement learning field.
 An exhaustive presentation of both techniques is provided in the following sections.
@@
 	At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation.
-Nevertheless, in the last years various approaches have been studied to investigate their applicability in tuning operations.
+Nevertheless, in the last years, various approaches have been studied to investigate their applicability in tuning operations.
 \\
-A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure.
-The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.\\
-The external optical laser signal provide is contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy.
-The modulation in energy is the converted into a charge modulation in the dispersive section, and finally the density modulated beams radiation is amplified in the radiators section.
+	A free-electron laser is a fourth-generation light source where the lasing medium consists of a very-high-speed electron moving freely through a magnetic structure.
+The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronisation with external timing systems.\\
+	The external optical laser signal provides is a contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy.
+The modulation in energy is converted into a charge modulation in the dispersive section, and finally, the density modulated beams radiation is amplified in the section of the radiator.
 The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.\\
-For this reason the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
-\subsection{\NB{our system: the modulator and the seed laser}}
+	For this reason, the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
 
-The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
-The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
+	\subsection{The training environment}
+
+%	The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
+%	The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
 	A schematic overview of the set-up is provided in \cref{fig:schematic_FEL}.
 	Two mirrors, TT1 and TT2, are used to control the trajectory of the laser by tilting and inclining, which gives a total of four degrees of freedom (DOF).
-In turn the laser overlaps with the electron beam between the two screens, CCD1 and CCD2.
-Lastly the monitor measures the intensity $I_0$.
-The final problem faced consists in optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.
+In turn, the laser overlaps with the electron beam between the two screens, CCD1 and CCD2.
+Lastly, the monitor measures the intensity, $I_0$.\\
+	The final problem faced consists of optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.
+	The RL algorithm accesses the problem via an \emph{openai gym} environment \cite{Brockman2016}, which interfaces the RL algorithm with the hardware control system.\\
+	The state $\bs$ is a four dimensional vector holding the current voltage values applied to the piezo motors, where each component lies the normalzed interval [-1, 1].
+Two values correspond to the two DOF of each mirror.
+The action $\ba$ is four dimensional, is a delta step on the voltages at step $t$ $\bs_t$:
+	\begin{align}
+		\bs_{t+1} = \bs_t+\ba_t.
+	\end{align} 
+	The training is done in episodes with a uniformly randomize initial state and a specific maximal length, called the horizon.
+For reason as discussed later, the horizon varies in the applications between either 10 or 50 steps.
+An early termination, when a specific threshold is obtained, lies in all cases at around 95\% of the maximal achievable intensity of the system.
+During the training phase exploration and the training of the function approximator takes place.\\
+	After a defined number of training episodes, a verification of 50 episodes is performed. In the second phase, the RL only applies the learned policy.
 
@@
 	\section{Deep Reinforcement learning}
-Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
+	Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and a reward function $\reward : \states \times \actions \rightarrow \real$, a scalar $\discount \in (0,1]$ called discount factor and an initial state distribution $\initstate$.
+$\transitions(\bs_{t+1}|\bs_t,\ba_t)$ characterises the probability to end in a state $\bs_{t+1}$ if an action $\ba_t$ is taken at state $\bs_t$.
+The tuple \mbox{$\mdp = (\states,\actions,\transitions,\initstate,\reward,\discount)$} defines a Markov decision process.\\
 	Goal of reinforcement learning is to find a $\policy^*: \bs\mapsto \ba$, which is the solution of:
 	\begin{align}
-\policy^*\arg\!\max J (\policy)  = 
+		\policy(\bs)^*=\arg\!\max J (\policy)  = 
 		\mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
@@
 	where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
-$\transitions(\bs_{t+1}|\bs_t,\ba_t)$ characterises the probability to end in a state $\bs_{t+1}$ if an action $\ba_t$ is taken at state $\bs_t$.
 	In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
-a deep neural network.
-In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{RichardS.UniversityofAlbertaSutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}.
-In the latter about approximate dynamic programming, which we now discuss.
-\section{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
-We do not provide an exhaustive treatment of state of the art MFRL algorithms.
-Only the main principles are covered for the later statements and an overview is given e.g.
-in \cite{RichardS.UniversityofAlbertaSutton2018,Levine2020}.
-We discuss in detail the normalised advantage function algorithm, which has good characteristics for accelerator control problems as shown in \cref{KAIN}.
-\subsection{Approximate dynamic programming}
-The state-value function $Q(\bs,\ba)$:
+a deep neural network.\\
+	 In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{Sutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}.
+To train these methods, a large number of direct interactions with the system is required, because they are on-policy.
+In the latter one deals with approximate dynamic programming, which we discuss in \cref{s:Model free reinforcement learning}.
+One advantage of these methods is, that they can use previously stored data as they are off-policy.
+A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010}.
+	\subsection{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
+	We cannot provide an exhaustive treatment of state of the art MFRL algorithms, but we focus on the main principles.
+An overview can be found e.g.
+in \cite{Sutton2018,Levine2020}.
+In the following, we discuss the normalised advantage function algorithm in some detail, which has good characteristics for accelerator control problems as shown in \cite{Kain2020,Hirlaender2020a}.
+	\subsubsection{Approximate dynamic programming}
+	The state-value function $Q^\policy(\bs,\ba)$, which tells us how good an action $\ba$ in a state $\bs$ following $\policy$ in terms of the expected return is:
 	\begin{align}
@@
 	is expressed as $Q^\policy_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
-By satisfying the Bellmann-optimality equation $Q^\policy_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$:
+By satisfying the Bellmann-optimality equation $Q^\policy_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$ by minimizing the Bellman error:
 	\begin{align}
@@
 	\end{align}
-The form of this equation can cause overestimation and other complications, when using a function approximator.
-Several methods exist which try to mitigate the problems \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
-One way to reduce the effect is to take a simple analytical form of the $Q$-function.
-\subsection{Normalized advantage function}
+	The form of this equation can cause overestimation and other complications, when using a function approximator and several attempts exist to reduce the problems \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
+	\subsubsection{Normalized advantage function}
+	One way to avoid the sensitivity to the mentioned complications is to take a simple analytical form with an explicitly calculable maximum of the $Q$-function.
 	If a specific quadratic form of the $Q$ function is assumed \cite{Gu2016}:
@@
 	\end{align}
-
-One modification, which is used in these tests is a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
-It is motivated by double Q-learning \cite{NIPS2010_091d584f,Hasselt2015}.
-The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality}:
+	One modification, which is used in these tests is the use of a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
+It is motivated by double Q-learning \cite{NIPS2010_091d584f,Hasselt2015,fujimoto2018addressing}.
+	The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality} the loss $L$ can be formulated as:
 	\begin{align}
-	\min_\theta\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
+		&L(\theta)=\\\notag&\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
 	\end{align}
 	$\theta_\text{targ}$ are the weights of a target network, which is softly updated.
-To stabilize the network training a small artificial noise is added to the actions in the update.
-This algorithm has an extremely good sample-efficiency for suitable problems as can be found often in accelerators and is used as the baseline for the considered control problem, as it shows very good results.
+To stabilize the network training a small clipped artificial noise is added to the actions called \emph{smoothing} as described in \cref{appendix:naf2}.\\
+	 This implementation has an extremely good sample-efficiency for suitable problems as can be found often in accelerators.
+It is used as the baseline for the considered control problem, as it achieves good results.
 An illustration and additional changes to the original proposal \cite{Gu2016} and a previous implementation used for experiments using a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
@@
 
-\section{Uncertainty aware Dyna-style reinforcement learning}
-The original Dyna algorithm \cite{Kurutach2018} is modified here in several aspects.
-Generally, Dyna style algorithms \cite{Sutton1991} denote algorithms, where a MFRL algorithm is trained on purely synthetic data from an approximate dynamics model or on a mixture of synthetic and real data.
-We use only synthetic data to reduce the interaction with the real environment to a minimum.
+	\subsection{Uncertainty aware DYNA-style reinforcement learning}
+	%The original DYNA algorithm \cite{Kurutach2018} is modified here in several aspects.
+
+	A nice overview of MBRL methods is provided in \cite{Wang2019}.
+Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and backpropagation in the dynamics model or \cite{Chua2018,Wang2019a}, a highly efficient sample based method.
+We focus on model-based data generation.\\
+		In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$:
+	\begin{align}
+		\hat{f_\theta}(\bs_t,\ba_t) := \{\transitions_\theta(\bs_t,\ba_t),r_\theta(\bs_t,\ba_t)\}.
+	\end{align}
+	Generally, DYNA style algorithms \cite{Sutton1991} denote algorithms, where an MFRL algorithm is trained on purely synthetic data from an approximate dynamics model $\hat \transitions$ or a mixture of synthetic and real data.
+We use only synthetic data to reduce the interaction with the real environment to a minimum.\\
 	An overview of the used method is shown in \cref{fig:MBRL_overview}.
-At the beginning the data is collected or read in from a pre-collection.
-An uncertainty aware model is trained, using anchored ensembles \cite{Pearce2018} on the data, which allows to take the allegorical (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
-Subsequently a MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data, by taking the uncertainty into account.
-After a defined number of training steps the controller is testes on each model individually.
-If there is no improvement in a specific ratio$<1$ of models, the controller is tested on the real environment for a number of episodes.
+In the beginning, the data is collected or read in from a pre-collection.
+An uncertainty aware model of the dynamics is trained, using anchored ensembles \cite{Pearce2018} on the data, which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
+It was motivated by the fact that pure epistemic ensemble techniques as \emph{ME-TRPO} were reported to be very sensitive to aleatoric noise \cite{Wang2019}.
+Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take the aleatoric and the epistemic uncertainties into account using probabilistic ensembles each model holding the parameters of the negative log likelihood of probability distribution function.\\
+	 Subsequently, an MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data by taking the ensemble uncertainty into account, as explained in the next section.\\
+	  After a defined number of training steps, the controller is testes on each model individually.
+If there is no improvement in a number smaller than the total number of the models, the controller is tested on the real environment for several episodes.
+In this way 'over-fitting' is avoided, as discussed in \cite{Kurutach2018}.
 The training is stopped if the controller solves the problem on the real environment.
-
+Otherwise, a batch of new data is collected.
 	\begin{figure}[!h]
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/MBRL_overview}
-	\caption{A schematic overview of the AE-Dyna approach used in this paper.}
+		\caption{A schematic overview of the AE-DYNA approach used in this paper.}
 		\label{fig:MBRL_overview}
@@
 	\subsection{Critical design decisions}\label{ss:critical_design}
-In the following the most important aspects of a successful application of a Dyna-AE algorithm are listed:
-\begin{itemize}
-	\item The ANN - including the prior - number of models - noise level - early stopping
-	\item The number of data points and the policy.
-	\item The uncertainty - how is this included?
-	\item The episodic design - avoid long trajectories.
-	\item The MFRL agent as e.g. the TRPO and PPO or the SAC and its training
-	\item Tuning the algorithm on a model.
-\end{itemize}
-We try to discuss all items in some detail in the following.
-The anchoring ensemble methods usually yields good results already with a small number of different networks.
-The idea behind it, is to mimic the posterior probability of the dynamics model to capture first of all its own uncertainty due to the lack of data.
-Empirical results show that three to five models were sufficient to see clear advantages over a single network approach and the main goal is not to calculate the exact posterior of the model.
-A small two-layer network with around 15-25 nodes and the $tanh$ activation were used.
-The last layer was linear and the inputs were normalized to the interval [-1,1].
-The prior is controlled by the initialization of the weights, which were normalized by the number of nodes in each layer.
-The noise level is also respected during the initialization.
-For the moment only homeoscetastic errors are used.
-Several methods to set the training method were implemented.
-Early stopping, a standard techniques showed good results, mainly to the fact, that the number of training steps is increasing with more data.
-The advantage is that the uncertainty at regions without data shrink in this way leading to a better training performance.
-A fixed loss value threshold was also tested, as also used in the original anchore-ensemble technique.
-In the experiments a combination of both was taken.\\
-One of the most crucial points is how this uncertainties are taken into account by the RL algorithms.
+	In the following the most important aspects of a successful application of model-based RL in accelerator problems are discussed, where we made some algorithmic design decisions which we think are beneficial for problems typically found in accelerator operation.\\
+%	 For the DYNA-AE algorithm  the important points are:
+%	\begin{itemize}
+%		\item The ANN - including the prior - number of models - noise level - early stopping.
+%		\item The number of data points and the policy.
+%		\item The uncertainty - how is this included?
+%		\item The episodic design - avoid long trajectories.
+%		\item The MFRL agent as, e.g. the TRPO and PPO or the SAC and its training.
+%		\item Tuning the algorithm on a model.
+%	\end{itemize}
+	The anchoring ensemble methods usually yield good results already with a small number of different networks.
+The idea is to mimic the posterior probability of the dynamics model to capture the uncertainty due to the lack of data.
+Empirical results showed that three to five models were satisfactory to see definite improvements over a single network approach, and the main purpose is not to determine the exact posterior of the model.
+A small two-layer network with around 15-25 nodes with a $tanh$ activation and a learning rate of $10^{-3}$ were used.
+The last layer was linear, and the inputs were normalised to the interval [-1,1].\\
+	 The prior is controlled by the initialisation of the weights of the network and defines how large the variation of the approximated function is assumed.
+The weights of the last layer were normalised by the number of nodes in each layer.
+The noise level is also respected during the initialisation.
+For the moment, only homoscedastic errors are used.\\
+	  Several methods of network training were implemented.
+Early stopping was employed successfully with a waiting time of about 20 steps.
+Mainly due to the fact that the number of training steps is increasing with more data in close-by domains.
+The benefit is that the modelled uncertainty at regions where no data is available is reduced in this way, leading to better training performance.
+A fixed loss value threshold was tested, as also used.
+In the experiments, a combination of both was taken.\\
+	One of the most crucial points is how these uncertainties are taken into account by the RL algorithms.
 Several different approaches were tried.
-It is possible to take the average of the models and add Gaussian noise leveled by the standard deviation of the models.
-Another straight forward way is to randomly select a model to provide at each training step, which is the original implementation of the ME-TRPO algorithm and was used in the experiments labeled \emph{ME-TRPO}.
+It is possible to take the average of the models and add Gaussian noise levelled by the standard deviation of the models.
+Another straight forward way is to randomly select a model to provide at each training step, which is used the original implementation of the \emph{ME-TRPO} algorithm and was used in the experiments labelled \emph{ME-TRPO}.
 A pessimistic setting only would select the model resulting in the lowest predicted reward.
-Good results were obtained by following a randomly selected single model each full episode and were used in the experiments labeled \emph{AE-Dyna}.\\
-The number of data-points taken every time the model as improved was firstly determined by the number of initial data points, collect using a random policy.
-The initialization phase has to be chosen not to small to minimize the chance of getting trapped in a local minimum for too long.
-Afterwards at least on full episode should be taken.
-We decided to use a short total episode length to reduce the impact of the compound error, the accumulation error following a wrong model, as well known for Dyna-style approaches.
+Pessimistic selection showed constant but too slow improvement for our tests.
+	\\ Good results were also obtained by following a randomly selected single model each full episode in combination with the  soft-actor-critic (\emph{SAC}), and were used in the experiments labelled \emph{AE-DYNA SAC}.
+The approach is similar \\
+	The number of data-points taken every time the model as improved was firstly determined by the number of initial data points, collected using a random policy.
+The initialisation phase has to be chosen not to small to minimise the chance of getting trapped in a local minimum for too long.
+Afterwards, at least one full episode should be taken.
+We decided to use a short horizon to reduce the impact of the compound error, the accumulation error following a wrong model, as well known for DYNA-style approaches.
 The maximal number of steps was ten.
@@
 	To decide which MFRL algorithm to use, two main algorithm classes have to be considered: the on and the off-policy algorithms.
-Online algorithm show a more stable convergence in general, while off-policy algorithms have the advantage that the data buffer can be filled with real data as well.
-A very attractive off-policy algorithm is the Soft-actor-critic.
-This algorithm tries to maximize the entropy of the actions to find a good trade-off between exploration and exploitation.
-On the other hand the on-policy algorithm TRPO provides some theoretical improvement guarantees.
-In the latter case the policy was improved over the whole training, while for the SAC the policy was reset, to take advantage of the exploration features.
-
-\subsection{Experiment results from FERMI RL tests}
-As discussed in the previous sections, several tests were performed on the FERMI XFEL.
-The main purpose was to test the newly implemented algorithms on a real system to evaluate their operational feasibility.
-The \emph{NAF2} algorithm, as a representative for highly sample efficient MFRL algorithms, was tested first.
+The online algorithm shows a more stable convergence in general, while off-policy algorithms have the advantage that the data buffer can be filled with real data as well.
+An attractive off-policy algorithm is \emph{SAC}.
+The \emph{SAC} not only tries to maximise $J$ \cref{eq:cumulative_reward} but also simultaneously is regularised to maximise the entropy in the action space.
+In this way, exploration is encouraged to avoid getting stuck in a local optimum and has a good trade-off between exploration and exploitation.\\
+	 On the other hand, the on-policy algorithm TRPO provides some theoretical improvement guarantees.
+As reported in \cite{Kurutach2018} using proximal policy optimization \cite{Schulman2017} as an on-line MFRL algorithm does not show the same performance, which was confirmed in our tests.
+In the latter case, the policy was improved over the whole training.
+For the \emph{SAC} the policy was reset after retraining the model, to profit from the exploration features.
+	\section{Experiment results from FERMI RL tests}
+	As discussed in the previous sections, several tests were performed on the FERMI FEL.
+	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility.
+A simplified simulation (details in \cref{a})was used to pre-tune the algorithms.
+The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, was tested first.
 	\subsection{MFRL tests}
-In total four tests were carried out, two using a single network and two using the double network architecture. In both cases \emph{smoothing} was applied, as described in \cref{appendix:naf2}.\\
-\Cref{fig:NAF_training} displays the results, averaged over the two test.
-A training of 100 episodes was done.
-In the upper figure the number of iterations per episode is plotted including the cumulative number of steps.
-In the verification \cref{fig:NAF_verification} both algorithms show a similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\
-Additionally the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps.
-The blue curves shows the Bellmann error, which is comparable.
-The state-value function, which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network tends to overestimate the value.
-In this case convergence is reached ~1400 steps.\\
-\subsection{MBRL tests}
-The second test campaign was employing the \emph{AE-DYNA} algorithm, as representative for pure a MBRL algorithms.
-Two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA SAC} variant.
-In both cases an ensemble of three network was used and the epistemic and aleatoric uncertainty is measured via the anchor-ensemble technique as discussed in \cref{ss:critical_design}.
-On top early stopping was used.
-\\ The first uses the trust region policy optimization -\emph{TRPO}- \cite{Schulman2015} to train the controller.
-The TRPO monotonically converges to better policy and this property is exploited in the training.
-The convergence property can be seen in \cref{fig:ME-TRPO_observables}.
-In the upper figure the total reward per batch is shown, as well as the number of data points used in the training of the dynamics model.
-In the lower plot the average cumulative reward as achieved by the TRPO on the individual models in the ensemble on a number of 10 episodes is drawn.
-The shade area shows the corresponding standard deviation to indicate the uncertainty of the dynamics model.
-To measure the convergence of the TRPO the logarithm of the standard deviation of the distribution drawn trajectories $p_\traj$ is also visualized.
-Here the training was stopped after 450 steps acquiring 25 steps each dynamics training.
-In the verification, as can be seen in \cref{fig:ME-TRPO_verification}, all of the 50 episodes where successfully finished after a few steps.
-\\
-Secondly, the \emph{AE-DYNA SAC} was tested, which uses the soft actor critic -SAC- algorithm \cite{Haarnoja2018a}.
-The SAC not only tries to maximise $J$ \cref{eq:cumulative_reward}, but also simultaneously but weights on the maximisation of the entropy in the action space.
-In this way exploration is encouraged to avoid getting stuck in a local optimum.
-In this test the controller is reset each time, when new data for the dynamics model training is acquired.
-Consequently, the performance drops each time, the dynamics model was retrained, as shown in \cref{fig:AE-DYNA_observables}.
-Chances to get trapped in a local optimum are smaller by using this strategy of training.
-In difference to the first MBRL test, the batches of when acquiring new data are 50 with an initial random walk of 100 steps.
-The number of initial steps was chosen carefully large enough, because the convergence is slowed down and there is the risk that the training becomes unfeasibly.
-The training was stopped after the acquisition of 500 data points and a verification was done as in the first test.
-Again the success is 100\%, but the number of iterations per step is smaller.
-It might be a result of the bigger number of data points, but in general, this method showed a better asymptotic performance than the $ME-TRPO$ variant.
 		\begin{figure}[!h]
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
-	\caption{The verification of the AE-DYNA-SAC on the FERMI FEL.}
+		\caption{The verification of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
 		\label{fig:NAF_convergence}
 	\end{figure}
+	In total, four tests were carried out, two using a single network and two using the double network architecture.
+A motivation for the used algorithmic set-up is discussed in \cref{appendix:naf2} on a classical example from control theory.
+In both cases \emph{smoothing} (explanation in \cref{appendix:naf2}) was applied.\\
+	\Cref{fig:NAF_training} displays the results, averaged over the two tests.
+A training of 100 episodes was accomplished.
+In the upper figure, the number of iterations per episode is plotted, including the cumulative number of steps.
+	In the verification \cref{fig:NAF_verification} both algorithms show similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\
+	Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps.
+The blue curves show the Bellmann error, which is comparable in both cases.
+The state-value function, which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value.
+In the single-network case, convergence is reached ~1400 steps.\\
+	\subsection{MBRL tests}
 
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
-	\caption{The training of the AE-DYNA-SAC on the FERMI FEL.}
+		\caption{The training of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
 		\label{fig:AE-DYNA_observables}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_verification.pdf}
-	\caption{The verification of the AE-DYNA-SAC on the FERMI FEL.}
+		\caption{The verification of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
 		\label{fig:AE-DYNA_verification}
@@
 	\end{figure}
+	The second test campaign was employing the \emph{AE-DYNA} algorithm, as a representative for pure MBRL algorithms.
+Two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA SAC} variant.
+The algorithmic design details are discussed in \cref{ss:critical_design}.
+	\\ The first uses the trust region policy optimisation -\emph{TRPO}- \cite{Schulman2015} to train the controller.
+The TRPO monotonically converges to better policy, and this property is exploited in training.
+The convergence property can be seen in \cref{fig:ME-TRPO_observables}.
+In the upper figure, the total reward per batch and the number of data points are shown as used in the training of the dynamics model.
+In the lower plot, the average cumulative reward as achieved by the TRPO on the individual models in the ensemble on ten episodes as in dependence of the epochs is drawn.
+During an epoch \emph{TRPO} is trained for 10000 steps on the synthetic data.
+The shaded area shows the corresponding standard deviation to indicate the uncertainty of the dynamics model.
+As a measure of convergence of the TRPO, the logarithm of the standard deviation of $p_\traj$ is visualised.
+The training was stopped after 450 steps collecting 25 steps each dynamics training.
+As shown in \cref{fig:ME-TRPO_verification}, all of the 50 verification episodes were successfully finished after a few steps.
+To verify the impact of the ensemble technique, a test with a single network using the same hyperparameters has been applied, where no convergence within 500 data points has been observed.
+\\
+	Secondly, the \emph{AE-DYNA SAC} was tested, which uses the \emph{SAC}- algorithm \cite{Haarnoja2018a}.
+In this test, the controller is reset each time, when new data for the dynamics model training is acquired.
+Consequently, the performance drops each time; the dynamics model was retrained, as shown in \cref{fig:AE-DYNA_observables}.
+Chances to get trapped in a local optimum are smaller by using this strategy of training.\\
+	In contrast to the first MBRL test, the new data batches are 50 with an initial random walk of 100 steps.
+The number of initial steps was chosen carefully high enough because the convergence is slowed down strongly, so that the training becomes unfeasible on a real machine.
+Each epoch consists of 2500 steps on the model.
+The training was stopped after the acquisition of 500 data points.
+The verification was executed as in the first test.
+Again the success is 100\%, but the number of needed iterations per episodes is less than for the ME-TRPO.
+It might be a result of the higher number of data points (50), but in general, this method exhibited better asymptotic performance than the $ME-TRPO$ variant.
 
 	\section{Discussion and outlook}
+	In this paper, the applicability of the method of deep reinforcement learning on the optimization of the FERMI FEL intensity was presented.
+Two different approaches were tested: a model free and a model based method.
+Both experiments yielded satisfactory results showing that this non-linear and noisy problem could be solved in a feasibly number of training steps.\\
 
 	\section{Conclusions}
-
+		 The used methods hold tremendous promise for automatizing setups common in accelerators.
+One purpose of this work was to provide some suggestions how to make advanced deep reinforcement learning techniques work on a real set-up by adjusting available methods.
+	\newpage
 	\appendix
 	\section{A non-linear standard control problem}
-To provide some transparency of these studies for other labs we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.
+	To provide some transparency of these studies for other labs, we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.\\
 	 It is a non-linear low dimensional unsolved continuous control problem.
 Unsolved means there is no threshold for the reward to terminate an episode.
-The episode length is set to 200 steps.
+The episode length, the horizon, is set to 200 steps.
 In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly in terms of noise handling.
-This is of importance when dealing with measurements on real systems.
+It is of importance when dealing with measurements on real systems.\\
+	 For statistical significance, all shown results were performed with five different seeds.
+The average value and the standard deviation (shaded) are plotted.
+For this study, we assume that the problem is successfully solved, if the cumulative reward surpasses a threshold of -200.
+The threshold is indicated by a dashed green line in the corresponding figures.
+%	  and a rolling average of 15 steps is used for a better visualization.
 
@@
 	%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
- We compare the different \emph{NAF} variants: \emph{smoothing-double}, \emph{smoothing single}, \emph{standart}.
-The smoothing adds a small clipped noise on the actions to stabilize the network training as:
+	We compare the different \emph{NAF} variants: \emph{Clipping}, \emph{No-clipping-smoothing}, \emph{No-clipping-no-smoothing}, where \emph{clipping} indicates the use of the double network and in all other cases a single network is used.
+	 The term \emph{smoothing} indicates that a small clipped noise is added on the actions to stabilize the network training as:
 	\begin{equation}
- 	a'(s') = \text{clip}\left(\mu_{\theta_{\text{targ}}}(s') + \text{clip}(\epsilon,-c,c), a_{Low}, a_{High}\right),
+		\ba(\bs) = \text{clip}\left(\mu_{\theta_{\text{targ}}}(\bs) + \text{clip}(\epsilon,-c,c), \ba_{Low}, \ba_{High}\right),
 	\end{equation}
- 	where $\epsilon \sim \mathcal{N}(0, \sigma)$. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
+	where $\epsilon \sim \mathcal{N}(0, \sigma)$.
+$c>0$ denotes the clipping coefficient and $ \ba_{Low}, \ba_{High}$ the minimal and maximal possible action.
+This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
 	The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
 	\begin{align}
-y(r,s',d) = r + \gamma (1 - d) \min_{i=1,2} Q_{\phi_{i, \text{targ}}}(s', a'(s'))
+		y(\reward_t,\bs_{t+1}, d_t) = \reward_t + \gamma (1 - d_t) \min_{i=1,2} V_{\theta_{i, \text{targ}}}(\bs_{t+1}),
 	\end{align}
-and then both are learned by regressing to this target:
-\begin{align}
-	L(\phi_1, {\mathcal D}) = \mathbb E_{(s,a,r,s',d) \sim {\mathcal D}}{
-		\Bigg( Q_{\phi_1}(s,a) - y(r,s',d) \Bigg)^2
-	}
-\end{align}
-and the policy is obtained via:
-\begin{align}
-\max_{\theta} \underset{s \sim {\mathcal D}}{{\mathrm E}}\left[ Q_{\phi_1}(s, \mu_{\theta}(s)) \right].
-\end{align}
-The results are shown in \cref{fig:comparsion_smoothing}.
-One sees the cumulative reward per episode for a training of a total of 50 episodes.
-The curve labelled \emph{clipping} corresponds to the double network including the smoothing method and shows the best stability during the training yielding quickly a high reward.
-Also the single network shows good and comparable performance, except for the stability.
-The worst performance is achieved without smoothing and a single network, nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
+	with $d$ is one if the episode is finished and 0 otherwise.
+	Then both are learned by regressing to this target (using the tuples from the data buffer $\mathcal D$ ):
+	\begin{multline}
+		L(\theta_i, {\mathcal D}) \\= \mathbb E_{(\bs_{t},\ba_{t},r_t,\bs_{t+1},d_t) \sim {\mathcal D}}{
+			\Bigg( Q_{\theta_i}(\bs_{t},\ba_{t}) - y(r_t,\bs_{t+1},d_t) \Bigg)^2
+		},
+	\\
+	i \in \{1,2\}
+	\end{multline}
+	and the policy is obtained via $\max_\ba Q_{\theta_1}(\bs,\ba) = \mu_{\theta_1}$.\\
+	The results are shown in \cref{fig:comparsion_smoothing_small}.
+One sees the cumulative reward per episode for a training of a total of 100 episodes.
+As mentioned, the curve labelled \emph{clipping} corresponds to the double network including \emph{smoothing} and shows the best overall stability during the training yielding quickly a high reward.
+Also the \emph{smoothed} single network, labelled \emph{No-clipping-smoothing}, shows good and comparable performance, except for the sightly decreased stability.
+The worst performance is achieved without smoothing and a single network (\emph{No-clipping-no-smoothing}), nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
+%	\begin{figure}[!h]
+%		\centering
+%		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
+%		\caption{Cumulative reward of different NAF implementations as discussed in the text.}
+%		\label{fig:comparsion_smoothing}
+%	\end{figure}
 	\begin{figure}[!h]
 	\centering
-	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
-	\caption{Cumulative reward of different NAF implementations as discussed in the text.}
-	\label{fig:comparsion_smoothing}
+	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_naf}
+	\caption{Cumulative reward of different \emph{NAF} implementations as discussed in the text without noise.}
+	\label{fig:comparsion_smoothing_small}
 \end{figure}
@@
 There the difference of the three methods becomes even more evident.The results are shown in \cref{fig:comparsion_noise}.
-After around 65 episodes the single network without smoothing decreases before reaching the final performance at around 95 episodes, while the smoothing prevents this performance drop.
+After around 65 episodes the single network without \emph{smoothing} (\emph{No-clipping-no-smoothing}) decreases before reaching the final performance at around 95 episodes, while  \emph{smoothing} prevents this performance drop in the other cases.
 	\begin{figure}[!h]
@@
 	\end{figure}
-Using the anchors in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}.
+	\\Using the anchors (the noise level is homeoscetastic) and known in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}.
 It the upper plot the mean cumulative reward on 10 test episodes on the real environment is shown during the training.
-It should indicate the result if the training is stopped at this training iteration.
+It should indicate the result if the training is stopped at this training epoch.
 One cannot observe this during a real training, unless one does costly performance measurements during the training.
-The blue curve reaches the target much quicker exhibiting less variation.\\
+The blue curve reaches the target much quicker exhibiting less variation.
+An epoch consists of 3000 iterations of the \emph{SAC}.\\
 	The lower plot of \cref{fig:comparsion_noise_ae_dyna} shows the batch rewards a measured during the data collection, which is observable during the training.
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise_ae_dyna}
-	\caption{Cumulative reward of AE-DYNA with artificial noise using the anchor.}
+		\caption{Cumulative reward of \emph{AE-DYNA} with artificial noise using the anchor.}
 		\label{fig:comparsion_noise_ae_dyna}
 	\end{figure}
-
+\subsection{NAF versus AE-DYNA}
+	Finally, \cref{fig:comparsion_NAF_AE-DYNA} should demonstrate the sample efficiency of the \emph{AE-DYNA-SAC} and the \emph{NAF} algorithm on the noisy \emph{inverted pendulum}.
+\emph{AE-DYNA} converges below 2000 data points (without noise even below 800) and the \emph{NAF2} starts to perform equally 10000 data points surpassing the -200 reward threshold.
+One clearly sees the increased sample efficiency on this problem using the \emph{AE-DYNA} in contrast to the \emph{NAF2}.
 		\begin{figure}[!h]
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_ae_dyna}
-	\caption{The comparison of the NAF2 and the AE-DYNA on the noisy pendulum.}
-	\label{fig:comparsion_noise_ae_dyna}
+		\caption{The comparison of the NAF2 and the \emph{AE-DYNA} on the noisy pendulum.}
+		\label{fig:comparsion_NAF_AE-DYNA}
 	\end{figure}
 
-\section{Theoretical aspects on the AE-DYNA on FERMI FEL}
-In this section we want to discuss some theoretical aspects concerning the AE-DYNA approach. A simulation was used to obtain the presented analysis. \\
-One issue of MBRL methods is the asymptotic performance.
-Altough good results were obtained using the AE-DYNA, there is a difference between the MFRL and MBRL.
+%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
+%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
+%	One issue of MBRL methods is asymptotic performance.
+Although good results were obtained using the \emph{AE-DYNA}, there is a difference between the MFRL and MBRL.
 We call it the \emph{reality gap}.
-There are ideas to attack the problem e.g.
-by learning a meta-policy \cite{Clavera2018} followed by a fine tuning on the residual physics, which minimizes the \emph{reality gap} ot training on a simulator and closing than with a small amount of training iterations on the real system the gap \cite{Zeng2019}.
-\begin{figure}[!h]
-	\centering
-	\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
-	\caption{The training .}
-	\label{fig:Learning_evolution}
-\end{figure}
+There are ideas to attack the problem, e.g.
+by learning a meta-policy \cite{Clavera2018} followed by a fine-tuning on the residual physics, which minimises the \emph{reality gap} of training on a simulator and closing than with a small number of training iterations on the real system the gap \cite{Zeng2019}.
+%	\begin{figure}[!h]
+%		\centering
+%		\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
+%		\caption{The training .}
+%		\label{fig:Learning_evolution}
+%	\end{figure}
+\section{Still open}
+	\begin{itemize}
+%	\item Noise benchmarks with pendulum - done
+%	\item AE details from papers
+%	\item Benchmark NAF2 with pendulum - done
+%	\item Compare NAF2 and AE-DYNA
+%	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
+%	\item Tailored to accelerators - short horizons are standard?
+	\item Long waiting time in MBRL	
+	\item Vary the model size
+%	\item Noise sensitivity as mentioned in Benchmark paper
+%	\item Improvements of the algorithms:
+%	\begin{itemize}
+%		\item Adding the noise feature
+%		\item Jumping feature using \emph{SAC}
+%		\item Noise in the policy
+%		\item discuss different approaches to, e.g. dynamic waiting time
+%		\item usually prior not discussed in papers
+%	\end{itemize}
+\end{itemize}
 
 	% The \nocite command causes all entries in a bibliography to be printed out
-% whether or not they are actually referenced in the text. This is appropriate
+	% whether or not they are referenced in the text. This is appropriate
 	% for the sample file to show the different styles of references, but authors
@@
 	%
-%\bibitem{multiobjective-optimization} A.
+	%\bibitem{multiobjective-optimisation} A.
 Edelen, N.
@@
 Mayes, and A.
-Adelmann, "Machine learning for orders of magnitude speedup in multiobjective optimization of particle accelerator systems", Phys.
+Adelmann, "Machine learning for orders of magnitude speedup in multiobjective optimisation of particle accelerator systems", Phys.
 Rev.
@@
 	%
-%\bibitem{SBP} A. Nagabandi, G. Kahn, R. S. Fearing, S. Levine, "Neural Network Dynamics for Model-Based Deep Reinforcement Learning with Model-Free Fine-Tuning", arXiv:1708.02596, 2017. 
+	%\bibitem{SBP} A. Nagabandi, G. Kahn, R. S. Fearing, S. Levine, "Neural Network DYNAmics for Model-Based Deep Reinforcement Learning with Model-Free Fine-Tuning", arXiv:1708.02596, 2017. 
 	%
-%\bibitem{DYNA} R. Sutton, "Dyna, an integrated architecture for learning, planning, and reacting", AAAI Spring Symposium, pp. 151-155, 1991, \url{https://doi.org/10.1145/122344.122377}.
+	%\bibitem{DYNA} R. Sutton, "DYNA, an integrated architecture for learning, planning, and reacting", AAAI Spring Symposium, pp. 151-155, 1991, \url{https://doi.org/10.1145/122344.122377}.
 	%
@@
 	%
-%\bibitem{deeppilco} Y. Gal, R. McAllister, and C. E. Rasmussen, “Improving PILCO with Bayesian neural network dynamics models”, in Data-Efficient Machine Learning workshop, ICML, 2016, vol. 4, p. 34.
+	%\bibitem{deeppilco} Y.
+Gal, R.
+McAllister, and C.
+E.
+Rasmussen, “Improving PILCO with Bayesian neural network dynamics models”, in Data-Efficient Machine Learning workshop, ICML, 2016, vol.
+4, p.
+34.
 	%
-%\bibitem{per} T. Schaul, J. Quan, I. Antonoglou, D. Silver, "Prioritized Experience Replay", arXiv:1511.05952, 2015-11-18.
+	%\bibitem{per} T. Schaul, J. Quan, I. Antonoglou, D. Silver, "Prioritised Experience Replay", arXiv:1511.05952, 2015-11-18.
 	%
@@
 	%\bibitem{linac4} G.
-Bellodi, "Linac4 Commissioning Status and Challenges to Nominal Operation", 61$^st$ ICFA Advanced Beam Dynamics Workshop on High-Intensity and High-Brightness Hadron Beams, Daejeon, Korea, 17 - 22 Jun 2018, pp.MOA1PL03, \url{https://doi.org/10.18429/JACoW-HB2018-MOA1PL03}.
+Bellodi, "Linac4 Commissioning Status and Challenges to Nominal Operation", 61$^st$ ICFA Advanced Beam DYNAmics Workshop on High-Intensity and High-Brightness Hadron Beams, Daejeon, Korea, 17 - 22 Jun 2018, pp.MOA1PL03, \url{https://doi.org/10.18429/JACoW-HB2018-MOA1PL03}.
 	%
-%\bibitem{auto_matching} F. Velotti, B. Goddard et al., "Automatic AWAKE electron beamline setup using unsupervised machine learning", to be submitted to Phys. Rev. Accel. Beams.
+	%\bibitem{auto_matching} F. Velotti, B. Goddard et al., "Automatic AWAKE electron beamline set-up using unsupervised machine learning", to be submitted to Phys. Rev. Accel. Beams.
 	%
@@
 	%
-%\bibitem{Kumar_2} A. Kumar, J. Fu, G. Tucker, S. Levine, “Stabilizing Off-Policy Q-Learning via Bootstrapping Error Reduction,” arXiv:1906.00949, 2019-06-03.
+	%\bibitem{Kumar_2} A. Kumar, J. Fu, G. Tucker, S. Levine, “Stabilising Off-Policy Q-Learning via Bootstrapping Error Reduction,” arXiv:1906.00949, 2019-06-03.
 	%

### 2020-11-30 AEDYNA.py

@@ -28,13 +28,13 @@ config = None
 
 steps_per_env = 201
 init_random_steps = 201
-total_steps = 2009
-num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
-
+# total_steps = 1207
+# num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
+num_epochs = 5
 print('Number of epochs: ', num_epochs)
 
 max_training_iterations = 50
-delay_before_convergence_check = 1
+delay_before_convergence_check = 5
 
 # Select the rl-algorithm
 
@@ -51,7 +51,7 @@ else:
 simulated_steps = 2500
 
 model_batch_size = 100
-num_ensemble_models = 5
+num_ensemble_models = 3
 
 early_stopping = True
 model_iter = 10
@@ -60,7 +60,7 @@ model_iter = 10
 
 # How often to check the progress of the network training
 # e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
-dynamic_wait_time = lambda it, ep: (it + 1) % 1 == 0  #
+dynamic_wait_time = lambda it, ep: (it + 1) % (ep+1)*2 == 0  #
 
 # Learning rate as function of ep:
 lr_start = 1e-3
@@ -175,7 +175,6 @@ class MonitoringEnv(gym.Wrapper):
             self.max_steps = env.max_steps
         except:
             self.max_steps = max_steps
-            print('Max steps set to:', self.max_steps)
         self.current_step = 0
 
     def reset(self, **kwargs):
@@ -487,7 +486,7 @@ class NetworkEnv(gym.Wrapper):
         state = np.zeros(self.env.observation_space.shape)
         maximum = 0
         if data is not None:
-            action = [data[1]]
+            # action = [data[1]]
             state = data[0]
             maximum = (data[2] - 1) / 2
         delta = 0.05
@@ -509,7 +508,8 @@ class NetworkEnv(gym.Wrapper):
                 # print(self.number_models)
                 for i1 in range(len(x)):
                     for j1 in range(len(y)):
-                        state[0] = x[i1]
+                        state[0] = np.cos(x[i1])
+                        state[1] = np.sin(x[j1])
                         state[1] = y[j1]
                         rewards[i1, j1] = (self.model_func(state, [np.squeeze(action)],
                                                            nr))[1] / num_ensemble_models
@@ -1207,9 +1207,12 @@ if __name__ == '__main__':
         # clipped_double_q
[…]

### 2020-11-30 AE_dyna_results.py

@@ -5,7 +5,7 @@ import pickle
 import matplotlib.pyplot as plt
 import numpy as np
 
-label = ["ME-TRPO", "AE-DYNA", 'Simulation'][0]
+label = ["ME-TRPO", "AE-DYNA", 'Simulation'][1]
 
 if label == "ME-TRPO":
     # ME-TRPO results
@@ -198,7 +198,7 @@ def plot_observables(data, label='Experiment', **kwargs):
     ax.set_ylabel('rewards tests (arb. units)')
     # plt.tw
     ax.grid(True)
-    ax.set_xlabel('no. iterations')
+    ax.set_xlabel('no. epochs')
     if length_all:
         ax2 = ax.twinx()
         color = 'lime'

### 2020-11-30 modelling_keras.py

@@ -1,74 +1,0 @@
-import numpy as np
-import tensorflow as tf
-from tensorflow import keras
-from tensorflow.keras import layers
-
-
-def fc(x, hidden_size, activation=tf.nn.tanh,
-       kernel_initializer=tf.compat.v1.random_uniform_initializer(-0.01, 0.01),
-       name=None):
-    layer = keras.layers.Dense(hidden_size, activation=activation,
-                               kernel_initializer=kernel_initializer,
-                               kernel_regularizer=None,
-                               bias_initializer=tf.compat.v1.constant_initializer(0.0), name=name)
-    return layer(x)
-
-
-kernel_initializer = tf.compat.v1.random_uniform_initializer(-0.01, 0.01)
-
-obs_dim = 2
-act_dim = 2
-hidden_sizes = (100, 100)
-inputs_state = keras.Input(shape=(obs_dim,), name="state_input")
-
-# h = inputs[:, 0:obs_dim]
-h = inputs_state
-for hidden_dim in hidden_sizes:
-    h = fc(h, hidden_dim, kernel_initializer=kernel_initializer)
-V = fc(h, 1, activation=None, kernel_initializer=kernel_initializer, name='V')
-
-l = fc(h, (act_dim * (act_dim + 1) / 2),
-       kernel_initializer=kernel_initializer, name='l')
-mu = fc(h, act_dim, kernel_initializer=kernel_initializer, name='mu')
-state_out_put = [V, l, mu]
-
-value_model = keras.Model(inputs_state, V, name='value_model')
-action_model = keras.Model(inputs_state, mu, name='action_model')
-
-inputs_action = keras.Input(shape=(act_dim,), name="action_input")
-
-pivot = 0
-rows = []
-for idx in range(act_dim):
-    count = act_dim - idx
-    diag_elem = tf.exp(tf.slice(l, (0, pivot), (-1, 1)))
-    non_diag_elems = tf.slice(l, (0, pivot + 1), (-1, count - 1))
-    row = tf.pad(tensor=tf.concat((diag_elem, non_diag_elems), 1), paddings=((0, 0), (idx, 0)))
-    rows.append(row)
-    pivot += count
-L = tf.transpose(a=tf.stack(rows, axis=1), perm=(0, 2, 1))
-P = tf.matmul(L, tf.transpose(a=L, perm=(0, 2, 1)))
[…]

### 2020-11-30 naf2_new.py

@@ -440,7 +440,7 @@ class QModel:
 class NAF(object):
     def __init__(self, env, training_info=dict(), pre_tune=None,
                  noise_info=dict(), save_frequency=500, directory=None, is_continued=False,
-                 clipped_double_q=False, q_smoothing=0.01, **nafnet_kwargs):
+                 clipped_double_q=False, q_smoothing_sigma=0.02, q_smoothing_clip=0.05, **nafnet_kwargs):
         """
         :param env: open gym environment to be solved
         :dict training_info: dictionary containing info for the training of the network
@@ -450,12 +450,15 @@ class NAF(object):
         :param directory: directory were weights are saved
         :param is_continued: continue a training, otherwise given directory deleted if existing
         :param clipped_double_q: use the clipped double q trick with switching all clipped_double_q steps
+        :param q_smoothing_clip: add small noise on actions to smooth the training
         :param q_smoothing: add small noise on actions to smooth the training
         :param nafnet_kwargs: keywords to handle the network and training
         """
 
         self.clipped_double_q = clipped_double_q
-        self.q_smoothing = q_smoothing
+        self.q_smoothing_clip = q_smoothing_clip
+        self.q_smoothing_sigma = q_smoothing_sigma
+
         self.losses2 = []
         self.vs2 = []
         self.model_switch = 1
@@ -552,12 +555,11 @@ class NAF(object):
         # Add small noise on the controller
         elif is_train:
             action = self.noise_function(np.squeeze(model.get_action([state])),self.idx_episode)
-            if self.q_smoothing is None:
+            if self.q_smoothing_clip is None:
                 return_value = np.clip(action, -1, 1)
             else:
-                sigma = 0.01
-                return_value = np.clip(action + np.clip(sigma * np.random.randn(self.action_size),
-                                                        -self.q_smoothing, self.q_smoothing), -1, 1)
+                return_value = np.clip(action + np.clip(self.sigma * np.random.randn(self.action_size),
+                                                        -self.q_smoothing_clip, self.q_smoothing_clip), -1, 1)
             return return_value
         else:
             action = model.get_action([state])

### 2020-11-30 read_naf_tests.py

@@ -8,9 +8,9 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 
 
-def read_experiment(root_dir):
+def read_experiment(root_dir, label = ''):
     data_all = pd.DataFrame()
-    n_average = 15
+    n_average = 25
     for root, dirs, files in os.walk(root_dir):
         for file in files:
             if file == 'plot_data_0.pkl':
@@ -19,7 +19,8 @@ def read_experiment(root_dir):
                 openfile = open(file_name, 'rb')
                 data = np.sum(pickle.load(openfile), axis=1)
                 # data_frame = pd.DataFrame(data, columns=[root.split('/')[0]])
-                data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
+                column_names = [root.split('_')[1] + label]
+                data_frame = pd.DataFrame(data, columns=column_names)
                 data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
                 data_all = pd.concat([data_all, data_frame.T], sort=False)
                 openfile.close()
@@ -35,20 +36,41 @@ except:
     # double = read_experiment(root_dir=root_dir)
 
     # data_all = pd.concat([single, double], sort=False)
-    root_dir = "Experiments-150-200-50-noisy/"
+    root_dir = "Experiments-150-200-50-long/"
     data_all = read_experiment(root_dir=root_dir)
 
 
-figure_name = '../Figures/Comparison_noise.pdf'
+# root_dir = "Experiments-150-200-50-long-0.05/"
+# data_all_1 = read_experiment(root_dir=root_dir)
+root_dir = "Experiments-150-200-50-noisy/"
+data_all_2 = read_experiment(root_dir=root_dir)
+
+# data_all = data_all_1
+# data_all = pd.concat([data_all, data_all_2])
+figure_name = '../Figures/Comparison___.pdf'
 current_test_name = ''
 
-fig, axs = plt.subplots(1, 1)
+fig, axs = plt.subplots(2, 1, sharex=True)
+ax = axs[0]
 
-# data_all = data_all.loc[:, :length]
 print(data_all)
+
 sns.set_palette(sns.color_palette("bright", 8))
-sns.lineplot(ax=axs, data=data_all.T)
+sns.lineplot(ax=ax, data=data_all.T)
[…]

### 2020-11-30 read_paper_tests.py

@@ -24,10 +24,11 @@ def read_experiment_observables(root_dir, key = 'tests_all'):
             if not key in ['step_counts_all']:
                 data = pd.DataFrame(pickle.load(openfile)[key]).sum()
                 # print(data)
-                data_frame = pd.DataFrame(data, columns=[root.split('_')[4]])
+                data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
                 # print(data_frame)
             else:
                 data = pickle.load(openfile)[key]
+                print(root.split('_')[1])
                 data_frame = pd.DataFrame(data, columns=[root.split('_')[4]])
 
             data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
@@ -78,47 +79,52 @@ def read_experiment_final(root_dir):
     return data_all
 
 
-root_dir = 'Data_logging/Experiment_noise_pendulum/'
-
-figure_name = '../../Figures/Comparison_noise_ae_dyna.pdf'
-
-fig, axs = plt.subplots(3, 1, sharex=True)
-ax=axs[0]
-data_all_observables = read_experiment_observables(root_dir=root_dir, key='tests_all')
-data_all_observables = data_all_observables.loc[:, :60]
-sns.set_palette(sns.color_palette("bright", 3))
-sns.lineplot(ax=ax, data=data_all_observables.T, legend=False)
-# ax.set_title('Verification on real environment')
-ax.axhline(y=-200, ls=':', color = 'lime')
-ax.set_ylabel('mean cum. reward\n (arb. units.)')
-
-ax = axs[1]
-data_all_observables = read_experiment_observables(root_dir=root_dir, key='step_counts_all')
-data_all_observables = data_all_observables.loc[:, :60]
-sns.set_palette(sns.color_palette("bright", 8))
-sns.lineplot(ax=ax, data=data_all_observables.T)
-ax.set_ylabel('no. data points')
-
-ax=axs[2]
-data_all_observables = read_experiment_observables(root_dir=root_dir, key='batch_rews_all')
-data_all_observables = data_all_observables.loc[:, :60]
-sns.set_palette(sns.color_palette("bright", 8))
-sns.lineplot(ax=ax, data=data_all_observables.T, legend=False)
-ax.set_ylabel('batch reward\n (arb. units.)')
[…]

### 2020-11-30 run_aedyna_noise_test_pendulum.py

@@ -3,22 +3,23 @@ import shutil
 
 name = 'random_seeds'
 # root_dir = "DOUBLE/"
-root_dir = 'Data/Simulation/noisy/'
+root_dir = 'Data/Simulation/ModelSize/'
 
 # if os.path.exists(root_dir):
 #     shutil.rmtree(root_dir)
 
 parameter_list = [
-    'Noise-off-non-aleatoric',
-    'Noise-on-non-aleatoric',
-    'Noise-on-aleatoric',
+    'Single',
+    'Three',
+    'Five',
+    'Ten'
 ]
 
 for index in range(len(parameter_list)):
     name = 'Setting'
     name += '_' + str(parameter_list[index])
     print(name)
-    for seed in [10, 20, 30, 40, 50]:
+    for seed in [10]:
         command = "python AEDYNA.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
         os.system(command)
 

### 2020-11-30 run_paper_naf_tests.py

@@ -3,7 +3,7 @@ import shutil
 
 name = 'random_seeds'
 # root_dir = "DOUBLE/"
-root_dir = "Experiments-150-200-50-noisy/"
+root_dir = "Experiments-400-100-small-smoothing/"
 
 # if os.path.exists(root_dir):
 #     shutil.rmtree(root_dir)
@@ -18,16 +18,17 @@ root_dir = "Experiments-150-200-50-noisy/"
 #     'False-high-noisy'
 # ]
 parameter_list = [
-    'Clipping-noisy',
-    'NocNoS-noisy',
-    'NoCS-noisy',
+    'Cls',
+    'Cns',
+    'Ncls',
+    'Ncns',
 ]
 
 for index in range(len(parameter_list)):
     name = 'Setting'
     name += '_' + str(parameter_list[index])
     print(name)
-    for seed in [10, 20, 30, 40, 50]:
+    for seed in [10]:
         command = "python run_naf2_for_tests.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
         os.system(command)
 

### 2020-12-01 main.tex
- paper changers

@@
 
-	\title{Model based and model-free reinforcement learning at FERMI FEL}
+	\title{Online Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Applications to FERMI FEL}
 
@@
 derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015} or model-based optimisations as Gaussian processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.
-Another promising way is to apply reinforcement learning \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020} which unveils several advantages over optimisation methods:
+Another promising way is to apply reinforcement learning (RL) \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020} which unveils several advantages over optimisation methods:
 	\begin{itemize}
@@
 	\end{align}
-	The Bellman operator $\bellman^*$ has a unique fixed point but is non-linear, due to the $\max$ - operator:
-	\begin{align}
-		\bellman^*\vec Q_\theta(\bs_t,\ba_t) := r(\bs_t,\ba_t)+\gamma\max_\ba\left( Q_\theta(\bs_{t+1},\ba) - Q_\theta(\bs_t,\ba_t)\right).
-	\end{align}
+	The Bellman operator $\bellman^*$ has a unique fixed point but is non-linear, due to the $\max$ - operator \cite{Sutton2018}:
+	\begin{multline}
+		\bellman^*\vec Q_\theta(\bs_t,\ba_t) := \\r(\bs_t,\ba_t)+\gamma\max_\ba\left( Q_\theta(\bs_{t+1},\ba) - Q_\theta(\bs_t,\ba_t)\right).
+	\end{multline}
 	The form of this equation can cause overestimation and other complications, when using a function approximator and several attempts exist to reduce the problems \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
@@
 Pessimistic selection showed constant but too slow improvement for our tests.
-	\\ Good results were also obtained by following a randomly selected single model each full episode in combination with the  soft-actor-critic (\emph{SAC}), and were used in the experiments labelled \emph{AE-DYNA SAC}.
-The approach is similar \\
+	\\ Good results were also obtained by following a randomly selected single model each full episode in combination with the  soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}, and were used in the experiments labelled \emph{AE-DYNA SAC}.
+\\
 	The number of data-points taken every time the model as improved was firstly determined by the number of initial data points, collected using a random policy.
@@
 Both experiments yielded satisfactory results showing that this non-linear and noisy problem could be solved in a feasibly number of training steps.\\
-
+	The model-free methods were slightly better in the final performance but also more sample were taken.
+The model-based methods had to be stopped due to operational reasons, hence additional studies regarding a long time performance would be interesting.
+One big issue when applying model-based methods is the computational effort.
+In our tests the data acquisition time was only a fraction of the time needed to train the controller on the learned model.
+In case, the used methods could be parallelized to decrease the computational time.
 	\section{Conclusions}
  The used methods hold tremendous promise for automatizing setups common in accelerators.
-One purpose of this work was to provide some suggestions how to make advanced deep reinforcement learning techniques work on a real set-up by adjusting available methods.
+One purpose of this work was to provide some suggestions how to make advanced deep reinforcement learning techniques work on a real set-up by adjusting available methods.\\
+ Generally control problems in accelerators have a short horizon, which makes the usage of \emph{DYNA-style} algorithms attractive as the \emph{Me-TRPO} or our  \emph{AE-DYNA} attractive choices for highly sample efficient algorithms.\\
+
+ To provide the possibility for other labs to use the proposed methods the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020}.
 	\newpage
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_naf}
-	\caption{Cumulative reward of different \emph{NAF} implementations as discussed in the text without noise.}
+	\caption{Cumulative reward of different \emph{NAF} implementations as discussed in the text on the \emph{inverted pendulum} without noise.}
 	\label{fig:comparsion_smoothing_small}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise}
-		\caption{Cumulative reward of different NAF implementations with artificial noise as discussed in the text.}
+		\caption{Cumulative reward of different \emph{NAF} implementations on the \emph{inverted pendulum} with artificial noise as discussed in the text.}
 		\label{fig:comparsion_noise}
@@
 One cannot observe this during a real training, unless one does costly performance measurements during the training.
-The blue curve reaches the target much quicker exhibiting less variation.
+Respecting the aleatoric (\emph{Noise-on-aleatoric}) helps to reach the target much quicker exhibiting less variation compared to the standard use of an ensemble (\emph{Noise-on-non-aleatoric}).
 An epoch consists of 3000 iterations of the \emph{SAC}.\\
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise_ae_dyna}
-		\caption{Cumulative reward of \emph{AE-DYNA} with artificial noise using the anchor.}
+		\caption{Cumulative reward of \emph{AE-DYNA} on the \emph{inverted pendulum} with artificial noise using the anchor.}
 		\label{fig:comparsion_noise_ae_dyna}
+	\end{figure}
+\Cref{fig:Compare_models_sizes} shows the impact of the number of models onto the performance on the \emph{inverted pendulum}.
+The maximum cumulative reward averaged over five different runs tested on the real environment during the training is visualized in dependence of the number of data points.
+A number of three models shows a good trade-off between performance and
+\begin{figure}[!h]
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/Compare_models_sizes}
+	\caption{Varying the number of models in the ensemble of the \emph{AE-DYNA} on the \emph{inverted pendulum}.}
+	\label{fig:Compare_models_sizes}
 \end{figure}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_ae_dyna}
-		\caption{The comparison of the NAF2 and the \emph{AE-DYNA} on the noisy pendulum.}
+		\caption{The comparison of the NAF2 and the \emph{AE-DYNA} on the noisy \emph{inverted pendulum}.}
 		\label{fig:comparsion_NAF_AE-DYNA}
@@
 	\begin{itemize}
+	\item mention the stable baseline.
 %	\item Noise benchmarks with pendulum - done
@@
 %	\item Tailored to accelerators - short horizons are standard?
-	\item Long waiting time in MBRL	
-	\item Vary the model size
+%	\item Long waiting time in MBRL	
+%	\item Vary the model size
 %	\item Noise sensitivity as mentioned in Benchmark paper
@@
 
-	% \bibliography{bibliography}
-	%\addcontentsline{toc}{chapter}{Bibliography}
-	%
-	%\begin{thebibliography}{9}
-	%
-	%\bibitem{lsa} D.
-Jacquet, R.
-Gorbonosov, G.
-Kruk, "LSA - The High Level Application Software of the LHC and its Performance during the first 3 years of Operation", ICALEPS, San Francisco, CA, USA, 6 - 11 Oct 2013, pp.thppc058.
-	%
-	%\bibitem{multiobjective-optimisation} A.
-Edelen, N.
-Neveu, M.
-Frey, Y.
-Huber, C.
-Mayes, and A.
-Adelmann, "Machine learning for orders of magnitude speedup in multiobjective optimisation of particle accelerator systems", Phys.
-Rev.
-Accel.
-Beams 23, 044601, 2020.
-	%
-	%\bibitem{collimator-alignment} G.
-Azzopardi, A.
-Muscat, G.
-Valentino, S.
-Redaelli, B.
-Salvachua, "Operational results of LHC collimator alignment using machine learning", Proc.
-IPAC'19, Melbourne, Australia, pp.
-1208-1211, 2019.
-	%
-	%\bibitem{SPSAlignment} %InProceedings (Hirlaender2019)
-	% S. Hirlaender, M. Fraser, B. Goddard , V. Kain, J. Prieto, L. Stoel, M. Szakaly, F. Velotti, "Automatisation of the SPS ElectroStatic Septa Alignment",  
-	%in 10th Int. Partile Accelerator Conf.(IPAC'19), 2019, p. 4001-4004.
-	%
-	%\bibitem{liu} E.
-Shaposhnikova et al., "LHC Injectors Upgrade (LIU) Project at CERN", 7$^{th}$ International Particle Accelerator Conference, Busan, Korea, 8 - 13 May 2016, pp.MOPOY059, \url{https://doi.org/10.18429/JACoW-IPAC2016-MOPOY059}.
-	%
-	%\bibitem{awake} E.
-Adli, A.
-Ahuja, O.
-Apsimon, R.
-Apsimon, A.-M.
-Bachmann, D.
-Barrientos, F.
-Batsch, J.
-Bauche, V.
-B.
-Olsen, M.
-Bernardini \textit{et al.}, “Acceleration of electrons in the plasma wakefield of a proton bunch,” Nature 561, 363–367 (2018), \url{https://doi.org/10.1038/s41586-018-0485-4)}.
-	%
-	%\bibitem{scheinker} A.
-Scheinker \textit{et al.}, "Online multi-objective particle accelerator optimization of the AWAKE electron beam line for simultaneous emittance and orbit control", AIP Advances 10, 055320 (2020), \url{https://doi.org/10.1063/5.0003423}.
-	%
-	%\bibitem{barto-sutton} R. Sutton, A. Barto, "Introduction to Reinforcement Learning",  MIT Press, Cambridge, MA, USA, 2018.
-	%
-	%\bibitem{LQR} H. Kwakernaak, R. Sivan, "Linear Optimal Control Systems", Wiley-Interscience, 1972.
-	%
-	%\bibitem{SBP} A. Nagabandi, G. Kahn, R. S. Fearing, S. Levine, "Neural Network DYNAmics for Model-Based Deep Reinforcement Learning with Model-Free Fine-Tuning", arXiv:1708.02596, 2017. 
-	%
-	%\bibitem{DYNA} R. Sutton, "DYNA, an integrated architecture for learning, planning, and reacting", AAAI Spring Symposium, pp. 151-155, 1991, \url{https://doi.org/10.1145/122344.122377}.
-	%
-	%\bibitem{benchmarkmbrl} T. Wang et al., “Benchmarking Model-Based Reinforcement Learning”,  \url{https://arxiv.org/abs/1907.02057v1}, 2019-07-03.
-	%
-	%\bibitem{dqn} V. Mnih \textit{et al.}, "Human-level control through deep reinforcement learning", Nature 518, 529–533 (2015).
-	%
-	%\bibitem{ddpg} T. Lillicrap~\textit{et al.}, "Continuous control with deep reinforcement learning", Proc. ICLR 2016 \url{https://arxiv.org/abs/1509.02971}.
-	%
-	%\bibitem{naf} S.
-Gu, T.
-Lillicrap, I.
-Sutskever, S.
-Levine, "Continuous Deep Q-Learning with Model-based Acceleration", in Proc.
-33rd International Conference on Machine Learning, New York, NY, USA, 2016.
-	%
-	%\bibitem{td3} S. Fujimoto, H. van Hoof, D. Meger, "Addressing Function Approximation Error in Actor-Critic Methods", arXiv:1802.09477, 2018.
-	%
-	%\bibitem{deeppilco} Y.
-Gal, R.
-McAllister, and C.
-E.
-Rasmussen, “Improving PILCO with Bayesian neural network dynamics models”, in Data-Efficient Machine Learning workshop, ICML, 2016, vol.
-4, p.
-34.
-	%
-	%\bibitem{per} T. Schaul, J. Quan, I. Antonoglou, D. Silver, "Prioritised Experience Replay", arXiv:1511.05952, 2015-11-18.
-	%
-	%\bibitem{pernafgit} S. Hirlaender, \textit{PER-NAF} available at \url{https://github.com/MathPhysSim/PER-NAF/} and
-	%\url{https://pypi.org/project/pernaf/}.
-	%\bibitem{pyjapc} "pyjapc" available at \url{https://pypi.org/project/pyjapc/}.
-	%
-	%\bibitem{openai_gym} \url{http://gym.openai.com}
-	%
-	%\bibitem{madx} MAD-X documentation and source code available at \url{https://mad.web.cern.ch/mad/}.
-	%
-	%\bibitem{linac4} G.
-Bellodi, "Linac4 Commissioning Status and Challenges to Nominal Operation", 61$^st$ ICFA Advanced Beam DYNAmics Workshop on High-Intensity and High-Brightness Hadron Beams, Daejeon, Korea, 17 - 22 Jun 2018, pp.MOA1PL03, \url{https://doi.org/10.18429/JACoW-HB2018-MOA1PL03}.
-	%
-	%\bibitem{auto_matching} F. Velotti, B. Goddard et al., "Automatic AWAKE electron beamline set-up using unsupervised machine learning", to be submitted to Phys. Rev. Accel. Beams.
-	%
-	%\bibitem{Kumar_1} A. Kumar, A. Gupta, S. Levine, “DisCor: Corrective Feedback in Reinforcement Learning via Distribution Correction,” arXiv:2003.07305, 2020-03-16.
-	%
-	%\bibitem{Kumar_2} A. Kumar, J. Fu, G. Tucker, S. Levine, “Stabilising Off-Policy Q-Learning via Bootstrapping Error Reduction,” arXiv:1906.00949, 2019-06-03.
-	%
-	%% NEW REFERENCES
-	%\bibitem[bruchon2020basic]{Bruchon, Niky, et al. "Basic Reinforcement Learning Techniques to Control the Intensity of a Seeded Free-Electron Laser." Electronics 9.5 (2020): 781.}
-	%% https://www.mdpi.com/2079-9292/9/5/781/htm
-	%
-	%
-	%
-	%\end{thebibliography}
-
 	%%%%% CLEAR DOUBLE PAGE!

### 2020-12-01 AEDYNA.py

@@ -30,7 +30,7 @@ steps_per_env = 201
 init_random_steps = 201
 # total_steps = 1207
 # num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
-num_epochs = 5
+num_epochs = 10
 print('Number of epochs: ', num_epochs)
 
 max_training_iterations = 50
@@ -448,7 +448,7 @@ class NetworkEnv(gym.Wrapper):
         self.verification = False
         if 'verification' in kwargs:
             self.verification = kwargs.get('verification')
-        self.visualize()
+        # self.visualize()
 
     def reset(self, **kwargs):
         self.current_model = np.random.randint(0, max(self.number_models, 1))
@@ -1084,7 +1084,7 @@ def aedyna(real_env, num_epochs=50, steps_per_env=100, algorithm='SAC',
                 f'data points: {len(model_buffer)}, ' + \
                 f'ep: {ep}, max: {data}\n' + hyp_str_all
         # TODO: Think about visualizations:
-        sim_env.visualize(data=data, label=label)
+        # sim_env.visualize(data=data, label=label)
 
         best_sim_test = -1e16 * np.ones(num_ensemble_models)
         agent = Agent(MlpPolicy, sim_env, verbose=1)

### 2020-12-01 read_paper_tests.py

@@ -8,7 +8,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 
 
-def read_experiment_observables(root_dir, key = 'tests_all'):
+def read_experiment_observables(root_dir, key='tests_all'):
     data_all = pd.DataFrame()
     n_average = 5
     for root, dirs, files in os.walk(root_dir):
@@ -22,20 +22,57 @@ def read_experiment_observables(root_dir, key = 'tests_all'):
             file_name = root + os.sep + file_name
             openfile = open(file_name, 'rb')
             if not key in ['step_counts_all']:
-                data = pd.DataFrame(pickle.load(openfile)[key]).sum()
-                # print(data)
+                object = pickle.load(openfile)
+                data = pd.DataFrame(object[key]).sum()
+                # index = object['step_counts_all']
+                # data.index = index
+
                 data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
-                # print(data_frame)
+                data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
             else:
                 data = pickle.load(openfile)[key]
                 print(root.split('_')[1])
-                data_frame = pd.DataFrame(data, columns=[root.split('_')[4]])
+                data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
 
-            data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
             data_all = pd.concat([data_all, data_frame.T], sort=False)
             openfile.close()
 
     return data_all
+
+
+def read_experiment_observables(root_dir, key='tests_all'):
+    data_all = pd.DataFrame()
+    n_average = 3
+    for root, dirs, files in os.walk(root_dir):
+        file_names = []
+        for file in files:
+            if 'observables' in file and not 'Noise-off' in root:
+                file_names.append(file)
+        if file_names:
+            file_names.sort()
+            file_name = file_names[-1]
+            file_name = root + os.sep + file_name
+            openfile = open(file_name, 'rb')
+            if not key in ['step_counts_all']:
+
+                object = pickle.load(openfile)
+                data = pd.DataFrame(object[key][0], columns=[root.split('_')[1]])
+                index = object['step_counts_all']
[…]

### 2020-12-01 run_aedyna_noise_test_pendulum.py

@@ -3,7 +3,7 @@ import shutil
 
 name = 'random_seeds'
 # root_dir = "DOUBLE/"
-root_dir = 'Data/Simulation/ModelSize/'
+root_dir = 'Data/Simulation/ModelSizeLong/'
 
 # if os.path.exists(root_dir):
 #     shutil.rmtree(root_dir)
@@ -19,7 +19,7 @@ for index in range(len(parameter_list)):
     name = 'Setting'
     name += '_' + str(parameter_list[index])
     print(name)
-    for seed in [10]:
+    for seed in [10, 20, 30, 40, 50]:
         command = "python AEDYNA.py " + name + ' ' + str(seed) + ' ' + root_dir + ' ' + str(index)
         os.system(command)
 

### 2020-12-01 main.tex
- paper changers

@@
 	\begin{abstract}
-		In this paper, we compare a model-based reinforcement learning approach in comparison to a model-free reinforcement learning approach applied at the FERMI FEL system.
-Both approaches hold tremendous promise and the main purpose of this paper is to show how reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems.
-In terms of sample-complexity, the model-based approach is faster, while asymptotic performance of the model-free method is sightly superior.
+		In this paper, we compare a model-based reinforcement learning to a model-free reinforcement learning method applied to the FERMI FEL system.
+Both approaches hold tremendous promise, and the principal goal of this paper is to show how reinforcement learning can be utilised on an operational level in a reasonable training time on real accelerator physics problems.
+In terms of sample-complexity, the model-based approach is better, while the asymptotic performance of the model-free method is slightly superior.
 The model-based algorithm is done in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
-In both cases existing methods were adapted to exhibit increased noise resistivity as incumbent in accelerator control problems.
+In both cases, the techniques were adapted to show increased noise resistivity as omnipresent in accelerator control problems.
 Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.
-		%The implementation is available at \cite{Hirlaender2020}.
 
@@
 		\item Already existing data can be used.
+		\item Adaption to non-stationary systems.
 		\item The underlying structure of the problem might be deduced.
@@
 	\begin{align}
-		\policy(\bs)^*=\arg\!\max J (\policy)  = 
+		J(\policy^*)=\max J (\policy)  = 
 		\mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
@@
 In the latter one deals with approximate dynamic programming, which we discuss in \cref{s:Model free reinforcement learning}.
-One advantage of these methods is, that they can use previously stored data as they are off-policy.
+One advantage of the latter approach is, that it can use previously stored data as it is an off-policy algorithm and generally needs less data to be trained.
 A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010}.
 	\subsection{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
-	We cannot provide an exhaustive treatment of state of the art MFRL algorithms, but we focus on the main principles.
+	We focus on the main principles of MBRL.
 An overview can be found e.g.
@@
 	\subsubsection{Approximate dynamic programming}
-	The state-value function $Q^\policy(\bs,\ba)$, which tells us how good an action $\ba$ in a state $\bs$ following $\policy$ in terms of the expected return is:
+	The state value function $V^\policy(\bs)$ tells us how good a state in terms of the expected return is following a specific $\policy$:
+		\begin{align}
+		V^\policy(\bs) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
+	\end{align}
+	The state-action-value function $Q^\policy(\bs,\ba)$, which tells us how good an action $\ba$ in a state $\bs$ following $\policy$ in terms of the expected return is:
 	\begin{align}
@@
 	\begin{align}
-		Q_\theta(\bs,\ba) = -\frac{1}{2}(\ba-\mu_\theta(\bs))P_\theta(\ba-\mu_\theta(\bs))^T+V_\theta(\bs).\label{eq:state-action-value-approxiation}
+		Q_\theta(\bs,\ba) = -\frac{1}{2}(\ba-\mu_\theta(\bs))P_\theta(\bs)(\ba-\mu_\theta(\bs))^T+V_\theta(\bs).\label{eq:state-action-value-approxiation}
 	\end{align}
+	$P_\theta(\bs)$ is a positive semi-infinite matrix.
 	One modification, which is used in these tests is the use of a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
@@
 
-	A nice overview of MBRL methods is provided in \cite{Wang2019}.
-Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and backpropagation in the dynamics model or \cite{Chua2018,Wang2019a}, a highly efficient sample based method.
+	A nice overview of model-based reinforcement learning (MBRL) methods is provided in \cite{Wang2019}.
+Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample based methods.
 We focus on model-based data generation.\\
-		In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$:
+		In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$.
+Additionally the reward is included in the model:
 	\begin{align}
-		\hat{f_\theta}(\bs_t,\ba_t) := \{\transitions_\theta(\bs_t,\ba_t),r_\theta(\bs_t,\ba_t)\}.
+		\hat{f_\theta}(\bs_t,\ba_t) := \{\hat\transitions_\theta(\bs_t,\ba_t),r_\theta(\bs_t,\ba_t)\}.
 	\end{align}
@@
 	An overview of the used method is shown in \cref{fig:MBRL_overview}.
-In the beginning, the data is collected or read in from a pre-collection.
+In the beginning, the data is collected or read from a previously collected data.
 An uncertainty aware model of the dynamics is trained, using anchored ensembles \cite{Pearce2018} on the data, which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
 It was motivated by the fact that pure epistemic ensemble techniques as \emph{ME-TRPO} were reported to be very sensitive to aleatoric noise \cite{Wang2019}.
-Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take the aleatoric and the epistemic uncertainties into account using probabilistic ensembles each model holding the parameters of the negative log likelihood of probability distribution function.\\
+Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take both uncertainties into account using probabilistic ensembles each model holding the parameters of the negative log likelihood of probability distribution function.
+Our implementation is simpler while showing similar performance.\\
 	 Subsequently, an MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data by taking the ensemble uncertainty into account, as explained in the next section.\\
-	  After a defined number of training steps, the controller is testes on each model individually.
-If there is no improvement in a number smaller than the total number of the models, the controller is tested on the real environment for several episodes.
+	  After a defined number of training steps, called epoch, the controller is validated on each model individually.
+If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes.
 In this way 'over-fitting' is avoided, as discussed in \cite{Kurutach2018}.
 The training is stopped if the controller solves the problem on the real environment.
-Otherwise, a batch of new data is collected.
+Otherwise, a batch of new data is collected and the model is improved and consecutively a new epoch starts.
 	\begin{figure}[!h]
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/MBRL_overview}
-		\caption{A schematic overview of the AE-DYNA approach used in this paper.}
+		\caption{A schematic overview of the \emph{AE-DYNA} approach used in this paper.}
 		\label{fig:MBRL_overview}
@@
 	\subsection{Critical design decisions}\label{ss:critical_design}
-	In the following the most important aspects of a successful application of model-based RL in accelerator problems are discussed, where we made some algorithmic design decisions which we think are beneficial for problems typically found in accelerator operation.\\
-%	 For the DYNA-AE algorithm  the important points are:
+In the following, the most important aspects of a successful application of model-based RL in accelerator problems are discussed, where we made some algorithmic design decisions which we think are beneficial for problems typically found in accelerator operation.\\
+%	 For the DYNA-AE algorithm, the important points are:
 %	\begin{itemize}
@@
 The idea is to mimic the posterior probability of the dynamics model to capture the uncertainty due to the lack of data.
-Empirical results showed that three to five models were satisfactory to see definite improvements over a single network approach, and the main purpose is not to determine the exact posterior of the model.
-A small two-layer network with around 15-25 nodes with a $tanh$ activation and a learning rate of $10^{-3}$ were used.
-The last layer was linear, and the inputs were normalised to the interval [-1,1].\\
-	 The prior is controlled by the initialisation of the weights of the network and defines how large the variation of the approximated function is assumed.
-The weights of the last layer were normalised by the number of nodes in each layer.
-The noise level is also respected during the initialisation.
-For the moment, only homoscedastic errors are used.\\
+Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes})).
+The main goal is not to determine the exact posterior probability of the model, but to provide the uncertainty to not over-train the controller at areas without sufficient data.\\ A small two-layer network with around 15-25 nodes with a $tanh$ activation and a learning rate of $10^{-3}$ were used.
+The last layer was linear, and the inputs $\{\bs,\ba\}$ were normalised to the interval [-1,1].\\
+The prior is controlled by the initialisation of the weights of the network.
+It defines how large the variation of the approximated function is expected.
+The weights of the last layer were normalised by the number of its nodes.
+The noise level is also respected during the initialisation and has played a crucial role as can be seen in \cref{fig:comparsion_noise_ae_dyna}.
+For the moment, only homoscedastic errors are considered.\\
 Several methods of network training were implemented.
-Early stopping was employed successfully with a waiting time of about 20 steps.
-Mainly due to the fact that the number of training steps is increasing with more data in close-by domains.
-The benefit is that the modelled uncertainty at regions where no data is available is reduced in this way, leading to better training performance.
-A fixed loss value threshold was tested, as also used.
+Early stopping was employed successfully with a waiting time of about 20 steps since the number of training steps is increasing with more data in close-by areas.
+This increase is followed by a shrinking of the uncertainty at regions where no data is available, leading to better training performance.
+A fixed maximal loss value threshold was also beneficially tested.
 In the experiments, a combination of both was taken.\\
-	One of the most crucial points is how these uncertainties are taken into account by the RL algorithms.
+One of the most crucial points is how model uncertainties are taken into account by the MFRL controller.
 Several different approaches were tried.
-It is possible to take the average of the models and add Gaussian noise levelled by the standard deviation of the models.
-Another straight forward way is to randomly select a model to provide at each training step, which is used the original implementation of the \emph{ME-TRPO} algorithm and was used in the experiments labelled \emph{ME-TRPO}.
-A pessimistic setting only would select the model resulting in the lowest predicted reward.
-Pessimistic selection showed constant but too slow improvement for our tests.
+One way is to take the average of the models and add Gaussian noise levelled by the standard deviation of the models.
+Another straight forward way is to randomly select a model to provide at each training step, which was showed the best performance the original implementation of the \emph{ME-TRPO} algorithm and was used in the experiments labelled \emph{ME-TRPO}.
+A pessimistic selection only would select the model resulting in the lowest predicted reward and showed constant but too slow improvement for our tests.
 \\ Good results were also obtained by following a randomly selected single model each full episode in combination with the soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}, and were used in the experiments labelled \emph{AE-DYNA SAC}.
 \\
-	The number of data-points taken every time the model as improved was firstly determined by the number of initial data points, collected using a random policy.
+The number of data-points in a batch taken every time of a new epoch has to be chosen carefully.
+It is influenced by the number of initial data points, collected using a random policy.
 The initialisation phase has to be chosen not to small to minimise the chance of getting trapped in a local minimum for too long.
-Afterwards, at least one full episode should be taken.
-We decided to use a short horizon to reduce the impact of the compound error, the accumulation error following a wrong model, as well known for DYNA-style approaches.
+Afterwards, from our experience, at least one full episode should be taken.
+We decided to use a short horizon to diminish the impact of the compound error, the accumulation error following a wrong model, as well known for DYNA-style approaches.
 The maximal number of steps was ten.
-During the data collected on the real system, the latest policy was taken with some small noise added to improve the exploration.\\
+During the data collected on the real system, the latest learned policy was taken with some small noise added to improve the exploration.\\
 To decide which MFRL algorithm to use, two main algorithm classes have to be considered: the on and the off-policy algorithms.
@@
 On the other hand, the on-policy algorithm TRPO provides some theoretical improvement guarantees.
-As reported in \cite{Kurutach2018} using proximal policy optimization \cite{Schulman2017} as an on-line MFRL algorithm does not show the same performance, which was confirmed in our tests.
+As reported in \cite{Kurutach2018} using proximal policy optimisation \cite{Schulman2017} as an online MFRL algorithm does not show the same performance, which was confirmed in our tests.
 In the latter case, the policy was improved over the whole training.
 For the \emph{SAC} the policy was reset after retraining the model, to profit from the exploration features.
-	\section{Experiment results from FERMI RL tests}
+	\section{Experimental results from FERMI RL online tests}
 	As discussed in the previous sections, several tests were performed on the FERMI FEL.
 	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility.
-A simplified simulation (details in \cref{a})was used to pre-tune the algorithms.
-The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, was tested first.
+The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
 	\subsection{MFRL tests}
-		\begin{figure}[!h]
+		\begin{figure}
 		\centering
@@
 
-	\begin{figure}[!h]
+	\begin{figure}
 		\centering
@@
 A motivation for the used algorithmic set-up is discussed in \cref{appendix:naf2} on a classical example from control theory.
-In both cases \emph{smoothing} (explanation in \cref{appendix:naf2}) was applied.\\
+In both cases \emph{smoothing} to increase the noise robustness as discussed in \cref{appendix:naf2}) was applied.\\
 	\Cref{fig:NAF_training} displays the results, averaged over the two tests.
@@
 In the upper figure, the number of iterations per episode is plotted, including the cumulative number of steps.
-	In the verification \cref{fig:NAF_verification} both algorithms show similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\
+The green line in the lower plot of \cref{fig:NAF_training} indicates the 95\% intensity reward threshold, to accept an episode to be successfully finished.
+	In the verification of 50 episodes (\cref{fig:NAF_verification}) both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance.
+\\
 	Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps.
@@
 The state-value function, which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value.
-In the single-network case, convergence is reached ~1400 steps.\\
+In the single-network case, convergence is reached ~1400 steps.
 	\subsection{MBRL tests}
-
-	\begin{figure}[!h]
+	\begin{figure}
 		\centering
@@
 	\end{figure}
-
-	\begin{figure}[!h]
+	\begin{figure}
 		\centering
@@
 	\end{figure}
-	\begin{figure}[!h]
+	\begin{figure}
 		\centering
@@
 The convergence property can be seen in \cref{fig:ME-TRPO_observables}.
-In the upper figure, the total reward per batch and the number of data points are shown as used in the training of the dynamics model.
+In the upper figure, the total reward per batch and the number of data points are shown as used to train the dynamics model against the number of epochs.
 In the lower plot, the average cumulative reward as achieved by the TRPO on the individual models in the ensemble on ten episodes as in dependence of the epochs is drawn.
@@
 As shown in \cref{fig:ME-TRPO_verification}, all of the 50 verification episodes were successfully finished after a few steps.
-To verify the impact of the ensemble technique, a test with a single network using the same hyperparameters has been applied, where no convergence within 500 data points has been observed.
+To verify the impact of the ensemble technique, a test with a single network using the same hyper-parameters has been applied, where no convergence within 500 data points has been observed.
 \\
@@
 In this test, the controller is reset each time, when new data for the dynamics model training is acquired.
-Consequently, the performance drops each time; the dynamics model was retrained, as shown in \cref{fig:AE-DYNA_observables}.
-Chances to get trapped in a local optimum are smaller by using this strategy of training.\\
-	In contrast to the first MBRL test, the new data batches are 50 with an initial random walk of 100 steps.
-The number of initial steps was chosen carefully high enough because the convergence is slowed down strongly, so that the training becomes unfeasible on a real machine.
-Each epoch consists of 2500 steps on the model.
+Consequently, the performance drops each time the dynamics model is retrained, as shown in \cref{fig:AE-DYNA_observables}.
+Chances to get trapped in a local optimum are littler by using this strategy of training.\\
+	In contrast to the first MBRL test, the new data batches consist of 50 data points with an initial random walk of 200 steps.
+The number of initial steps was chosen high enough because the convergence is slowed down strongly, so that the training becomes unfeasible on a real machine as discussed in \cref{ss:critical_design}.
+Each epoch consists of 2500 steps of controller training on the model.
 The training was stopped after the acquisition of 500 data points.
@@
 It might be a result of the higher number of data points (50), but in general, this method exhibited better asymptotic performance than the $ME-TRPO$ variant.
-
 	\section{Discussion and outlook}
 	In this paper, the applicability of the method of deep reinforcement learning on the optimization of the FERMI FEL intensity was presented.
-Two different approaches were tested: a model free and a model based method.
+Two different approaches were tested: a model free and a model based method.\\
 	 Both experiments yielded satisfactory results showing that this non-linear and noisy problem could be solved in a feasibly number of training steps.\\
+	Usage of the proposed methods in an operational way is attractive and would replace the current optimization method using the destructive measurement of the screens and additionally would save valuable time.\\
 	The model-free methods were slightly better in the final performance but also more sample were taken.
@@
 In case, the used methods could be parallelized to decrease the computational time.
+
 \section{Conclusions}
@@
  Generally control problems in accelerators have a short horizon, which makes the usage of \emph{DYNA-style} algorithms attractive as the \emph{Me-TRPO} or our  \emph{AE-DYNA} attractive choices for highly sample efficient algorithms.\\
+ Complementary, the \emph{NAF} algorithms provides a good alternative, revealing as a model-free method good asymptotic performance.
+As most cases in accelerator control could be covered using the assumption of the quadratic dependence of the actions of the state-action value function, there is a wide range of possible applications.\\
 
  To provide the possibility for other labs to use the proposed methods the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020}.
+ \vfill
+ \newpage
+ \vfill
   \newpage
@@
 The maximum cumulative reward averaged over five different runs tested on the real environment during the training is visualized in dependence of the number of data points.
-A number of three models shows a good trade-off between performance and
+A number of three models (label \emph{Three}) shows a good trade-off between performance and training time.
+A single network might not converge or too slowly (label \emph{Single}) and a model of ten shown fast and stable performance in this case (label \emph{Ten}).
 \begin{figure}[!h]
@@
 	\begin{itemize}
-	\item mention the stable baseline.
+	\item reference clean up.
 %	\item Noise benchmarks with pendulum - done

### 2020-12-01 read_paper_tests.py

@@ -42,7 +42,7 @@ def read_experiment_observables(root_dir, key='tests_all'):
 
 def read_experiment_observables(root_dir, key='tests_all'):
     data_all = pd.DataFrame()
-    n_average = 3
+    n_average = 5
     for root, dirs, files in os.walk(root_dir):
         file_names = []
         for file in files:
@@ -53,7 +53,7 @@ def read_experiment_observables(root_dir, key='tests_all'):
             file_name = file_names[-1]
             file_name = root + os.sep + file_name
             openfile = open(file_name, 'rb')
-            if not key in ['step_counts_all']:
+            if not key in ['step_counts_all'] and not root.split('_')[1] == 'Five':
 
                 object = pickle.load(openfile)
                 data = pd.DataFrame(object[key][0], columns=[root.split('_')[1]])
@@ -61,14 +61,13 @@ def read_experiment_observables(root_dir, key='tests_all'):
                 data.index = index
                 data = data.groupby(data.index).max()
                 # print(data)
-                data = data.rolling(n_average).mean().shift(int(n_average / 2))
-            else:
-                data = pickle.load(openfile)[key]
-                print(root.split('_')[1])
-                data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
+            # else:
+            #     data = pickle.load(openfile)[key]
+            #     print(root.split('_')[1])
+            #     data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
 
-            data_all = pd.concat([data_all, data], axis=1, sort=False)
-            print(data_all)
+                data_all = pd.concat([data_all, data], axis=1, sort=False)
+
             openfile.close()
 
     return data_all
@@ -153,10 +152,13 @@ def read_experiment_final(root_dir):
 # plt.show()
 
 root_dir = 'Data/Simulation/ModelSizeLong/'
-
+n_average = 30
 figure_name = 'Figures/Compare_models_sizes.pdf'
-data_all = read_experiment_observables(root_dir=root_dir).T
-print(data_all)
+data_all = read_experiment_observables(root_dir=root_dir)
+data_all = data_all.reindex(np.linspace(201, 2010, num=151)).interpolate('values')
+
+data_all = data_all.rolling(n_average).mean().shift(int(n_average / 2)).T
+print(data_all.T)
 fig, axs = plt.subplots(1, 1)
 
 sns.set_palette(sns.color_palette("bright", 8))

### 2020-12-02 main.tex
- paper changers

@@
 Especially if there is no way to model the physics a priori, one might use optimisation techniques as, e.g.
-derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015} or model-based optimisations as Gaussian processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.
+derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015,Albright2019} or model-based optimisations as Gaussian processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.
 Another promising way is to apply reinforcement learning (RL) \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020} which unveils several advantages over optimisation methods:
@@
 		\item Already existing data can be used.
-		\item Adaption to non-stationary systems.
+		\item Adaption to non-stationary systems is possible.
 		\item The underlying structure of the problem might be deduced.
@@
 	The external optical laser signal provides is a contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy.
-The modulation in energy is converted into a charge modulation in the dispersive section, and finally, the density modulated beams radiation is amplified in the section of the radiator.
+The modulation in energy is converted into a charge modulation in the dispersive section, and finally, the density modulated beams radiation is amplified in the section of the radiator.\\
 	 The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.\\
@@
 		\begin{align}
-		V^\policy(\bs) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
+		V^\policy(\bs_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
 	\end{align}
-	The state-action-value function $Q^\policy(\bs,\ba)$, which tells us how good an action $\ba$ in a state $\bs$ following $\policy$ in terms of the expected return is:
+	The state-action-value function $Q^\policy(\bs,\ba)$, which tells us how good an action $\ba$ in a state $\bs$ following $\policy$ is, in terms of the expected return:
 	\begin{align}
@@
 It is motivated by double Q-learning \cite{NIPS2010_091d584f,Hasselt2015,fujimoto2018addressing}.
-	The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality} the loss $L$ can be formulated as:
+	The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality} the loss $\mathcal L$ can be formulated as:
 	\begin{align}
-		&L(\theta)=\\\notag&\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
+		&\mathcal L(\theta)=\\\notag&\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
 	\end{align}
@@
 To stabilize the network training a small clipped artificial noise is added to the actions called \emph{smoothing} as described in \cref{appendix:naf2}.\\
-	 This implementation has an extremely good sample-efficiency for suitable problems as can be found often in accelerators.
-It is used as the baseline for the considered control problem, as it achieves good results.
+	 This implementation has an extremely good sample-efficiency for suitable problems as can be found often in accelerators.We use it as the baseline for the FERMI-FEL control problem, as it achieves good results.
 An illustration and additional changes to the original proposal \cite{Gu2016} and a previous implementation used for experiments using a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
@@
 		In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$.
-Additionally the reward is included in the model:
+Additionally in our case the reward is included in the model \footnote{In some applications the reward can be deduced from the current state or the state change.
+In such a situation the modelling can be simplified to model directly $\hat\transitions$}:
 	\begin{align}
-		\hat{f_\theta}(\bs_t,\ba_t) := \{\hat\transitions_\theta(\bs_t,\ba_t),r_\theta(\bs_t,\ba_t)\}.
+		\hat{f_\theta}(\bs_t,\ba_t) := \{\hat\transitions_\theta(\bs_t,\ba_t),r_\theta(\bs_t,\ba_t)\}.\label{eq:dynamics_model}
 	\end{align}
@@
 %	\end{itemize}
-The anchoring ensemble methods usually yield good results already with a small number of different networks.
+The anchor ensembling method, as treated in \cref{appendix:The impact of noise}, usually yields good results already with a small number of different networks.
 The idea is to mimic the posterior probability of the dynamics model to capture the uncertainty due to the lack of data.
@@
 The weights of the last layer were normalised by the number of its nodes.
-The noise level is also respected during the initialisation and has played a crucial role as can be seen in \cref{fig:comparsion_noise_ae_dyna}.
-For the moment, only homoscedastic errors are considered.\\
+For the moment, only homoscedastic Gaussian errors are considered with a standard deviation is $\sigma_\epsilon$ is also respected during the initialisation \cref{appendix:The impact of noise} and improves the training in the presence of noise as can be seen in \cref{fig:comparsion_noise_ae_dyna}.
+\\
 Several methods of network training were implemented.
@@
 We decided to use a short horizon to diminish the impact of the compound error, the accumulation error following a wrong model, as well known for DYNA-style approaches.
-The maximal number of steps was ten.
+The maximal number of steps in our experimental runs was ten.
 During the data collected on the real system, the latest learned policy was taken with some small noise added to improve the exploration.\\
@@
 In this way, exploration is encouraged to avoid getting stuck in a local optimum and has a good trade-off between exploration and exploitation.\\
-On the other hand, the on-policy algorithm TRPO provides some theoretical improvement guarantees.
+On the other hand, the on-policy algorithm TRPO \cite{Schulman2015} provides some theoretical improvement guarantees.
 As reported in \cite{Kurutach2018} using proximal policy optimisation \cite{Schulman2017} as an online MFRL algorithm does not show the same performance, which was confirmed in our tests.
@@
 One purpose of this work was to provide some suggestions how to make advanced deep reinforcement learning techniques work on a real set-up by adjusting available methods.\\
- Generally control problems in accelerators have a short horizon, which makes the usage of \emph{DYNA-style} algorithms attractive as the \emph{Me-TRPO} or our  \emph{AE-DYNA} attractive choices for highly sample efficient algorithms.\\
+ Generally control problems in accelerators have a short horizon, which makes the usage of \emph{DYNA-style} algorithms attractive as the \emph{ME-TRPO} or our  \emph{AE-DYNA} attractive choices for highly sample efficient algorithms.\\
  Complementary, the \emph{NAF} algorithms provides a good alternative, revealing as a model-free method good asymptotic performance.
@@
 \end{figure}
-	\subsection{The impact of noise}
+
+%Note a subscript has been introduced, $j \in \{1 ... M\}$, with the view of an ensemble of $M$ ANNs, each with a distinct draw of ${\theta}_{anc}$.
+
+% \tp{possible to just state equations you know, save space}
+	\subsection{The impact of noise}\label{appendix:The impact of noise}
 	A test adding artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma)$ with $\sigma=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}.
@@
 	\end{figure}
-	\\Using the anchors (the noise level is homeoscetastic) and known in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}.
-It the upper plot the mean cumulative reward on 10 test episodes on the real environment is shown during the training.
+\subsubsection{Regression assuming homoskedastic Gaussian noise using `Anchored ensembling'}
+The Bayesian community and the RL community is developping
+new approaches to Bayesian inference. It is done by adding a regularisation
+term to a loss function. It returns a a point estimate of
+the Bayesian posterior the maximum a posteriori (MAP) parameter estimate. 
+Injecting noise into this loss, either to targets or regularisation term, and sampling
+repeatedly (i.e. ensembling), produces a distribution of MAP solutions mimicking that of the true posterior \cite{Gu2007, Chen2011, Bardsley2012, Pearce2018}.\\
+Consider an artificial neural net (ANN) containing parameters, ${\theta}$, making predictions for the dynamics and the reward \cref{eq:dynamics_model}, $\hat{f}_\theta$, with $N$ data points.
+If the prior is given by $P(\theta) = \mathcal{N}(\pmb{\mu}_{prior}, \pmb{\Sigma}_{prior})$, maximising the following returns MAP parameter estimates (details see \cite{Pearce2018}):
+\begin{equation}
+	\label{eq_MAP_loglike_anc}
+	{\theta}_{MAP}  = \text{argmax}_{{\theta}} \log( P_{\mathcal{D}}( \mathcal{D} | {\theta} ) ) - 
+	\frac{1}{2} 
+	\lVert \pmb{\Sigma}_{prior}^{-1/2} \cdot
+	({\theta} - \pmb{\mu}_{prior}) \rVert^2_2
+\end{equation}
+We replace $\pmb{\mu}_{prior}$ with some random variable $ {\theta}_{anc}$.
+%\begin{equation}
+%\label{eq_MAP_loglike_anc}
+%\pmb{f}_\text{MAP}(\pmb{\theta}_{anc})  = \text{argmax}_{\pmb{\theta}} \log( P_{\mathcal{D}}( \mathcal{D} | \pmb{\theta} ) ) - 
+% \frac{1}{2} 
+% \lVert \pmb{\Sigma}_{prior}^{-1/2} \cdot
+% (\pmb{\theta} - \pmb{\theta}_{anc}) \rVert^2_2
+%\end{equation}
+${\theta}_{anc} \sim \mathcal{N}(\pmb{\mu}_{prior},\pmb{\Sigma}_{prior} )$, to make it practical and
+assuming homoskedastic Gaussian noise of variance $\sigma^2_\epsilon$, we can give parametric form of data likelihood.
+We take $M$ models, where each model is indexed by$j \in \{1 ...
+M\}$.
+The MAP estimates are found by minimising ($\mathcal L_{j}$ denotes the loss of the $j^\text{th}$ model),
+\begin{equation}
+	\label{eqn_anch_loss_matrix}
+	\mathcal L_{j} =  
+	\frac{1}{N} \lvert \lvert f - \hat{f}_{\theta, j} \rvert \rvert ^2_2
+	+ \frac{1}{N} \lvert \lvert \pmb{\Gamma}^{1/2} \cdot (\theta_j - \theta_{anc,j}) \rvert \rvert ^2_2.
+\end{equation}
+We have defined a diagonal regularisation matrix, $\pmb{\Gamma}$, where the $i^{th}$ diagonal element is the ratio of data noise of the target variable to prior variance for parameter $\theta_i$:
+\begin{equation}
+	\label{eqn_anch_loss_init}
+	\text{diag}(\pmb{\Gamma})_i = \frac{\sigma^2_\epsilon}{\sigma^2_{prior_i}},
+\end{equation}
+	Using the anchors and $\sigma_\epsilon=\sigma$ in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}.
+In the upper plot the mean cumulative reward on 10 test episodes on the real environment is shown during the training.
 It should indicate the result if the training is stopped at this training epoch.
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/Compare_models_sizes}
-	\caption{Varying the number of models in the ensemble of the \emph{AE-DYNA} on the \emph{inverted pendulum}.}
+	\caption{Varying number of models in the ensemble of the \emph{AE-DYNA} on the \emph{inverted pendulum}.}
 	\label{fig:Compare_models_sizes}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_ae_dyna}
-		\caption{The comparison of the NAF2 and the \emph{AE-DYNA} on the noisy \emph{inverted pendulum}.}
+		\caption{The comparison of the \emph{NAF2} and the \emph{AE-DYNA} on the noisy \emph{inverted pendulum}.}
 		\label{fig:comparsion_NAF_AE-DYNA}
@@
 	\item reference clean up.
+	\item Conclusion and outlook\dots
 %	\item Noise benchmarks with pendulum - done

### 2020-12-02 AEDYNA.py

@@ -9,26 +9,24 @@ import numpy as np
 import tensorflow as tf
 from inverted_pendulum import PendulumEnv
 
-# from stable_baselines.common.noise import NormalActionNoise
-
-# from naf2 import NAF
+'''This script includes the AE-DYNA algorithm. '''
 
 # set random seed
 random_seed = 111
 np.random.seed(random_seed)
-config = tf.ConfigProto(
-    device_count={'GPU': 0}
-)
-config = None
+
+# config = tf.ConfigProto(
+#     device_count={'GPU': 0}
+# )
+# config = None
 
 ############################################################
 # Hyperparameters
 ############################################################
 
 
-steps_per_env = 201
+steps_per_epoch = 201
 init_random_steps = 201
-# total_steps = 1207
 # num_epochs = int((total_steps - init_random_steps) / steps_per_env) + 1
 num_epochs = 10
 print('Number of epochs: ', num_epochs)
@@ -36,27 +34,13 @@ print('Number of epochs: ', num_epochs)
 max_training_iterations = 50
 delay_before_convergence_check = 5
 
-# Select the rl-algorithm
-
-algorithm = 'SAC'
-# algorithm = 'PPO'
-
-if algorithm == 'SAC':
-    from stable_baselines.sac.policies import MlpPolicy
-    from stable_baselines import SAC as Agent
-else:
-    from stable_baselines.common.policies import MlpPolicy
-    from stable_baselines import PPO2 as Agent
-
 simulated_steps = 2500
 
 model_batch_size = 100
 num_ensemble_models = 3
 
 early_stopping = True
-model_iter = 10
-
-# model_training_iterations = 10
+model_iter = 30
 
 # How often to check the progress of the network training
 # e.g. lambda it, episode: (it + 1) % max(3, (ep+1)*2) == 0
@@ -67,7 +51,7 @@ lr_start = 1e-3
 lr_end = 1e-3
 lr = lambda ep: max(lr_start + ep / 30 * (lr_end - lr_start), lr_end)
 
-# Set max episode length manually for the pendulum
+# Set max episode length manually here for the pendulum
 max_steps = 200
 
 # Class for data storage during the tests
@@ -622,12 +606,11 @@ def restore_model(old_model_variables, m_variables):
 
 def aedyna(real_env, num_epochs=50, steps_per_env=100, algorithm='SAC',
            simulated_steps=1000, num_ensemble_models=2, model_iter=15, model_batch_size=512,
[…]

### 2020-12-02 main.tex
- paper changes

@@
 \subsubsection{Regression assuming homoskedastic Gaussian noise using `Anchored ensembling'}
-The Bayesian community and the RL community is developping
-new approaches to Bayesian inference. It is done by adding a regularisation
-term to a loss function. It returns a a point estimate of
+The Bayesian community and the RL community is spending more and more attention to
+new approaches to Bayesian inference. A regularisation
+term is added to a loss function, which returns a point estimate of
 the Bayesian posterior the maximum a posteriori (MAP) parameter estimate. 
@@
 ${\theta}_{anc} \sim \mathcal{N}(\pmb{\mu}_{prior},\pmb{\Sigma}_{prior} )$, to make it practical and
-assuming homoskedastic Gaussian noise of variance $\sigma^2_\epsilon$, we can give parametric form of data likelihood.
+assuming homoskedastic Gaussian noise of variance $\sigma^2_\epsilon$, we obtain a parametric form of the data likelihood.
 We take $M$ models, where each model is indexed by$j \in \{1 ...
@@
 	\centering
-	\includegraphics*[width=0.5\textwidth]{Figures/Compare_models_sizes}
+	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_models_sizes}
 	\caption{Varying number of models in the ensemble of the \emph{AE-DYNA} on the \emph{inverted pendulum}.}
@@
 		\centering
-		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_ae_dyna}
+		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_AE-DYNA}
 		\caption{The comparison of the \emph{NAF2} and the \emph{AE-DYNA} on the noisy \emph{inverted pendulum}.}
@@
 	\begin{itemize}
-	\item reference clean up.
+	\item Reference clean up.
+	\item Rethink the concept of an epoch.
 	\item Conclusion and outlook\dots

### 2020-12-02 compare_naf_ae_dyna.py
- paper changes

@@ -1,0 +1,93 @@
+import os
+import sys
+import pickle
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+# Read NAF data
+# def read_experiment(root_dir):
+#     data_all = pd.DataFrame()
+#     n_average = 15
+#     for root, dirs, files in os.walk(root_dir):
+#         for file in files:
+#             if file == 'plot_data_0.pkl':
+#                 file_name = root + os.sep + file
+#                 print(file_name)
+#                 openfile = open(file_name, 'rb')
+#                 data = np.sum(pickle.load(openfile), axis=1)
+#                 # data_frame = pd.DataFrame(data, columns=[root.split('/')[0]])
+#                 data_frame = pd.DataFrame(data, columns=[root.split('_')[1]])
+#                 # data_frame = data_frame.rolling(n_average).mean().shift(int(n_average / 2))
+#                 data_all = pd.concat([data_all, data_frame.T], sort=False)
+#                 openfile.close()
+#
+#     return data_all
+#
+# root_dir = "Experiments-150-200-50-noisy/"
+# data_all_naf = read_experiment(root_dir=root_dir)
+#
+# data_all_naf.to_csv('NAF.csv')
+
+# def read_experiment_observables(root_dir, key = 'tests_all'):
+#     data_all = pd.DataFrame()
+#     n_average = 5
+#     for root, dirs, files in os.walk(root_dir):
+#         file_names = []
+#         for file in files:
+#             if 'observables' in file and not 'Noise-off' in root:
+#                 file_names.append(file)
+#         if file_names:
+#             file_names.sort()
+#             file_name = file_names[-1]
+#             file_name = root + os.sep + file_name
+#             openfile = open(file_name, 'rb')
+#             if not key in ['step_counts_all']:
+#                 data = pd.DataFrame(pickle.load(openfile)[key]).sum()
+#                 print(data)
+#                 data_frame = pd.DataFrame(data, columns=[root.split('_')[4]])
+#                 print(data_frame)
+#             else:
+#                 data = pickle.load(openfile)[key]
[…]

### 2020-12-02 run_naf2_for_tests.py

@@ -1,0 +1,290 @@
+import os
+import pickle
+import random
+import sys
+
+import gym
+import matplotlib.pyplot as plt
+import numpy as np
+import tensorflow as tf
+
+from inverted_pendulum import PendulumEnv
+
+from naf2_new import NAF
+
+# set random seed
+random_seed = 111
+np.random.seed(random_seed)
+random.seed(random_seed)
+
+
+def plot_results(env, file_name):
+    # plotting
+    print('Now plotting')
+    rewards = env.rewards
+    initial_rewards = env.init_rewards
+    # print('initial_rewards :', initial_rewards)
+
+    iterations = []
+    final_rews = []
+    starts = []
+    sum_rews = []
+    mean_rews = []
+
+    for i in range(len(rewards)):
+        if (len(rewards[i]) > 0):
+            final_rews.append(rewards[i][len(rewards[i]) - 1])
+
+            iterations.append(len(rewards[i]))
+            sum_rews.append(np.sum(rewards[i]))
+            mean_rews.append(np.mean(rewards[i]))
+
+            try:
+                starts.append(initial_rewards[i])
+            except:
+                pass
+    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
+
+    fig, axs = plt.subplots(2, 1)
+
+    ax = axs[0]
+    color = 'blue'
+    ax.plot(iterations, c=color)
+    ax.set_ylabel('steps', color=color)
+    ax.tick_params(axis='y', labelcolor=color)
+    ax1 = plt.twinx(ax)
+    color = 'k'
+    ax1.plot(np.cumsum(iterations), c=color)
+    ax1.set_ylabel('cumulative steps', color=color)
+    ax.set_title('Iterations' + plot_suffix)
+    # fig.suptitle(label, fontsize=12)
+
+    ax = axs[1]
+    color = 'red'
+    ax.plot(starts, c=color)
+    ax.set_ylabel('starts', color=color)
+    ax.axhline(-0.05, ls=':', color='r')
+    ax.tick_params(axis='y', labelcolor=color)
+    ax.set_title('Final reward per episode')  # + plot_suffix)
+    ax.set_xlabel('# episode')
+
+    ax1 = plt.twinx(ax)
+    color = 'lime'
+    ax1.set_ylabel('finals', color=color)
[…]

### 2020-12-03 main.tex
- paper changes

@@
 		\begin{align}
-		V^\policy(\bs_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
+		V^\policy(\bs_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],\label{eq:state-value-function}
 	\end{align}
@@
 For the \emph{SAC} the policy was reset after retraining the model, to profit from the exploration features.
-	\section{Experimental results from FERMI RL online tests}
+
+	\subsection{Comparison of the used algorithms}
+	\Cref{tab:overview_algorithms} shows a comparison of the discussed algorithms.
+Details are provided in the appendix.
+All algorithms are suitable for many control tasks as common in accelerator operation.
+Their strength lies in the high sample efficiency, but they differ in representational power and noise robustness.
+\\
+	In \cref{sec:Experimental results from FERMI RL online tests} the \emph{NAF2},
+	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment.
+		\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
+	\caption{\label{tab:overview_algorithms}%
+		An overview of the algorithms discussed in this paper.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lllll}
+			\textrm{Algorithm}&
+			\textrm{Class}&
+			\textrm{NR\footnote{Noise robustness}}&
+			\textrm{RP\footnote{Representational power}}&
+			\textrm{SE\footnote{Sample efficiency}}\\
+			\colrule
+			\emph{NAF} & model-free & low &low &high\\
+			\emph{NAF2} & model-free & high &low& high\\
+			\emph{ME-TRPO} & model-based & low& high& high\\
+			\emph{AE-DYNA} & model-based & high& high& high\\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
+	\section{Experimental results from FERMI RL online tests}\label{sec:Experimental results from FERMI RL online tests}
 	As discussed in the previous sections, several tests were performed on the FERMI FEL.
-	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility.
+	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility.\\ 
+	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments.
+Since there was a positive outcome, it is evident, that the utilised methods are at an operational level.\\
 	The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
-		\caption{The training of the \emph{NAF2} on the FERMI FEL.}
+		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
 		\label{fig:NAF_training}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
-		\caption{The verification of the \emph{NAF2} on the FERMI FEL.}
+		\caption{The verification of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
 		\label{fig:NAF_verification}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
-		\caption{The verification of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
+		\caption{The training metrics of the \emph{AE-DYNA-SAC} on the FERMI FEL using a single network (dashed) and a double network (solid).
+The Bellman error (\cref{eq:minimize_bellmann_optimality}) and the state-value function (\cref{eq:state-value-function}) are shown.
+}
 		\label{fig:NAF_convergence}
@@
 	Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps.
-The blue curves show the Bellmann error, which is comparable in both cases.
-The state-value function, which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value.
-In the single-network case, convergence is reached ~1400 steps.
+The blue curves show the Bellmann error \cref{eq:minimize_bellmann_optimality}, which is comparable in both cases.
+The state-value function $V$ (\cref{eq:state-value-function}), which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value.
+In the single-network case, convergence is reached in ~1400 steps.
+$V$ was averaged over 200 states.
 	\subsection{MBRL tests}
@@
 	\item Reference clean up.
-	\item Rethink the concept of an epoch.
 	\item Conclusion and outlook\dots
+	\item Clean discussion of the results. What is superior and why? A table with numbers...
 %	\item Noise benchmarks with pendulum - done

### 2020-12-03 main.tex
- paper changes

@@
 
-	\title{Online Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Applications to FERMI FEL}
+	\title{Online Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
 
@@
 Both approaches hold tremendous promise, and the principal goal of this paper is to show how reinforcement learning can be utilised on an operational level in a reasonable training time on real accelerator physics problems.
-In terms of sample-complexity, the model-based approach is better, while the asymptotic performance of the model-free method is slightly superior.
-The model-based algorithm is done in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
+In terms of representational power and sample-efficiency, the model-based approach is better, while the asymptotic performance of the model-free method is slightly superior.
+The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
 In both cases, the techniques were adapted to show increased noise resistivity as omnipresent in accelerator control problems.
@@
 	In particle accelerator operation, one main goal is to provide stable and reproducible performance.
-In order to achieve this, several control problems have to be considered simultaneously \cite{Kain2020}.
+Achieving this demands to consider several control problems simultaneously \cite{Kain2020}.
 Especially if there is no way to model the physics a priori, one might use optimisation techniques as, e.g.
-derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015,Albright2019} or model-based optimisations as Gaussian processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.
-Another promising way is to apply reinforcement learning (RL) \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020} which unveils several advantages over optimisation methods:
+derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015,Albright2019} or model-based optimisations as Gaussian processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.\\
+	 Recently, the community has begun to explore a novel approach to solve control problems.
+Reinforcement learning (RL) \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020} unveils several advantages over optimisation methods:
 	\begin{itemize}
@@
 	\end{itemize}
-	One demanding aspect using RL is the number of iterations needed to train a controller, since RL methods are known to be data-hungry \cite{Sutton2018}, and a second critical aspect is the robustness of the training in real world problems itself.\\
+	One demanding aspect using RL is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018}, and a second critical aspect is the robustness of the training in real world problems itself.\\
 	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.\\
-	 Two different algorithm classes were applied successfully to the FERMI FEL problem and reveal different advantages.
-The critical aspects are addressed, and tailored solutions, generally applicable to problems showing a sample complexity as in accelerator control problems are presented.
+	 Two different algorithm classes (\emph{NAF} and \emph{AE-DYNA}) were applied successfully to the FERMI FEL problem and reveal different advantages.
+The critical aspects are addressed, and tailored solutions, generally applicable to problems showing a sample-complexity as in accelerator control problems are presented.
 	The paper is organised as follows:
@@
 		\item Design decisions of the implementations used in these studies and theoretical concerns
-		\item Results 
+		\item Results of the experiments 
 		\item Summary and outlook
@@
 
-	\section{The physical set-up of the studied problem}
-	\begin{figure}[!h]
-		\centering
-		\input{Figures/SL_Alignment_Scheme}
-		\caption{A schematic view on the set-up of the FERMI FEL.}
-		\label{fig:schematic_FEL}
-	\end{figure}
+	\section{The set-up of the studied problem}
+	\subsection{The physical set-up}
 	In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator.\\
@@
 	For this reason, the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
-
 	\subsection{The training environment}
-
+		\begin{figure}
+		\centering
+		\input{Figures/SL_Alignment_Scheme}
+		\caption{A schematic view on the set-up of the FERMI FEL.}
+		\label{fig:schematic_FEL}
+	\end{figure}
 %	The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
@@
 Two values correspond to the two DOF of each mirror.
-The action $\ba$ is four dimensional, is a delta step on the voltages at step $t$ $\bs_t$:
+The action $\ba$ is four dimensional, namely a delta step on the voltages at time-step $t$ $\bs_t$:
 	\begin{align}
@@
 	\end{align} 
-	The training is done in episodes with a uniformly randomize initial state and a specific maximal length, called the horizon.
-For reason as discussed later, the horizon varies in the applications between either 10 or 50 steps.
-An early termination, when a specific threshold is obtained, lies in all cases at around 95\% of the maximal achievable intensity of the system.
-During the training phase exploration and the training of the function approximator takes place.\\
+	The training is done in episodes with a uniformly\footnote{in the normalized state-space} randomized initial state and a specific maximal length of taken steps afterwards - the horizon.
+For reason discussed later, the horizon varies in the applications between either 10 or 50 steps.
+An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system.\\
+	 During the training phase exploration and the training of the function approximator takes place.
 	After a defined number of training episodes, a verification of 50 episodes is performed. In the second phase, the RL only applies the learned policy.
@@
 	\end{align}
-	where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
-	In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
+	where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.\\
+	In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba) \approx Q_\theta(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
 a deep neural network.\\
 	 In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{Sutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}.
-To train these methods, a large number of direct interactions with the system is required, because they are on-policy.
-In the latter one deals with approximate dynamic programming, which we discuss in \cref{s:Model free reinforcement learning}.
+To train these methods, a large number of direct interactions with the system is required, because they belong to on-policy algorithms.
+In the latter one deals with approximate dynamic programming, which we shortly treat in \cref{s:Model free reinforcement learning}.
 One advantage of the latter approach is, that it can use previously stored data as it is an off-policy algorithm and generally needs less data to be trained.
@@
 in \cite{Sutton2018,Levine2020}.
-In the following, we discuss the normalised advantage function algorithm in some detail, which has good characteristics for accelerator control problems as shown in \cite{Kain2020,Hirlaender2020a}.
+In the following, we discuss the normalised advantage function (\emph{NAF}) algorithm in some detail, as it has good characteristics for accelerator control problems \cite{Kain2020,Hirlaender2020a}.
 	\subsubsection{Approximate dynamic programming}
@@
 	\end{align}
-	And $\policy$ can be calculated via:
+	$\policy$ can be calculated via:
 	\begin{align}
@@
 	\end{multline}
-	The form of this equation can cause overestimation and other complications, when using a function approximator and several attempts exist to reduce the problems \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
+	The nature of this equation can cause overestimation and other complications, when using a function approximator and several attempts exist to overcome the difficulties \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
 	\subsubsection{Normalized advantage function}
-	One way to avoid the sensitivity to the mentioned complications is to take a simple analytical form with an explicitly calculable maximum of the $Q$-function.
+	One way to avoid the sensitivity to the mentioned complications is to choose a simple analytical form with an explicitly calculable maximum of the $Q$-function.
 	If a specific quadratic form of the $Q$ function is assumed \cite{Gu2016}:
@@
 	\end{align}
-	$P_\theta(\bs)$ is a positive semi-infinite matrix.
-	One modification, which is used in these tests is the use of a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
+	$P_\theta(\bs)$ is a state-dependent, positive-definite
+	square matrix, parametrised in a specific way \cite{Gu2016}.
+One modification, which we made in our experiments is the application of a twin network (weights for network $i$ denoted by $\theta^i$).
+Only one network is used to obtain the policy, while the other is employed for the update rule to avoid over-estimation.
 It is motivated by double Q-learning \cite{NIPS2010_091d584f,Hasselt2015,fujimoto2018addressing}.
@@
 To stabilize the network training a small clipped artificial noise is added to the actions called \emph{smoothing} as described in \cref{appendix:naf2}.\\
-	 This implementation has an extremely good sample-efficiency for suitable problems as can be found often in accelerators.We use it as the baseline for the FERMI-FEL control problem, as it achieves good results.
-An illustration and additional changes to the original proposal \cite{Gu2016} and a previous implementation used for experiments using a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
+	 This implementation has an extremely good sample-efficiency for suitable problems as can be found often in accelerators.
+We use it as the baseline for the FERMI-FEL control problem, as it yields good results.
+Additional changes to the original proposal \cite{Gu2016} and a previous implementation used for experiments using a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
 	%\subsection{New NAF network architecture}
@@
 		In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$.
-Additionally in our case the reward is included in the model \footnote{In some applications the reward can be deduced from the current state or the state change.
+Additionally in our case the reward is included in the model \footnote{In some applications $\reward$ can be deduced from the current state or the state change.
 In such a situation the modelling can be simplified to model directly $\hat\transitions$}:
@@
 	An overview of the used method is shown in \cref{fig:MBRL_overview}.
-In the beginning, the data is collected or read from a previously collected data.
-An uncertainty aware model of the dynamics is trained, using anchored ensembles \cite{Pearce2018} on the data, which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
+In the beginning, the data is collected or read from previously collected data.
+An uncertainty aware model of the dynamics is trained, using 'anchored ensembling' \cite{Pearce2018} on the data, which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
 It was motivated by the fact that pure epistemic ensemble techniques as \emph{ME-TRPO} were reported to be very sensitive to aleatoric noise \cite{Wang2019}.
-Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take both uncertainties into account using probabilistic ensembles each model holding the parameters of the negative log likelihood of probability distribution function.
+Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take these uncertainties into account using probabilistic ensembles.
 Our implementation is simpler while showing similar performance.\\
-	 Subsequently, an MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data by taking the ensemble uncertainty into account, as explained in the next section.\\
+	 Subsequently, an MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data by taking the model uncertainty into account, as explained in the next section.\\
 	  After a defined number of training steps, called epoch, the controller is validated on each model individually.
 If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes.
-In this way 'over-fitting' is avoided, as discussed in \cite{Kurutach2018}.
+In this way 'over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}.
 The training is stopped if the controller solves the problem on the real environment.
@@
 	\subsection{Critical design decisions}\label{ss:critical_design}
-In the following, the most important aspects of a successful application of model-based RL in accelerator problems are discussed, where we made some algorithmic design decisions which we think are beneficial for problems typically found in accelerator operation.\\
+In the following, the most important aspects of a successful application of model-based RL are discussed.
+We made some algorithmic design decisions, which we think are beneficial for problems typically found in accelerator operation.\\
 %	 For the DYNA-AE algorithm, the important points are:
@@
 %		\item The ANN - including the prior - number of models - noise level - early stopping.
-%		\item The number of data points and the policy.
+%		\item The number of data-points and the policy.
 %		\item The uncertainty - how is this included?
@@
 %	\end{itemize}
-The anchor ensembling method, as treated in \cref{appendix:The impact of noise}, usually yields good results already with a small number of different networks.
-The idea is to mimic the posterior probability of the dynamics model to capture the uncertainty due to the lack of data.
+The 'anchored ensembling' method, as treated in \cref{appendix:The impact of noise}, usually yields good results already with a small number of models.
 Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes})).
-The main goal is not to determine the exact posterior probability of the model, but to provide the uncertainty to not over-train the controller at areas without sufficient data.\\ A small two-layer network with around 15-25 nodes with a $tanh$ activation and a learning rate of $10^{-3}$ were used.
+The main goal is not to determine the exact posterior probability of the model, but, as already mentioned, providing the uncertainty to not over-train the controller at areas without sufficient data.\\ A small densely connected two-layer network with around 15-25 nodes each and $tanh$ activation functions were used.
 The last layer was linear, and the inputs $\{\bs,\ba\}$ were normalised to the interval [-1,1].\\
-The prior is controlled by the initialisation of the weights of the network.
+The prior (determined by the initial weights of the network) is controlled by the initialisation of the weights of the network.
 It defines how large the variation of the approximated function is expected.
 The weights of the last layer were normalised by the number of its nodes.
-For the moment, only homoscedastic Gaussian errors are considered with a standard deviation is $\sigma_\epsilon$ is also respected during the initialisation \cref{appendix:The impact of noise} and improves the training in the presence of noise as can be seen in \cref{fig:comparsion_noise_ae_dyna}.
+For the moment, only homoscedastic Gaussian errors are considered with a standard deviation is $\sigma_\epsilon$ is also respected in a regularisation term \cref{appendix:The impact of noise} and improves the training in the presence of noise as can be seen in \cref{fig:comparsion_noise_ae_dyna}.
 \\
 Several methods of network training were implemented.
-Early stopping was employed successfully with a waiting time of about 20 steps since the number of training steps is increasing with more data in close-by areas.
-This increase is followed by a shrinking of the uncertainty at regions where no data is available, leading to better training performance.
-A fixed maximal loss value threshold was also beneficially tested.
+Early stopping, with a learning rate of $10^{-3}$ to $10^{-4}$  was employed  with a waiting time of about 20 steps.
+The number of training steps is increasing with more data in close-by areas.
+It is followed by a shrinking of the uncertainty at regions where no data is available, leading to better training performance.
+A fixed maximal loss threshold was also beneficially tested.
 In the experiments, a combination of both was taken.\\
@@
 Several different approaches were tried.
-One way is to take the average of the models and add Gaussian noise levelled by the standard deviation of the models.
-Another straight forward way is to randomly select a model to provide at each training step, which was showed the best performance the original implementation of the \emph{ME-TRPO} algorithm and was used in the experiments labelled \emph{ME-TRPO}.
-A pessimistic selection only would select the model resulting in the lowest predicted reward and showed constant but too slow improvement for our tests.
-\\ Good results were also obtained by following a randomly selected single model each full episode in combination with the soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}, and were used in the experiments labelled \emph{AE-DYNA SAC}.
+One way is to take the mean of the models $\mu_m$ and the standard deviation $\sigma_m$ and sample from $\gauss(\mu_m,\sigma_m)$.
+Another strategy is to randomly select a model to provide at each training step, which demonstrated the best performance the original implementation of the \emph{ME-TRPO} algorithm and was used in the experiments labelled \emph{ME-TRPO} \cite{Kurutach2018}.
+A pessimistic selection would only select the model resulting in the lowest predicted reward and showed constant but too slow improvement for our tests.
+Good results were also obtained by following a randomly selected single model each full episode in combination with the soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}, and were used in the experiments labelled \emph{AE-DYNA SAC}.
 \\
-The number of data-points in a batch taken every time of a new epoch has to be chosen carefully.
-It is influenced by the number of initial data points, collected using a random policy.
-The initialisation phase has to be chosen not to small to minimise the chance of getting trapped in a local minimum for too long.
+The number of data-points in a new batch to improve the model has to be chosen carefully.
+It is influenced by the number of randomly collected initial data-points (and $\policy$).
+The initialisation phase has to be selected not to small to minimise chances getting trapped in a local minimum for too long.
 Afterwards, from our experience, at least one full episode should be taken.
-We decided to use a short horizon to diminish the impact of the compound error, the accumulation error following a wrong model, as well known for DYNA-style approaches.
+We decided to use a short horizon to diminish the impact of the compound error \cite{Janner2019}, the accumulation error following a wrong model, as well known in DYNA-style approaches.
 The maximal number of steps in our experimental runs was ten.
-During the data collected on the real system, the latest learned policy was taken with some small noise added to improve the exploration.\\
+During the data collected on the real system, the latest learned policy was taken with some small additional Gaussian noise to improve exploration.\\
 To decide which MFRL algorithm to use, two main algorithm classes have to be considered: the on and the off-policy algorithms.
-The online algorithm shows a more stable convergence in general, while off-policy algorithms have the advantage that the data buffer can be filled with real data as well.
+The online algorithm shows a more stable convergence in general, while off-policy algorithms have the advantage that the data buffer can be filled with real data.
 An attractive off-policy algorithm is \emph{SAC}.
@@
 In the latter case, the policy was improved over the whole training.
-For the \emph{SAC} the policy was reset after retraining the model, to profit from the exploration features.
+For the \emph{AE-DYNA-SAC} the policy was reset after retraining the model, to profit from the exploration features.
 
-	\subsection{Comparison of the used algorithms}
+	\subsubsection{Overview of the used algorithms}
 	\Cref{tab:overview_algorithms} shows a comparison of the discussed algorithms.
@@
 All algorithms are suitable for many control tasks as common in accelerator operation.
-Their strength lies in the high sample efficiency, but they differ in representational power and noise robustness.
+Their strength lies in the high sample-efficiency, but they differ in representational power and noise robustness.
 \\
@@
 			\textrm{Class}&
-			\textrm{NR\footnote{Noise robustness}}&
-			\textrm{RP\footnote{Representational power}}&
-			\textrm{SE\footnote{Sample efficiency}}\\
+			\textrm{ \begin{tabular}{@{}c@{}}Noise\\robustness\end{tabular}}&
+			\textrm{ \begin{tabular}{@{}c@{}}Representational\\power\end{tabular}}&
+			\textrm{ \begin{tabular}{@{}c@{}}Sample\\efficiency\end{tabular}}\\
 			\colrule
@@
 	As discussed in the previous sections, several tests were performed on the FERMI FEL.
-	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility.\\ 
+	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility. 
 	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments.
@@
 	In total, four tests were carried out, two using a single network and two using the double network architecture.
-A motivation for the used algorithmic set-up is discussed in \cref{appendix:naf2} on a classical example from control theory.
-In both cases \emph{smoothing} to increase the noise robustness as discussed in \cref{appendix:naf2}) was applied.\\
+A motivation for the used algorithmic set-up is given in \cref{appendix:naf2} on a classical example from control theory, the \emph{inverted pendulum}.
+In \cref{fig:comparsion_noise} the role \emph{smoothing} to increase the noise robustness can be seen.
+Therefore \emph{smoothing} was applied in all experiments.\\
 	\Cref{fig:NAF_training} displays the results, averaged over the two tests.
@@
 The state-value function $V$ (\cref{eq:state-value-function}), which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value.
-In the single-network case, convergence is reached in ~1400 steps.
-$V$ was averaged over 200 states.
+In the single-network case, convergence is reached after around 1400 steps.
+In the plot $V$ was averaged over 200 states.
 	\subsection{MBRL tests}
@@
 The convergence property can be seen in \cref{fig:ME-TRPO_observables}.
-In the upper figure, the total reward per batch and the number of data points are shown as used to train the dynamics model against the number of epochs.
+In the upper figure, the total reward per batch and the number of data-points are shown as used to train the dynamics model against the number of epochs.
 In the lower plot, the average cumulative reward as achieved by the TRPO on the individual models in the ensemble on ten episodes as in dependence of the epochs is drawn.
-During an epoch \emph{TRPO} is trained for 10000 steps on the synthetic data.
+During an epoch the \emph{TRPO} is trained for 10000 steps on the synthetic data.
 The shaded area shows the corresponding standard deviation to indicate the uncertainty of the dynamics model.
@@
 The training was stopped after 450 steps collecting 25 steps each dynamics training.
-As shown in \cref{fig:ME-TRPO_verification}, all of the 50 verification episodes were successfully finished after a few steps.
-To verify the impact of the ensemble technique, a test with a single network using the same hyper-parameters has been applied, where no convergence within 500 data points has been observed.
+	\\ As shown in \cref{fig:ME-TRPO_verification}, all of the 50 verification episodes were successfully finished after a few steps.
+To verify the impact of the ensemble technique, a test with a single network using the same hyper-parameters has been applied, where no convergence within 500 data-points has been observed.
 \\
@@
 In this test, the controller is reset each time, when new data for the dynamics model training is acquired.
-Consequently, the performance drops each time the dynamics model is retrained, as shown in \cref{fig:AE-DYNA_observables}.
-Chances to get trapped in a local optimum are littler by using this strategy of training.\\
-	In contrast to the first MBRL test, the new data batches consist of 50 data points with an initial random walk of 200 steps.
+Consequently, the performance drops each time the dynamics model is retrained, as shown in \cref{fig:AE-DYNA_observables}.\\
+	In contrast to the \emph{ME-TRPO} training, the data batches consisted of 50 data-points with an initial random walk of 200 steps.
 The number of initial steps was chosen high enough because the convergence is slowed down strongly, so that the training becomes unfeasible on a real machine as discussed in \cref{ss:critical_design}.
 Each epoch consists of 2500 steps of controller training on the model.
-The training was stopped after the acquisition of 500 data points.
+The training was stopped after the acquisition of 500 data-points.
 The verification was executed as in the first test.
 Again the success is 100\%, but the number of needed iterations per episodes is less than for the ME-TRPO.
-It might be a result of the higher number of data points (50), but in general, this method exhibited better asymptotic performance than the $ME-TRPO$ variant.
+It might be a result of the higher number of data-points (50), but in general, this method exhibited better asymptotic performance than the \emph{ME-TRPO} variant.
 	\section{Discussion and outlook}
@@
 Two different approaches were tested: a model free and a model based method.\\
-	 Both experiments yielded satisfactory results showing that this non-linear and noisy problem could be solved in a feasibly number of training steps.\\
+	 Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasibly number of training steps.
+In the MFRL, 700 data-points were used and in the MBRL 450-500 data-points were used.
+In all experiments the 50 verification episodes were finished successfully.\\
 	Usage of the proposed methods in an operational way is attractive and would replace the current optimization method using the destructive measurement of the screens and additionally would save valuable time.\\
-	The model-free methods were slightly better in the final performance but also more sample were taken.
-The model-based methods had to be stopped due to operational reasons, hence additional studies regarding a long time performance would be interesting.
-One big issue when applying model-based methods is the computational effort.
-In our tests the data acquisition time was only a fraction of the time needed to train the controller on the learned model.
+	The MFRL methods were slightly better in the final performance but also more samples have been collected during the training.
+The MBRL methods had to be stopped due to lack of available experimental time, hence additional studies regarding a long time performance would be interesting.\\ One big issue in applying MBRL methods can be the computational time to train the controller on the model.
+In our tests the data acquisition time was only a fraction of the time needed for the controller-training.
 In case, the used methods could be parallelized to decrease the computational time.
@@
 \section{Conclusions}
- The used methods hold tremendous promise for automatizing setups common in accelerators.
-One purpose of this work was to provide some suggestions how to make advanced deep reinforcement learning techniques work on a real set-up by adjusting available methods.\\
- Generally control problems in accelerators have a short horizon, which makes the usage of \emph{DYNA-style} algorithms attractive as the \emph{ME-TRPO} or our  \emph{AE-DYNA} attractive choices for highly sample efficient algorithms.\\
- Complementary, the \emph{NAF} algorithms provides a good alternative, revealing as a model-free method good asymptotic performance.
-As most cases in accelerator control could be covered using the assumption of the quadratic dependence of the actions of the state-action value function, there is a wide range of possible applications.\\
+ The presented methods hold tremendous promise for automatizing set-ups typical in accelerators.
+One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real set-up by adapting available methods.\\
+Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA-style} algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency attractive choices.\\
+Complementary, the \emph{NAF} algorithm presents a good alternative, revealing, as a model-free method, good asymptotic performance.
+As numerous cases in accelerator control can be treated with the assumption of the quadratic dependence of the actions of the state-action value function, there is a broad spectrum of potential applications.\\
 
- To provide the possibility for other labs to use the proposed methods the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020}.
+ To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020}.
  \vfill
@@
  \vfill
+
   \newpage
@@
 	\end{multline}
-	and the policy is obtained via $\max_\ba Q_{\theta_1}(\bs,\ba) = \mu_{\theta_1}$.\\
+	and the policy is obtained via $\max_\ba Q_{\theta_1}(\bs,\ba) = \mu_{\theta_1}(\bs)$.\\
 	The results are shown in \cref{fig:comparsion_smoothing_small}.
@@
 	\subsection{The impact of noise}\label{appendix:The impact of noise}
-	A test adding artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma)$ with $\sigma=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}.
+	A test adding large artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma_\epsilon)$ with $\sigma_\epsilon=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}.
 There the difference of the three methods becomes even more evident.The results are shown in \cref{fig:comparsion_noise}.
@@
 The Bayesian community and the RL community is spending more and more attention to
-new approaches to Bayesian inference. A regularisation
-term is added to a loss function, which returns a point estimate of
+new approaches to Bayesian inference. In the `Anchored ensembling' technique a regularisation
+term is added to the loss function, which returns a point estimate of
 the Bayesian posterior the maximum a posteriori (MAP) parameter estimate. 
-Injecting noise into this loss, either to targets or regularisation term, and sampling
-repeatedly (i.e. ensembling), produces a distribution of MAP solutions mimicking that of the true posterior \cite{Gu2007, Chen2011, Bardsley2012, Pearce2018}.\\
-Consider an artificial neural net (ANN) containing parameters, ${\theta}$, making predictions for the dynamics and the reward \cref{eq:dynamics_model}, $\hat{f}_\theta$, with $N$ data points.
+Using an ensemble of networks and adding noise on either the targets of the regularisation term a distribution of MAP solutions  \cite{Gu2007, Chen2011, Bardsley2012, Pearce2018} is obtained, which mimics true posterior.\\
+Consider an artificial neural net containing parameters, ${\theta}$, making predictions for the dynamics and the reward \cref{eq:dynamics_model}, $\hat{f}_\theta$ ($f$ denotes the true unknown function), with $N$ data-points.
 If the prior is given by $P(\theta) = \mathcal{N}(\pmb{\mu}_{prior}, \pmb{\Sigma}_{prior})$, maximising the following returns MAP parameter estimates (details see \cite{Pearce2018}):
@@
 assuming homoskedastic Gaussian noise of variance $\sigma^2_\epsilon$, we obtain a parametric form of the data likelihood.
-We take $M$ models, where each model is indexed by$j \in \{1 ...
-M\}$.
-The MAP estimates are found by minimising ($\mathcal L_{j}$ denotes the loss of the $j^\text{th}$ model),
+Taking $M$ models, where each model is indexed by$j \in \{1 ...
+M\}$, the MAP estimates are found by minimising ($\mathcal L_{j}$ denotes the loss of the $j^\text{th}$ model),
 \begin{equation}
@@
 \end{equation}
-We have defined a diagonal regularisation matrix, $\pmb{\Gamma}$, where the $i^{th}$ diagonal element is the ratio of data noise of the target variable to prior variance for parameter $\theta_i$:
+$\pmb{\Gamma}$ is diagonal regularisation matrix. The $i^{th}$ diagonal element is the ratio of data noise of the target variable to prior variance for parameter $\theta_i$:
 \begin{equation}
 	\label{eqn_anch_loss_init}
-	\text{diag}(\pmb{\Gamma})_i = \frac{\sigma^2_\epsilon}{\sigma^2_{prior_i}},
+	\text{diag}(\pmb{\Gamma})_i = \frac{\sigma^2_\epsilon}{\sigma^2_{prior_i}}.
 \end{equation}
-	Using the anchors and $\sigma_\epsilon=\sigma$ in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}.
+	Using the anchors and $\sigma_\epsilon=0.05$ in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}.
 In the upper plot the mean cumulative reward on 10 test episodes on the real environment is shown during the training.
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise_ae_dyna}
-		\caption{Cumulative reward of \emph{AE-DYNA} on the \emph{inverted pendulum} with artificial noise using the anchor.}
+		\caption{Cumulative reward of \emph{AE-DYNA-SAC} on the \emph{inverted pendulum} with artificial noise using the anchor.}
 		\label{fig:comparsion_noise_ae_dyna}
@@
 \Cref{fig:Compare_models_sizes} shows the impact of the number of models onto the performance on the \emph{inverted pendulum}.
-The maximum cumulative reward averaged over five different runs tested on the real environment during the training is visualized in dependence of the number of data points.
+The maximum cumulative reward averaged over five different runs tested on the real environment during the training is visualized in dependence of the number of data-points.
 A number of three models (label \emph{Three}) shows a good trade-off between performance and training time.
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_models_sizes}
-	\caption{Varying number of models in the ensemble of the \emph{AE-DYNA} on the \emph{inverted pendulum}.}
+	\caption{Varying number of models in the ensemble of the \emph{AE-DYNA-SAC} on the \emph{inverted pendulum}.}
 	\label{fig:Compare_models_sizes}
@@
 \subsection{NAF versus AE-DYNA}
-	Finally, \cref{fig:comparsion_NAF_AE-DYNA} should demonstrate the sample efficiency of the \emph{AE-DYNA-SAC} and the \emph{NAF} algorithm on the noisy \emph{inverted pendulum}.
-\emph{AE-DYNA} converges below 2000 data points (without noise even below 800) and the \emph{NAF2} starts to perform equally 10000 data points surpassing the -200 reward threshold.
-One clearly sees the increased sample efficiency on this problem using the \emph{AE-DYNA} in contrast to the \emph{NAF2}.
+	Finally, \cref{fig:comparsion_NAF_AE-DYNA} demonstrates the sample-efficiency of the \emph{AE-DYNA-SAC} and the \emph{NAF} algorithm on the noisy \emph{inverted pendulum}.
+\emph{AE-DYNA-SAC} converges below 2000 data-points (without noise even below 800) and the \emph{NAF2} starts to perform equally 10000 data-points surpassing the -200 reward threshold.
+One clearly sees the increased sample-efficiency on this problem using the \emph{AE-DYNA} in contrast to the \emph{NAF2}.
 		\begin{figure}[!h]
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_AE-DYNA}
-		\caption{The comparison of the \emph{NAF2} and the \emph{AE-DYNA} on the noisy \emph{inverted pendulum}.}
+		\caption{The comparison of the \emph{NAF2} and the \emph{AE-DYNA-SAC} on the noisy \emph{inverted pendulum}.}
 		\label{fig:comparsion_NAF_AE-DYNA}
@@
 	\begin{itemize}
-	\item Reference clean up.
+	\item Reference and code clean up.
 	\item Conclusion and outlook\dots
-	\item Clean discussion of the results. What is superior and why? A table with numbers...
+	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
+	\item Check the plots.
 %	\item Noise benchmarks with pendulum - done

### 2020-12-03 AEDYNA.py

@@ -1014,8 +1014,8 @@ def aedyna(real_env, num_epochs=50, steps_per_env=100, algorithm='SAC',
 
             if ep == 0:
                 # Sample random action during the first epoch
-                if step_count % 5 == 0:
-                    env.reset()
+                # if step_count % 5 == 0:
+                #     env.reset()
                 act = np.random.uniform(-1, 1, size=act_dim)
                 act = np.squeeze(act)
             else:
@@ -1252,7 +1252,7 @@ if __name__ == '__main__':
             :param kwargs: noise
             :param render: flag to render
             """
-            self.render = render
+            self.showing_render = render
             self.current_step = 0
             if 'noise' in kwargs:
                 self.noise = kwargs.get('noise')
@@ -1271,14 +1271,14 @@ if __name__ == '__main__':
             obs, reward, done, info = self.env.step(action)
             if self.current_step >= 200:
                 done = True
-            if self.render:
+            if self.showing_render:
                 # Simulate and visualise the environment
                 self.env.render()
             obs = obs + self.noise * np.random.randn(self.env.observation_space.shape[-1])
             # reward = reward / 10
             return obs, reward, done, info
-    real_env = TestWrapperEnv(PendulumEnv(), render=False, noise=parameters.get('noise'))
-
+    env = TestWrapperEnv(PendulumEnv(), render=True, noise=parameters.get('noise'))
+    real_env = gym.wrappers.Monitor(env, "recordings_new", force=True)
     ############################################################
     # Setting the network parameters
     ############################################################

### 2020-12-03 run_naf2.py

@@ -182,7 +182,7 @@ class MonitoringEnv(gym.Wrapper):
         ob = self.scale_state_env(ob)
         reward = self.scale_rew(reward)
 
-        env.render()
+        # env.render()
         # print(action, ob, reward)
         return ob, reward, done, info
 
@@ -259,7 +259,7 @@ if __name__ == '__main__':
 
     # We normalize in a MonitoringEnv state action and reward to [-1,1] for the agent and plot results
     env = MonitoringEnv(env=PendulumEnv(), plot_progress=True)
-    # env = gym.wrappers.Monitor(env, "recording2", force=True)
+    env = gym.wrappers.Monitor(env, "recording2", force=True, video_callable=lambda episode_id: episode_id%10==0)
 
     nafnet_kwargs = dict(hidden_sizes=[100, 100], activation=tf.nn.tanh
                          , kernel_initializer=tf.random_normal_initializer(0, 0.05, seed=random_seed))

### 2020-12-03 run_naf2_for_tests.py

@@ -159,7 +159,7 @@ class MonitoringEnv(gym.Wrapper):
                 self.plot_results(self.current_episode)
         if self.plot_label:
             # Show the environment
-            self.env.render()
+            self.env.showing_render()
         obs = obs + self.noise * np.random.randn(self.env.observation_space.shape[-1])
         reward = reward/10
         return obs, reward, done, info

### 2020-12-03 read_naf_tests.py
- paper changes

@@ -43,15 +43,15 @@ except:
 # root_dir = "Experiments-150-200-50-long-0.05/"
 # data_all_1 = read_experiment(root_dir=root_dir)
 root_dir = "Experiments-150-200-50-noisy/"
-data_all_2 = read_experiment(root_dir=root_dir)
+data_all = read_experiment(root_dir=root_dir)
 
 # data_all = data_all_1
 # data_all = pd.concat([data_all, data_all_2])
-figure_name = '../Figures/Comparison___.pdf'
+figure_name = '../Figures/Comparison_noise'
 current_test_name = ''
 
-fig, axs = plt.subplots(2, 1, sharex=True)
-ax = axs[0]
+fig, axs = plt.subplots(1, 1, sharex=True)
+ax = axs
 
 print(data_all)
 
@@ -62,15 +62,36 @@ plt.ylabel('cum. reward (arb. units.)')
 ax.set_ylim(-1250,-100)
 print(data_all)
 
-root_dir = "Experiments-300-20-large-smoothing/"
-data_all = read_experiment(root_dir=root_dir)
-# sns.set_palette(sns.color_palette("bright", 8))
-ax = axs[1]
-sns.lineplot(ax=ax, data=data_all.T)
 ax.axhline(y=-200, ls=':', c='lime')
-
+ax.legend(['Clipping','No-clipping-no-smoothing','No-clipping-smoothing'])
 plt.ylabel('cum. reward (arb. units.)')
 plt.xlabel('no. episode')
 ax.set_ylim(-1250,-100)
-# plt.savefig(root_dir + figure_name)
+plt.savefig(root_dir + figure_name+'.pdf')
+plt.savefig(root_dir + figure_name+'.png')
 plt.show()
+
+# fig, axs = plt.subplots(2, 1, sharex=True)
+# ax = axs[0]
+#
+# print(data_all)
+#
+# sns.set_palette(sns.color_palette("bright", 8))
+# sns.lineplot(ax=ax, data=data_all.T)
+# ax.axhline(y=-200, ls=':', c='lime')
+# plt.ylabel('cum. reward (arb. units.)')
+# ax.set_ylim(-1250,-100)
+# print(data_all)
+#
+# root_dir = "Experiments-300-20-large-smoothing/"
+# data_all = read_experiment(root_dir=root_dir)
+# # sns.set_palette(sns.color_palette("bright", 8))
+# ax = axs[1]
[…]

### 2020-12-04 main.tex
- paper changes

@@
 	\author{Niky Bruchon}%, Gianfranco Fenu, Felice Andrea Pellegrino, Erica Salvato}
-	\affiliation{University of Trieste, Trieste, IT}
+	\affiliation{University of Trieste, Trieste, Italy}
 	% \author{Gianluca Valentino}
@@
 			\textrm{ \begin{tabular}{@{}c@{}}Noise\\robustness\end{tabular}}&
-			\textrm{ \begin{tabular}{@{}c@{}}Representational\\power\end{tabular}}&
+			\textrm{ \begin{tabular}{@{}c@{}}Represent.\\power\end{tabular}}&
 			\textrm{ \begin{tabular}{@{}c@{}}Sample\\efficiency\end{tabular}}\\
@@
 \end{table}
+\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
+	\caption{\label{tab:overview_algorithms_results}%
+		An overview over the experimental results of the different algorithms.
+	}
+\begin{ruledtabular}
+\begin{tabular}{llrr}
+%\toprule
+&    &  mean &   std \\
+%\midrule
+ae\_dyna & cum & -1.44 &  1.10 \\
+me\_trpo & cum & -1.95 &  1.76 \\
+naf\_double & cum & -0.58 &  0.72 \\
+naf\_single & cum & -0.66 &  0.80 \\
+ae\_dyna & final &  0.04 &  0.06 \\
+me\_trpo & final &  0.01 &  0.04 \\
+naf\_double & final & -0.00 &  0.02 \\
+naf\_single & final &  0.00 &  0.02 \\
+ae\_dyna & nr &  4.28 &  1.26 \\
+me\_trpo & nr &  5.46 &  2.32 \\
+naf\_double & nr &  2.67 &  1.33 \\
+naf\_single & nr &  2.56 &  1.26 \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
 	\section{Experimental results from FERMI RL online tests}\label{sec:Experimental results from FERMI RL online tests}

### 2020-12-04 Experimental_comparison.py

@@ -1,0 +1,79 @@
+import pickle
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+def read_rewards(rewards_in, data_range_in=None):
+    iterations_all = []
+    final_rews_all = []
+    mean_rews_all = []
+    for k in range(len(rewards_in)):
+        rewards = rewards_in[k]
+        iterations = []
+        final_rews = []
+        mean_rews = []
+        if data_range_in is None:
+            data_range = range(len(rewards))
+        else:
+            data_range = range(data_range_in[0], min(len(rewards), data_range_in[1]))
+        for i in data_range:
+            if len(rewards[i]) > 0:
+                final_rews.append(rewards[i][len(rewards[i]) - 1])
+                iterations.append(len(rewards[i]))
+                try:
+                    mean_rews.append(np.sum(rewards[i][1:]))
+                except:
+                    mean_rews.append([])
+        iterations_all.append(iterations)
+        final_rews_all.append(final_rews)
+        mean_rews_all.append(mean_rews)
+    iterations = pd.DataFrame(iterations_all)
+    final_rews = pd.DataFrame(final_rews_all)
+    mean_rews = pd.DataFrame(mean_rews_all)
+    return pd.concat([iterations, mean_rews, final_rews], keys=['nr', 'cum', 'final'])
+
+
+project_directory = ''
+name = 'Rewards_NAF_double.pkl'
+
+filehandler = open(project_directory + name, 'rb')
+object = pickle.load(filehandler)
+
+naf_double = read_rewards(object, data_range_in=[99, 149]).mean(level=0)
+
+name = 'Rewards_NAF_single.pkl'
+
+filehandler = open(project_directory + name, 'rb')
+object = pickle.load(filehandler)
+naf_single = read_rewards(object, data_range_in=[99, 149]).mean(level=0)
+
+name = 'Rewards_AE-DYNA.pkl'
+
+filehandler = open(project_directory + name, 'rb')
+object = pickle.load(filehandler)
+ae_dyna = read_rewards([object]).mean(level=0)
+name = 'Rewards_ME-TRPO.pkl'
+
+filehandler = open(project_directory + name, 'rb')
+object = pickle.load(filehandler)
+me_trpo = read_rewards([object]).mean(level=0)
+
+# print(naf_double)
[…]

### 2020-12-04 experimental_results_AE_DYNAs.py

@@ -5,7 +5,7 @@ import pickle
 import matplotlib.pyplot as plt
 import numpy as np
 
-label = ["ME-TRPO", "AE-DYNA", 'Simulation'][1]
+label = ["ME-TRPO", "AE-DYNA", 'Simulation'][0]
 
 if label == "ME-TRPO":
     # ME-TRPO results
@@ -119,14 +119,16 @@ def plot_results(data, label=None, **kwargs):
     # ax.plot(iterations_s, c=color, ls=':')
     ax.set_ylabel('no. iterations', color=color)
     ax.tick_params(axis='y', labelcolor=color)
+    ax.set_ylim(1,10)
     ax1 = plt.twinx(ax)
     color = 'k'
     ax1.plot(np.cumsum(iterations), c=color)
     # ax1.plot(np.cumsum(iterations_s), c=color, ls=':')
+    # print(np.cumsum(iterations)[-1])
     ax1.set_ylabel('no. cumulative steps', color=color)
     ax.set_title(label)
     # fig.suptitle(label, fontsize=12)
-
+    ax1.set_ylim(0, 275)
     ax = axs[1]
     # ax.axvspan(0, 100, alpha=0.2, color='coral')
     color = 'blue'
@@ -215,6 +217,11 @@ def plot_observables(data, label='Experiment', **kwargs):
         plt.savefig(kwargs.get('save_name') + '.png')
     plt.show()
 
+def save_data(data, name):
+    '''logging function to save results to pickle'''
+    out_put_writer = open(name + '.pkl', 'wb')
+    pickle.dump(data, out_put_writer, -1)
+    out_put_writer.close()
 
 # plot verification
 if label in ["ME-TRPO", "AE-DYNA", "Simulation"]:
@@ -232,6 +239,11 @@ if label in ["ME-TRPO", "AE-DYNA", "Simulation"]:
     object = pickle.load(filehandler)
     save_name = 'Figures/' + label
     plot_results(object, label=label, save_name=save_name)
+
+    name = 'Rewards_'+label
+    print(name)
+    save_data(object['rews'], name=name)
+
 
 # plot observables
 

### 2020-12-04 experimental_results_NAF.py

@@ -7,7 +7,6 @@ import matplotlib.pyplot as plt
 
 
 def load_pickle_logging(file_name):
-    # directory = 'checkpoints/' + file_name + '/'
     directory = file_name + '/'
     files = []
     directory = directory + 'data/'
@@ -26,7 +25,6 @@ def load_pickle_logging(file_name):
 
 
 def load_pickle_final(file_name):
-    # directory = 'checkpoints/' + file_name + '/'
     directory = file_name + '/'
     file = 'plot_data_0.pkl'
 
@@ -38,17 +36,27 @@ def load_pickle_final(file_name):
     return rews, inits, losses, v_s
 
 
-file_name = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_double_q_Tango_11'
-states_0, actions_0, rewards_0, dones_0 = load_pickle_logging(file_name)
-file_name = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_double_q_Tango_11_bis'
-states_1, actions_1, rewards_1, dones_1 = load_pickle_logging(file_name)
-rewards = [rewards_0, rewards_1]
+# file_name = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_double_q_Tango_11'
+# states_0, actions_0, rewards_0, dones_0 = load_pickle_logging(file_name)
+# file_name = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_double_q_Tango_11_bis'
+# states_1, actions_1, rewards_1, dones_1 = load_pickle_logging(file_name)
+# rewards = [rewards_0, rewards_1]
+#
+# file_name_s = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_single_q_Tango_11'
+# states_s0, actions_s0, rewards_s0, dones_s0 = load_pickle_logging(file_name_s)
+# file_name_s = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_single_q_Tango_11_bis'
+# states_s, actions_s, rewards_s, dones_s = load_pickle_logging(file_name_s)
+# rewards_s = [rewards_s, rewards_s0]
 
-file_name_s = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_single_q_Tango_11'
-states_s0, actions_s0, rewards_s0, dones_s0 = load_pickle_logging(file_name_s)
[…]

### 2020-12-04 naf2_new.py

@@ -4,6 +4,13 @@ import shutil
 import matplotlib.pyplot as plt
 import gym
 
+
+'''This is a version of the NAF Article (Gu2016)
+ Gu, S.; Lillicrap, T.; Sutskever, I. & Levine, S.
+ Continuous Deep Q-Learning with Model-based Acceleration 2016
+with some modifications as explained in the paper.
+We use tensorflow 2.3'''
+
 import tensorflow as tf
 from tensorflow import keras
 # Turn off warnings form tensorflow

### 2020-12-07 main.tex
Updates from Overleaf
- paper changes

@@
 The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
-In both cases, the techniques were adapted to show increased noise resistivity as omnipresent in accelerator control problems.
+In both cases, the techniques were adapted to show increased noise robustness as omnipresent in accelerator control problems.
 Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.
@@
 One advantage of the latter approach is, that it can use previously stored data as it is an off-policy algorithm and generally needs less data to be trained.
-A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010}.
+A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010,Lillicrap2015,pmlr-v32-silver14}.
 	\subsection{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
@@
 	\begin{align}
-		Q^\policy(\bs,\ba) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
+		Q^\policy(\bs_t,\ba_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
 	\end{align}
@@
 	A nice overview of model-based reinforcement learning (MBRL) methods is provided in \cite{Wang2019}.
-Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample based methods.
+Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample based methods (e.g.
+the cross entropy method \cite{Boer2005}).
 We focus on model-based data generation.\\
@@
 Several methods of network training were implemented.
-Early stopping, with a learning rate of $10^{-3}$ to $10^{-4}$  was employed  with a waiting time of about 20 steps.
+\emph{Early stopping} \cite{Goodfellow2016}, with a learning rate of $10^{-3}$ to $10^{-4}$  was employed  with a waiting time of about 20 steps and a validation ratio of 20\%.
 The number of training steps is increasing with more data in close-by areas.
@@
 	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
-	\item Check the plots.
+	\item Check the plots, which might be already okay!
 %	\item Noise benchmarks with pendulum - done

### 2020-12-07 main.tex
- paper changes

@@
 	The nature of this equation can cause overestimation and other complications, when using a function approximator and several attempts exist to overcome the difficulties \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
-	\subsubsection{Normalized advantage function}
+	\subsubsection{Normalized advantage function}\label{ss:Normalized advantage function}
 	One way to avoid the sensitivity to the mentioned complications is to choose a simple analytical form with an explicitly calculable maximum of the $Q$-function.
@@
 			\colrule
-			\emph{NAF} & model-free & low &low &high\\
-			\emph{NAF2} & model-free & high &low& high\\
-			\emph{ME-TRPO} & model-based & low& high& high\\
-			\emph{AE-DYNA} & model-based & high& high& high\\
-		\end{tabular}
-	\end{ruledtabular}
-\end{table}
-\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
-	\caption{\label{tab:overview_algorithms_results}%
-		An overview over the experimental results of the different algorithms.
-	}
-\begin{ruledtabular}
-\begin{tabular}{llrr}
-%\toprule
-&    &  mean &   std \\
-%\midrule
-ae\_dyna & cum & -1.44 &  1.10 \\
-me\_trpo & cum & -1.95 &  1.76 \\
-naf\_double & cum & -0.58 &  0.72 \\
-naf\_single & cum & -0.66 &  0.80 \\
-ae\_dyna & final &  0.04 &  0.06 \\
-me\_trpo & final &  0.01 &  0.04 \\
-naf\_double & final & -0.00 &  0.02 \\
-naf\_single & final &  0.00 &  0.02 \\
-ae\_dyna & nr &  4.28 &  1.26 \\
-me\_trpo & nr &  5.46 &  2.32 \\
-naf\_double & nr &  2.67 &  1.33 \\
-naf\_single & nr &  2.56 &  1.26 \\
+			\emph{NAF} & MFRL & low &low &high\\
+			\emph{NAF2} &MFRL & high &low& high\\
+			\emph{ME-TRPO} & MBRL & low& high& high\\
+			\emph{AE-DYNA} & MBRL & high& high& high\\
 		\end{tabular}
@@
 	\section{Discussion and outlook}
+	\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
+		\caption{\label{tab:overview_verification}%
+			An overview over the verification results of the different algorithms.
+		}
+		\begin{ruledtabular}
+			\begin{tabular}{lrrrr}
+				%\toprule
+				&  data&  nr&  cum &   final \\
+				%\midrule
+				\emph{AE-DYNA} &500 &  4.28 $\pm$  1.26 & -1.44 $\pm$  1.10 &  0.04 $\pm$  0.06\\
+				\emph{ME-TRPO} &450&  5.46 $\pm$  2.32 & -1.95 $\pm$  1.76 &  0.01 $\pm$  0.04\\
+				\emph{NAF} &1 074&  2.67 $\pm$  1.33 & -0.58 $\pm$  0.72 &0.00 $\pm$  0.02\\
+				\emph{NAF2} &824& 2.56 $\pm$  1.26 & -0.66 $\pm$  0.80 &0.00 $\pm$  0.02\\
+				%ae\_dyna & final &  0.04 &  0.06 \\
+				%me\_trpo & final &  0.01 &  0.04 \\
+				%naf\_double & final & -0.00 &  0.02 \\
+				%naf\_single & final &  0.00 &  0.02 \\
+				%ae\_dyna & nr &  4.28 &  1.26 \\
+				%me\_trpo & nr &  5.46 &  2.32 \\
+				%naf\_double & nr &  2.67 &  1.33 \\
+				%naf\_single & nr &  2.56 &  1.26 \\
+			\end{tabular}
+		\end{ruledtabular}
+	\end{table}
 	In this paper, the applicability of the method of deep reinforcement learning on the optimization of the FERMI FEL intensity was presented.
@@
 	 Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasibly number of training steps.
-In the MFRL, 700 data-points were used and in the MBRL 450-500 data-points were used.
-In all experiments the 50 verification episodes were finished successfully.\\
+The results of the verification is summarised in \cref{tab:overview_verification}.
+In the MFRL, around 800-1000 and in the MBRL 450-500 data-points were used.
+In all experiments the 50 verification episodes were finished successfully.
+\emph{NAF} performs better in terms of episode length while the average reward in the \emph{DYNA}-style algorithms is higher.
+The experiments were done on different days, hence under slightly different conditions.\\
 	Usage of the proposed methods in an operational way is attractive and would replace the current optimization method using the destructive measurement of the screens and additionally would save valuable time.\\
@@
 	%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
-	We compare the different \emph{NAF} variants: \emph{Clipping}, \emph{No-clipping-smoothing}, \emph{No-clipping-no-smoothing}, where \emph{clipping} indicates the use of the double network and in all other cases a single network is used.
+	We compare the different \emph{NAF} variants: \emph{Clipping}, \emph{No-clipping-smoothing}, \emph{No-clipping-no-smoothing}, where \emph{clipping} indicates the use of the double network, as introduced in \cref{ss:Normalized advantage function} and in all other cases a single network is used.
 	 The term \emph{smoothing} indicates that a small clipped noise is added on the actions to stabilize the network training as:
@@
 	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
+	\item Add the noise policy of NAF
 	\item Check the plots, which might be already okay!

### 2020-12-07 Experimental_comparison.py

@@ -69,7 +69,7 @@ data_all = pd.concat([naf_single, naf_double, ae_dyna, me_trpo],
 
 idx = pd.IndexSlice
 print(data_all.loc[idx[:, 'nr'], :].T.describe().T)
-# print(data_all.T.describe().T[['mean', 'std']].apply(lambda x: np.round(x,2)).to_latex())
+print(data_all.T.describe().T[['mean', 'std']].apply(lambda x: np.round(x,2)).to_latex())
 # print(data_all.loc[idx[:, 'final'], :].mean(axis=1))
 # print(data_all)
 # plt.show()

### 2020-12-07 main.tex
- Niky last fermi test

@@
 
-	\title{Online Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
+	\title{Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
 
@@
 	% \affiliation{University of Malta}
-	\affiliation{Univertsity of Salzburg, Salzburg, Austria}
+	\affiliation{University of Salzburg, Salzburg, Austria}
 	\author{Niky Bruchon}%, Gianfranco Fenu, Felice Andrea Pellegrino, Erica Salvato}
@@
 	\end{itemize}
-	One demanding aspect using RL is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018}, and a second critical aspect is the robustness of the training in real world problems itself.\\
+	One demanding aspect using RL is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018,DulacArnold2019}, and a second critical aspect is the robustness of the training in real world problems itself.\\
 	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.\\
@@
 
- To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020}.
+ To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
  \vfill
@@
 	\item Check the plots, which might be already okay!
+	\item Acknowledgements
 %	\item Noise benchmarks with pendulum - done

### 2020-12-07 local_fel_simulated_env.py

@@ -218,7 +218,8 @@ class FelLocalEnv(gym.Env):
                 new_state = kwargs.get('set_state')
                 print('set_state')
             else:
-                new_state = self.observation_space.sample()
+                # new_state = self.observation_space.sample()
+                new_state = np.array([.1,.3,.7,.2])
 
             self.set_state(new_state)
             state = self.get_state()

### 2020-12-07 optimization_interface.py

@@ -1,0 +1,115 @@
+import numpy as np
+import scipy.optimize as opt
+import matplotlib.pyplot as plt
+import matplotlib
+import pandas as pd
+import time
+
+# import PyQt5
+matplotlib.use("Qt5Agg")
+
+from local_fel_simulated_env import FelLocalEnv
+from simulated_tango import SimTangoConnection
+
+tango = SimTangoConnection()
+
+
+# real_env = FelLocalEnv(tango=tango)
+
+class EnvironmentWrapper():
+    def __init__(self, **kwargs):
+        self.real_env = FelLocalEnv(tango=tango)
+        self.real_env.reset()
+        self.action_space_dimensions = self.real_env.action_space.shape[0]
+        self.action_bounds_high, self.action_bounds_low = self.real_env.action_space.high, self.real_env.action_space.low
+        print(self.action_bounds_high)
+        # print(self.action_bounds_high, self.action_bounds_low)
+        self.action_names = ['a1', 'a2', 'a3', 'a4']
+        self.action_history = pd.DataFrame(columns=self.action_names)
+        self.state_names = ['s1', 's2', 's3', 's4']
+        self.state_history = pd.DataFrame(columns=self.state_names)
+
+        # self.state_names = list(self.awake_env.state_names)
+        # state_index_h, state_index_v = [], []
+        #
+        # for element in self.state_names:
+        #     name, plane = element.split('_')
+        #     if plane == 'horizontal':
+        #        state_index_h.append(name)
+        #     else:
+        #         state_index_v.append(name)
+        # self.state_h_history = pd.DataFrame(columns=state_index_h)
+        # self.state_v_history = pd.DataFrame(columns=state_index_v)
+        # print(self.state_h_history, self.state_v_history)
+
+        self.fig = plt.figure()
+        self.ax1 = self.fig.add_subplot(211)
+        self.ax2 = self.fig.add_subplot(212, sharex=self.ax1)
+        # plt.show(block=False)
+
+        self.fig1, self.axs = plt.subplots(2, 1, sharex=True)
+        # plt.show(block=False)
+
+        self.counter = -1
+
+        self.session_name = 'Niky_has_fun.h5'
+
+        self.algorithm_name = None
+
+        self.current_action = np.zeros(self.action_space_dimensions)
+
+    def reset(self, state=None):
+        # self.current_state, self.current_reward = self.real_env.reset() NIKY
+        self.current_state = self.real_env.reset()
+        self.current_reward = np.nan
[…]

### 2020-12-09 main.tex
- papers change

@@
 \newcommand{\NB}[1]{\textcolor{red}{#1}}
+\newcommand{\tocheck}[1]{\textcolor{blue}{#1}}
 \include{tex/defs}
@@
 	In particle accelerator operation, one main goal is to provide stable and reproducible performance.
-Achieving this demands to consider several control problems simultaneously \cite{Kain2020}.
+Achieving this demands to consider several control problems simultaneously \cite{Kain2020,Scheinker2018}.
 Especially if there is no way to model the physics a priori, one might use optimisation techniques as, e.g.
@@
 	One demanding aspect using RL is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018,DulacArnold2019}, and a second critical aspect is the robustness of the training in real world problems itself.\\
-	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste.\\
-	 Two different algorithm classes (\emph{NAF} and \emph{AE-DYNA}) were applied successfully to the FERMI FEL problem and reveal different advantages.
+	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}.\\
+	 Two different algorithm classes, a modification of the normalized advantage function \emph{NAF} (\cref{ss:Normalized advantage function}) and a novel anchored ensembling model-based algorithm, the \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), were applied successfully to the FERMI FEL problem and reveal different advantages.
 The critical aspects are addressed, and tailored solutions, generally applicable to problems showing a sample-complexity as in accelerator control problems are presented.
+	 \subsection{An Overview of the Main Results}
+	 We demonstrate optimization techniques that simultaneously adjust 4 parameters, the tilt and incline of 2 mirrors that control the trajectory and overlap of the seed laser in the FERMI FEL relative to the beam in the undulator.
+We show that our algorithms require at most 1000 training sets, after which they are able to maximize the FERMI FEL’s output intensity to > that 95\% of the maximum within only ~3 to ~5 steps depending on which algorithm is used.
+Also we show the operational applicability of purely model-based reinforcement learning in a accelerator control problem.\\
+	  Furthermore, our algorithms can perform real-time learning based on data and therefore can be re-run and re-trained for time-varying systems (is this true?).
 	The paper is organised as follows:
@@
 	A free-electron laser is a fourth-generation light source where the lasing medium consists of a very-high-speed electron moving freely through a magnetic structure.
-The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronisation with external timing systems.\\
-	The external optical laser signal provides is a contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy.
+By using an external seeding source, the FERMI FEL has several advantages relative to standard FEL approaches in terms of increased stability in pulse and photon energy, reduced size, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.\\
+The external optical laser signal interacts with the relativistic electron beam in the undulator, introducing an energy modulation that aids in the FEL process.
 The modulation in energy is converted into a charge modulation in the dispersive section, and finally, the density modulated beams radiation is amplified in the section of the radiator.\\
-	 The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.\\
+	 The importance of ensuring the best possible overlap between the seed laser and the electron beam in the modulator is therefore evident.\\
 	For this reason, the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
@@
 	The RL algorithm accesses the problem via an \emph{openai gym} environment \cite{Brockman2016}, which interfaces the RL algorithm with the hardware control system.\\
-	The state $\bs$ is a four dimensional vector holding the current voltage values applied to the piezo motors, where each component lies the normalzed interval [-1, 1].
+	The state $\bs$ is a four dimensional vector holding the current voltage values applied to the piezo motors, where each component lies within the normalized interval [-1, 1].
 Two values correspond to the two DOF of each mirror.
@@
 For reason discussed later, the horizon varies in the applications between either 10 or 50 steps.
-An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system.\\
+An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system \tocheck{discussion with Niky}.\\
 	 During the training phase exploration and the training of the function approximator takes place.
@@
 	\section{Deep Reinforcement learning}
-	Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and a reward function $\reward : \states \times \actions \rightarrow \real$, a scalar $\discount \in (0,1]$ called discount factor and an initial state distribution $\initstate$.
+	In what follows, we denote the set of all states $\bs$ by $\states$ and the set of all actions $\ba$ by $\actions$.
+Relative to state and action vectors, $(\bs,\ba)$, we define a reward function $\reward : \states \times \actions \rightarrow \real$, a scalar $\discount \in (0,1]$ called discount factor and an initial state distribution $\initstate$.
 $\transitions(\bs_{t+1}|\bs_t,\ba_t)$ characterises the probability to end in a state $\bs_{t+1}$ if an action $\ba_t$ is taken at state $\bs_t$.
@@
 	\subsection{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
+	\tocheck{Why Q learning?}
 	We focus on the main principles of MBRL.
@@
 in \cite{Sutton2018,Levine2020}.
-In the following, we discuss the normalised advantage function (\emph{NAF}) algorithm in some detail, as it has good characteristics for accelerator control problems \cite{Kain2020,Hirlaender2020a}.
+In the following, we discuss the normalised advantage function (\emph{NAF}) algorithm in some detail, as it has good characteristics for accelerator control problems as its high sample efficiency\cite{Kain2020,Hirlaender2020a}.
 	\subsubsection{Approximate dynamic programming}
@@
 	\begin{align}
-		Q^\policy(\bs_t,\ba_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
+		Q^\policy(\bs_t,\ba_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right].
 	\end{align}
-	is expressed as $Q^\policy_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator.
+	is expressed as $Q^\policy_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator, such as the weights of a neural network.
 By satisfying the Bellmann-optimality equation $Q^\policy_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$ by minimizing the Bellman error:
@@
 	\begin{align}
-		\policy_\theta(\bs)=(\delta(\ba)-\arg\!\max_\ba Q_\theta(\bs,\ba)).
+		\policy_\theta(\bs)=\delta(\ba-\arg\!\max_\ba Q_\theta(\bs,\ba)).
 	\end{align}
@@
 
-	\subsection{Uncertainty aware DYNA-style reinforcement learning}
+	\subsection{Uncertainty aware DYNA-style reinforcement learning}\label{ss:Uncertainty aware DYNA-style reinforcement learning}
 	%The original DYNA algorithm \cite{Kurutach2018} is modified here in several aspects.
@@
 The 'anchored ensembling' method, as treated in \cref{appendix:The impact of noise}, usually yields good results already with a small number of models.
-Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes})).
-The main goal is not to determine the exact posterior probability of the model, but, as already mentioned, providing the uncertainty to not over-train the controller at areas without sufficient data.\\ A small densely connected two-layer network with around 15-25 nodes each and $tanh$ activation functions were used.
+Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes}).
+The main goal is not to determine the exact posterior probability of the model, but, as already mentioned, providing the uncertainty to not over-train the controller at areas without sufficient data.\\
+A small densely connected two-layer network with around 15-25 nodes each and $tanh$ activation functions were used.
 The last layer was linear, and the inputs $\{\bs,\ba\}$ were normalised to the interval [-1,1].\\
@@
 Another strategy is to randomly select a model to provide at each training step, which demonstrated the best performance the original implementation of the \emph{ME-TRPO} algorithm and was used in the experiments labelled \emph{ME-TRPO} \cite{Kurutach2018}.
-A pessimistic selection would only select the model resulting in the lowest predicted reward and showed constant but too slow improvement for our tests.
-Good results were also obtained by following a randomly selected single model each full episode in combination with the soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}, and were used in the experiments labelled \emph{AE-DYNA SAC}.
+A pessimistic selection would only select the model resulting in the lowest predicted reward and showed constant but too slow improvement for our online-training.
+Good results were also obtained by following a randomly selected single model each full episode in combination with the soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}, and were used in the experiments labelled \emph{AE-DYNA-SAC}.
 \\
@@
 An attractive off-policy algorithm is \emph{SAC}.
-The \emph{SAC} not only tries to maximise $J$ \cref{eq:cumulative_reward} but also simultaneously is regularised to maximise the entropy in the action space.
+The \emph{SAC} not only tries to maximise $J$ \cref{eq:cumulative_reward} but also simultaneously is regularised to maximise the entropy in the action space \cite{Haarnoja2018a}.
 In this way, exploration is encouraged to avoid getting stuck in a local optimum and has a good trade-off between exploration and exploitation.\\
@@
 	\caption{\label{tab:overview_algorithms}%
-		An overview of the algorithms discussed in this paper.
+		Overview of the algorithms.
 	}
@@
 Therefore \emph{smoothing} was applied in all experiments.\\
-	\Cref{fig:NAF_training} displays the results, averaged over the two tests.
+	\Cref{fig:NAF_training} displays the results, averaged over the two tests \footnote{Considering the training is stochastic with random initial conditions, by chance, the training of the double network variant started above the threshold in both cases.}.
 A training of 100 episodes was accomplished.
 In the upper figure, the number of iterations per episode is plotted, including the cumulative number of steps.
-The green line in the lower plot of \cref{fig:NAF_training} indicates the 95\% intensity reward threshold, to accept an episode to be successfully finished.
+The green lines in the lower plot of \cref{fig:NAF_training} indicate the 95\% intensity reward threshold, to accept an episode to be successfully finished.
+The blue lines show the cumulative reward during the epoch.
 	In the verification of 50 episodes (\cref{fig:NAF_verification}) both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance.
@@
 In the single-network case, convergence is reached after around 1400 steps.
-In the plot $V$ was averaged over 200 states.
+In the plot $V$ was averaged over 200 randomly selected states.
 	\subsection{MBRL tests}
@@
 	The second test campaign was employing the \emph{AE-DYNA} algorithm, as a representative for pure MBRL algorithms.
-Two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA SAC} variant.
+Two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA-SAC} variant.
 The algorithmic design details are discussed in \cref{ss:critical_design}.
@@
 \\
-	Secondly, the \emph{AE-DYNA SAC} was tested, which uses the \emph{SAC}- algorithm \cite{Haarnoja2018a}.
+	Secondly, the \emph{AE-DYNA-SAC} was tested, which uses the \emph{SAC}- algorithm \cite{Haarnoja2018a}.
 In this test, the controller is reset each time, when new data for the dynamics model training is acquired.

### 2020-12-09 experimental_results_NAF.py

@@ -117,12 +117,31 @@ def plot_results(rewards, rewards_single, label=None, **kwargs):
         if verification:
             ax1.set_ylim(0, 275)
             ax.set_ylim(1, 10)
-    color = 'k'
+    color = 'lime'
     ax1.plot(np.cumsum(iterations), c=color)
     ax1.plot(np.cumsum(iterations_s), c=color, ls=':')
     ax1.set_ylabel('no. cumulative steps', color=color)
+    ax1.tick_params(axis='y', labelcolor=color)
     ax.set_title(label)
     # fig.suptitle(label, fontsize=12)
+    import matplotlib.lines as mlines
+
+    single_line = mlines.Line2D([], [], color='k', ls=':', label='single network')
+    double_line = mlines.Line2D([], [], color='k', ls='-', label='double network')
+    plt.legend(handles=[single_line,double_line])
+
+    # import pandas as pd
+    # df = pd.DataFrame([iterations[:10], iterations[-10:]], index=['initial','final']).T
+    # df_d =df.describe().T[['mean']]
+    # df = pd.DataFrame([iterations_s[:10], iterations_s[-10:]], index=['initial','final']).T
+    # df_s = df.describe().T[['mean']]
+    # df = pd.concat([df_s,df_d], axis=1)
+    # df.columns =['single network', 'double network']
+    # pd.plotting.table(ax, np.round(df, 2), loc='center right', colWidths=[0.15, 0.15],
+    #                 colLabels=[single_line,double_line], label='av')
+    # New subplot
+    # plt.legend(['double network', 'single network'])
+
 
     ax = axs[1]
     # ax.axvspan(0, 100, alpha=0.2, color='coral')
@@ -152,6 +171,7 @@ def plot_results(rewards, rewards_single, label=None, **kwargs):
         save_name = kwargs.get('save_name')
         plt.savefig(save_name + '_episodes.pdf')
         plt.savefig(save_name + '_episodes.png')
+        plt.savefig(save_name + '_episodes.pgf')
     plt.show()
 
 
@@ -170,29 +190,6 @@ def read_losses_v_s(losses0, v_s0, max_length):
     losses = np.mean(losses_all, axis=0)
     v_s = np.mean(v_s_all, axis=0)
[…]

### 2020-12-11 main.tex
- paper changes

@@
 	In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator.\\
-	At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation.
-Nevertheless, in the last years, various approaches have been studied to investigate their applicability in tuning operations.
-\\
+%	At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation. 
+In the last years, various approaches have been studied to investigate their applicability in tuning operations. \\
 	A free-electron laser is a fourth-generation light source where the lasing medium consists of a very-high-speed electron moving freely through a magnetic structure.
@@
 	The final problem faced consists of optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.
-	The RL algorithm accesses the problem via an \emph{openai gym} environment \cite{Brockman2016}, which interfaces the RL algorithm with the hardware control system.\\
+	The RL algorithm accesses the problem via an \emph{openai gym} environment \cite{Brockman2016}, which interfaces the RL algorithm with the control system.\\
 	The state $\bs$ is a four dimensional vector holding the current voltage values applied to the piezo motors, where each component lies within the normalized interval [-1, 1].
@@
 For reason discussed later, the horizon varies in the applications between either 10 or 50 steps.
-An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system \tocheck{discussion with Niky}.\\
+An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system \tocheck{discussion with Niky}.
+In our experiments the maximal intensity was provided by a manual optimization done of an expert.
+\\
 	 During the training phase exploration and the training of the function approximator takes place.
@@
 	\end{figure}
-
+	\begin{figure}
+		\centering
+		\includegraphics*[width=0.5\textwidth]{Figures/Evolution_double.pdf}
+		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
+		\label{fig:NAF_evolution_double}
+	\end{figure}
+	\begin{figure}
+	\centering
+	\includegraphics*[width=0.5\textwidth]{Figures/Evolution_single.pdf}
+	\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
+	\label{fig:NAF_evolution_single}
+\end{figure}
 	\begin{figure}
@@
 		\centering
-		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_verification.pdf}
-		\caption{The verification of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
+		\includegraphics*[width=0.5\textwidth]{Figures/Verification_DYNA_all_episodes.pdf}
+		\caption{The verification of the \emph{ME-TRPO} and the \emph{AE-DYNA-SAC} on the FERMI FEL.}
 		\label{fig:AE-DYNA_verification}
 	\end{figure}
+
 	\begin{figure}
@@
 	\end{figure}
+
 	\begin{figure}
 	\centering
-		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
-		\caption{The verification of the ME-TRPO on the FERMI FEL.}
-		\label{fig:ME-TRPO_verification}
+	\includegraphics*[width=0.5\textwidth]{Figures/Worst_episode_MBRL.pdf}
+	\caption{The verification of the \emph{ME-TRPO} and the \emph{AE-DYNA-SAC} on the FERMI FEL.}
+	\label{fig:Worst_episode_MBRL}
 	\end{figure}
+%	\begin{figure}
+%		\centering
+%		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
+%		\caption{The verification of the ME-TRPO on the FERMI FEL.}
+%		\label{fig:ME-TRPO_verification}
+%	\end{figure}
 	The second test campaign was employing the \emph{AE-DYNA} algorithm, as a representative for pure MBRL algorithms.
@@
 				%\midrule
-				\emph{AE-DYNA} &500 &  4.28 $\pm$  1.26 & -1.44 $\pm$  1.10 &  0.04 $\pm$  0.06\\
-				\emph{ME-TRPO} &450&  5.46 $\pm$  2.32 & -1.95 $\pm$  1.76 &  0.01 $\pm$  0.04\\
-				\emph{NAF} &1 074&  2.67 $\pm$  1.33 & -0.58 $\pm$  0.72 &0.00 $\pm$  0.02\\
-				\emph{NAF2} &824& 2.56 $\pm$  1.26 & -0.66 $\pm$  0.80 &0.00 $\pm$  0.02\\
+				\emph{AE-DYNA} &500 &  3.28 $\pm$  1.26 & -1.44 $\pm$  1.10 &  0.04 $\pm$  0.06\\
+				\emph{ME-TRPO} &450&  4.46 $\pm$  2.32 & -1.95 $\pm$  1.76 &  0.01 $\pm$  0.04\\
+				\emph{NAF} &1 074&  2.56 $\pm$  1.96 & -0.66 $\pm$  1.22 &0.00 $\pm$  0.02\\
+				\emph{NAF2} &824& 2.64 $\pm$  1.65 & -0.57 $\pm$  0.92 &0.00 $\pm$  0.03\\
 				%ae\_dyna & final &  0.04 &  0.06 \\
@@
  To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
+
+ \section{Acknowledgements}
+ Giulio Gaio
  \vfill

### 2020-12-11 Experimental_comparison.py

@@ -19,10 +19,12 @@ def read_rewards(rewards_in, data_range_in=None):
             data_range = range(data_range_in[0], min(len(rewards), data_range_in[1]))
         for i in data_range:
             if len(rewards[i]) > 0:
-                final_rews.append(rewards[i][len(rewards[i]) - 1])
-                iterations.append(len(rewards[i]))
+                current_rewards = (pd.DataFrame(rewards[i]).dropna().values).flatten()
+                print(current_rewards)
+                final_rews.append(current_rewards[-1])
+                iterations.append(len(current_rewards))
                 try:
-                    mean_rews.append(np.sum(rewards[i][1:]))
+                    mean_rews.append(np.sum(current_rewards))
                 except:
                     mean_rews.append([])
         iterations_all.append(iterations)
@@ -40,13 +42,15 @@ name = 'Rewards_NAF_double.pkl'
 filehandler = open(project_directory + name, 'rb')
 object = pickle.load(filehandler)
 
-naf_double = read_rewards(object, data_range_in=[99, 149]).mean(level=0)
+# naf_double = read_rewards(object, data_range_in=[99, 149]).mean(level=0)
+naf_double = read_rewards(object).mean(level=0)
 
 name = 'Rewards_NAF_single.pkl'
 
 filehandler = open(project_directory + name, 'rb')
 object = pickle.load(filehandler)
-naf_single = read_rewards(object, data_range_in=[99, 149]).mean(level=0)
+# naf_single = read_rewards(object, data_range_in=[99, 149]).mean(level=0)
+naf_single = read_rewards(object).mean(level=0)
 
 name = 'Rewards_AE-DYNA.pkl'
 
@@ -59,10 +63,10 @@ filehandler = open(project_directory + name, 'rb')
 object = pickle.load(filehandler)
 me_trpo = read_rewards([object]).mean(level=0)
 
-# print(naf_double)
-# print(naf_single)
-# print(ae_dyna)
-# print(me_trpo)
+print(naf_double)
+print(naf_single)
+print(ae_dyna)
+print(me_trpo)
 
 data_all = pd.concat([naf_single, naf_double, ae_dyna, me_trpo],
                      keys=['naf_single', 'naf_double', 'ae_dyna', 'me_trpo'])

### 2020-12-11 experimental_results_AE_DYNAs.py

@@ -4,6 +4,7 @@ import os
 import pickle
 import matplotlib.pyplot as plt
 import numpy as np
+import pandas as pd
 
 label = ["ME-TRPO", "AE-DYNA", 'Simulation'][0]
 
@@ -27,14 +28,16 @@ def read_rewards(rewards):
     final_rews = []
     mean_rews = []
     for i in range(len(rewards)):
-        if len(rewards[i]) > 0:
-            final_rews.append(rewards[i][len(rewards[i]) - 1])
-            iterations.append(len(rewards[i]))
+        current_rewards = rewards[i][1:]
+        if len(current_rewards) > 0:
+            # print(current_rewards)
+            final_rews.append(rewards[i][-1])
+            iterations.append(len(current_rewards))
             try:
-                mean_rews.append(np.sum(rewards[i][1:]))
+                mean_rews.append(np.sum(current_rewards))
             except:
                 mean_rews.append([])
-            stds.append(np.std(rewards[i][1:]))
+            stds.append(np.std(current_rewards))
 
     # iterations = np.mean(np.array(iterations_all), axis=0)
     # final_rews = np.mean(np.array(final_rews_all), axis=0)
@@ -159,6 +162,111 @@ def plot_results(data, label=None, **kwargs):
         plt.savefig(save_name + '_verification.png')
     plt.show()
 
+def plot_results_verification(data_array, label=None, **kwargs):
+    if 'data_range_in' in kwargs:
+        data_range_in = kwargs.get('data_range_in')
+    else:
+        data_range_in = None
+
+    rewards = data_array[0]['rews']
+    iterations, final_rews, mean_rews, _ = read_rewards(rewards)
+
+    rewards = data_array[1]['rews']
+    iterations_s, final_rews_s, mean_rews_s, iter_all_s = read_rewards(rewards)
+    plot_suffix = ""  # f', number of iterations: {env.TOTAL_COUNTER}, Linac4 time: {env.TOTAL_COUNTER / 600:.1f} h'
+    fig, axs = plt.subplots(1, 1, sharex=True)
+    fig.set_size_inches(fig.get_size_inches()[0],fig.get_size_inches()[1]*0.6)
+    ax = axs
+    # ax.axvspan(0, 100, alpha=0.2, color='coral')
+    color = 'blue'
[…]

### 2020-12-11 experimental_results_NAF.py

@@ -36,6 +36,30 @@ def load_pickle_final(file_name):
     return rews, inits, losses, v_s
 
 
+def load_buffer_final(file_name):
+    directory = file_name + '/'
+    file = 'buffer_data_verification.pkl'
+
+    with open(directory + file, 'rb') as f:
+        buffer = pickle.load(f)
+    return buffer
+
+
+def load_buffer(file_name):
+    directory = file_name + '/'
+    files = []
+    directory = directory + 'data/'
+    for f in os.listdir(directory):
+        if 'buffer_data' in f and 'pkl' in f:
+            files.append(f)
+    files.sort()
+    print(files[-1])
+
+    with open(directory + files[-1], 'rb') as f:
+        buffer = pickle.load(f)
+    return buffer
+
+
 # file_name = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_double_q_Tango_11'
 # states_0, actions_0, rewards_0, dones_0 = load_pickle_logging(file_name)
 # file_name = 'Data_Experiments/2020_07_20_NAF@FERMI/FEL_training_100_double_q_Tango_11_bis'
@@ -76,7 +100,8 @@ def read_rewards(rewards_in, data_range_in=None):
         for i in data_range:
             # print(i)
             if len(rewards[i]) > 0:
-                final_rews.append(rewards[i][len(rewards[i]) - 1])
+
+                final_rews.append(rewards[i][-1])
                 iterations.append(len(rewards[i]))
                 try:
                     mean_rews.append(np.sum(rewards[i][1:]))
@@ -86,10 +111,15 @@ def read_rewards(rewards_in, data_range_in=None):
         final_rews_all.append(final_rews)
         mean_rews_all.append(mean_rews)
 
-    iterations = np.mean(np.array(iterations_all), axis=0)
-    final_rews = np.mean(np.array(final_rews_all), axis=0)
-    mean_rews = np.mean(np.array(mean_rews_all), axis=0)
-    return iterations, final_rews, mean_rews
+    iterations_mean = np.mean(np.array(iterations_all), axis=0)
+    iterations_std = np.std(np.array(iterations_all), axis=0)
+    final_rews_mean = np.mean(np.array(final_rews_all), axis=0)
[…]

### 2020-12-14 main.tex
- paper changes

@@
 
-	\title{Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
+	\title{Uncertainty Aware Model-Based Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
 
@@
 	 We demonstrate optimization techniques that simultaneously adjust 4 parameters, the tilt and incline of 2 mirrors that control the trajectory and overlap of the seed laser in the FERMI FEL relative to the beam in the undulator.
-We show that our algorithms require at most 1000 training sets, after which they are able to maximize the FERMI FEL’s output intensity to > that 95\% of the maximum within only ~3 to ~5 steps depending on which algorithm is used.
+We show that our algorithms require at most 1000 training sets, after which they are able to maximize the FERMI FEL’s output intensity to $>$ that 95\% of the maximum within only $\sim 3$ to $\sim 5$ steps depending on which algorithm is used.
 Also we show the operational applicability of purely model-based reinforcement learning in a accelerator control problem.\\
-	  Furthermore, our algorithms can perform real-time learning based on data and therefore can be re-run and re-trained for time-varying systems (is this true?).
+	  Furthermore, our algorithms can perform real-time learning based on data and therefore can be re-run and re-trained for time-varying systems.
 	The paper is organised as follows:
@@
 	The RL algorithm accesses the problem via an \emph{openai gym} environment \cite{Brockman2016}, which interfaces the RL algorithm with the control system.\\
-	The state $\bs$ is a four dimensional vector holding the current voltage values applied to the piezo motors, where each component lies within the normalized interval [-1, 1].
+	The state $\bs$ is a four dimensional vector holding the current voltage values applied to the piezo motors, where each component lies within the normalized interval [0, 1].
 Two values correspond to the two DOF of each mirror.
@@
 	The training is done in episodes with a uniformly\footnote{in the normalized state-space} randomized initial state and a specific maximal length of taken steps afterwards - the horizon.
-For reason discussed later, the horizon varies in the applications between either 10 or 50 steps.
-An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system \tocheck{discussion with Niky}.
-In our experiments the maximal intensity was provided by a manual optimization done of an expert.
+For reason discussed later, the horizon varies in the applications between either 10 or 500 steps.
+An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system.
+In our experiments the maximal intensity was determined by a manual expert optimization.
 \\
@@
 	 Subsequently, an MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data by taking the model uncertainty into account, as explained in the next section.\\
-	  After a defined number of training steps, called epoch, the controller is validated on each model individually.
+	  After a defined number of training steps, called epoch, the controller is validated on each model of the ensemble individually.
 If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes.
@@
 The last layer was linear, and the inputs $\{\bs,\ba\}$ were normalised to the interval [-1,1].\\
-The prior (determined by the initial weights of the network) is controlled by the initialisation of the weights of the network.
+The prior is controlled by the initialisation of the weights of the network.
 It defines how large the variation of the approximated function is expected.
@@
 The initialisation phase has to be selected not to small to minimise chances getting trapped in a local minimum for too long.
-Afterwards, from our experience, at least one full episode should be taken.
+Afterwards, from our experience, at least one full episode per epoch should be taken.
 We decided to use a short horizon to diminish the impact of the compound error \cite{Janner2019}, the accumulation error following a wrong model, as well known in DYNA-style approaches.
@@
 In the latter case, the policy was improved over the whole training.
-For the \emph{AE-DYNA-SAC} the policy was reset after retraining the model, to profit from the exploration features.
+For the \emph{AE-DYNA-SAC} the policy was reset after retraining the model, to profit from its exploration features.
 
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Evolution_double.pdf}
-		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
+		\caption{The training of the \emph{NAF2} on the FERMI FEL using a double network.}
 		\label{fig:NAF_evolution_double}
@@
 	\includegraphics*[width=0.5\textwidth]{Figures/Evolution_single.pdf}
-	\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
+	\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network.}
 	\label{fig:NAF_evolution_single}
@@
 A training of 100 episodes was accomplished.
-In the upper figure, the number of iterations per episode is plotted, including the cumulative number of steps.
-The green lines in the lower plot of \cref{fig:NAF_training} indicate the 95\% intensity reward threshold, to accept an episode to be successfully finished.
-The blue lines show the cumulative reward during the epoch.
-	In the verification of 50 episodes (\cref{fig:NAF_verification}) both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance.
+The number of iterations per episode is plotted, including the cumulative number of steps.
+In all episodes during the training the 95\% intensity reward threshold was surpassed, and therewith successfully finished.
+The evolution of the states is provided in \cref{fig:NAF_evolution_single} and \cref{fig:NAF_evolution_double} for the single and double network.
+At the boundary of the domain, it takes time to learn, since many actions are mapped onto the same state\footnote{The states are clipped to $[0,1]$.}.
+The behaviour indicates that the double network converges faster spending less time at the boundary.\\
+	In the verification of 100 episodes (\cref{fig:NAF_verification}) both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance.
 \\
@@
 	The second test campaign was employing the \emph{AE-DYNA} algorithm, as a representative for pure MBRL algorithms.
-Two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA-SAC} variant.
+As discussed, two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA-SAC} variant.
 The algorithmic design details are discussed in \cref{ss:critical_design}.
@@
 The convergence property can be seen in \cref{fig:ME-TRPO_observables}.
-In the upper figure, the total reward per batch and the number of data-points are shown as used to train the dynamics model against the number of epochs.
+In the upper figure, the total reward per batch gathered from the real environment and the number of data-points are shown as used to train the dynamics model against the number of epochs.
 In the lower plot, the average cumulative reward as achieved by the TRPO on the individual models in the ensemble on ten episodes as in dependence of the epochs is drawn.
@@
 The training was stopped after 450 steps collecting 25 steps each dynamics training.
-	\\ As shown in \cref{fig:ME-TRPO_verification}, all of the 50 verification episodes were successfully finished after a few steps.
+	\\ As shown in \cref{fig:AE-DYNA_verification}, all of the 50 verification episodes were successfully finished after a few steps.
 To verify the impact of the ensemble technique, a test with a single network using the same hyper-parameters has been applied, where no convergence within 500 data-points has been observed.
@@
 The verification was executed as in the first test.
-Again the success is 100\%, but the number of needed iterations per episodes is less than for the ME-TRPO.
-It might be a result of the higher number of data-points (50), but in general, this method exhibited better asymptotic performance than the \emph{ME-TRPO} variant.
+Again the success of all 50 episodes is 100\%, but the number of needed iterations per episodes is less than for the ME-TRPO, wich can be seen in \cref{fig:AE-DYNA_verification}.
+The somewhat better performance might be a result of the higher number of data-points (50), but in general, this method exhibited better asymptotic performance than the \emph{ME-TRPO} variant.\\
+	In both experiments the training was using a small number of data-points while still successfully solving the problem.
+in \cref{fig:Worst_episode_MBRL} the worst verification episode of both experiments is plotted.
+Different experiment days influence the boundary conditions and hence the performance.
+However, a longer training would increase the performance in both cases, especially for the \emph{ME-TRPO}.
+\Cref{fig:Worst_episode_MBRL} shows that the tt2 incline and tilt parameters for the \emph{ME-TRPO} move up and then down towards the optimal position, which would be cured by a better dynamics model.
 	\section{Discussion and outlook}
@@
  \section{Acknowledgements}
- Giulio Gaio
+ We want to thank Giulio Gaio, who made this experiments possible and helped to setup and restore the FERMI FEL.
+Also we are thankful for the valuable feedback from Alexander Scheinker on the draft of this publication.
  \vfill

### 2020-12-14 experimental_results_AE_DYNAs.py

@@ -424,13 +424,15 @@ ax1.plot(data_r.index.values, data_r.values, c=color, drawstyle="steps-post")
 ax1.set_ylabel('FEL intensity (% of max)', color=color)
 ax1.tick_params(axis='y', labelcolor=color)
 
-ax.plot(data_r.index.values, data_s, ls=':', drawstyle="steps-post")
+ax.plot(data_r.index.values, data_s, drawstyle="steps-post")
 
 ax.set_xlabel('no. step')
 ax.set_ylabel('norm. state (arb. units)')
 ax.legend(labels=data_s.columns, bbox_to_anchor=(0., 1.22, 1., .102), loc='lower left',
           ncol=2, mode="expand", borderaxespad=0.)
 ax.set_ylim(0, 1)
+ax.set_zorder(ax1.get_zorder()+1)
+ax.patch.set_visible(False)
 
 object = object_ae
 states_1 = object['obss']
@@ -448,9 +450,7 @@ data_s = pd.DataFrame(states_1[selected_item])
 data_a = pd.DataFrame(actions_1[selected_item][1:], columns=['1','2','3','4'])
 data_s.columns = data_a.columns
 data_correct = pd.concat([pd.DataFrame(data_s.iloc[0,:]).T, data_a], ignore_index=True)
-print(data_s.head())
-print(data_correct.cumsum().head())
-# print(data_a.iloc[:,:])
+
 data_s.columns = ['tt1 tilt', 'tt1 incline', 'tt2 tilt', 'tt2 incline']
 data_r = pd.DataFrame(rewards_1[selected_item])
 data_r = data_r + 1
@@ -460,7 +460,7 @@ max = data_r[1:][dones_1[selected_item][1:]].values
 ax_0 = ax_2
 # ax_0.text(0, 1.02, r'Worst training episode among the last ten.', fontdict=font)
 ax_0.set_title(r'Worst verification episode AE-DYNA', fontdict=font)
-ax_0.plot(data_s.index.values, data_s.values, ls=':', drawstyle="steps-post")
+ax_0.plot(data_s.index.values, data_s.values, drawstyle="steps-post")
 # plt.legend(loc='upper right')
 ax_0.set_xlabel('no. step')
 ax_0.set_ylabel('norm. state (arb. units)')
@@ -472,6 +472,8 @@ ax_1.set_ylabel('FEL intensity (% of max)', color=color)
[…]

### 2020-12-14 experimental_results_NAF.py

@@ -452,7 +452,7 @@ fig, (ax_1, ax_2) = plt.subplots(2)
 
 ax = ax_1
 ax1 = ax.twinx()
-ax1.set_title(r'Worst training episode among the first ten.', fontdict=font, zorder=10)
+ax1.set_title(r'Worst training episode among the first ten.', fontdict=font, zorder=0)
 color = 'lime'
 
 ax1.axhline(y=0.95, c=color, ls='--')
@@ -461,7 +461,7 @@ ax1.plot(data_r.index.values, data_r.values, c=color, drawstyle="steps-post")
 ax1.set_ylabel('FEL intensity (% of max)', color=color)
 ax1.tick_params(axis='y', labelcolor=color)
 
-ax.plot(data_r.index.values, data_s, ls=':', drawstyle="steps-post")
+ax.plot(data_r.index.values, data_s, drawstyle="steps-post")
 
 ax.set_xlabel('no. step')
 ax.set_ylabel('norm. state (arb. units)')
@@ -469,18 +469,23 @@ ax.legend(labels=data_s.columns, bbox_to_anchor=(0., 1.22, 1., .102), loc='lower
           ncol=2, mode="expand", borderaxespad=0.)
 ax.set_ylim(0, 1)
 
-shift = 2
+ax.set_zorder(ax1.get_zorder()+1)
+ax.patch.set_visible(False)
+shift = 10
 selected_item = np.argmax([len(states_1[i]) for i
                            in range(len(states_1) - shift, len(states_1))]) + len(states_1) - shift
 print('selected_item', selected_item)
 # selected_item = 55
 # for max_lengths in range(len(states_1)):
 data_s = pd.DataFrame(states_1[selected_item])
+print('data_s', data_s)
+data_s = data_s.drop_duplicates()
+print('data_s', data_s)
+
 data_a = pd.DataFrame(actions_1[selected_item][1:], columns=['1','2','3','4'])
-data_s.columns = data_a.columns
-data_correct = pd.concat([pd.DataFrame(data_s.iloc[0,:]).T, data_a], ignore_index=True)
-print(data_s.head())
-print(data_correct.cumsum().head())
+# data_s.columns = data_a.columns
+# data_correct = pd.concat([pd.DataFrame(data_s.iloc[0,:]).T, data_a], ignore_index=True)
+
 # print(data_a.iloc[:,:])
[…]

### 2020-12-16 main.tex
- paper changes

@@
 
-	\title{Uncertainty Aware Model-Based Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
+	\title{Model-Based and Model-Free Deep Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
 
-	\author{Simon Hirlaender}
+	\author{Simon Hirlaender}\email{simon.hirlaender@sbg.ac.at}
 	%\thanks{These two authors contributed equally}
 	% \affiliation{University of Malta}
-	\affiliation{University of Salzburg, Salzburg, Austria}
+	\affiliation{IDA Lab, University of Salzburg, 5020 Salzburg, Austria}
 	\author{Niky Bruchon}%, Gianfranco Fenu, Felice Andrea Pellegrino, Erica Salvato}
@@
 	% %\thanks{These two authors contributed equally}
-	% \email{verena.kain@cern.ch}
 
@@
 	\begin{abstract}
-		In this paper, we compare a model-based reinforcement learning to a model-free reinforcement learning method applied to the FERMI FEL system.
-Both approaches hold tremendous promise, and the principal goal of this paper is to show how reinforcement learning can be utilised on an operational level in a reasonable training time on real accelerator physics problems.
-In terms of representational power and sample-efficiency, the model-based approach is better, while the asymptotic performance of the model-free method is slightly superior.
+		Reinforcement learning holds tremendous promise in accelerator controls, and the principal goal of this paper is to show how reinforcement learning can be utilised on an operational level on accelerator physics problems.
+		We compare a model-based reinforcement to a model-free reinforcement learning method applied to the FERMI FEL system.
+In terms of representational power and sample-efficiency, the model-based approach is overtaking, while the asymptotic performance of the model-free method is slightly superior.
 The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
@@
 	\end{itemize}
-	One demanding aspect using RL is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018,DulacArnold2019}, and a second critical aspect is the robustness of the training in real world problems itself.\\
+	One demanding aspect using RL in real world applications is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018,DulacArnold2019}.
+A second critical aspect to be considered is the robustness of the training in a noisy data limited environment.\\
 	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}.\\
-	 Two different algorithm classes, a modification of the normalized advantage function \emph{NAF} (\cref{ss:Normalized advantage function}) and a novel anchored ensembling model-based algorithm, the \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), were applied successfully to the FERMI FEL problem and reveal different advantages.
-The critical aspects are addressed, and tailored solutions, generally applicable to problems showing a sample-complexity as in accelerator control problems are presented.
+	 We successfully applied two different algorithm classes, a highly sample-efficient variant of continuous Q-learning (\cref{ss:Normalized advantage function}) and a newly proposed 
+	  model-based algorithm, which we call \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), to the FERMI FEL problem.\\ The methods reveal different characteristics addressing the stated critical aspects.
+They are generally applicable to problems showing a sample-complexity as in common accelerator control problems.
 	 \subsection{An Overview of the Main Results}
 	 We demonstrate optimization techniques that simultaneously adjust 4 parameters, the tilt and incline of 2 mirrors that control the trajectory and overlap of the seed laser in the FERMI FEL relative to the beam in the undulator.
-We show that our algorithms require at most 1000 training sets, after which they are able to maximize the FERMI FEL’s output intensity to $>$ that 95\% of the maximum within only $\sim 3$ to $\sim 5$ steps depending on which algorithm is used.
-Also we show the operational applicability of purely model-based reinforcement learning in a accelerator control problem.\\
-	  Furthermore, our algorithms can perform real-time learning based on data and therefore can be re-run and re-trained for time-varying systems.
+We show that our algorithms require significantly less than 1000 training sets, after which they are able to maximize the FERMI FEL’s output intensity to $>$ that 95\% of the maximum within only $\sim 3$ to $\sim 5$ steps depending on which algorithm is used.
+Also we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\
+	  Furthermore, our algorithms can perform real-time learning based on data and therefore can be re-run and re-trained for time-varying systems.\\
 	The paper is organised as follows:
@@
 		\item Results of the experiments 
-		\item Summary and outlook
+		\item Discussion, outlook and summary
 	\end{itemize}
@@
 	 The importance of ensuring the best possible overlap between the seed laser and the electron beam in the modulator is therefore evident.\\
-	For this reason, the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
+%	For this reason, the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
 	\subsection{The training environment}
@@
 	\begin{align}
-		\bs_{t+1} = \bs_t+\ba_t.
+		\bs_{t+1} = \bs_t+\ba_t,
 	\end{align} 
+	and is limited by an $ |\ba_t|\leq \ba_\text{max}=\frac{1}{12}$.
 	The training is done in episodes with a uniformly\footnote{in the normalized state-space} randomized initial state and a specific maximal length of taken steps afterwards - the horizon.
@@
 	 During the training phase exploration and the training of the function approximator takes place.
-	After a defined number of training episodes, a verification of 50 episodes is performed. In the second phase, the RL only applies the learned policy.
+	After a defined number of training episodes, a verification of 50 or 100 episodes is performed. In the second phase, the RL only applies the learned policy.
 
@@
 	\section{Deep Reinforcement learning}
+    Most commonly two main categories of deep RL algorithms are distinguished: model-free RL (MFRL) and model-based RL (MBRL).
+MFRL algorithms learn directly from the interaction with a system, while MBRL algorithm learn a dynamics model from these interactions first.
+Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, there applications is mainly restricted to simulated environments due to their low sample-efficiency.
+Yet putting constrains on the algorithms, as discussed in \cref{s:Model free reinforcement learning}, partly cures the sample inefficiency turning them into applicable tools for real world problems.\\
+      On the other hand MBRL showing a much better sample efficiency.
+The challenge for these methods is to learn an accurate model, which turns out to be rather hard and leads to the so called model-bias problem \cite{Deisenroth2011}.
+In \cref{ss:Uncertainty aware DYNA-style reinforcement learning} we discuss attempts to capture the model uncertainty to alleviate this obstacle by making use of model ensembles and Bayesian neural networks resulting in algorithms, exhibiting a high sample-efficiency and a low model-bias.\\
 	In what follows, we denote the set of all states $\bs$ by $\states$ and the set of all actions $\ba$ by $\actions$.
@@
 	\end{align}
-	where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.\\
-	In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba) \approx Q_\theta(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
+	where
+	\begin{align}
+	p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)
+	\label{eq:trajectory_distribution}
+	\end{align}
+
+is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
+
+	\subsection{Model free reinforcement learning}\label{s:Model free reinforcement learning}
+		In the modern field of RL one differentiates if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba) \approx Q_\theta(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
 a deep neural network.\\
 	In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{Sutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}.
-To train these methods, a large number of direct interactions with the system is required, because they belong to on-policy algorithms.
-In the latter one deals with approximate dynamic programming, which we shortly treat in \cref{s:Model free reinforcement learning}.
-One advantage of the latter approach is, that it can use previously stored data as it is an off-policy algorithm and generally needs less data to be trained.
-A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010,Lillicrap2015,pmlr-v32-silver14}.
-	\subsection{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
-	\tocheck{Why Q learning?}
-	We focus on the main principles of MBRL.
-An overview can be found e.g.
+To train these methods, a large number of direct interactions with the system is required, since they belong to \emph{on-policy} algorithms.
+Hence, we rule them out for direct online training but make use of their characteristics in MBRL (\cref{ss:critical_design}).\\
+	 In the second case we speak about Q-learning, which belongs to approximate dynamic programming.
+One advantage is the possible usage of use previously stored data as we use it in an \emph{off-policy} fashion, which generally needs less data to be trained than \emph{on-policy} methods.
+A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010,Lillicrap2015,Silver2014}.\\
+	 As already adduced, MFRL algorithms learn directly from interactions with the system.
+An exhaustive overview can be found e.g.
 in \cite{Sutton2018,Levine2020}.
-In the following, we discuss the normalised advantage function (\emph{NAF}) algorithm in some detail, as it has good characteristics for accelerator control problems as its high sample efficiency\cite{Kain2020,Hirlaender2020a}.
+In the following, we discuss the normalised advantage function (\emph{NAF}) algorithm \cite{Gu2016} in some detail, as it has good characteristics for accelerator control problems.
+It is highly sample-efficiency  and its representational power is sufficient for most common cases\cite{Kain2020,Hirlaender2020a}.
+Modifications to increase the stability are subject of \cref{ss:Normalized advantage function}.
 	\subsubsection{Approximate dynamic programming}
@@
 		\begin{align}
-		V^\policy(\bs_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],\label{eq:state-value-function}
+		V^\policy(\bs_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right].\label{eq:state-value-function}
 	\end{align}
-	The state-action-value function $Q^\policy(\bs,\ba)$, which tells us how good an action $\ba$ in a state $\bs$ following $\policy$ is, in terms of the expected return:
+	The state-action-value function - short Q function - $Q^\policy(\bs,\ba)$, which tells us how good an action $\ba$ in a state $\bs$ following $\policy$ is, in terms of the expected return:
 	\begin{align}
-		Q^\policy(\bs_t,\ba_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right].
+		Q^\policy(\bs_t,\ba_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
 	\end{align}
 	is expressed as $Q^\policy_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator, such as the weights of a neural network.
-By satisfying the Bellmann-optimality equation $Q^\policy_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$ by minimizing the Bellman error:
+By satisfying the Bellmann-optimality equation $Q^\policy_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$ by minimizing the Bellman error\footnote{$\vec{Q}_\theta$ denotes the Q-function $Q_\theta$ represented as a vector of length $|\states|\times |\actions|$.}:
 	\begin{align}
@@
 	\begin{align}
-		\policy_\theta(\bs)=\delta(\ba-\arg\!\max_\ba Q_\theta(\bs,\ba)).
+		\policy_\theta(\bs_t)=\delta(\ba_t-\arg\!\max_\ba Q_\theta(\bs_t,\ba_t)).
 	\end{align}
-	The Bellman operator $\bellman^*$ has a unique fixed point but is non-linear, due to the $\max$ - operator \cite{Sutton2018}:
+	The Bellman operator $\bellman^*$ has a unique fixed point but is non-linear, due to the involved $\max$ - operator \cite{Sutton2018}:
 	\begin{multline}
@@
 	\end{multline}
-	The nature of this equation can cause overestimation and other complications, when using a function approximator and several attempts exist to overcome the difficulties \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
-	\subsubsection{Normalized advantage function}\label{ss:Normalized advantage function}
+	The nature of this equation can cause overestimation and other complications, when using a function approximator and several attempts exist to overcome its difficulties \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
+	\subsubsection{Design decisions for MFRL}\label{ss:Normalized advantage function}
 	One way to avoid the sensitivity to the mentioned complications is to choose a simple analytical form with an explicitly calculable maximum of the $Q$-function.
-	If a specific quadratic form of the $Q$ function is assumed \cite{Gu2016}:
+	If a specific quadratic form of the Q-function is assumed \cite{Gu2016}:
 	\begin{align}
@@
 	\begin{align}
-		&\mathcal L(\theta)=\\\notag&\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
+		&\mathcal L(\theta)=\\\notag&\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2.
 	\end{align}
-	$\theta_\text{targ}$ are the weights of a target network, which is softly updated.
-To stabilize the network training a small clipped artificial noise is added to the actions called \emph{smoothing} as described in \cref{appendix:naf2}.\\
-	 This implementation has an extremely good sample-efficiency for suitable problems as can be found often in accelerators.
+	$\theta_\text{targ}$ are the weights of a target network, which is softly updated \cite{Lillicrap2015, Gu2016,Silver2014}.
+To further stabilize the network training a small clipped artificial noise is added to the actions called \emph{smoothing} \cite{fujimoto2018addressing}.
+The effect is demonstrated in \cref{appendix:naf2} on a classical example from control theory, the \emph{inverted pendulum}.
+In \cref{fig:comparsion_noise} its role to increase the noise robustness can be seen.
+Therefore \emph{smoothing} was applied in all MFRL experiments.\\
+	 Our implementation has a high sample-efficiency and noise resistivity.
 We use it as the baseline for the FERMI-FEL control problem, as it yields good results.
-Additional changes to the original proposal \cite{Gu2016} and a previous implementation used for experiments using a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
+Additional changes to the original proposal \cite{Gu2016} and a previous implementation used for accelerator control using to a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
+We refer to our modified implementation as \emph{NAF2} and the implementation close to the original, as in \cite{Gu2016}, as \emph{NAF}.
 	%\subsection{New NAF network architecture}
@@
 
-	A nice overview of model-based reinforcement learning (MBRL) methods is provided in \cite{Wang2019}.
+	A nice overview of state of the art MBRL methods is provided in \cite{Wang2019}.
 Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample based methods (e.g.
@@
 		In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$.
-Additionally in our case the reward is included in the model \footnote{In some applications $\reward$ can be deduced from the current state or the state change.
+In addition to the next state, the reward is included in the model \footnote{In some applications $\reward$ can be deduced from the current state or the state change.
 In such a situation the modelling can be simplified to model directly $\hat\transitions$}:
@@
 We use only synthetic data to reduce the interaction with the real environment to a minimum.\\
-	An overview of the used method is shown in \cref{fig:MBRL_overview}.
-In the beginning, the data is collected or read from previously collected data.
-An uncertainty aware model of the dynamics is trained, using 'anchored ensembling' \cite{Pearce2018} on the data, which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
-It was motivated by the fact that pure epistemic ensemble techniques as \emph{ME-TRPO} were reported to be very sensitive to aleatoric noise \cite{Wang2019}.
-Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take these uncertainties into account using probabilistic ensembles.
-Our implementation is simpler while showing similar performance.\\
-	 Subsequently, an MFRL algorithm is deployed to learn the controller on the learned model by only using the synthetic data by taking the model uncertainty into account, as explained in the next section.\\
-	  After a defined number of training steps, called epoch, the controller is validated on each model of the ensemble individually.
+	A schematic overview of our used method is shown in \cref{fig:MBRL_overview}.
+Initially, data is gathered by interaction with the real environment or read from a previous collection.
+An uncertainty aware model of the dynamics is trained, using 'anchored ensembling' \cite{Pearce2018} on the data (details provided in \cref{appendix:The impact of noise,ss:The uncertainty aware dynamics model}), which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
+It was motivated by the fact that pure epistemic ensemble techniques as \cite{Kurutach2018} were reported to be very sensitive to aleatoric noise \cite{Wang2019}.
+Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take these uncertainties into consideration using probabilistic ensembles.
+Our implementation is simpler while showing excellent performance in considered experiments.\\
+	 Subsequently, an MFRL algorithm is deployed to learn the controller on the dynamics model by only using the synthetic data and interoperating the model uncertainty, as explained in \cref{ss:The uncertainty aware dynamics model}.\\
+	  After a defined number of training steps, called epoch, the controller is validated on each model of the ensemble, individually.
 If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes.
 In this way 'over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}.
-The training is stopped if the controller solves the problem on the real environment.
+The training is stopped if the controller produces satisfying results on the real environment.
 Otherwise, a batch of new data is collected and the model is improved and consecutively a new epoch starts.
@@
 	\end{figure}
-	\subsection{Critical design decisions}\label{ss:critical_design}
-In the following, the most important aspects of a successful application of model-based RL are discussed.
+	\subsection{Critical design decisions in MBRL}\label{ss:critical_design}
+In the following, the most important aspects of a successful application of MBRL are discussed.
+To our knowledge uncertainty aware MBRL is applied for the first time with success on an accelerator problem and might be, despite its incredible potential, demanding.
 We made some algorithmic design decisions, which we think are beneficial for problems typically found in accelerator operation.\\
@@
 %	\end{itemize}
-The 'anchored ensembling' method, as treated in \cref{appendix:The impact of noise}, usually yields good results already with a small number of models.
+\subsubsection{The uncertainty aware dynamics model}\label{ss:The uncertainty aware dynamics model}
+The 'anchored ensembling' method usually yields good results already with a small number of models.
 Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes}).
@@
 The prior is controlled by the initialisation of the weights of the network.
-It defines how large the variation of the approximated function is expected.
+It defines how large the variation of the approximated function is expected and were randomized uniformly in a small interval $[-\Delta, \Delta]$.
 The weights of the last layer were normalised by the number of its nodes.
-For the moment, only homoscedastic Gaussian errors are considered with a standard deviation is $\sigma_\epsilon$ is also respected in a regularisation term \cref{appendix:The impact of noise} and improves the training in the presence of noise as can be seen in \cref{fig:comparsion_noise_ae_dyna}.
+For the moment, only homoskedastic Gaussian errors are considered with a standard deviation is $\sigma_\epsilon$ is also respected in a regularisation term \cref{appendix:The impact of noise} and significantly improves the training in the presence of noise, as shown in \cref{fig:comparsion_noise_ae_dyna}.
 \\
@@
 A fixed maximal loss threshold was also beneficially tested.
-In the experiments, a combination of both was taken.\\
+In the experiments, a combination of both was taken.
+\subsubsection{The controller algorithm}
+To decide which MFRL algorithm to use, two already mentioned main algorithm classes are considered  (\cref{s:Model free reinforcement learning}): \emph{on-} and  \emph{off-policy} algorithms \cite{Sutton2018}.
+\emph{On-policy} algorithm show a stabler convergence in general, while \emph{off-policy} algorithms have the advantage that the data buffer can be filled with real data.\\
+The \emph{on-policy} algorithm trust region policy optimization (TRPO) \cite{Schulman2015} provides the theoretical guarantee of a monotonic improvement of the policy.
+We use this algorithm in experiments labelled as \emph{ME-TRPO}, which stands for model-ensemble TRPO as proposed in \cite{Kurutach2018}.
+As reported in \cite{Kurutach2018}, using proximal policy optimisation \cite{Schulman2017}, another prominent \emph{on-policy} algorithm, reveals minor performance, which was confirmed in our tests.
+\\
+An attractive off-policy algorithm is the soft-actor-critic (\emph{SAC}) algorithm.
+\cite{fujimoto2018addressing,Hill2018}.
+\emph{SAC} not only tries to maximise \cref{eq:cumulative_reward} but also simultaneously is regularised to maximise the entropy in the action space \cite{Haarnoja2018a}.
+In this way, exploration is encouraged to avoid getting stuck in a local optimum and a good trade-off between exploration and exploitation is achieved.
+We refer to experiments using the \emph{SAC} as \emph{AE-DYNA-SAC}.
+For \emph{AE-DYNA-SAC} the controller was reset after each retraining of the dynamics model, to profit from its exploration features, while for \emph{ME-TRPO} the policy was constantly improved to exploit its monotony.
+\subsubsection{Handling of the uncertainty}
 One of the most crucial points is how model uncertainties are taken into account by the MFRL controller.
@@
 One way is to take the mean of the models $\mu_m$ and the standard deviation $\sigma_m$ and sample from $\gauss(\mu_m,\sigma_m)$.
-Another strategy is to randomly select a model to provide at each training step, which demonstrated the best performance the original implementation of the \emph{ME-TRPO} algorithm and was used in the experiments labelled \emph{ME-TRPO} \cite{Kurutach2018}.
-A pessimistic selection would only select the model resulting in the lowest predicted reward and showed constant but too slow improvement for our online-training.
-Good results were also obtained by following a randomly selected single model each full episode in combination with the soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}, and were used in the experiments labelled \emph{AE-DYNA-SAC}.
-\\
+Another strategy is to randomly select a model to provide at each training step, which demonstrated the best performance the original implementation of the model ensemble \emph{TRPO} (\emph{ME-TRPO}) algorithm and was used in the experiments labelled \emph{ME-TRPO} \cite{Kurutach2018}.
+A pessimistic selection would only select the model resulting in the lowest predicted reward (as used e.g.
+in \cite{kidambi2020morel}) and showed constant but too slow improvement for our online-training.
+Good results were also obtained by following a randomly selected single model each full episode in combination, which was used in the \emph{AE-DYNA-SAC}. 
+\subsubsection{The data acquisition}
 The number of data-points in a new batch to improve the model has to be chosen carefully.
-It is influenced by the number of randomly collected initial data-points (and $\policy$).
-The initialisation phase has to be selected not to small to minimise chances getting trapped in a local minimum for too long.
+It is influenced by the number of randomly collected initial data-points (and then by the used policy $\policy$).
+The initialisation phase has to be selected not to small to minimise risks getting trapped in a local minimum for too long.
 Afterwards, from our experience, at least one full episode per epoch should be taken.
 We decided to use a short horizon to diminish the impact of the compound error \cite{Janner2019}, the accumulation error following a wrong model, as well known in DYNA-style approaches.
-The maximal number of steps in our experimental runs was ten.
-During the data collected on the real system, the latest learned policy was taken with some small additional Gaussian noise to improve exploration.\\
-To decide which MFRL algorithm to use, two main algorithm classes have to be considered: the on and the off-policy algorithms.
-The online algorithm shows a more stable convergence in general, while off-policy algorithms have the advantage that the data buffer can be filled with real data.
-An attractive off-policy algorithm is \emph{SAC}.
-The \emph{SAC} not only tries to maximise $J$ \cref{eq:cumulative_reward} but also simultaneously is regularised to maximise the entropy in the action space \cite{Haarnoja2018a}.
-In this way, exploration is encouraged to avoid getting stuck in a local optimum and has a good trade-off between exploration and exploitation.\\
-On the other hand, the on-policy algorithm TRPO \cite{Schulman2015} provides some theoretical improvement guarantees.
-As reported in \cite{Kurutach2018} using proximal policy optimisation \cite{Schulman2017} as an online MFRL algorithm does not show the same performance, which was confirmed in our tests.
-In the latter case, the policy was improved over the whole training.
-For the \emph{AE-DYNA-SAC} the policy was reset after retraining the model, to profit from its exploration features.
-
-	\subsubsection{Overview of the used algorithms}
+The maximal number of steps in our experimental runs per episode was ten.
+During the data acquisition on the real system, the latest learned policy $\policy$ was taken with some small additional Gaussian noise to improve exploration.
+%\tocheck{Niky discuss the averaging of the data}
+	\section{Experimental results from FERMI RL online tests}\label{sec:Experimental results from FERMI RL online tests}
+	As discussed in the previous sections, several tests were performed on the FERMI FEL.
+	The main purpose was to try the newly implemented algorithms on a real system to evaluate their operational feasibility. 
+	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments.\\ 
+%	Since there was a positive outcome, we deduce that the utilised methods are at an operational level.\\
 	\Cref{tab:overview_algorithms} shows a comparison of the discussed algorithms.
 Details are provided in the appendix.
-All algorithms are suitable for many control tasks as common in accelerator operation.
+All algorithms are suitable for control tasks as common in accelerator operation.
 Their strength lies in the high sample-efficiency, but they differ in representational power and noise robustness.
-\\
 In \cref{sec:Experimental results from FERMI RL online tests} the \emph{NAF2},
 	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment.
-\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
+	\begin{table}[h!]%The best place to locate the table environment is directly after its first reference in text
 		\caption{\label{tab:overview_algorithms}%
@@
 		\begin{ruledtabular}
-		\begin{tabular}{lllll}
+			\begin{tabular}{l p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
 				\textrm{Algorithm}&
 				\textrm{Class}&
-			\textrm{ \begin{tabular}{@{}c@{}}Noise\\robustness\end{tabular}}&
-			\textrm{ \begin{tabular}{@{}c@{}}Represent.\\power\end{tabular}}&
-			\textrm{ \begin{tabular}{@{}c@{}}Sample\\efficiency\end{tabular}}\\
+				\textrm{Noise \newline robustness}&
+				\textrm{Represent. power}&
+				\textrm{Sample \newline efficiency}\\
 				\colrule
@@
 	\end{table}
-
-	\section{Experimental results from FERMI RL online tests}\label{sec:Experimental results from FERMI RL online tests}
-	As discussed in the previous sections, several tests were performed on the FERMI FEL.
-	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility. 
-	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments.
-Since there was a positive outcome, it is evident, that the utilised methods are at an operational level.\\
 	The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
@@
 	In total, four tests were carried out, two using a single network and two using the double network architecture. 
-A motivation for the used algorithmic set-up is given in \cref{appendix:naf2} on a classical example from control theory, the \emph{inverted pendulum}.
-In \cref{fig:comparsion_noise} the role \emph{smoothing} to increase the noise robustness can be seen.
-Therefore \emph{smoothing} was applied in all experiments.\\
 	\Cref{fig:NAF_training} displays the results, averaged over the two tests \footnote{Considering the training is stochastic with random initial conditions, by chance, the training of the double network variant started above the threshold in both cases.}.
@@
 The number of iterations per episode is plotted, including the cumulative number of steps.
-In all episodes during the training the 95\% intensity reward threshold was surpassed, and therewith successfully finished.
-The evolution of the states is provided in \cref{fig:NAF_evolution_single} and \cref{fig:NAF_evolution_double} for the single and double network.
+In all episodes during the training the 95\% intensity reward threshold was surpassed, and therewith they were successfully finished.
+The evolution of the states is provided in \cref{fig:NAF_evolution_single,fig:NAF_evolution_double} for the single and double network.
 At the boundary of the domain, it takes time to learn, since many actions are mapped onto the same state\footnote{The states are clipped to $[0,1]$.}.
-The behaviour indicates that the double network converges faster spending less time at the boundary.\\
-	In the verification of 100 episodes (\cref{fig:NAF_verification}) both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance.
-\\
+The behaviour indicates that the double network converges faster and spends less time at the boundary.\\
+	In the verification of 100 episodes, \cref{fig:NAF_verification}, both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance.
 Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps.
@@
 As discussed, two variants were implemented: the \emph{ME-TRPO} variant and the \emph{AE-DYNA-SAC} variant.
-The algorithmic design details are discussed in \cref{ss:critical_design}.
-	\\ The first uses the trust region policy optimisation -\emph{TRPO}- \cite{Schulman2015} to train the controller.
-The TRPO monotonically converges to better policy, and this property is exploited in training.
-The convergence property can be seen in \cref{fig:ME-TRPO_observables}.
-In the upper figure, the total reward per batch gathered from the real environment and the number of data-points are shown as used to train the dynamics model against the number of epochs.
-In the lower plot, the average cumulative reward as achieved by the TRPO on the individual models in the ensemble on ten episodes as in dependence of the epochs is drawn.
+The algorithmic design details were discussed in \cref{ss:critical_design}.
+	\\ To exploit the convergence properties of the \emph{TRPO}, the policy was never reset as can be seen in \cref{fig:ME-TRPO_observables}.
+In the upper figure, the total reward per batch gathered from the real environment and the number of data-points used to train the dynamics model as a function of the number of epochs are shown.
+In the lower plot, the average cumulative reward of ten episodes as achieved by the \emph{TRPO} on the individual models of the ensemble in dependence of the epochs is drawn.
 During an epoch the \emph{TRPO} is trained for 10000 steps on the synthetic data.
 The shaded area shows the corresponding standard deviation to indicate the uncertainty of the dynamics model.
-As a measure of convergence of the TRPO, the logarithm of the standard deviation of $p_\traj$ is visualised.
+After initial performance drops caused by the insufficiently trained model, the algorithm continuously improves.
+As a measure of convergence of the \emph{TRPO}, the logarithm of the standard deviation of $p_\policy$ (\cref{eq:trajectory_distribution}) is visualised.
 The training was stopped after 450 steps collecting 25 steps each dynamics training.
@@
 \\
-	Secondly, the \emph{AE-DYNA-SAC} was tested, which uses the \emph{SAC}- algorithm \cite{Haarnoja2018a}.
-In this test, the controller is reset each time, when new data for the dynamics model training is acquired.
-Consequently, the performance drops each time the dynamics model is retrained, as shown in \cref{fig:AE-DYNA_observables}.\\
+	Secondly, the \emph{AE-DYNA-SAC} was tested.
+As discussed in \cref{ss:critical_design}, in this test, the controller was reset each time, when the model was retrained and consequently, the performance drops each time as shown in \cref{fig:AE-DYNA_observables}.\\
 	In contrast to the \emph{ME-TRPO} training, the data batches consisted of 50 data-points with an initial random walk of 200 steps.
-The number of initial steps was chosen high enough because the convergence is slowed down strongly, so that the training becomes unfeasible on a real machine as discussed in \cref{ss:critical_design}.
+The number of initial steps was chosen high enough because otherwise the convergence is slowed down strongly, so that the training becomes unfeasible on a real machine (discussed in \cref{ss:critical_design}).
 Each epoch consists of 2500 steps of controller training on the model.
@@
 The verification was executed as in the first test.
-Again the success of all 50 episodes is 100\%, but the number of needed iterations per episodes is less than for the ME-TRPO, wich can be seen in \cref{fig:AE-DYNA_verification}.
+Again, the success of all 50 episodes is 100\%, but the number of needed iterations per episodes is less than for the ME-TRPO, wich can be seen in \cref{fig:AE-DYNA_verification}.
 The somewhat better performance might be a result of the higher number of data-points (50), but in general, this method exhibited better asymptotic performance than the \emph{ME-TRPO} variant.\\
 	In both experiments the training was using a small number of data-points while still successfully solving the problem.
-in \cref{fig:Worst_episode_MBRL} the worst verification episode of both experiments is plotted.
-Different experiment days influence the boundary conditions and hence the performance.
+In \cref{fig:Worst_episode_MBRL} the worst verification episodes of both experiments are plotted.
+The experiments, taken at different days, may influence the boundary conditions and hence the performance.
 However, a longer training would increase the performance in both cases, especially for the \emph{ME-TRPO}.
-\Cref{fig:Worst_episode_MBRL} shows that the tt2 incline and tilt parameters for the \emph{ME-TRPO} move up and then down towards the optimal position, which would be cured by a better dynamics model.
+\Cref{fig:Worst_episode_MBRL} shows that the tt2 incline and tilt parameters for the \emph{ME-TRPO} move up and then down towards the optimal position, which would be improved by a better dynamics model.
 	\section{Discussion and outlook}
-	\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
+	In this paper, the applicability of the method of deep reinforcement learning on the optimization of the FERMI FEL intensity was presented.
+Two different approaches were tested: a model free and a model based method.\\
+	 Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasibly number of training steps.
+The results of the verification is summarised in \cref{tab:overview_verification}.
+In the model-free case, around 800-1000 and in the model-based 450-500 data-points were used.
+In all experiments all verification episodes were finished successfully.
+\emph{NAF2} performs better in terms of episode length while the average reward in the \emph{DYNA}-style algorithms is higher.
+The experiments were done on different days, hence under slightly different conditions.\\
+	Usage of the proposed methods in an operational way is attractive and would replace the current optimization method using the destructive measurement of the screens and additionally would save valuable time.\\
+	The MFRL methods were slightly better in the final performance but also more samples have been collected during the training.
+The MBRL methods had to be stopped due to lack of available experimental time, hence additional studies regarding a long time performance would be interesting.\\ One big issue in applying MBRL methods can be the computational time to train the controller on the model.
+In our tests the data acquisition time was only a fraction of the time needed for the controller-training.
+In case, the used methods could be parallelized to reduce the computational time.
+		\begin{table}[h]%The best place to locate the table environment is directly after its first reference in text
 		\caption{\label{tab:overview_verification}%
@@
 		\begin{ruledtabular}
-			\begin{tabular}{lrrrr}
+			\begin{tabular}{l>{\raggedleft} p{1cm}>{\raggedleft}p{1.8cm}>{\raggedleft} p{1.8cm} >{\raggedleft\arraybackslash} p{1.8cm}}
 				%\toprule
-				&  data&  nr&  cum &   final \\
-				%\midrule
+				&  \centering Data points&   \centering Episode length&   \centering Cumulative reward &   \centering Final\\reward\tabularnewline
+				\hline
 				\emph{AE-DYNA} &500 &  3.28 $\pm$  1.26 & -1.44 $\pm$  1.10 &  0.04 $\pm$  0.06\\
 				\emph{ME-TRPO} &450&  4.46 $\pm$  2.32 & -1.95 $\pm$  1.76 &  0.01 $\pm$  0.04\\
-				\emph{NAF} &1 074&  2.56 $\pm$  1.96 & -0.66 $\pm$  1.22 &0.00 $\pm$  0.02\\
+				\emph{NAF} &1074&  2.56 $\pm$  1.96 & -0.66 $\pm$  1.22 &0.00 $\pm$  0.02\\
 				\emph{NAF2} &824& 2.64 $\pm$  1.65 & -0.57 $\pm$  0.92 &0.00 $\pm$  0.03\\
@@
 	\end{table}
-	In this paper, the applicability of the method of deep reinforcement learning on the optimization of the FERMI FEL intensity was presented.
-Two different approaches were tested: a model free and a model based method.\\
-	 Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasibly number of training steps.
-The results of the verification is summarised in \cref{tab:overview_verification}.
-In the MFRL, around 800-1000 and in the MBRL 450-500 data-points were used.
-In all experiments the 50 verification episodes were finished successfully.
-\emph{NAF} performs better in terms of episode length while the average reward in the \emph{DYNA}-style algorithms is higher.
-The experiments were done on different days, hence under slightly different conditions.\\
-	Usage of the proposed methods in an operational way is attractive and would replace the current optimization method using the destructive measurement of the screens and additionally would save valuable time.\\
-	The MFRL methods were slightly better in the final performance but also more samples have been collected during the training.
-The MBRL methods had to be stopped due to lack of available experimental time, hence additional studies regarding a long time performance would be interesting.\\ One big issue in applying MBRL methods can be the computational time to train the controller on the model.
-In our tests the data acquisition time was only a fraction of the time needed for the controller-training.
-In case, the used methods could be parallelized to decrease the computational time.
-
 \section{Conclusions}
- The presented methods hold tremendous promise for automatizing set-ups typical in accelerators.
+ The presented reinforcement learning methods hold tremendous promise for automatizing set-ups typical in accelerators.
 One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real set-up by adapting available methods.\\
-Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA-style} algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency attractive choices.\\
-Complementary, the \emph{NAF} algorithm presents a good alternative, revealing, as a model-free method, good asymptotic performance.
-As numerous cases in accelerator control can be treated with the assumption of the quadratic dependence of the actions of the state-action value function, there is a broad spectrum of potential applications.\\
-
+Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA-style} algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency appealing choices.\\
+Complementary, the \emph{NAF2} algorithm presents a good alternative, revealing, as a model-free method, good asymptotic performance.
+As numerous cases in accelerator control can be captured assuming a quadratic dependence on the actions of the state-action value function, there is a broad spectrum of potential applications.\\
 To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
@@
  We want to thank Giulio Gaio, who made this experiments possible and helped to setup and restore the FERMI FEL.
-Also we are thankful for the valuable feedback from Alexander Scheinker on the draft of this publication.
- \vfill
- \newpage
- \vfill
+Also, we are thankful for the valuable feedback from Alexander Scheinker on the draft of this publication.
+Finally we want to thank the Land Salzburg for their financial support to create the context of the IDA lab, which made this work possible.
 
-  \newpage
 \appendix
@@
 The episode length, the horizon, is set to 200 steps.
-In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly in terms of noise handling.
+In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly concerning the noise handling.
 It is of importance when dealing with measurements on real systems.\\
-	 For statistical significance, all shown results were performed with five different seeds.
+For statistical significance, all shown results were obtained using five different seeds.
 The average value and the standard deviation (shaded) are plotted.
@@
 \end{align}
-	with $d$ is one if the episode is finished and 0 otherwise.
+with $d$ is 1 if the episode is finished and 0 otherwise.
 Then both are learned by regressing to this target (using the tuples from the data buffer $\mathcal D$ ):
@@
 A test adding large artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma_\epsilon)$ with $\sigma_\epsilon=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}.
-There the difference of the three methods becomes even more evident.The results are shown in \cref{fig:comparsion_noise}.
+There the difference of the three methods becomes even more evident.
+The results are shown in \cref{fig:comparsion_noise}.
 After around 65 episodes the single network without \emph{smoothing} (\emph{No-clipping-no-smoothing}) decreases before reaching the final performance at around 95 episodes, while  \emph{smoothing} prevents this performance drop in the other cases.
-	\begin{figure}[!h]
+\begin{figure}
 	\centering
@@
 It should indicate the result if the training is stopped at this training epoch.
-One cannot observe this during a real training, unless one does costly performance measurements during the training.
+One cannot observe this quantity during a real training, unless one does costly performance measurements while training.
 Respecting the aleatoric (\emph{Noise-on-aleatoric}) helps to reach the target much quicker exhibiting less variation compared to the standard use of an ensemble (\emph{Noise-on-non-aleatoric}).
@@
 The lower plot of \cref{fig:comparsion_noise_ae_dyna} shows the batch rewards a measured during the data collection, which is observable during the training.
-	\begin{figure}[!h]
+\begin{figure}
 	\centering
@@
 A single network might not converge or too slowly (label \emph{Single}) and a model of ten shown fast and stable performance in this case (label \emph{Ten}).
-\begin{figure}[!h]
+\begin{figure}
 	\centering
@@
 One clearly sees the increased sample-efficiency on this problem using the \emph{AE-DYNA} in contrast to the \emph{NAF2}.
-		\begin{figure}[!h]
+\begin{figure}
 	\centering
@@
 %	\end{figure}
-\section{Still open}
-	\begin{itemize}
-	\item Reference and code clean up.
-	\item Conclusion and outlook\dots
-	\item Clean discussion of the results. What is superior and why? A table with numbers \dots
-	\item Add the noise policy of NAF
-	\item Check the plots, which might be already okay!
-	\item Acknowledgements
+%\section{Still open}
+%	\begin{itemize}
+%	\item Reference and code clean up.
+%	\item Add the noise policy of NAF
 %	\item Noise benchmarks with pendulum - done
@@
 %	\end{itemize}
-\end{itemize}
+%\end{itemize}
 
@@
 	% \bibliography{apssamp}% Produces the bibliography via BibTeX.
+		%%%%% CLEAR DOUBLE PAGE!
+	\newpage{\pagestyle{empty}\cleardoublepage}
 	\bibliography{tex/Bibliography}

### 2020-12-16 naf2.py
- paper changes

@@ -1,671 +1,0 @@
-import os
-import pickle
-import shutil
-import time
-
-import tensorflow as tf
-from tensorflow import keras
-
-tf.keras.backend.set_floatx('float64')
-import numpy as np
-from tqdm import tqdm
-
-
-class ReplayBuffer:
-    """
-    A simple FIFO experience replay buffer for NAF_debug agents.
-    """
-
-    def __init__(self, obs_dim, act_dim, size):
-        self.obs1_buf = np.zeros([size, obs_dim], dtype=np.float64)
-        self.obs2_buf = np.zeros([size, obs_dim], dtype=np.float64)
-        self.acts_buf = np.zeros([size, act_dim], dtype=np.float64)
-        self.rews_buf = np.zeros(size, dtype=np.float64)
-        self.done_buf = np.zeros(size, dtype=np.float64)
-        self.ptr, self.size, self.max_size = 0, 0, size
-
-    def store(self, obs, act, rew, next_obs, done):
-        self.obs1_buf[self.ptr] = obs
-        self.obs2_buf[self.ptr] = next_obs
-        self.acts_buf[self.ptr] = act
-        self.rews_buf[self.ptr] = rew
-        self.done_buf[self.ptr] = done
-        self.ptr = (self.ptr + 1) % self.max_size
-        self.size = min(self.size + 1, self.max_size)
-
-    def sample_batch(self, batch_size=32):
-        if self.size < batch_size:
-            idxs = np.arange(self.size)
-        else:
-            idxs = np.random.randint(0, self.size, size=batch_size)
-
-        return dict(obs1=self.obs1_buf[idxs],
-                    obs2=self.obs2_buf[idxs],
-                    acts=self.acts_buf[idxs],
-                    rews=self.rews_buf[idxs],
-                    done=self.done_buf[idxs])
-
-    def save_to_pkl(self, name, directory):
-        idxs = np.arange(self.size)
-        buffer_data = dict(obs1=self.obs1_buf[idxs],
-                           obs2=self.obs2_buf[idxs],
-                           acts=self.acts_buf[idxs],
-                           rews=self.rews_buf[idxs],
-                           done=self.done_buf[idxs])
-        f = open(directory + name, "wb")
-        pickle.dump(buffer_data, f)
-        f.close()
-
-    def read_from_pkl(self, name, directory):
-        with open(directory + name, 'rb') as f:
-            buffer_data = pickle.load(f)
[…]

### 2020-12-16 experimental_results_AE_DYNAs.py
- paper changes

@@ -1,481 +1,0 @@
-'''This script generates the plots for the AE-DYNA tests at FERMI FEL'''
-
-import os
-import pickle
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-
-label = ["ME-TRPO", "AE-DYNA", 'Simulation'][0]
-
-if label == "ME-TRPO":
-    # ME-TRPO results
-    project_directory = 'Data_Experiments/2020_10_06_ME_TRPO_stable@FERMI/run2/'
-elif label == "AE-DYNA":
-    # AE-Dyna results
-    project_directory = 'Data_Experiments/2020_11_05_AE_Dyna@FERMI/-nr_steps_25-cr_lr-n_ep_13-m_bs_100-sim_steps_3000-m_iter_35-ensnr_3-init_200/'
-else:
-    project_directory = 'Data_logging/temp/-nr_steps_25-cr_lr-n_ep_7-m_bs_100-sim_steps_2500-m_iter_30-ensnr_5-init_100/'
-
-
-def read_rewards(rewards):
-    iterations_all = []
-    final_rews_all = []
-    mean_rews_all = []
-    stds = []
-
-    iterations = []
-    final_rews = []
-    mean_rews = []
-    for i in range(len(rewards)):
-        current_rewards = rewards[i][1:]
-        if len(current_rewards) > 0:
-            # print(current_rewards)
-            final_rews.append(rewards[i][-1])
-            iterations.append(len(current_rewards))
-            try:
-                mean_rews.append(np.sum(current_rewards))
-            except:
-                mean_rews.append([])
-            stds.append(np.std(current_rewards))
-
-    # iterations = np.mean(np.array(iterations_all), axis=0)
-    # final_rews = np.mean(np.array(final_rews_all), axis=0)
-    # mean_rews = np.mean(np.array(mean_rews_all), axis=0)
-
-    return np.array(iterations), np.array(final_rews), np.array(mean_rews), np.array(stds)
-
-
-# def plot_results(data, label='Verification', **kwargs):
-#         '''plotting'''
-#         rewards = data['rews']
-#         # iterations = []
-#         # finals = []
-#         # means = []
-#         # stds = []
-#         #
-#         # for i in range(len(rewards)):
-#         #     if (len(rewards[i]) > 1):
-#         #         finals.append(rewards[i][-1])
[…]

### 2020-12-17 main.tex
- paper changes

@@
 
-	\title{Model-Based and Model-Free Deep Reinforcement Learning in Accelerator Operation with Application to FERMI FEL}
+	\title{Model-free and Bayesian Neural Network Model-based Deep Reinforcement Learning for Particle Accelerator Control Demonstrated on the FERMI FEL}
 
@@
 	%\thanks{These two authors contributed equally}
-	% \affiliation{University of Malta}
+
 	\affiliation{IDA Lab, University of Salzburg, 5020 Salzburg, Austria}
-	\author{Niky Bruchon}%, Gianfranco Fenu, Felice Andrea Pellegrino, Erica Salvato}
+	\author{Niky Bruchon}
 	\affiliation{University of Trieste, Trieste, Italy}	
-	% \author{Gianluca Valentino}
-	% \affiliation{University of Malta, Msida, MT}
-	% \author{Marco Lonza, Giulio Gaio}
-	% \affiliation{Elettra Sincrotrone Trieste, Trieste, IT}
-	% \author{Verena Kain}
-	% %\thanks{These two authors contributed equally}
-
 	\date{\today}% It is always \today, today,
-	%  but any date may be explicitly specified
 
 	\begin{abstract}
-		Reinforcement learning holds tremendous promise in accelerator controls, and the principal goal of this paper is to show how reinforcement learning can be utilised on an operational level on accelerator physics problems.
-		We compare a model-based reinforcement to a model-free reinforcement learning method applied to the FERMI FEL system.
-In terms of representational power and sample-efficiency, the model-based approach is overtaking, while the asymptotic performance of the model-free method is slightly superior.
+		Reinforcement learning holds tremendous promise in accelerator controls.
+The primary goal of this paper is to show how this approach can be utilised on an operational level on accelerator physics problems.
+Despite the success of model-free reinforcement learning in several domains, sample-efficiency still is a bottle-neck, which might be encompassed by model-based methods.
+		We compare well-suited purely model-based to model-free reinforcement learning applied to the intensity optimisation on the FERMI FEL system.
+We find that the model-based approach demonstrates higher representational power and sample-efficiency, while the asymptotic performance of the model-free method is slightly superior.
 The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
-In both cases, the techniques were adapted to show increased noise robustness as omnipresent in accelerator control problems.
+In both cases, the algorithms were implemented in a way, which presents increased noise robustness as omnipresent in accelerator control problems.
 Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.
@@
 	In particle accelerator operation, one main goal is to provide stable and reproducible performance.
-Achieving this demands to consider several control problems simultaneously \cite{Kain2020,Scheinker2018}.
+Achieving this demands the consideration of several control problems simultaneously \cite{Kain2020,Scheinker2018}.
 Especially if there is no way to model the physics a priori, one might use optimisation techniques as, e.g.
@@
 	Recently, the community has begun to explore a novel approach to solve control problems.
-Reinforcement learning (RL) \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020} unveils several advantages over optimisation methods:
+Reinforcement learning (RL) \cite{Bruchon2020,Bruchon2019,Kain2020,Pang2020,John2020} unveils several advantages over optimisation methods:
 	\begin{itemize}
@@
 	\end{itemize}
-	One demanding aspect using RL in real world applications is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018,DulacArnold2019}.
-A second critical aspect to be considered is the robustness of the training in a noisy data limited environment.\\
+	One demanding aspect using RL in real-world applications is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018,DulacArnold2019}.
+A second critical aspect to be considered is the robustness of the training in a noisy data-limited environment.\\
 	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}.\\
-	 We successfully applied two different algorithm classes, a highly sample-efficient variant of continuous Q-learning (\cref{ss:Normalized advantage function}) and a newly proposed 
-	  model-based algorithm, which we call \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), to the FERMI FEL problem.\\ The methods reveal different characteristics addressing the stated critical aspects.
+	We successfully applied two different algorithm classes, a highly sample-efficient variant of continuous Q-learning (\cref{ss:Normalized advantage function}) and a newly proposed model-based algorithm, which we call \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), to the FERMI FEL problem.\\ The methods reveal different characteristics addressing the stated critical aspects.
 They are generally applicable to problems showing a sample-complexity as in common accelerator control problems.
 	\subsection{An Overview of the Main Results}
-	 We demonstrate optimization techniques that simultaneously adjust 4 parameters, the tilt and incline of 2 mirrors that control the trajectory and overlap of the seed laser in the FERMI FEL relative to the beam in the undulator.
-We show that our algorithms require significantly less than 1000 training sets, after which they are able to maximize the FERMI FEL’s output intensity to $>$ that 95\% of the maximum within only $\sim 3$ to $\sim 5$ steps depending on which algorithm is used.
-Also we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\
-	  Furthermore, our algorithms can perform real-time learning based on data and therefore can be re-run and re-trained for time-varying systems.\\
+	We demonstrate optimisation techniques that simultaneously adjust four parameters, the tilt and incline of two mirrors that control the trajectory and overlap of the seed laser in the FERMI FEL relative to the beam in the undulator.
+%	We show that our algorithms require significantly less than 1000 training sets, after which they can maximise the FERMI FEL's output intensity to $>$ that 95\% of the maximum within only $\sim 3$ to $\sim 5$ steps depending on which algorithm is used.
+	By making use of model ensembles and Bayesian neural networks our algorithms exhibit a high sample-efficiency (learning using $<1000$ data sets), a low model-bias, and are extremely fast as they are able to maximize the FERMI FEL's output intensity to greater than 95\% of the maximum within only 3-5 steps.
+	 Also, we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\
+	Furthermore, our algorithms can perform real-time learning based on data and therefore, can be re-run and re-trained for time-varying systems.\\
 	The paper is organised as follows:
@@
 	In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator.\\
-%	At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation. 
-In the last years, various approaches have been studied to investigate their applicability in tuning operations. \\
+	%	At FERMI, several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the primary purpose of guaranteeing a steady intensity of the light radiation. 
+	In the last years, various approaches have been studied to investigate their applicability in machine tuning. \\
 	A free-electron laser is a fourth-generation light source where the lasing medium consists of a very-high-speed electron moving freely through a magnetic structure.
-By using an external seeding source, the FERMI FEL has several advantages relative to standard FEL approaches in terms of increased stability in pulse and photon energy, reduced size, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.\\
+By using an external seeding source, the FERMI FEL has several advantages relative to standard FEL approaches in terms of increased stability in pulse and photon energy, reduced size, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronisation with external timing systems.\\
 	The external optical laser signal interacts with the relativistic electron beam in the undulator, introducing an energy modulation that aids in the FEL process.
@@
 	\end{figure}
-%	The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
+	%	The most critical parameters in a seeded free-electron laser is the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
 	%	The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
@@
 	Two mirrors, TT1 and TT2, are used to control the trajectory of the laser by tilting and inclining, which gives a total of four degrees of freedom (DOF).
-In turn, the laser overlaps with the electron beam between the two screens, CCD1 and CCD2.
+In turn, the laser overlaps with the electron beam between the two removable screens, CCD1 and CCD2.
 Lastly, the monitor measures the intensity, $I_0$.\\
@@
 	\end{align} 
-	and is limited by an $ |\ba_t|\leq \ba_\text{max}=\frac{1}{12}$.
-	The training is done in episodes with a uniformly\footnote{in the normalized state-space} randomized initial state and a specific maximal length of taken steps afterwards - the horizon.
-For reason discussed later, the horizon varies in the applications between either 10 or 500 steps.
+	and is limited by an $ |\ba_t|\leq \ba_\text{max}=1/12$.
+	The training is done in episodes with a uniformly\footnote{in the normalised state-space} randomised initial state and a specific maximal length of taken steps afterwards - the horizon.
+For a reason discussed later, the horizon varies in the applications between either 10 or 500 steps.
 An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system.
-In our experiments the maximal intensity was determined by a manual expert optimization.
+In our experiments, the maximal intensity was determined by manual expert optimisation.
 \\
-	 During the training phase exploration and the training of the function approximator takes place.
-	After a defined number of training episodes, a verification of 50 or 100 episodes is performed. In the second phase, the RL only applies the learned policy.
+	During the training phase exploration and the training of the function, approximator takes place.
+	After a defined number of training episodes, verification of 50 or 100 episodes is performed. In the second phase, the RL only applies the learned policy.
 
@@
 	\section{Deep Reinforcement learning}
-    Most commonly two main categories of deep RL algorithms are distinguished: model-free RL (MFRL) and model-based RL (MBRL).
-MFRL algorithms learn directly from the interaction with a system, while MBRL algorithm learn a dynamics model from these interactions first.
-Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, there applications is mainly restricted to simulated environments due to their low sample-efficiency.
-Yet putting constrains on the algorithms, as discussed in \cref{s:Model free reinforcement learning}, partly cures the sample inefficiency turning them into applicable tools for real world problems.\\
-      On the other hand MBRL showing a much better sample efficiency.
-The challenge for these methods is to learn an accurate model, which turns out to be rather hard and leads to the so called model-bias problem \cite{Deisenroth2011}.
+	Most commonly, two main categories of deep RL algorithms are distinguished: model-free RL (MFRL) and model-based RL (MBRL).
+MFRL algorithms learn directly from the interaction with a system, while the MBRL algorithm learns a dynamics model from these interactions first.
+Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, their applications is mainly restricted to simulated environments due to their low sample-efficiency.
+However, putting constraints on the algorithms, as discussed in \cref{s:Model-free reinforcement learning}, partly cures the sample inefficiency, turning them into practical tools for real-world problems.\\
+	On the other hand, MBRL showing a much better sample efficiency.
+The challenge for these methods is to learn an accurate model, which turns out to be rather hard and leads to the so-called model-bias problem \cite{Deisenroth2011}.
 In \cref{ss:Uncertainty aware DYNA-style reinforcement learning} we discuss attempts to capture the model uncertainty to alleviate this obstacle by making use of model ensembles and Bayesian neural networks resulting in algorithms, exhibiting a high sample-efficiency and a low model-bias.\\
@@
 
-	\subsection{Model free reinforcement learning}\label{s:Model free reinforcement learning}
+	\subsection{Model free reinforcement learning}\label{s:Model-free reinforcement learning}
 	In the modern field of RL one differentiates if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba) \approx Q_\theta(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
@@
 Hence, we rule them out for direct online training but make use of their characteristics in MBRL (\cref{ss:critical_design}).\\
-	 In the second case we speak about Q-learning, which belongs to approximate dynamic programming.
-One advantage is the possible usage of use previously stored data as we use it in an \emph{off-policy} fashion, which generally needs less data to be trained than \emph{on-policy} methods.
+	In the second case, we speak about Q-learning, which belongs to approximate dynamic programming.
+One advantage is the possible usage of use previously stored data as we use it in an \emph{off-policy} fashion, which generally needs fewer data to be trained than \emph{on-policy} methods.
 A combination of both is also used, known as actor-critic methods \cite{Szepesvari2010,Lillicrap2015,Silver2014}.\\
 	As already adduced, MFRL algorithms learn directly from interactions with the system.
-An exhaustive overview can be found e.g.
+An exhaustive overview can be found, e.g.
 in \cite{Sutton2018,Levine2020}.
 In the following, we discuss the normalised advantage function (\emph{NAF}) algorithm \cite{Gu2016} in some detail, as it has good characteristics for accelerator control problems.
-It is highly sample-efficiency  and its representational power is sufficient for most common cases\cite{Kain2020,Hirlaender2020a}.
+It is highly sample-efficiency, and its representational power is sufficient for most common cases\cite{Kain2020,Hirlaender2020a}.
 Modifications to increase the stability are subject of \cref{ss:Normalized advantage function}.
@@
 Therefore \emph{smoothing} was applied in all MFRL experiments.\\
-	 Our implementation has a high sample-efficiency and noise resistivity.
+	Our implementation has high sample-efficiency and noise resistivity.
 We use it as the baseline for the FERMI-FEL control problem, as it yields good results.
@@
 
-	A nice overview of state of the art MBRL methods is provided in \cite{Wang2019}.
-Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample based methods (e.g.
+	An excellent overview of the state of the art MBRL methods is provided in \cite{Wang2019}.
+Suitable methods are \cite{Gal2016,6654139}, based on Gaussian processes and Bayesian Neural Networks and back-propagation in the dynamics model or \cite{Chua2018,Wang2019a}, using highly efficient sample-based methods (e.g.
 the cross entropy method \cite{Boer2005}).
@@
 In addition to the next state, the reward is included in the model \footnote{In some applications $\reward$ can be deduced from the current state or the state change.
-In such a situation the modelling can be simplified to model directly $\hat\transitions$}:
+In such a situation, the modelling can be simplified to model directly $\hat\transitions$}:
 	\begin{align}
@@
 	Subsequently, an MFRL algorithm is deployed to learn the controller on the dynamics model by only using the synthetic data and interoperating the model uncertainty, as explained in \cref{ss:The uncertainty aware dynamics model}.\\
-	  After a defined number of training steps, called epoch, the controller is validated on each model of the ensemble, individually.
+	After a defined number of training steps, called an epoch, the controller is validated on each model of the ensemble, individually.
 If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes.
-In this way 'over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}.
+In this way, 'over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}.
 The training is stopped if the controller produces satisfying results on the real environment.
-Otherwise, a batch of new data is collected and the model is improved and consecutively a new epoch starts.
+Otherwise, a batch of new data is collected, and the model is improved, and consecutively a new epoch starts.
 	\begin{figure}[!h]
@@
 	The prior is controlled by the initialisation of the weights of the network.
-It defines how large the variation of the approximated function is expected and were randomized uniformly in a small interval $[-\Delta, \Delta]$.
+It defines how large the variation of the approximated function is expected and they were randomised uniformly in an interval $[-\Delta, \Delta]$, where $|\Delta|\leq0.1$.
 The weights of the last layer were normalised by the number of its nodes.
-For the moment, only homoskedastic Gaussian errors are considered with a standard deviation is $\sigma_\epsilon$ is also respected in a regularisation term \cref{appendix:The impact of noise} and significantly improves the training in the presence of noise, as shown in \cref{fig:comparsion_noise_ae_dyna}.
+For the moment, only homoskedastic Gaussian errors are considered with a standard deviation $\sigma_\epsilon$.
+This is respected in a regularisation term (details see \cref{appendix:The impact of noise}) and significantly improves the training in the presence of noise, as shown in \cref{fig:comparsion_noise_ae_dyna}.
 \\
@@
 	\subsubsection{The controller algorithm}
-To decide which MFRL algorithm to use, two already mentioned main algorithm classes are considered  (\cref{s:Model free reinforcement learning}): \emph{on-} and  \emph{off-policy} algorithms \cite{Sutton2018}.
+	To decide which MFRL algorithm to use, two already mentioned main algorithm classes are considered  (\cref{s:Model-free reinforcement learning}): \emph{on-} and  \emph{off-policy} algorithms \cite{Sutton2018}.
 \emph{On-policy} algorithm show a stabler convergence in general, while \emph{off-policy} algorithms have the advantage that the data buffer can be filled with real data.\\
-The \emph{on-policy} algorithm trust region policy optimization (TRPO) \cite{Schulman2015} provides the theoretical guarantee of a monotonic improvement of the policy.
+	The \emph{on-policy} algorithm trust region policy optimisation (TRPO) \cite{Schulman2015} provides the theoretical guarantee of a monotonic improvement of the policy.
 We use this algorithm in experiments labelled as \emph{ME-TRPO}, which stands for model-ensemble TRPO as proposed in \cite{Kurutach2018}.
@@
 \\
-An attractive off-policy algorithm is the soft-actor-critic (\emph{SAC}) algorithm.
-\cite{fujimoto2018addressing,Hill2018}.
+	An attractive \emph{off-policy} algorithm is the soft-actor-critic (\emph{SAC}) \cite{fujimoto2018addressing,Hill2018}.
 \emph{SAC} not only tries to maximise \cref{eq:cumulative_reward} but also simultaneously is regularised to maximise the entropy in the action space \cite{Haarnoja2018a}.
-In this way, exploration is encouraged to avoid getting stuck in a local optimum and a good trade-off between exploration and exploitation is achieved.
+In this way, exploration is encouraged to avoid getting stuck in a local optimum, and a good trade-off between exploration and exploitation is achieved.
 We refer to experiments using the \emph{SAC} as \emph{AE-DYNA-SAC}.
-For \emph{AE-DYNA-SAC} the controller was reset after each retraining of the dynamics model, to profit from its exploration features, while for \emph{ME-TRPO} the policy was constantly improved to exploit its monotony.
+	For \emph{AE-DYNA-SAC} the controller was reset after each re-training of the dynamics model, to profit from its exploration features, while for \emph{ME-TRPO} the policy was improved continuously to exploit its monotony.
 	\subsubsection{Handling of the uncertainty}
@@
 Another strategy is to randomly select a model to provide at each training step, which demonstrated the best performance the original implementation of the model ensemble \emph{TRPO} (\emph{ME-TRPO}) algorithm and was used in the experiments labelled \emph{ME-TRPO} \cite{Kurutach2018}.
-A pessimistic selection would only select the model resulting in the lowest predicted reward (as used e.g.
+A pessimistic selection would only select the model resulting in the lowest predicted reward (as used, e.g.
 in \cite{kidambi2020morel}) and showed constant but too slow improvement for our online-training.
@@
 	As discussed in the previous sections, several tests were performed on the FERMI FEL.
-	The main purpose was to try the newly implemented algorithms on a real system to evaluate their operational feasibility. 
+	The primary purpose was to try the newly implemented algorithms on a real system to evaluate their operational feasibility. 
 	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments.\\ 
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
-		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
+		\caption{The training of different variants of the \emph{NAF2} algorithm  on the FERMI FEL, averaged over two complete trainings (the standard-deviations are indicated by the shaded areas).
+The number of iterations (blue) shows the steps until the intensity is optimised, starting from a random initial position.}
 		\label{fig:NAF_training}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Evolution_double.pdf}
-		\caption{The training of the \emph{NAF2} on the FERMI FEL using a double network.}
+		\caption{The evolution of the states of the \emph{NAF2} algorithm on the FERMI FEL using a double network during the training.}
 		\label{fig:NAF_evolution_double}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Evolution_single.pdf}
-	\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network.}
+		\caption{The evolution of the states of the \emph{NAF2} algorithm on the FERMI FEL using a single network during the training.}
 		\label{fig:NAF_evolution_single}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
-		\caption{The verification of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
+		\caption{The verification episodes of the variants of the trained model-free \emph{NAF2} algorithm on the FERMI FEL.
+The number of iterations (blue) shows the steps until the intensity is optimised, starting from a random initial position.}
 		\label{fig:NAF_verification}
@@
 The number of iterations per episode is plotted, including the cumulative number of steps.
-In all episodes during the training the 95\% intensity reward threshold was surpassed, and therewith they were successfully finished.
+In all episodes, during the training, the 95\% intensity reward threshold was surpassed, and therewith they were successfully finished.
 The evolution of the states is provided in \cref{fig:NAF_evolution_single,fig:NAF_evolution_double} for the single and double network.
@@
 In the single-network case, convergence is reached after around 1400 steps.
-In the plot $V$ was averaged over 200 randomly selected states.
+In the plot, $V$ was averaged over 200 randomly selected states.
 	\subsection{MBRL tests}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
-		\caption{The training of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
+		\caption{The training observables of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
 		\label{fig:AE-DYNA_observables}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Verification_DYNA_all_episodes.pdf}
-		\caption{The verification of the \emph{ME-TRPO} and the \emph{AE-DYNA-SAC} on the FERMI FEL.}
+		\caption{The verification episodes of the trained model-based methods: \emph{ME-TRPO} and \emph{AE-DYNA-SAC} on the FERMI FEL.
+The number of iterations (blue) shows the steps until the intensity is optimised, starting from a random initial position.}
 		\label{fig:AE-DYNA_verification}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
-		\caption{The training of the ME-TRPO on the FERMI FEL.}
+		\caption{The training observables of the \emph{ME-TRPO} on the FERMI FEL.}
 		\label{fig:ME-TRPO_observables}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Worst_episode_MBRL.pdf}
-	\caption{The verification of the \emph{ME-TRPO} and the \emph{AE-DYNA-SAC} on the FERMI FEL.}
+		\caption{The evolution of the states during the worst verification episodes of the trained \emph{ME-TRPO} and the \emph{AE-DYNA-SAC} on the FERMI FEL.}
 		\label{fig:Worst_episode_MBRL}
@@
 	\\ To exploit the convergence properties of the \emph{TRPO}, the policy was never reset as can be seen in \cref{fig:ME-TRPO_observables}.
-In the upper figure, the total reward per batch gathered from the real environment and the number of data-points used to train the dynamics model as a function of the number of epochs are shown.
-In the lower plot, the average cumulative reward of ten episodes as achieved by the \emph{TRPO} on the individual models of the ensemble in dependence of the epochs is drawn.
+In the upper figure, the total reward per batch gathered from the real environment, and the number of data-points used to train the dynamics model as a function of the number of epochs is shown.
+In the lower plot, the average cumulative reward of ten episodes as achieved by the \emph{TRPO} on the individual models of the ensemble independence of the epochs is drawn.
 During an epoch the \emph{TRPO} is trained for 10000 steps on the synthetic data.
@@
 	Secondly, the \emph{AE-DYNA-SAC} was tested.
-As discussed in \cref{ss:critical_design}, in this test, the controller was reset each time, when the model was retrained and consequently, the performance drops each time as shown in \cref{fig:AE-DYNA_observables}.\\
+As discussed in \cref{ss:critical_design}, in this test, the controller was reset each time, when the model was re-trained and consequently, the performance drops each time as shown in \cref{fig:AE-DYNA_observables}.\\
 	In contrast to the \emph{ME-TRPO} training, the data batches consisted of 50 data-points with an initial random walk of 200 steps.
-The number of initial steps was chosen high enough because otherwise the convergence is slowed down strongly, so that the training becomes unfeasible on a real machine (discussed in \cref{ss:critical_design}).
+The number of initial steps was chosen high enough because otherwise the convergence is slowed down strongly so that the training becomes unfeasible on a real machine (discussed in \cref{ss:critical_design}).
 Each epoch consists of 2500 steps of controller training on the model.
@@
 The verification was executed as in the first test.
-Again, the success of all 50 episodes is 100\%, but the number of needed iterations per episodes is less than for the ME-TRPO, wich can be seen in \cref{fig:AE-DYNA_verification}.
+Again, the success of all 50 episodes is 100\%.
+The number of needed iterations per episodes is less than for the ME-TRPO, which can be seen in \cref{fig:AE-DYNA_verification}.
 The somewhat better performance might be a result of the higher number of data-points (50), but in general, this method exhibited better asymptotic performance than the \emph{ME-TRPO} variant.\\
-	In both experiments the training was using a small number of data-points while still successfully solving the problem.
+	In both experiments, the training was using a small number of data-points while still successfully solving the problem.
 In \cref{fig:Worst_episode_MBRL} the worst verification episodes of both experiments are plotted.
 The experiments, taken at different days, may influence the boundary conditions and hence the performance.
-However, a longer training would increase the performance in both cases, especially for the \emph{ME-TRPO}.
+However, longer training would increase the performance in both cases, especially for the \emph{ME-TRPO}.
 \Cref{fig:Worst_episode_MBRL} shows that the tt2 incline and tilt parameters for the \emph{ME-TRPO} move up and then down towards the optimal position, which would be improved by a better dynamics model.
 	\section{Discussion and outlook}
-	In this paper, the applicability of the method of deep reinforcement learning on the optimization of the FERMI FEL intensity was presented.
-Two different approaches were tested: a model free and a model based method.\\
-	 Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasibly number of training steps.
-The results of the verification is summarised in \cref{tab:overview_verification}.
+	In this paper, the applicability of deep reinforcement learning on the optimisation of the FERMI FEL intensity was presented.
+Two different approaches were tested: a model-free and a model-based method.\\
+	Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasible number of training steps.
+The results of the verification are summarised in \cref{tab:overview_verification}.
 In the model-free case, around 800-1000 and in the model-based 450-500 data-points were used.
-In all experiments all verification episodes were finished successfully.
+In all experiments, all verification episodes were finished successfully.
 \emph{NAF2} performs better in terms of episode length while the average reward in the \emph{DYNA}-style algorithms is higher.
 The experiments were done on different days, hence under slightly different conditions.\\
-	Usage of the proposed methods in an operational way is attractive and would replace the current optimization method using the destructive measurement of the screens and additionally would save valuable time.\\
-	The MFRL methods were slightly better in the final performance but also more samples have been collected during the training.
-The MBRL methods had to be stopped due to lack of available experimental time, hence additional studies regarding a long time performance would be interesting.\\ One big issue in applying MBRL methods can be the computational time to train the controller on the model.
-In our tests the data acquisition time was only a fraction of the time needed for the controller-training.
-In case, the used methods could be parallelized to reduce the computational time.
+	Usage of the proposed methods in an operational way is attractive and could replace in the future the current optimisation method, which needs the destructive screen measurement.
+Hence an online retraining could be done and sacred time could be saved.\\
+	The MFRL methods were slightly better in the final performance, but also more samples have been collected during the training.
+The MBRL methods had to be stopped due to lack of available experimental time.
+Hence additional studies regarding a long time performance would be interesting.\\ One big issue in applying MBRL methods can be the computational time to train the controller on the model.
+In our tests, the data acquisition time was only a fraction of the time needed for the controller-training.
+In case, the used methods could be parallelised to reduce the computational time.
 	\begin{table}[h]%The best place to locate the table environment is directly after its first reference in text
 		\caption{\label{tab:overview_verification}%
-			An overview over the verification results of the different algorithms.
+			An overview over the verification performance of the different trained algorithms on the FERMI FEL including their standard deviation.
 		}
@@
 				%\toprule
-				&  \centering Data points&   \centering Episode length&   \centering Cumulative reward &   \centering Final\\reward\tabularnewline
+				&  \centering Data points (counts)&   \centering Episode length (counts)&   \centering Cumulative reward\newline (arb. units)&   \centering Final\\reward \newline (arb. units)\tabularnewline
 				\hline
@@
 	\section{Acknowledgements}
- We want to thank Giulio Gaio, who made this experiments possible and helped to setup and restore the FERMI FEL.
+	We want to thank Giulio Gaio, who made these experiments possible and helped to set up and restore the FERMI FEL.
 Also, we are thankful for the valuable feedback from Alexander Scheinker on the draft of this publication.
-Finally we want to thank the Land Salzburg for their financial support to create the context of the IDA lab, which made this work possible.
+Finally, we want to thank the Land Salzburg for their financial support to create the context of the IDA lab, which made this work possible.
 
@@
 The average value and the standard deviation (shaded) are plotted.
-For this study, we assume that the problem is successfully solved, if the cumulative reward surpasses a threshold of -200.
-The threshold is indicated by a dashed green line in the corresponding figures.
-%	  and a rolling average of 15 steps is used for a better visualization.
+For this study, we assume that the problem is successfully solved if the cumulative reward surpasses a threshold of -200.
+A dashed green line indicates the threshold in the corresponding figures.
+	%	  and a rolling average of 15 steps is used for better visualisation.
 
@@
 One sees the cumulative reward per episode for a training of a total of 100 episodes.
-As mentioned, the curve labelled \emph{clipping} corresponds to the double network including \emph{smoothing} and shows the best overall stability during the training yielding quickly a high reward.
-Also the \emph{smoothed} single network, labelled \emph{No-clipping-smoothing}, shows good and comparable performance, except for the sightly decreased stability.
-The worst performance is achieved without smoothing and a single network (\emph{No-clipping-no-smoothing}), nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
+As mentioned, the curve labelled \emph{clipping} corresponds to the double network, including \emph{smoothing} and shows the best overall stability during the training yielding a high reward quickly.
+Also, the \emph{smoothed} single network, labelled \emph{No-clipping-smoothing}, shows good and comparable performance, except for the slightly decreased stability.
+The worst performance is achieved without smoothing and a single network (\emph{No-clipping-no-smoothing}), nevertheless the result is competing with state of the art model-free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
 	%	\begin{figure}[!h]
@@
 	The lower plot of \cref{fig:comparsion_noise_ae_dyna} shows the batch rewards a measured during the data collection, which is observable during the training.
-\begin{figure}
+	\begin{figure}[h]
 		\centering
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise_ae_dyna}
-	\caption{Cumulative reward of \emph{AE-DYNA-SAC} on the \emph{inverted pendulum} with artificial noise using the anchor.}
+		\caption{Cumulative reward of \emph{AE-DYNA-SAC} on the \emph{inverted pendulum} with artificial noise using the 'anchor ensembling'.}
 		\label{fig:comparsion_noise_ae_dyna}
@@
 A single network might not converge or too slowly (label \emph{Single}) and a model of ten shown fast and stable performance in this case (label \emph{Ten}).
-\begin{figure}
+	\begin{figure}[h]
 		\centering
@@
 	\subsection{NAF versus AE-DYNA}
-	Finally, \cref{fig:comparsion_NAF_AE-DYNA} demonstrates the sample-efficiency of the \emph{AE-DYNA-SAC} and the \emph{NAF} algorithm on the noisy \emph{inverted pendulum}.
-\emph{AE-DYNA-SAC} converges below 2000 data-points (without noise even below 800) and the \emph{NAF2} starts to perform equally 10000 data-points surpassing the -200 reward threshold.
-One clearly sees the increased sample-efficiency on this problem using the \emph{AE-DYNA} in contrast to the \emph{NAF2}.
-\begin{figure}
+		\begin{figure}[h]
 		\centering
@@
 
+	Finally, \cref{fig:comparsion_NAF_AE-DYNA} demonstrates the sample-efficiency of the \emph{AE-DYNA-SAC} and the \emph{NAF} algorithm on the noisy \emph{inverted pendulum}.
+\emph{AE-DYNA-SAC} converges below 2000 data-points (without noise even below 800) and the \emph{NAF2} starts to perform equally 10000 data-points surpassing the -200 reward threshold.
+One clearly sees the increased sample-efficiency on this problem using the \emph{AE-DYNA} in contrast to the \emph{NAF2}.
+
 	%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
-%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
+	%	In this section, we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
 	%	One issue of MBRL methods is asymptotic performance.

### 2020-12-17 run_naf2.py

@@ -262,6 +262,7 @@ if __name__ == '__main__':
 
     # We normalize in a MonitoringEnv state action and reward to [-1,1] for the agent and plot results
     env = MonitoringEnv(env=PendulumEnv(), plot_progress=True)
+    # If you want a video:
     env = gym.wrappers.Monitor(env, "recording2", force=True, video_callable=lambda episode_id: episode_id%10==0)
 
     nafnet_kwargs = dict(hidden_sizes=[100, 100], activation=tf.nn.tanh

### 2020-12-17 main.tex
- paper changes

@@
 \documentclass[
-reprint,
+reprint,nofootinbib,
 %superscriptaddress,
@@
 
-	\title{Model-free and Bayesian Neural Network Model-based Deep Reinforcement Learning for Particle Accelerator Control Demonstrated on the FERMI FEL}
+	\title{Model-free and Bayesian Ensembling Model-based Deep Reinforcement Learning for Particle Accelerator Control Demonstrated on the FERMI FEL}
 
@@
 	\affiliation{IDA Lab, University of Salzburg, 5020 Salzburg, Austria}
-	\author{Niky Bruchon}
-	\affiliation{University of Trieste, Trieste, Italy}	
+	\author{Niky Bruchon}\email{niky.bruchon@phd.units.it}
+	\affiliation{University of Trieste, 34127 Trieste, Italy}	
 	\date{\today}% It is always \today, today,
@@
 A second critical aspect to be considered is the robustness of the training in a noisy data-limited environment.\\
-	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}.\\
+	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}.
+\Cref{fig:elletra_research} shows an aerial
+	overview of the Elettra site and the FERMI free-electron laser buildings.\\
 	We successfully applied two different algorithm classes, a highly sample-efficient variant of continuous Q-learning (\cref{ss:Normalized advantage function}) and a newly proposed model-based algorithm, which we call \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), to the FERMI FEL problem.\\ The methods reveal different characteristics addressing the stated critical aspects.
 They are generally applicable to problems showing a sample-complexity as in common accelerator control problems.
+	\begin{figure}[h]
+		\centering
+		\includegraphics*[width=0.45\textwidth]{Figures/FERMI_AerialView}
+		\caption{The Elettra research centre hosting the FERMI free electron laser \cite{Brochon2020}.}
+		\label{fig:elletra_research}
+	\end{figure}
 	\subsection{An Overview of the Main Results}
@@
 	By making use of model ensembles and Bayesian neural networks our algorithms exhibit a high sample-efficiency (learning using $<1000$ data sets), a low model-bias, and are extremely fast as they are able to maximize the FERMI FEL's output intensity to greater than 95\% of the maximum within only 3-5 steps.
-	 Also, we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\
+	 Thus, we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\
 	Furthermore, our algorithms can perform real-time learning based on data and therefore, can be re-run and re-trained for time-varying systems.\\
@@
 	\section{The set-up of the studied problem}
-	\subsection{The physical set-up}
+	\subsection{The Physical Set-up}
 	In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator.\\
@@
 	%	For this reason, the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
-	\subsection{The training environment}
+	\subsection{The Training Environment}
 	\begin{figure}
@@
 	\end{align} 
-	and is limited by an $ |\ba_t|\leq \ba_\text{max}=1/12$.
-	The training is done in episodes with a uniformly\footnote{in the normalised state-space} randomised initial state and a specific maximal length of taken steps afterwards - the horizon.
+	and is limited by an $ |\ba_t|\leq \ba_\text{max}=1/12$\footnote{The value was determined experimentally to guarantee reproducibility.}.
+	The training is done in episodes with a uniformly\footnote{Sampled in the normalised state-space.} randomised initial state and a specific maximal length of taken steps afterwards - the horizon.
 For a reason discussed later, the horizon varies in the applications between either 10 or 500 steps.
 An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system.
-In our experiments, the maximal intensity was determined by manual expert optimisation.
+In our experiments, the maximal intensity was determined through manual expert optimisation.
 \\
@@
 	Most commonly, two main categories of deep RL algorithms are distinguished: model-free RL (MFRL) and model-based RL (MBRL).
-MFRL algorithms learn directly from the interaction with a system, while the MBRL algorithm learns a dynamics model from these interactions first.
+MFRL algorithms learn directly from the interaction with a system, while MBRL algorithms learn a dynamics model from these interactions first.
 Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, their applications is mainly restricted to simulated environments due to their low sample-efficiency.
@@
 
-	\subsection{Model free reinforcement learning}\label{s:Model-free reinforcement learning}
+	\subsection{Model-free Reinforcement Learning}\label{s:Model-free reinforcement learning}
 	In the modern field of RL one differentiates if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba) \approx Q_\theta(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g.
@@
 
-	\subsection{Uncertainty aware DYNA-style reinforcement learning}\label{ss:Uncertainty aware DYNA-style reinforcement learning}
+	\subsection{Uncertainty Aware DYNA-style Reinforcement Learning}\label{ss:Uncertainty aware DYNA-style reinforcement learning}
 	%The original DYNA algorithm \cite{Kurutach2018} is modified here in several aspects.
@@
 In addition to the next state, the reward is included in the model \footnote{In some applications $\reward$ can be deduced from the current state or the state change.
-In such a situation, the modelling can be simplified to model directly $\hat\transitions$}:
+In such a situation, the modelling can be simplified to model directly $\hat\transitions$.}:
 	\begin{align}
@@
 Initially, data is gathered by interaction with the real environment or read from a previous collection.
-An uncertainty aware model of the dynamics is trained, using 'anchored ensembling' \cite{Pearce2018} on the data (details provided in \cref{appendix:The impact of noise,ss:The uncertainty aware dynamics model}), which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
+An uncertainty aware model of the dynamics is trained, using `anchored ensembling' \cite{Pearce2018} on the data (details provided in \cref{appendix:The impact of noise,ss:The uncertainty aware dynamics model}), which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account.
 It was motivated by the fact that pure epistemic ensemble techniques as \cite{Kurutach2018} were reported to be very sensitive to aleatoric noise \cite{Wang2019}.
@@
 If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes.
-In this way, 'over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}.
+In this way, `over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}.
 The training is stopped if the controller produces satisfying results on the real environment.
@@
 	\end{figure}
-	\subsection{Critical design decisions in MBRL}\label{ss:critical_design}
+	\subsection{Critical Design Decisions in MBRL}\label{ss:critical_design}
 	In the following, the most important aspects of a successful application of MBRL are discussed.
@@
 	\subsubsection{The uncertainty aware dynamics model}\label{ss:The uncertainty aware dynamics model}
-	The 'anchored ensembling' method usually yields good results already with a small number of models.
+	The `anchored ensembling' method usually yields good results already with a small number of models.
 Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes}).
@@
 Afterwards, from our experience, at least one full episode per epoch should be taken.
-We decided to use a short horizon to diminish the impact of the compound error \cite{Janner2019}, the accumulation error following a wrong model, as well known in DYNA-style approaches.
+We decided to use a short horizon to diminish the impact of the compound error \cite{Janner2019}, the accumulation error following a wrong model, as well known in \emph{DYNA}-style approaches.
 The maximal number of steps in our experimental runs per episode was ten.
@@
 In \cref{sec:Experimental results from FERMI RL online tests} the \emph{NAF2},
-	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment.
+	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment. The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
 	\begin{table}[h!]%The best place to locate the table environment is directly after its first reference in text
@@
 	\end{table}
-	The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
-	\subsection{MFRL tests}
+
+	\subsection{MFRL Tests}
 	\begin{figure}
@@
 	\end{figure}
-
 	\begin{figure}
@@
 In the plot, $V$ was averaged over 200 randomly selected states.
-	\subsection{MBRL tests}
+	\subsection{MBRL Tests}
 	\begin{figure}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
-		\caption{The training observables of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
+		\caption{The training observables of the \emph{AE-DYNA-SAC} on the FERMI FEL. Detail are provided in the text.}
 		\label{fig:AE-DYNA_observables}
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
-		\caption{The training observables of the \emph{ME-TRPO} on the FERMI FEL.}
+		\caption{The training observables of the \emph{ME-TRPO} on the FERMI FEL. Detail are provided in the text.}
 		\label{fig:ME-TRPO_observables}
@@
 	\section{Discussion and outlook}
-	In this paper, the applicability of deep reinforcement learning on the optimisation of the FERMI FEL intensity was presented.
-Two different approaches were tested: a model-free and a model-based method.\\
+	In this paper, the applicability of deep reinforcement learning on the optimisation of the FERMI FEL intensity was presented. Two different approaches were tested: model-free and model-based.\\
 	Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasible number of training steps.
@@
 	Usage of the proposed methods in an operational way is attractive and could replace in the future the current optimisation method, which needs the destructive screen measurement.
-Hence an online retraining could be done and sacred time could be saved.\\
+Hence an online retraining could be done and valuable time could be saved.\\
 	The MFRL methods were slightly better in the final performance, but also more samples have been collected during the training.
@@
 In case, the used methods could be parallelised to reduce the computational time.
-	\begin{table}[h]%The best place to locate the table environment is directly after its first reference in text
+	\begin{table}[!h]%The best place to locate the table environment is directly after its first reference in text
 		\caption{\label{tab:overview_verification}%
@@
 	The presented reinforcement learning methods hold tremendous promise for automatizing set-ups typical in accelerators.
-One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real set-up by adapting available methods.\\
-	Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA-style} algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency appealing choices.\\
+One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real set-up by adapting available methods.
+This was demonstrated on the FERMI FEL intensity optimization problem.\\
+	Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA}-style algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency appealing choices.\\
 	Complementary, the \emph{NAF2} algorithm presents a good alternative, revealing, as a model-free method, good asymptotic performance.
 As numerous cases in accelerator control can be captured assuming a quadratic dependence on the actions of the state-action value function, there is a broad spectrum of potential applications.\\
-	To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
+	To provide the possibility for other laboratories to profit form the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
 
@@
 Also, we are thankful for the valuable feedback from Alexander Scheinker on the draft of this publication.
-Finally, we want to thank the Land Salzburg for their financial support to create the context of the IDA lab, which made this work possible.
+Finally, we want to thank the Land Salzburg for their financial support to create the IDA Lab, which made this work possible.
 
 	\appendix
-	\section{A non-linear standard control problem}
+	\section{A Non-linear Standard Control Problem}
 	To provide some transparency of these studies for other labs, we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.\\
@@
 
-	\subsection{NAF2 details}\label{appendix:naf2}
+	\subsection{NAF2 Details}\label{appendix:naf2}
 	%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
@@
 	% \tp{possible to just state equations you know, save space}
-	\subsection{The impact of noise}\label{appendix:The impact of noise}
+	\subsection{The Impact of Noise}\label{appendix:The impact of noise}
 	A test adding large artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma_\epsilon)$ with $\sigma_\epsilon=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}.
@@
 	\end{figure}
-	\subsubsection{Regression assuming homoskedastic Gaussian noise using `Anchored ensembling'}
+	\subsubsection{Regression assuming homoskedastic Gaussian noise using `anchored ensembling'}
 	The Bayesian community and the RL community is spending more and more attention to
-	new approaches to Bayesian inference. In the `Anchored ensembling' technique a regularisation
+	new approaches to Bayesian inference. In the `anchored ensembling' technique a regularisation
 	term is added to the loss function, which returns a point estimate of
@@
 		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise_ae_dyna}
-		\caption{Cumulative reward of \emph{AE-DYNA-SAC} on the \emph{inverted pendulum} with artificial noise using the 'anchor ensembling'.}
+		\caption{Cumulative reward of \emph{AE-DYNA-SAC} on the \emph{inverted pendulum} with artificial noise using the `anchor ensembling'.}
 		\label{fig:comparsion_noise_ae_dyna}

### 2025-12-29 main.tex
docs: Add `requirements.txt` and enhance `README.md` with installation instructions and improved code descriptions.

@@
 	\begin{abstract}
-		Reinforcement learning holds tremendous promise in accelerator controls.
-The primary goal of this paper is to show how this approach can be utilised on an operational level on accelerator physics problems.
-Despite the success of model-free reinforcement learning in several domains, sample-efficiency still is a bottle-neck, which might be encompassed by model-based methods.
+		Reinforcement learning holds significant potential in accelerator controls.
+The primary goal of this paper is to demonstrate the operational application of this approach to accelerator physics problems.
+Despite the success of model-free reinforcement learning in several domains, sample efficiency remains a limiting factor, a challenge that model-based methods may mitigate.
 		We compare well-suited purely model-based to model-free reinforcement learning applied to the intensity optimisation on the FERMI FEL system.
@@
 The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
-In both cases, the algorithms were implemented in a way, which presents increased noise robustness as omnipresent in accelerator control problems.
+In both cases, the algorithms were implemented in a way, which presents increased noise robustness as prevalent in accelerator control problems.
 Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.
@@
 	In particle accelerator operation, one main goal is to provide stable and reproducible performance.
-Achieving this demands the consideration of several control problems simultaneously \cite{Kain2020,Scheinker2018}.
-Especially if there is no way to model the physics a priori, one might use optimisation techniques as, e.g.
-derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015,Albright2019} or model-based optimisations as Gaussian processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.\\
+Achieving this requires addressing several control problems simultaneously \cite{Kain2020,Scheinker2018}.
+Particularly when a priori physics modeling is infeasible, one may employ optimisation techniques such as derivative-free optimisers (DFOs) \cite{Huang2013,Bruchon2017,Scheinker2020,Hirlaender2019,Welsch2015,Albright2019} or model-based optimisations as Gaussian Processes \cite{Hanuka2020,Roussel2020} to restore or maintain the performance.\\
 	Recently, the community has begun to explore a novel approach to solve control problems.
@@
 		\item It covers a larger class of problems, as RL optimises a sequence of decisions.
-		\item It memorises the problem and does not always begin from zero as a DFO.
+		\item It retains information about the problem structure and does not always begin from zero as a DFO.
 		\item Already existing data can be used.
@@
 	\end{itemize}
-	One demanding aspect using RL in real-world applications is the number of iterations needed to train a controller - the sample-efficiency, since RL methods are known to be data-hungry \cite{Sutton2018,DulacArnold2019}.
-A second critical aspect to be considered is the robustness of the training in a noisy data-limited environment.\\
+	One demanding aspect using RL in real-world applications is the number of iterations needed to train a controller - the sample efficiency, since RL methods are known to be data-intensive \cite{Sutton2018,DulacArnold2019}.
+A second critical aspect to be considered is the robustness of training in a noisy, data-limited environment.\\
 	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}.
@@
 	overview of the Elettra site and the FERMI free-electron laser buildings.\\
-	We successfully applied two different algorithm classes, a highly sample-efficient variant of continuous Q-learning (\cref{ss:Normalized advantage function}) and a newly proposed model-based algorithm, which we call \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), to the FERMI FEL problem.\\ The methods reveal different characteristics addressing the stated critical aspects.
+	We successfully applied two different algorithm classes, a highly sample-efficient variant of continuous Q-learning (\cref{ss:Normalized advantage function}) and a newly proposed model-based algorithm, which we call \emph{AE-DYNA} (\cref{ss:Uncertainty aware DYNA-style reinforcement learning}), to the FERMI FEL problem.\\ The methods exhibit distinct characteristics in addressing the aforementioned critical aspects.
 They are generally applicable to problems showing a sample-complexity as in common accelerator control problems.
@@
 	By making use of model ensembles and Bayesian neural networks our algorithms exhibit a high sample-efficiency (learning using $<1000$ data sets), a low model-bias, and are extremely fast as they are able to maximize the FERMI FEL's output intensity to greater than 95\% of the maximum within only 3-5 steps.
-	 Thus, we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\
+	 Thus, we demonstrate the operational applicability of purely model-based reinforcement learning to an accelerator control problem.\\
 	Furthermore, our algorithms can perform real-time learning based on data and therefore, can be re-run and re-trained for time-varying systems.\\
@@
 	\begin{itemize}
-		\item Description of the problem set-up at FERMI
+		\item Description of the problem setup at FERMI
 		\item Overview of RL
@@
 
-	\section{The set-up of the studied problem}
-	\subsection{The Physical Set-up}
+	\section{The setup of the studied problem}
+	\subsection{The Physical Setup}
 	In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator.\\
@@
 		\input{Figures/SL_Alignment_Scheme}
-		\caption{A schematic view on the set-up of the FERMI FEL.}
+		\caption{A schematic view on the setup of the FERMI FEL.}
 		\label{fig:schematic_FEL}
@@
 	%	The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
-	A schematic overview of the set-up is provided in \cref{fig:schematic_FEL}.
+	A schematic overview of the setup is provided in \cref{fig:schematic_FEL}.
 	Two mirrors, TT1 and TT2, are used to control the trajectory of the laser by tilting and inclining, which gives a total of four degrees of freedom (DOF).
@@
 Two values correspond to the two DOF of each mirror.
-The action $\ba$ is four dimensional, namely a delta step on the voltages at time-step $t$ $\bs_t$:
+The action $\ba$ is four-dimensional, namely a delta step on the voltages at time-step $t$ $\bs_t$:
 	\begin{align}
@@
 However, putting constraints on the algorithms, as discussed in \cref{s:Model-free reinforcement learning}, partly cures the sample inefficiency, turning them into practical tools for real-world problems.\\
-	On the other hand, MBRL showing a much better sample efficiency.
+	On the other hand, MBRL shows much better sample efficiency.
 The challenge for these methods is to learn an accurate model, which turns out to be rather hard and leads to the so-called model-bias problem \cite{Deisenroth2011}.
-In \cref{ss:Uncertainty aware DYNA-style reinforcement learning} we discuss attempts to capture the model uncertainty to alleviate this obstacle by making use of model ensembles and Bayesian neural networks resulting in algorithms, exhibiting a high sample-efficiency and a low model-bias.\\
+In \cref{ss:Uncertainty aware DYNA-style reinforcement learning} we discuss attempts to capture the model uncertainty to alleviate this obstacle by making use of model ensembles and Bayesian neural networks resulting in algorithms, exhibiting a high sample efficiency and a low model-bias.\\
 	In what follows, we denote the set of all states $\bs$ by $\states$ and the set of all actions $\ba$ by $\actions$.
@@
 The tuple \mbox{$\mdp = (\states,\actions,\transitions,\initstate,\reward,\discount)$} defines a Markov decision process.\\
-	Goal of reinforcement learning is to find a $\policy^*: \bs\mapsto \ba$, which is the solution of:
+	The goal of reinforcement learning is to find a $\policy^*: \bs\mapsto \ba$, which is the solution of:
 	\begin{align}
@@
 a deep neural network.\\
-	In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{Sutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}.
+	In the first case $\policy$ is optimised directly and one speaks about policy gradient methods \cite{Sutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}.
 To train these methods, a large number of direct interactions with the system is required, since they belong to \emph{on-policy} algorithms.
@@
 In the following, we discuss the normalised advantage function (\emph{NAF}) algorithm \cite{Gu2016} in some detail, as it has good characteristics for accelerator control problems.
-It is highly sample-efficiency, and its representational power is sufficient for most common cases\cite{Kain2020,Hirlaender2020a}.
+It has high sample efficiency, and its representational power is sufficient for most common cases\cite{Kain2020,Hirlaender2020a}.
 Modifications to increase the stability are subject of \cref{ss:Normalized advantage function}.
@@
 All algorithms are suitable for control tasks as common in accelerator operation.
-Their strength lies in the high sample-efficiency, but they differ in representational power and noise robustness.
-In \cref{sec:Experimental results from FERMI RL online tests} the \emph{NAF2},
+Their strength lies in the high sample efficiency, but they differ in representational power and noise robustness.
+In \cref{sec:Experimental results from FERMI RL online tests}, the \emph{NAF2},
 	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment. The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
@@
 In the upper figure, the total reward per batch gathered from the real environment, and the number of data-points used to train the dynamics model as a function of the number of epochs is shown.
-In the lower plot, the average cumulative reward of ten episodes as achieved by the \emph{TRPO} on the individual models of the ensemble independence of the epochs is drawn.
+In the lower plot, the average cumulative reward of ten episodes as achieved by the \emph{TRPO} on the individual models of the ensemble independently of the epochs is drawn.
 During an epoch the \emph{TRPO} is trained for 10000 steps on the synthetic data.
@@
 The verification was executed as in the first test.
-Again, the success of all 50 episodes is 100\%.
+Again, the success rate of all 50 episodes is 100\%.
 The number of needed iterations per episodes is less than for the ME-TRPO, which can be seen in \cref{fig:AE-DYNA_verification}.
@@
 	In this paper, the applicability of deep reinforcement learning on the optimisation of the FERMI FEL intensity was presented. Two different approaches were tested: model-free and model-based.\\
-	Both experiments yielded satisfactory results showing that a non-linear and noisy problem could be solved in a feasible number of training steps.
+	Both experiments yielded satisfactory results, demonstrating that a non-linear, noisy problem can be solved within a feasible number of training steps.
 The results of the verification are summarised in \cref{tab:overview_verification}.
@@
 The MBRL methods had to be stopped due to lack of available experimental time.
-Hence additional studies regarding a long time performance would be interesting.\\ One big issue in applying MBRL methods can be the computational time to train the controller on the model.
+Hence additional studies regarding a long time performance would be interesting.\\ A significant challenge in applying MBRL methods is the computational time to train the controller on the model.
 In our tests, the data acquisition time was only a fraction of the time needed for the controller-training.
@@
 	\section{Conclusions}
-	The presented reinforcement learning methods hold tremendous promise for automatizing set-ups typical in accelerators.
-One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real set-up by adapting available methods.
-This was demonstrated on the FERMI FEL intensity optimization problem.\\
+	The presented reinforcement learning methods hold tremendous promise for automating setups typical in accelerators.
+One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real setup by adapting available methods.
+This was demonstrated on the FERMI FEL intensity optimisation problem.\\
 	Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA}-style algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency appealing choices.\\
@@
 As numerous cases in accelerator control can be captured assuming a quadratic dependence on the actions of the state-action value function, there is a broad spectrum of potential applications.\\
-	To provide the possibility for other laboratories to profit form the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
+	To provide the possibility for other laboratories to profit from the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.
 

### 2025-12-29 paper/main.tex
refactor: Relocate experiment data to a dedicated `data/` directory and add paper assets.

--- 2025-12-29 main.tex
@@
 The model-based algorithm is implemented in a DYNA-style using an uncertainty aware model, and the model-free algorithm is based on tailored deep Q-learning.
-In both cases, the algorithms were implemented in a way, which presents increased noise robustness as prevalent in accelerator control problems.
+In both cases, the algorithms were implemented in a manner that presents increased noise robustness as prevalent in accelerator control problems.
 Code is released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper}.
@@
 MFRL algorithms learn directly from the interaction with a system, while MBRL algorithms learn a dynamics model from these interactions first.
-Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, their applications is mainly restricted to simulated environments due to their low sample-efficiency.
+Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, their application is mainly restricted to simulated environments due to their low sample-efficiency.
 However, putting constraints on the algorithms, as discussed in \cref{s:Model-free reinforcement learning}, partly cures the sample inefficiency, turning them into practical tools for real-world problems.\\
@@
 One way is to take the mean of the models $\mu_m$ and the standard deviation $\sigma_m$ and sample from $\gauss(\mu_m,\sigma_m)$.
-Another strategy is to randomly select a model to provide at each training step, which demonstrated the best performance the original implementation of the model ensemble \emph{TRPO} (\emph{ME-TRPO}) algorithm and was used in the experiments labelled \emph{ME-TRPO} \cite{Kurutach2018}.
+Another strategy is to randomly select a model to provide at each training step, which demonstrated the best performance in the original implementation of the model ensemble \emph{TRPO} (\emph{ME-TRPO}) algorithm and was used in the experiments labelled \emph{ME-TRPO} \cite{Kurutach2018}.
 A pessimistic selection would only select the model resulting in the lowest predicted reward (as used, e.g.
@@
 	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment. The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
-	\begin{table}[h!]%The best place to locate the table environment is directly after its first reference in text
+	\begin{table*}[t]%The best place to locate the table environment is directly after its first reference in text
 		\caption{\label{tab:overview_algorithms}%
@@
 		}
-		\begin{ruledtabular}
-			\begin{tabular}{l p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
+			\begin{tabular}{lcccc}
+				\hline\hline
 				\textrm{Algorithm}&
 				\textrm{Class}&
-				\textrm{Noise \newline robustness}&
+				\textrm{Noise robustness}&
 				\textrm{Represent. power}&
-				\textrm{Sample \newline efficiency}\\
-				\colrule
+				\textrm{Sample efficiency}\\
+				\hline
 				\emph{NAF} & MFRL & low &low &high\\
@@
 				\emph{AE-DYNA} & MBRL & high& high& high\\
+				\hline\hline
 			\end{tabular}
-		\end{ruledtabular}
-	\end{table}
+	\end{table*}
 
@@
 In case, the used methods could be parallelised to reduce the computational time.
-	\begin{table}[!h]%The best place to locate the table environment is directly after its first reference in text
+	\begin{table*}[t]%The best place to locate the table environment is directly after its first reference in text
 		\caption{\label{tab:overview_verification}%
@@
 		}
-		\begin{ruledtabular}
-			\begin{tabular}{l>{\raggedleft} p{1cm}>{\raggedleft}p{1.8cm}>{\raggedleft} p{1.8cm} >{\raggedleft\arraybackslash} p{1.8cm}}
-				%\toprule
-				&  \centering Data points (counts)&   \centering Episode length (counts)&   \centering Cumulative reward\newline (arb. units)&   \centering Final\\reward \newline (arb. units)\tabularnewline
+			\begin{tabular}{lrrrr}
+				\hline\hline
+				&  \centering Data points (counts)&   \centering Episode length (counts)&   \centering Cumulative reward (arb. units)&   \centering Final reward (arb. units)\tabularnewline
 				\hline
@@
 				%naf\_single & nr &  2.56 &  1.26 \\
+				\hline\hline
 			\end{tabular}
-		\end{ruledtabular}
-	\end{table}
+	\end{table*}
 	\section{Conclusions}
@@
 	\subsubsection{Regression assuming homoskedastic Gaussian noise using `anchored ensembling'}
-	The Bayesian community and the RL community is spending more and more attention to
-	new approaches to Bayesian inference. In the `anchored ensembling' technique a regularisation
-	term is added to the loss function, which returns a point estimate of
-	the Bayesian posterior the maximum a posteriori (MAP) parameter estimate. 
-	Using an ensemble of networks and adding noise on either the targets of the regularisation term a distribution of MAP solutions  \cite{Gu2007, Chen2011, Bardsley2012, Pearce2018} is obtained, which mimics true posterior.\\
+	The Bayesian and RL communities are paying increasing attention to new approaches for Bayesian inference.
+In the `anchored ensembling' technique, a regularization term is added to the loss function, which yields a point estimate of the Bayesian posterior, specifically the maximum a posteriori (MAP) parameter estimate.
+	By employing an ensemble of networks and introducing noise to the regularization term's targets, a distribution of MAP solutions \cite{Gu2007, Chen2011, Bardsley2012, Pearce2018} is obtained, which mimics the true posterior.\\
 	Consider an artificial neural net containing parameters, ${\theta}$, making predictions for the dynamics and the reward \cref{eq:dynamics_model}, $\hat{f}_\theta$ ($f$ denotes the true unknown function), with $N$ data-points.
@@
 A number of three models (label \emph{Three}) shows a good trade-off between performance and training time.
-A single network might not converge or too slowly (label \emph{Single}) and a model of ten shown fast and stable performance in this case (label \emph{Ten}).
+A single network might not converge or too slowly (label \emph{Single}) and a model of ten showed fast and stable performance in this case (label \emph{Ten}).
 	\begin{figure}[h]
@@
 	Finally, \cref{fig:comparsion_NAF_AE-DYNA} demonstrates the sample-efficiency of the \emph{AE-DYNA-SAC} and the \emph{NAF} algorithm on the noisy \emph{inverted pendulum}.
-\emph{AE-DYNA-SAC} converges below 2000 data-points (without noise even below 800) and the \emph{NAF2} starts to perform equally 10000 data-points surpassing the -200 reward threshold.
+\emph{AE-DYNA-SAC} converges below 2000 data-points (without noise even below 800) and the \emph{NAF2} starts to perform equally after 10000 data-points surpassing the -200 reward threshold.
 One clearly sees the increased sample-efficiency on this problem using the \emph{AE-DYNA} in contrast to the \emph{NAF2}.

### 2025-12-29 paper/main.tex
docs: Clarify anchored ensembling in paper, update README abstract wording, image paths, and table formatting.

@@
 In the `anchored ensembling' technique, a regularization term is added to the loss function, which yields a point estimate of the Bayesian posterior, specifically the maximum a posteriori (MAP) parameter estimate.
-	By employing an ensemble of networks and introducing noise to the regularization term's targets, a distribution of MAP solutions \cite{Gu2007, Chen2011, Bardsley2012, Pearce2018} is obtained, which mimics the true posterior.\\
+	By employing an ensemble of networks and introducing noise to the regularization term's targets, a distribution of MAP solutions \cite{Gu2007, Chen2011, Bardsley2012, Pearce2018} is obtained, which mimics the true posterior.
+This approach addresses a key criticism of standard deep ensembles: that randomly initialized networks merely find diverse local optima rather than sampling from the true posterior distribution.
+The regularization term anchoring the weights to a prior draw provides the theoretical linkage to Bayesian inference that standard ensembles lack.\\
 	Consider an artificial neural net containing parameters, ${\theta}$, making predictions for the dynamics and the reward \cref{eq:dynamics_model}, $\hat{f}_\theta$ ($f$ denotes the true unknown function), with $N$ data-points.
